---
title: "`r paste0(study_label, ' — DRSP Cyclic GAM Reports')`"
author: "Generated by DRSP GAM Template"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

> **Template Purpose:** Automated, participant-centric analysis of DRSP symptom
> cyclicity using cyclic GAMs. This template works with any dataset containing
> some or all DRSP items (drsp_1 through drsp_21).
>
> 1. **Setup & Preparation** — load data, auto-detect DRSP items, compute composites, define outcomes.
> 2. **Automated Reporting** — one landscape PDF per participant with compact cover and paged plots.
> 3. **Overall Summary** — sample-level summary PDF with key metrics and visualizations.
> 4. **CSV Audit** — per-ID × outcome × time-var stats for triage/QC.

---

# Step 0: Load Configuration
```{r load-config}
# Source the user configuration file — EDIT drsp_gam_config.R BEFORE running.
source("drsp_gam_config.R")
```

# Step 1: Libraries
```{r libraries, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(ggtext)
library(mgcv)
library(patchwork)
library(gridExtra)
library(grid)
library(tibble)
library(tidyr)
library(scales)
library(glue)

options(scipen = 999)
set.seed(12345)
```

# Step 2: Load Template Functions
```{r load-functions}
source("drsp_gam_scoring.R")
source("drsp_gam_functions.R")
```

# Step 3: Load Data
```{r load-data}
# Load dataset
if (grepl("\\.rds$", data_file_path, ignore.case = TRUE)) {
  df <- readRDS(data_file_path)
} else if (grepl("\\.csv$", data_file_path, ignore.case = TRUE)) {
  df <- read.csv(data_file_path, stringsAsFactors = FALSE)
} else {
  stop("Unsupported file format. Please provide an .rds or .csv file.")
}

stopifnot("id" %in% names(df))
message(sprintf("Loaded data: %d rows, %d participants", nrow(df), length(unique(df$id))))
```

# Step 4: Score DRSP Composites & Build Outcome Map
```{r score-composites}
# Auto-detect DRSP items, score composites, build outcome map
scoring_result <- score_drsp_composites(df)
df                  <- scoring_result$data
items_found         <- scoring_result$items_found
items_missing       <- scoring_result$items_missing
composites_created  <- scoring_result$composites_created
composites_skipped  <- scoring_result$composites_skipped
composite_specs     <- scoring_result$composite_specs

# Build the named outcome vector (composites first, then individual items)
outcomes <- build_outcomes_map(items_found, composites_created)
message(sprintf("Total outcomes to model: %d (%d composites + %d individual items)",
                length(outcomes), length(composites_created), length(items_found)))

# Classify into affective vs physical
domain_info <- classify_outcomes(items_found, composite_specs)
physical_symptoms      <- domain_info$physical_items
all_affective_outcomes <- domain_info$all_affective
all_physical_outcomes  <- domain_info$all_physical
```

# Step 5: Setup Output, Colors, Axis Metadata
```{r setup-output}
# --- Create dated output folder ---
current_date <- format(Sys.Date(), "%Y%m%d")
output_folder <- file.path(gam_output_base, paste0("GAM_Reports_", current_date))
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)

# --- Verify time variables exist ---
time_vars <- intersect(time_vars_to_run, names(df))
if (length(time_vars) == 0) {
  stop("None of the specified time variables were found in the data: ",
       paste(time_vars_to_run, collapse = ", "))
}
if (length(time_vars) < length(time_vars_to_run)) {
  missing_tvs <- setdiff(time_vars_to_run, time_vars)
  warning("Time variables not found (skipped): ", paste(missing_tvs, collapse = ", "))
}

# --- Build color map ---
drsp_color_map <- build_drsp_color_map(outcomes)

# --- Build axis metadata ---
axis_meta <- build_axis_meta(time_vars, default_cycle_phasing)

# --- Build prediction grids ---
pred_grid_list <- build_pred_grids(time_vars)

# --- Build outcome plot settings ---
outcome_plot_settings <- build_outcome_plot_settings(outcomes)

# --- Normalization scale ranges ---
norm_scale_ranges <- outcome_plot_settings %>%
  dplyr::mutate(
    norm_min = dplyr::if_else(!is.na(y_min), y_min, NA_real_),
    norm_max = dplyr::if_else(!is.na(y_max), y_max, NA_real_)
  ) %>%
  dplyr::select(variable, norm_min, norm_max)

# --- DRSP variable pattern ---
DRSP_VAR_PATTERN <- "^drsp_\\d+$"
all_raw_drsp <- names(outcomes)[grepl(DRSP_VAR_PATTERN, names(outcomes))]

# --- Border key caption ---
BORDER_KEY_CAPTION <- "Border key:  \u25A0 Red = GAM range meets person-specific 30% change benchmark  |  No border = below benchmark"

# --- Max possible counts per category ---
n_physical_max <- length(intersect(all_raw_drsp, physical_symptoms))
n_affect_max   <- length(setdiff(all_raw_drsp, physical_symptoms))

# --- Outcome categories for page organization ---
outcome_categories <- tibble::tibble(
  variable = names(outcomes),
  outcome_label = unname(outcomes)
) %>%
  dplyr::mutate(
    category = dplyr::case_when(
      !grepl(DRSP_VAR_PATTERN, variable) ~ "DRSP Composites",
      TRUE ~ "Mood & Physical (DRSP)"
    )
  )

# --- Page definitions (composites page + pages of 9 individual items) ---
page_definitions <- c(
  list(
    Composites_Page = outcome_categories %>%
      dplyr::filter(category == "DRSP Composites") %>%
      dplyr::pull(outcome_label)
  ),
  {
    indiv_labels <- outcome_categories %>%
      dplyr::filter(category == "Mood & Physical (DRSP)") %>%
      dplyr::pull(outcome_label)
    if (length(indiv_labels) > 0) {
      split(indiv_labels, ceiling(seq_along(indiv_labels) / 9))
    } else {
      list()
    }
  }
)
if (length(page_definitions) > 1) {
  names(page_definitions)[2:length(page_definitions)] <-
    paste0("DRSP_Page_", seq_len(length(page_definitions) - 1))
}

# --- Pre-calculate log-transformations ---
for (var in names(outcomes)) {
  if (var %in% names(df)) {
    log_var_name <- paste0(var, "_Log")
    df[[log_var_name]] <- log(df[[var]] + 1)
  }
}
```

# Step 6: Per-ID Participant Reports
```{r per-id-reports, results='hide', message=FALSE}
if (!generate_per_id_reports) {
  message("Skipping per-ID reports (generate_per_id_reports = FALSE)")
} else {

report_folder <- file.path(output_folder, "Participant_Reports_PDF_byID")
dir.create(report_folder, showWarnings = FALSE, recursive = TRUE)

ids_to_run    <- unique(df$id)
all_ids_signif <- list()
id_global_ymax_map <- c()
id_benchmark_map   <- c()
max_plots_per_page <- plots_per_col * plots_per_row

table_theme <- gridExtra::ttheme_default(
  base_size = 7,
  core    = list(fg_params = list(hjust = 0, x = 0.05)),
  colhead = list(fg_params = list(hjust = 0, x = 0.05)),
  padding = grid::unit(c(3, 2), "mm")
)

for (current_id in ids_to_run) {
  message(sprintf("--- Processing ID: %s ---", current_id))
  datSX_id <- subset(df, id == current_id)

  # --- Compute person-specific C-PASS-equivalent change benchmark ---
  drsp_item_cols <- grep("^drsp_\\d+$", names(datSX_id), value = TRUE)
  all_drsp_vals <- unlist(datSX_id[, drsp_item_cols, drop = FALSE], use.names = FALSE)
  all_drsp_vals <- all_drsp_vals[is.finite(all_drsp_vals)]
  id_scale_range <- if (length(all_drsp_vals) > 0) {
    max(all_drsp_vals) - min(all_drsp_vals)
  } else {
    NA_real_
  }
  id_benchmark <- if (!is.na(id_scale_range)) id_scale_range * cpass_change_pct else NA_real_
  id_benchmark_map[as.character(current_id)] <- id_benchmark
  message(sprintf("    Scale range: %.1f, Benchmark: %.2f", id_scale_range, id_benchmark))

  # ---- Open landscape PDF per participant ----
  pdf_filename <- file.path(report_folder, sprintf("Report_ID_%s.pdf", current_id))
  if (isTRUE(skip_existing_reports) && file.exists(pdf_filename)) {
    message("    -> File exists; skipping.")
    next
  }
  grDevices::pdf(pdf_filename, width = pdf_width, height = pdf_height)

  # ---- Page 0: Raw data coverage ----
  if (length(time_vars) > 0) {
    tv0 <- time_vars[1]
    meta0 <- axis_meta[[tv0]]

    make_raw_panel <- function(dat, tv, meta, var, y_lab, color) {
      if (!var %in% names(dat)) {
        p <- ggplot2::ggplot() +
          ggplot2::annotate("text", x = 0, y = 0, label = paste0(var, " not in data"),
                            size = 5, color = "grey50") +
          ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = "n = 0 daily observations") +
          ggplot2::theme_minimal(base_size = 12)
        return(p)
      }
      vals <- dat[[var]]
      tv_vals <- dat[[tv]]
      valid <- !is.na(vals) & !is.na(tv_vals)
      n_obs <- sum(valid)
      qcounts <- quarter_bin_counts(tv_vals[valid])
      sub_label <- paste0(
        "n = ", n_obs, " daily observations",
        "  (Q1: ", qcounts["Q1"], ", Q2: ", qcounts["Q2"],
        ", Q3: ", qcounts["Q3"], ", Q4: ", qcounts["Q4"], ")"
      )
      p <- ggplot2::ggplot(dat, ggplot2::aes(x = .data[[tv]], y = .data[[var]])) +
        ggplot2::geom_vline(xintercept = c(-0.5, 0, 0.5), color = "black",
                            linewidth = 0.4, linetype = "solid") +
        ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                    labels = meta$labels) +
        ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = sub_label) +
        ggplot2::theme_minimal(base_size = 12)
      if (n_obs > 0) {
        p <- p + ggplot2::geom_point(color = color, alpha = 0.45, size = 1.8)
      } else {
        p <- p + ggplot2::annotate("text", x = 0, y = 0, label = "No data",
                                   size = 5, color = "grey50")
      }
      p
    }

    # Show coverage for last DRSP item available and first composite (or first item)
    coverage_var1 <- if ("drsp_21" %in% names(datSX_id)) "drsp_21" else tail(items_found, 1)
    coverage_var2 <- if ("drsp_NA" %in% names(datSX_id)) "drsp_NA" else {
      if (length(composites_created) > 0) composites_created[1] else items_found[1]
    }

    p_panel1 <- make_raw_panel(datSX_id, tv0, meta0, coverage_var1,
                               outcomes[coverage_var1], "#CC6677") +
      ggplot2::coord_cartesian(ylim = c(1, 6))
    p_panel2 <- make_raw_panel(datSX_id, tv0, meta0, coverage_var2,
                               outcomes[coverage_var2], "#D55E00") +
      ggplot2::coord_cartesian(ylim = c(1, 6))

    tv0_vals <- datSX_id[[tv0]]
    tv0_vals <- tv0_vals[!is.na(tv0_vals)]
    tv0_qcounts <- quarter_bin_counts(tv0_vals)
    tv0_sparse  <- is_quarter_sparse(tv0_qcounts, sparse_quarter_cutoff)

    coverage_subtitle <- paste0("Time variable: ", tv0)
    if (tv0_sparse) {
      n_empty_q  <- sum(tv0_qcounts == 0)
      n_sparse_q <- sum(tv0_qcounts > 0 & tv0_qcounts < sparse_quarter_cutoff)
      if (n_empty_q > 0) {
        warn_detail <- paste0(n_empty_q, " quarter(s) with 0 observations")
      } else {
        warn_detail <- paste0(n_sparse_q, " quarter(s) with <", sparse_quarter_cutoff,
                              " observations (max 1 allowed)")
      }
      coverage_subtitle <- paste0(
        coverage_subtitle,
        "\n\u26A0 Insufficient coverage for ", tv0, ": ", warn_detail
      )
    }

    p_cover <- p_panel1 / p_panel2 +
      patchwork::plot_annotation(
        title    = paste0("Raw Data Coverage \u2014 ID: ", current_id),
        subtitle = coverage_subtitle,
        theme    = ggplot2::theme(
          plot.title    = ggplot2::element_text(face = "bold", size = 12),
          plot.subtitle = ggplot2::element_text(
            size = if (tv0_sparse) 10 else 9,
            color = if (tv0_sparse) "red" else "grey30"
          )
        )
      )
    print(p_cover)

    # ---- Page 0b: Interpretation Key ----
    key_title <- grid::textGrob(
      paste0("How to Read This Report \u2014 ID: ", current_id),
      gp = grid::gpar(fontsize = 16, fontface = "bold")
    )
    key_body <- grid::textGrob(
      paste0(
        "VISUAL ELEMENTS\n\n",
        " \u25B6  Pink shaded region: Area above 4.5 on the y-axis, shown when the axis expands\n",
        "     beyond the default 1\u20134.5 range.\n\n",
        " \u25B6  Red dashed line:   Menses onset (cycle day 1)\n\n",
        " \u25B6  Green dashed line: Estimated ovulation\n\n",
        " \u25B6  Red horizontal line (at y = 4): Reference line of \"moderate\" severity\n\n",
        " \u25B6  Dots on GAM curve: Peak (highest) and trough (lowest) predicted values\n\n",
        " \u25B6  Faded points:     Raw daily data behind the GAM curve\n\n",
        " \u25B6  Annotation text:  Stars (*) indicate significance (p < .05); Range = max \u2212 min\n",
        "                     of predicted values; EDF = effective degrees of freedom\n\n",
        "PLOT BORDER COLORS (individual DRSP item plots)\n\n",
        " \u25B6  Red border:     GAM range meets this person\u2019s C-PASS-equivalent 30% change\n",
        "                     benchmark (", sprintf("%.2f", id_benchmark), " = scale range ",
        sprintf("%.0f", id_scale_range), " \u00d7 30%)\n\n",
        " \u25B6  No border:      GAM range below person-specific benchmark"
      ),
      gp = grid::gpar(fontsize = 9.5, lineheight = 1.1, fontfamily = "mono"),
      just = "left", x = 0.05
    )
    gridExtra::grid.arrange(key_title, key_body, ncol = 1,
                            heights = grid::unit(c(0.08, 0.92), "npc"))

    # ---- Page 0c: Composite Outcome Definitions ----
    if (length(composite_specs) > 0) {
      comp_title <- grid::textGrob(
        paste0("Composite Outcome Definitions \u2014 ID: ", current_id),
        gp = grid::gpar(fontsize = 16, fontface = "bold")
      )
      comp_lines <- "The following composite scores are computed as the row-wise mean\n(rowMeans, na.rm = TRUE) of their component DRSP items.\n\n"
      for (domain in c("AFFECTIVE", "PHYSICAL")) {
        comp_lines <- paste0(comp_lines, "\n", domain, " COMPOSITES\n\n")
        for (cname in names(composite_specs)) {
          spec <- composite_specs[[cname]]
          if (spec$domain != domain) next
          items_str <- paste(spec$components_display, collapse = ", ")
          wrapped <- strwrap(paste0("= mean of: ", items_str),
                             width = 90, indent = 5, exdent = 17)
          comp_lines <- paste0(
            comp_lines,
            " \u25B6  ", spec$label, " (", cname, ")\n",
            paste(wrapped, collapse = "\n"), "\n\n"
          )
        }
      }
      comp_body <- grid::textGrob(
        comp_lines,
        gp = grid::gpar(fontsize = 10, lineheight = 1.15, fontfamily = "mono"),
        just = "left", x = 0.05
      )
      gridExtra::grid.arrange(comp_title, comp_body, ncol = 1,
                              heights = grid::unit(c(0.08, 0.92), "npc"))
    }

    # ---- Page 0d: Score Distribution by Item (boxplots) ----
    box_vars <- names(outcomes)[names(outcomes) %in% names(datSX_id)]
    if (length(box_vars) > 0) {
      box_data <- tidyr::pivot_longer(
        datSX_id[, c("id", box_vars), drop = FALSE],
        cols      = dplyr::all_of(box_vars),
        names_to  = "variable",
        values_to = "score"
      ) %>%
        dplyr::filter(!is.na(score)) %>%
        dplyr::mutate(
          label  = outcomes[variable],
          domain = ifelse(variable %in% c(physical_symptoms,
                                          domain_info$physical_composites),
                          "Physical", "Affective")
        )

      if (nrow(box_data) > 0) {
        label_order <- box_data %>%
          dplyr::group_by(label) %>%
          dplyr::summarise(med = median(score, na.rm = TRUE), .groups = "drop") %>%
          dplyr::arrange(med) %>%
          dplyr::pull(label)
        box_data$label <- factor(box_data$label, levels = label_order)

        p_box <- ggplot2::ggplot(box_data,
                                  ggplot2::aes(x = label, y = score, fill = domain)) +
          ggplot2::geom_boxplot(width = 0.6, outlier.size = 1, alpha = 0.7) +
          ggplot2::coord_flip() +
          ggplot2::scale_fill_manual(
            values = c("Affective" = DOMAIN_COL_AFFECTIVE,
                       "Physical"  = DOMAIN_COL_PHYSICAL),
            name = NULL
          ) +
          ggplot2::scale_y_continuous(breaks = 1:6, limits = c(1, 6)) +
          ggplot2::labs(
            title    = paste0("Score Distribution by Item \u2014 ID: ", current_id),
            subtitle = "Boxplots of raw daily scores (1\u20136 DRSP scale); ordered by median score",
            x = NULL, y = "Daily Score",
            caption  = "Blue = affective/emotional; orange = physical/somatic. Composites are row-means of component items."
          ) +
          ggplot2::theme_minimal(base_size = 10) +
          ggplot2::theme(
            plot.title   = ggplot2::element_text(face = "bold", size = 13),
            plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
            legend.position = "bottom"
          )
        print(p_box)
      }
    }

  } else {
    grid::grid.newpage()
    grid::grid.text(
      label = paste0(
        "Raw Data Coverage \u2014 ID: ", current_id,
        "\n\nNo eligible time variables were available for GAM fitting."
      ),
      x = 0.5, y = 0.55, gp = grid::gpar(fontsize = 12, fontface = "bold")
    )
    grDevices::dev.off()
    next
  }

  # ---- Pass 1: fit & collect ----
  preds_by_tv <- list()
  sig_by_tv   <- list()

  for (tv in time_vars) {
    message(sprintf("  ... time variable: %s", tv))
    tv_vals_all <- datSX_id[[tv]]
    tv_vals_all <- tv_vals_all[!is.na(tv_vals_all)]
    tv_qcounts  <- quarter_bin_counts(tv_vals_all)
    if (is_quarter_sparse(tv_qcounts, sparse_quarter_cutoff)) {
      message(sprintf("    -> Skipping tv=%s: insufficient quarter coverage", tv))
      next
    }

    all_preds <- list()
    sig_list  <- list()

    for (outcome_var in names(outcomes)) {
      res <- fit_outcome(datSX_id, outcome_var, tv, outcomes,
                         pred_grid_list, sparse_cutoff = sparse_quarter_cutoff)
      if (is.null(res)) next
      all_preds[[length(all_preds) + 1]] <- res$pred
      sig_list[[length(sig_list) + 1]]  <- res$sig
    }

    if (length(all_preds) > 0) {
      preds_by_tv[[tv]] <- dplyr::bind_rows(all_preds)
      sig_by_tv[[tv]]   <- dplyr::bind_rows(sig_list)
    }
  }

  if (!length(preds_by_tv)) {
    message("    -> No successful models for this ID; skipping.")
    grDevices::dev.off()
    next
  }

  # ---- Compute global ymax ----
  global_max_val <- 4.5
  for (tv_name in names(preds_by_tv)) {
    est_vals <- preds_by_tv[[tv_name]]$estimate
    finite_est <- est_vals[is.finite(est_vals)]
    if (length(finite_est) > 0) global_max_val <- max(global_max_val, max(finite_est))
  }
  id_global_ymax <- expand_ymax(4.5, global_max_val)
  id_global_ylimits <- c(1, id_global_ymax)
  id_global_ymax_map[as.character(current_id)] <- id_global_ymax

  sig_all <- dplyr::bind_rows(sig_by_tv)
  sig_all$id <- current_id
  all_ids_signif[[as.character(current_id)]] <- sig_all

  # ---- COVER: Compact counts by time_var ----
  sig_all_clean <- sig_all %>%
    dplyr::mutate(
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    )

  cover_counts <- sig_all_clean %>%
    dplyr::group_by(time_var) %>%
    dplyr::summarise(
      tested = dplyr::n(),
      n_sig  = sum(stars != "", na.rm = TRUE),
      n_cpass = sum(is.finite(range) & range >= id_benchmark & stars != "", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::unite("count", n_sig, tested, sep = " / ") %>%
    dplyr::rename(`Sig / Tested` = count,
                  !!paste0("Range \u2265", sprintf("%.2f", id_benchmark), " sig") := n_cpass)

  cover_theme <- gridExtra::ttheme_default(
    base_size = 9,
    core = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"),
                padding = grid::unit(c(2, 2), "mm")),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
  )

  title_grob    <- grid::textGrob(paste("Compact Summary \u2014 ID:", current_id),
                                  gp = grid::gpar(fontsize = 16, fontface = "bold"))
  subtitle_grob <- grid::textGrob("Rows: # significant / # tested (p < .05) per time variable",
                                  gp = grid::gpar(fontsize = 9))
  cover_grob    <- gridExtra::tableGrob(cover_counts, rows = NULL, theme = cover_theme)
  gridExtra::grid.arrange(title_grob, subtitle_grob, cover_grob,
                          ncol = 1, heights = grid::unit(c(0.12, 0.08, 0.80), "npc"))

  # ---- Stars matrix (sig outcomes only) ----
  compact_stars <- sig_all_clean %>%
    dplyr::group_by(variable) %>%
    dplyr::mutate(any_sig = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(any_sig) %>%
    dplyr::mutate(Outcome = dplyr::coalesce(outcome_label, variable)) %>%
    dplyr::select(Outcome, time_var, stars) %>%
    tidyr::pivot_wider(names_from = time_var, values_from = stars, values_fill = "") %>%
    dplyr::arrange(Outcome)

  if (nrow(compact_stars) > 0) {
    compact_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(fg_params = list(hjust = 0, x = 0.02),
                  padding = grid::unit(c(1.5, 2), "mm")),
      colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
    )
    header <- grid::textGrob("Significant Outcomes Only (\u2605 = p < .05)",
                             gp = grid::gpar(fontsize = 12, fontface = "bold"))
    tbl <- gridExtra::tableGrob(compact_stars, rows = NULL, theme = compact_theme)
    gridExtra::grid.arrange(header, tbl, ncol = 1, heights = grid::unit(c(0.10, 0.90), "npc"))
  }

  # ---- Significant EDFs ranked by Range ----
  sig_ranked <- sig_all_clean %>%
    dplyr::filter(is.finite(p.value) & p.value < 0.05) %>%
    dplyr::mutate(Outcome = dplyr::coalesce(outcome_label, variable)) %>%
    dplyr::transmute(
      Outcome, TimeVar = time_var,
      EDF = round(edf, 2), Range = round(range, 2),
      F = round(F, 2), N = n
    ) %>%
    dplyr::arrange(dplyr::desc(Range), Outcome, TimeVar)

  if (nrow(sig_ranked) == 0) {
    header <- grid::textGrob(
      paste("Significant Smooths \u2014 ID:", current_id, "(none at p < .05)"),
      gp = grid::gpar(fontsize = 14, fontface = "bold"))
    sub <- grid::textGrob("No significant cyclic effects detected.",
                          gp = grid::gpar(fontsize = 10, fontface = "italic"))
    gridExtra::grid.arrange(header, sub, ncol = 1, heights = grid::unit(c(0.2, 0.8), "npc"))
  } else {
    ranked_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(fg_params = list(hjust = 0, x = 0.02),
                  padding = grid::unit(c(1.2, 2), "mm")),
      colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
    )
    title_grob <- grid::textGrob(
      paste("Significant Smooths (p < .05), Ranked by Range \u2014 ID:", current_id),
      gp = grid::gpar(fontsize = 14, fontface = "bold"))
    note_grob <- grid::textGrob(
      "Columns: EDF = effective degrees of freedom; Range = max(pred) \u2212 min(pred) on original scale.",
      gp = grid::gpar(fontsize = 9, fontface = "italic"))
    ranked_tbl <- gridExtra::tableGrob(sig_ranked, rows = NULL, theme = ranked_theme)
    gridExtra::grid.arrange(title_grob, note_grob, ranked_tbl,
                            ncol = 1, heights = grid::unit(c(0.12, 0.06, 0.82), "npc"))
  }

  # ---- Overlaid GAM pages ----
  for (tv in names(preds_by_tv)) {
    # Individual DRSP items overlay
    plot_overlay_page(
      sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
      tv = tv, meta = axis_meta[[tv]], outcomes = outcomes,
      drsp_color_map = drsp_color_map,
      var_list = all_raw_drsp,
      range_thresh = id_benchmark,
      title_prefix = "Overlaid GAM Curves (Individual DRSP Items)",
      subtitle_label = "Individual DRSP items",
      current_id = current_id,
      ylim_strategy = "fixed", fixed_ylim = id_global_ylimits,
      show_hline = TRUE, hline_y = 4,
      legend_ncol = 4, show_range_in_labels = TRUE
    )

    # Composite outcomes overlay
    composite_vars <- names(outcomes)[!grepl(DRSP_VAR_PATTERN, names(outcomes))]
    if (length(composite_vars) > 0) {
      plot_overlay_page(
        sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
        tv = tv, meta = axis_meta[[tv]], outcomes = outcomes,
        drsp_color_map = drsp_color_map,
        var_list = composite_vars,
        range_thresh = Inf,
        title_prefix = "Overlaid GAM Curves (DRSP Composites)",
        subtitle_label = "DRSP composite outcomes",
        current_id = current_id,
        ylim_strategy = "fixed", fixed_ylim = id_global_ylimits,
        show_hline = FALSE, legend_ncol = 3, show_range_in_labels = TRUE
      )
    }
  }

  # ---- Per time-var: significance table + plot pages ----
  for (tv in names(preds_by_tv)) {
    summary_df <- sig_by_tv[[tv]] %>%
      dplyr::left_join(dplyr::select(outcome_categories, variable, category), by = "variable") %>%
      dplyr::mutate(
        stars = dplyr::case_when(
          is.finite(p.value) & p.value < 0.001 ~ "***",
          is.finite(p.value) & p.value < 0.01  ~ "**",
          is.finite(p.value) & p.value < 0.05  ~ "*",
          TRUE ~ ""
        ),
        Outcome = dplyr::coalesce(outcome_label, variable)
      ) %>%
      dplyr::select(Category = category, Outcome, N = n, EDF = edf, F = F, Range = range, Sig = stars) %>%
      dplyr::arrange(Category, Outcome)

    tv_title <- grid::textGrob(paste("Significance \u2014 ID:", current_id, "| Time Variable:", tv),
                               gp = grid::gpar(fontsize = 14, fontface = "bold"))
    tv_table <- gridExtra::tableGrob(summary_df, rows = NULL, theme = table_theme)
    gridExtra::grid.arrange(tv_title, tv_table, ncol = 1,
                            heights = grid::unit(c(0.12, 0.88), "npc"))

    combined_plot_data <- preds_by_tv[[tv]]

    for (page_name in names(page_definitions)) {
      page_outcome_labels <- page_definitions[[page_name]]
      page_data <- dplyr::filter(combined_plot_data, outcome_label %in% page_outcome_labels)
      if (nrow(page_data) == 0) next

      plot_list <- lapply(page_outcome_labels, function(lab) {
        plot_df <- dplyr::filter(page_data, outcome_label == lab)
        if (nrow(plot_df) == 0) return(NULL)
        create_annotated_plot(plot_df, tv, axis_meta, outcome_plot_settings,
                              drsp_color_map, y_limits = id_global_ylimits,
                              raw_data = datSX_id, benchmark = id_benchmark)
      })
      plot_list <- Filter(Negate(is.null), plot_list)

      if (length(plot_list) > 0) {
        if (length(plot_list) < max_plots_per_page) {
          plot_list <- c(plot_list,
                         rep(list(patchwork::plot_spacer()), max_plots_per_page - length(plot_list)))
        }
        page_title_label <- gsub("_", " ", page_name)
        page_title <- paste(page_title_label, "\u2014 ID:", current_id, "| Time Variable:", tv)

        is_drsp_page <- grepl("^DRSP_Page_", page_name)
        border_key_caption <- if (is_drsp_page) BORDER_KEY_CAPTION else NULL

        final_page <- patchwork::wrap_plots(plotlist = plot_list, ncol = plots_per_col) +
          patchwork::plot_annotation(
            title = page_title,
            caption = border_key_caption,
            theme = ggplot2::theme(
              plot.title = ggplot2::element_text(size = 12, face = "bold"),
              plot.caption = ggplot2::element_text(size = 8, hjust = 0.5, color = "grey30")
            )
          )
        print(final_page)
      }
    }
  }

  grDevices::dev.off()
  message(sprintf("    --- \u2705 Saved Report: %s ---", pdf_filename))
}

# ---- Write CSV summary ----
if (generate_csv_summary && length(all_ids_signif)) {
  eval_csv <- file.path(output_folder, paste0("participant_significance_summary_", current_date, ".csv"))
  write.csv(dplyr::bind_rows(all_ids_signif), eval_csv, row.names = FALSE)
  message(sprintf("Wrote summary CSV: %s", eval_csv))
}

} # end if generate_per_id_reports
```

# Step 7: Overall Sample Summary Report
```{r overall-summary, results='hide', message=FALSE}
if (!generate_overall_summary) {
  message("Skipping overall summary (generate_overall_summary = FALSE)")
} else if (!exists("all_ids_signif") || length(all_ids_signif) == 0) {
  message("No per-ID results available; skipping overall summary.")
} else {

gam_summary_all <- dplyr::bind_rows(all_ids_signif)

# Attach person-specific benchmark
benchmark_df <- data.frame(
  id = names(id_benchmark_map),
  id_benchmark = unname(id_benchmark_map),
  stringsAsFactors = FALSE
)
gam_summary_all <- dplyr::left_join(
  gam_summary_all %>% dplyr::mutate(id = as.character(id)),
  benchmark_df, by = "id"
)

overall_pdf <- file.path(output_folder, paste0("Overall_Sample_Summary_", current_date, ".pdf"))
if (isTRUE(capabilities("cairo"))) {
  grDevices::cairo_pdf(overall_pdf, width = pdf_width, height = pdf_height)
} else {
  grDevices::pdf(overall_pdf, width = pdf_width, height = pdf_height)
}

# --- Page numbering ---
total_report_pages <- 7L
page_counter <- 0L
draw_page_number <- function() {
  page_counter <<- page_counter + 1L
  grid::grid.text(
    paste0("Page ", page_counter, " of ", total_report_pages),
    x = 0.98, y = 0.02, just = c("right", "bottom"),
    gp = grid::gpar(fontsize = 8, col = "grey50")
  )
}

# --- Affective classification helpers ---
affective_composites <- intersect(domain_info$affective_composites, composites_created)
phys_composites      <- intersect(domain_info$physical_composites, composites_created)

label_domain_colors <- function(labels) {
  vars <- names(outcomes)[match(labels, outcomes)]
  vars <- ifelse(is.na(vars), labels, vars)
  ifelse(vars %in% all_physical_outcomes, DOMAIN_COL_PHYSICAL, DOMAIN_COL_AFFECTIVE)
}

# --- Per-ID summary ---
id_outcome_summary <- gam_summary_all %>%
  dplyr::group_by(id, variable) %>%
  dplyr::summarise(
    best_p     = ifelse(any(is.finite(p.value)), min(p.value[is.finite(p.value)]), NA_real_),
    best_range = ifelse(any(is.finite(range)),   max(range[is.finite(range)]),     NA_real_),
    best_n     = ifelse(any(is.finite(n)),        max(n[is.finite(n)]),             NA_real_),
    sig        = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE),
    id_benchmark = dplyr::first(id_benchmark),
    .groups    = "drop"
  ) %>%
  dplyr::mutate(
    meets_benchmark = is.finite(best_range) & is.finite(id_benchmark) & best_range >= id_benchmark
  )

all_ids     <- unique(gam_summary_all$id)
n_gam_valid <- length(all_ids)
n_total_sample <- length(unique(df$id))

# ---- PAGE 1: Cover ----
cover_title <- grid::textGrob(
  paste0(study_label, " \u2014 How to Read This Report"),
  gp = grid::gpar(fontsize = 22, fontface = "bold"), vjust = 0.5
)
cover_purpose <- grid::textGrob(
  paste(strwrap(paste0(
    "This report summarizes cyclical symptom patterns across the menstrual cycle ",
    "for all participants in your sample. Each participant\u2019s daily symptom ratings ",
    "were modeled using Generalized Additive Models (GAMs) to capture smooth, ",
    "non-linear patterns of change across the cycle."
  ), width = 120), collapse = "\n"),
  gp = grid::gpar(fontsize = 11, lineheight = 1.3), just = "left", x = 0.04
)
cover_defs <- grid::textGrob(
  paste0(
    "KEY DEFINITIONS\n\n",
    " \u25B6  Range: The difference between the highest and lowest predicted values from each person\u2019s GAM curve.\n\n",
    " \u25B6  Significant GAM (p < .05): The cyclical pattern is statistically distinguishable from a flat line.\n\n",
    " \u25B6  Meets benchmark: Each person\u2019s benchmark is ", sprintf("%.0f%%", cpass_change_pct * 100),
    " of their personal range of DRSP scale used.\n",
    "     A GAM range meeting or exceeding this threshold is flagged as meeting the C-PASS-equivalent\n",
    "     change criterion.\n\n",
    " \u25B6  Affective outcomes: Mood/emotional symptoms and composites.\n\n",
    " \u25B6  Physical outcomes: Somatic symptoms and composites.\n\n",
    " \u25B6  GAM-valid: Participants with data in all 4 cycle quarters (\u22651 obs each) and at most\n",
    "     one quarter having fewer than ", sparse_quarter_cutoff, " observations."
  ),
  gp = grid::gpar(fontsize = 9.5, lineheight = 1.1), just = "left", x = 0.04
)
gridExtra::grid.arrange(
  cover_title, cover_purpose, cover_defs,
  ncol = 1, heights = grid::unit(c(0.07, 0.10, 0.52), "npc"),
  padding = grid::unit(1, "lines")
)
draw_page_number()

# ---- PAGE 2: Quarter Coverage ----
cov_tv <- time_vars[1]
all_sample_ids <- unique(df$id)
cov_rows <- lapply(all_sample_ids, function(pid) {
  pid_vals <- df[[cov_tv]][df$id == pid]
  pid_vals <- pid_vals[!is.na(pid_vals)]
  qc <- quarter_bin_counts(pid_vals)
  n_empty  <- sum(qc == 0)
  n_sparse <- sum(qc > 0 & qc < sparse_quarter_cutoff)
  passes   <- n_empty == 0 && n_sparse <= 1
  data.frame(id = pid, Q1 = qc["Q1"], Q2 = qc["Q2"], Q3 = qc["Q3"], Q4 = qc["Q4"],
             total_obs = sum(qc), included = passes, stringsAsFactors = FALSE)
})
cov_df <- dplyr::bind_rows(cov_rows)
n_included <- sum(cov_df$included)
n_excluded <- sum(!cov_df$included)

cov_long <- cov_df %>%
  tidyr::pivot_longer(cols = c("Q1", "Q2", "Q3", "Q4"),
                      names_to = "quarter", values_to = "obs_count") %>%
  dplyr::mutate(status = ifelse(included, "Included", "Excluded"),
                id_label = as.character(id))

id_order_incl <- cov_df %>% dplyr::filter(included) %>%
  dplyr::arrange(dplyr::desc(total_obs), id) %>% dplyr::pull(id) %>% as.character()
id_order_excl <- cov_df %>% dplyr::filter(!included) %>%
  dplyr::arrange(dplyr::desc(total_obs), id) %>% dplyr::pull(id) %>% as.character()
cov_long$id_label <- factor(cov_long$id_label, levels = c(id_order_excl, id_order_incl))
cov_long$status <- factor(cov_long$status, levels = c("Excluded", "Included"))

quarter_labels <- c("Q1\n[-1, -0.5)", "Q2\n[-0.5, 0)", "Q3\n[0, 0.5)", "Q4\n[0.5, 1]")
cov_long$quarter <- factor(cov_long$quarter, levels = c("Q1", "Q2", "Q3", "Q4"),
                            labels = quarter_labels)
max_obs_display <- min(max(cov_long$obs_count, na.rm = TRUE), 30)
cov_long$obs_capped <- pmin(cov_long$obs_count, max_obs_display)

p_cov <- ggplot2::ggplot(cov_long,
           ggplot2::aes(x = quarter, y = id_label, fill = obs_capped)) +
  ggplot2::geom_tile(color = "white", linewidth = 0.15) +
  ggplot2::geom_text(ggplot2::aes(label = obs_count), size = 1.6, color = "grey20") +
  ggplot2::facet_grid(status ~ ., scales = "free_y", space = "free_y", switch = "y") +
  ggplot2::scale_fill_gradient(
    low = "#FFF3E0", high = "#E65100", na.value = "grey90",
    name = "Observations", limits = c(0, max_obs_display),
    breaks = pretty(c(0, max_obs_display), n = 4)
  ) +
  ggplot2::labs(
    title = "Quarter-Based Data Coverage Across the Cycle by Participant",
    subtitle = paste0("Time variable: ", cov_tv,
                      "  |  Included: N = ", n_included,
                      "  |  Excluded: N = ", n_excluded,
                      "  |  Total: N = ", length(all_sample_ids)),
    x = "Cycle Quarter", y = NULL
  ) +
  ggplot2::theme_minimal(base_size = 8) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold", size = 12),
    plot.subtitle = ggplot2::element_text(size = 9),
    axis.text.y = ggplot2::element_text(size = 4),
    strip.text.y.left = ggplot2::element_text(face = "bold", size = 9, angle = 0),
    strip.placement = "outside", panel.grid = ggplot2::element_blank(),
    legend.position = "right"
  )
print(p_cov)
draw_page_number()

# ---- PAGE 3: Key Metrics ----
ids_any_benchmark <- id_outcome_summary %>%
  dplyr::filter(meets_benchmark) %>% dplyr::distinct(id) %>% dplyr::pull(id)
pct_any_benchmark <- length(ids_any_benchmark) / n_gam_valid * 100

ids_affective_benchmark <- id_outcome_summary %>%
  dplyr::filter(variable %in% all_affective_outcomes, meets_benchmark) %>%
  dplyr::distinct(id) %>% dplyr::pull(id)
pct_affective_benchmark <- length(ids_affective_benchmark) / n_gam_valid * 100

ids_any_sig <- id_outcome_summary %>%
  dplyr::filter(sig) %>% dplyr::distinct(id) %>% dplyr::pull(id)
pct_any_sig <- length(ids_any_sig) / n_gam_valid * 100

ids_affective_sig <- id_outcome_summary %>%
  dplyr::filter(variable %in% all_affective_outcomes, sig) %>%
  dplyr::distinct(id) %>% dplyr::pull(id)
pct_affective_sig <- length(ids_affective_sig) / n_gam_valid * 100

ids_affective_either <- id_outcome_summary %>%
  dplyr::filter(variable %in% all_affective_outcomes, sig | meets_benchmark) %>%
  dplyr::distinct(id) %>% dplyr::pull(id)
pct_affective_either <- length(ids_affective_either) / n_gam_valid * 100

key_metrics <- data.frame(
  Metric = c(
    "Total Participants in Data", "Participants with Valid GAM Data", "",
    "N with \u2265 1 significant GAM (any outcome)",
    "% with \u2265 1 significant GAM (any outcome)",
    "N with \u2265 1 significant GAM (affective only)",
    "% with \u2265 1 significant GAM (affective only)", "",
    "N meeting person-specific benchmark (any)",
    "% meeting person-specific benchmark (any)",
    "N meeting benchmark (affective only)",
    "% meeting benchmark (affective only)", "",
    "N with \u2265 1 affective meeting EITHER criterion",
    "% with \u2265 1 affective meeting EITHER criterion"
  ),
  Value = c(
    as.character(n_total_sample), as.character(n_gam_valid), "",
    as.character(length(ids_any_sig)), sprintf("%.1f%%", pct_any_sig),
    as.character(length(ids_affective_sig)), sprintf("%.1f%%", pct_affective_sig), "",
    as.character(length(ids_any_benchmark)), sprintf("%.1f%%", pct_any_benchmark),
    as.character(length(ids_affective_benchmark)), sprintf("%.1f%%", pct_affective_benchmark), "",
    as.character(length(ids_affective_either)), sprintf("%.1f%%", pct_affective_either)
  ),
  stringsAsFactors = FALSE
)

key_theme <- gridExtra::ttheme_default(
  base_size = 9.5,
  core    = list(fg_params = list(hjust = 0, x = 0.02)),
  colhead = list(fg_params = list(hjust = 0, x = 0.02, fontface = "bold"))
)

title_grob <- grid::textGrob(
  paste0(study_label, " \u2014 Cyclical Symptom Change Summary"),
  gp = grid::gpar(fontsize = 18, fontface = "bold"), vjust = 0.5
)
subtitle_grob <- grid::textGrob(
  paste0("N = ", n_total_sample, " total | N = ", n_gam_valid, " with valid GAM data"),
  gp = grid::gpar(fontsize = 11, fontface = "italic"), vjust = 0.5
)
key_tbl <- gridExtra::tableGrob(key_metrics, rows = NULL, theme = key_theme)
gridExtra::grid.arrange(
  title_grob, subtitle_grob, key_tbl,
  ncol = 1, heights = grid::unit(c(0.07, 0.04, 0.60), "npc")
)
draw_page_number()

# ---- PAGE 4: Per-Outcome Descriptives ----
outcome_desc <- id_outcome_summary %>%
  dplyr::group_by(variable) %>%
  dplyr::summarise(
    N = dplyr::n(),
    mean_range = mean(best_range, na.rm = TRUE),
    sd_range = sd(best_range, na.rm = TRUE),
    median_range = median(best_range, na.rm = TRUE),
    n_sig = sum(sig, na.rm = TRUE),
    pct_sig = sum(sig, na.rm = TRUE) / n_gam_valid * 100,
    n_benchmark = sum(meets_benchmark, na.rm = TRUE),
    pct_benchmark = sum(meets_benchmark, na.rm = TRUE) / n_gam_valid * 100,
    n_either = sum(sig | meets_benchmark, na.rm = TRUE),
    pct_either = sum(sig | meets_benchmark, na.rm = TRUE) / n_gam_valid * 100,
    .groups = "drop"
  ) %>%
  dplyr::left_join(
    tibble::enframe(outcomes, name = "variable", value = "label"), by = "variable"
  ) %>%
  dplyr::mutate(label = dplyr::coalesce(label, variable)) %>%
  dplyr::arrange(dplyr::desc(pct_either))

desc_display <- outcome_desc %>%
  dplyr::transmute(
    Outcome = label,
    N = N,
    `Mean Range` = round(mean_range, 2),
    `SD Range` = round(sd_range, 2),
    `Median Range` = round(median_range, 2),
    `% Sig` = sprintf("%.1f", pct_sig),
    `% Benchmark` = sprintf("%.1f", pct_benchmark),
    `% Either` = sprintf("%.1f", pct_either)
  )

desc_theme <- gridExtra::ttheme_default(
  base_size = 7.5,
  core = list(fg_params = list(hjust = 0, x = 0.02)),
  colhead = list(fg_params = list(hjust = 0, x = 0.02, fontface = "bold"))
)
desc_title <- grid::textGrob("Per-Outcome Descriptives (sorted by % Either)",
                              gp = grid::gpar(fontsize = 14, fontface = "bold"))
desc_note <- grid::textGrob(
  paste0("Percentages use GAM-valid N = ", n_gam_valid, " as denominator."),
  gp = grid::gpar(fontsize = 8, fontface = "italic"))
desc_tbl <- gridExtra::tableGrob(desc_display, rows = NULL, theme = desc_theme)
gridExtra::grid.arrange(desc_title, desc_note, desc_tbl, ncol = 1,
                        heights = grid::unit(c(0.06, 0.03, 0.85), "npc"))
draw_page_number()

# ---- PAGE 5: Range Distributions (histograms) ----
hist_data <- id_outcome_summary %>%
  dplyr::left_join(tibble::enframe(outcomes, name = "variable", value = "label"), by = "variable") %>%
  dplyr::mutate(label = dplyr::coalesce(label, variable))

p_hist <- ggplot2::ggplot(hist_data, ggplot2::aes(x = best_range, fill = sig)) +
  ggplot2::geom_histogram(bins = 20, alpha = 0.8, position = "stack") +
  ggplot2::facet_wrap(~ label, scales = "free_y", ncol = 4) +
  ggplot2::scale_fill_manual(
    values = c("TRUE" = DOMAIN_COL_AFFECTIVE, "FALSE" = "grey70"),
    labels = c("TRUE" = "Significant", "FALSE" = "Not sig."),
    name = NULL
  ) +
  ggplot2::labs(
    title = "Distribution of GAM Ranges by Outcome",
    subtitle = paste0("N = ", n_gam_valid, " GAM-valid participants"),
    x = "GAM Range (max - min predicted)", y = "Count"
  ) +
  ggplot2::theme_minimal(base_size = 8) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold", size = 12),
    strip.text = ggplot2::element_text(size = 6),
    legend.position = "bottom"
  )
print(p_hist)
draw_page_number()

# ---- PAGE 6: Per-Person Range Distributions (dot + box) ----
dot_data <- id_outcome_summary %>%
  dplyr::left_join(tibble::enframe(outcomes, name = "variable", value = "label"), by = "variable") %>%
  dplyr::mutate(label = dplyr::coalesce(label, variable))

# Order outcomes by median range
label_order_dot <- dot_data %>%
  dplyr::group_by(label) %>%
  dplyr::summarise(med = median(best_range, na.rm = TRUE), .groups = "drop") %>%
  dplyr::arrange(med) %>%
  dplyr::pull(label)
dot_data$label <- factor(dot_data$label, levels = label_order_dot)

p_dot <- ggplot2::ggplot(dot_data, ggplot2::aes(x = label, y = best_range)) +
  ggplot2::geom_boxplot(width = 0.4, outlier.shape = NA, alpha = 0.3) +
  ggplot2::geom_jitter(width = 0.15, alpha = 0.4, size = 1,
                       ggplot2::aes(color = sig)) +
  ggplot2::coord_flip() +
  ggplot2::scale_color_manual(
    values = c("TRUE" = DOMAIN_COL_AFFECTIVE, "FALSE" = "grey60"),
    labels = c("TRUE" = "Significant", "FALSE" = "Not sig."),
    name = NULL
  ) +
  ggplot2::labs(
    title = "Per-Person GAM Range Distributions by Outcome",
    subtitle = paste0("N = ", n_gam_valid, " GAM-valid participants"),
    x = NULL, y = "GAM Range"
  ) +
  ggplot2::theme_minimal(base_size = 8) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold", size = 12),
    legend.position = "bottom"
  )
print(p_dot)
draw_page_number()

# ---- PAGE 7: Significance & Benchmark Rates ----
rates_data <- outcome_desc %>%
  dplyr::select(label, pct_sig, pct_benchmark) %>%
  tidyr::pivot_longer(cols = c(pct_sig, pct_benchmark),
                      names_to = "metric", values_to = "pct") %>%
  dplyr::mutate(
    metric = dplyr::case_when(
      metric == "pct_sig" ~ "% Significant",
      metric == "pct_benchmark" ~ "% Meets Benchmark"
    )
  )

rates_data$label <- factor(rates_data$label,
                            levels = rev(outcome_desc$label))

p_rates <- ggplot2::ggplot(rates_data,
             ggplot2::aes(x = label, y = pct, fill = metric)) +
  ggplot2::geom_col(position = "dodge", width = 0.7, alpha = 0.85) +
  ggplot2::coord_flip() +
  ggplot2::scale_fill_manual(
    values = c("% Significant" = DOMAIN_COL_AFFECTIVE, "% Meets Benchmark" = DOMAIN_COL_PHYSICAL),
    name = NULL
  ) +
  ggplot2::labs(
    title = "Significance & Benchmark Rates by Outcome",
    subtitle = paste0("N = ", n_gam_valid, " GAM-valid participants (sorted by % Either)"),
    x = NULL, y = "Percentage of GAM-valid Sample"
  ) +
  ggplot2::theme_minimal(base_size = 8) +
  ggplot2::theme(
    plot.title = ggplot2::element_text(face = "bold", size = 12),
    legend.position = "bottom"
  )
print(p_rates)
draw_page_number()

grDevices::dev.off()
message(sprintf("\u2705 Overall summary saved: %s", overall_pdf))

} # end if generate_overall_summary
```

# Done
```{r done}
message(sprintf("\n\u2705 All reports complete! Output folder: %s", output_folder))
```
