title: "HERA_primary"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
---

# HERA Primary Data Preparation Pipeline

## Overview

Prepares CLEAN datasets for Raina Pang R01 looking at alcohol use across the cycle

**Key Outputs:** `hera_YYYYMMDD_clean.rds` (clean dataset with NO derived variables), diagnostic plots



```{r setup}
# Load packages, set options, create output folder
source("HERA_00_setup.R")

# Create subfolder for data prep outputs
data_prep_folder <- file.path(output_folder, "01_data_preparation")
if (!dir.exists(data_prep_folder)) {
  dir.create(data_prep_folder, recursive = TRUE)
  cat("Created subfolder:", data_prep_folder, "\n")
}
```

## Data Import

```{r import-survey-data}

library(readr)
library(dplyr)

file_path <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/02_datasets/HERA_DAILY/02_data_prep_workspace/V.1-V.5-FINAL.csv"

hera_raw <- read_csv(file_path, col_types = cols(.default = col_character()))

pm_surveys <- hera_raw %>%
  select(
    id,
    daterated,
    #wake_alcuse,
    #ran_alcuse,
    drsp_1:drsp_21,
    #night_alcuse,
       # night_alcplan,
        #stress,
       # wake_alcmotive_pain,
       # wake_alcmotive_anx,
       # wake_alcmotive_dep,
       # wake_alcmotive_enhance,
       # wake_alcmotive_celeb,
       # wake_alcmotive_fomo,
       # wake_alcmotive_other,
       # wake_alcmotive_na,
    # night_alcmotive_pain,
    # night_alcmotive_anx,
    # night_alcmotive_dep,
    # night_alcmotive_enhance,
    # night_alcmotive_celeb,
    # night_alcmotive_fomo,
    # night_alcmotive_other,
    # night_alcmotive_na,
    # alcuse,
    # sub_alc,
    # alc_duration,
    # wake_alcduration,
    # age,
    # height,
    # weight,
    # audit_consumption_sum,
    # audit_dependence_sum,
    # audit_problems_sum,
    # audit_sum,
    Mens_onset,
    Ov_day,
    TemperatureDeviationC,
    TemperatureScore,
    TemperatureTrendDeviation
  )

View(pm_surveys)
```


## Clean Survey Data

```{r clean-pm, warning=FALSE}
# Clean evening surveys: alcohol use (numdrinkstoday), menses, LHpos, drsp_1 to drsp_24
# Parses dates

# ---- 1. Process Dates ----
# NOTE: janitor::clean_names() has already been applied in load_and_clean_survey()
# so column names are already lowercase

#View(pm_surveys)

# remove observations where id or daterated are NA or where drsp_1 is NA

pm_surveys <- pm_surveys %>%
  filter(!is.na(id), !is.na(daterated), !is.na(drsp_1))
```


```{r clean-pm, warning=FALSE}
# Format date_rated as MM/DD/YYYY
pm_surveys <- pm_surveys %>%
  mutate(daterated = mdy(daterated))

# ---- 1c. Track Missing Data BEFORE Filtering ----
# Document missing data patterns before removing any rows
n_total_rows <- nrow(pm_surveys)
n_missing_id <- sum(is.na(pm_surveys$id))
n_missing_daterated <- sum(is.na(pm_surveys$daterated))
n_missing_both <- sum(is.na(pm_surveys$id) & is.na(pm_surveys$daterated))

cat("\n=== AM Survey Missing Data Report (Pre-filtering) ===\n")
cat(sprintf("Total rows: %d\n", n_total_rows))
cat(sprintf("Rows with missing id: %d (%.1f%%)\n", n_missing_id, 100 * n_missing_id / n_total_rows))
cat(sprintf("Rows with missing daterated: %d (%.1f%%)\n", n_missing_daterated, 100 * n_missing_daterated / n_total_rows))
cat(sprintf("Rows with both missing: %d (%.1f%%)\n", n_missing_both, 100 * n_missing_both / n_total_rows))

# ---- 1d. Filter Rows with Missing Keys ----
# Drop observations where id or daterated are missing (cannot be matched in merge)
pm_surveys <- pm_surveys %>%
  filter(!is.na(id), !is.na(daterated))

cat(sprintf("Rows after removing missing id/daterated: %d (removed %d rows)\n\n", 
            nrow(pm_surveys), n_total_rows - nrow(pm_surveys)))

# ---- 2. Clean Alcohol Variables ----

# Ensure numdrinkstoday is numeric
#pm_surveys <- pm_surveys %>% 
 # mutate(numdrinkstoday = as.numeric(numdrinkstoday))

# ---- 6. Visualization and Diagnostics ----
# Visualize Missing Data
cat("\n=== AM Survey Missing Data Pattern (Final) ===\n")
#vis_miss(pm_surveys) 

# Print summary
cat("\n=== AM Survey Cleaning Complete ===\n")
cat("Variables: All in snake_case format\n")
cat("Alcohol variables: numdrinkstoday (numeric)")

count_rows_ids(pm_surveys)

names(pm_surveys)
```


# Create DRSP variable list
```{r more-cleaning, warning=FALSE}
# Clean evening surveys: mood (DRSP)
# Recodes 

# ---- 0. Define Variable Lists ----
# These variable lists are needed for recoding DRSP and bdefs scales below
# Following refactoring guidelines: each chunk defines its own variable lists

drsp_vars <- c("drsp_1", "drsp_2", "drsp_3", "drsp_4", "drsp_5", "drsp_6", "drsp_7", "drsp_8", 
               "drsp_9", "drsp_10", "drsp_11", "drsp_12", "drsp_13", "drsp_14",
               "drsp_15", "drsp_16", "drsp_17", "drsp_18", "drsp_19",
               "drsp_20", "drsp_21")

```


# Create a new variable "ovtoday", subset by id, and code ovtoday=1 if the prior day (daterated) has a positive LH test (OvDay == 1), otherwise ovtoday=0. This indicates whether the participant is likely ovulating on the day they completed the PM survey.

```{r more-cleaning, warning=FALSE}
df <- pm_surveys %>%
  arrange(id, daterated) %>%
  group_by(id) %>%
  mutate(ovtoday = if_else(lag(Ov_day) == 1, 1, 0, missing = 0)) %>%
  ungroup()

```

# Check for duplicates and remove

```{r dups}


df <- df %>%
  distinct()

View(df)

count_rows_ids(df)
```



```{r verify-column-names}
#print column names to verify
cat("\n=== FINAL DATAFRAME COLUMNS ===\n")
print(colnames(df))
```


```{r calc-day-of-week}
#Create day of week variables including categorical, sin, and cos variables
df <- df %>%
  mutate(day_of_week = weekdays(as.Date(daterated)),
         day_of_week = factor(day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))

names(df)

```


```{r}
# if menses = NA, then set menses to 0, otherwise keep the original value of menses (which is 1 for period start dates) - make no assumptions about the variable type of mens_onset

df <- df %>%
  mutate(menses = if_else(is.na(Mens_onset), 0, as.numeric(Mens_onset)))

```



```{r visualize-cycle-events}
# Step 5: Visualize Period Start Dates Over Time with Color-Coded Tiles
periodstartplot <- df %>%
  ggplot(aes(x = daterated, y = id, fill = factor(menses))) +
  geom_tile() +
  labs(
    title = "Period Start Dates Over Time",
    x = "Date",
    y = "ID",
    fill = "Period Start"
  ) +
  scale_fill_manual(values = c("grey", "red")) +
  theme_minimal(base_size = 14)

periodstartplot

# Save to data prep subfolder, specify periodstartplot
save_png(plot = periodstartplot, 
         filename = "period_start_dates_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap

# print location of saved file
cat("Period start dates heatmap saved\n")

# Step 6: Visualize Ovulation Dates Over Time
ovulationplot <- df %>%
  ggplot(aes(x = daterated, y = id, fill = factor(ovtoday))) +
  geom_tile() +
  labs(
    title = "Ovulation Dates Over Time",
    x = "Date",
    y = "ID",
    fill = "Ovulation Date"
  ) +
  scale_fill_manual(values = c("grey", "blue")) +
  theme_minimal(base_size = 14)

# Save to data prep subfolder, specify ovulationplot
save_png(plot = ovulationplot, 
         filename = "ovulation_dates_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap

ovulationplot

#df %>% select(id, daterated, menses, ovtoday) %>% View()

##vis_miss(df)



```


# Calculate aggregate DRSP scores
```{r calc-drsp-aggregates}
# =============================================================================
# DRSP COMPOSITE SCORE CALCULATION (FOR HERA DIAGNOSTIC DATA)
# =============================================================================
# NOTE: These subscales are calculated here for the HERA dataset specifically,
# but they are ALSO calculated in HERA_03_processing_shared.Rmd for the 
# merged ADHD+HERA dataset to ensure both studies receive identical processing.
# =============================================================================

# set drsp variables to numeric
df <- df %>%
  mutate(across(c("drsp_1", "drsp_2", "drsp_3", "drsp_4", "drsp_5", "drsp_6", "drsp_7", "drsp_8", 
               "drsp_9", "drsp_10", "drsp_11", "drsp_12", "drsp_13", "drsp_14",
               "drsp_15", "drsp_16", "drsp_17", "drsp_18", "drsp_19",
               "drsp_20", "drsp_21"), as.numeric))

# Create total DRSP symptom score (means of all items)
df <- df %>%
  mutate(drsp_total = rowMeans(select(., 
      "drsp_1", "drsp_2", "drsp_3", "drsp_4", "drsp_5", "drsp_6", "drsp_7", "drsp_8", 
               "drsp_9", "drsp_10", "drsp_11", "drsp_12", "drsp_13", "drsp_14",
               "drsp_15", "drsp_16", "drsp_17", "drsp_18", "drsp_19",
               "drsp_20", "drsp_21"
  ), na.rm = TRUE))

# Create total DRSP negative affect score
df <- df %>%
  mutate(drsp_NA = rowMeans(select(., 
      "drsp_1", "drsp_2", "drsp_3", "drsp_4", "drsp_5", "drsp_6", "drsp_7", "drsp_8", 
               "drsp_9", "drsp_16", "drsp_17"
  ), na.rm = TRUE))

df <- df %>%
  mutate(drsp_NAmax = pmax(
    drsp_1, drsp_2, drsp_3, drsp_4, drsp_5, drsp_6, drsp_7, drsp_8,
    drsp_9, drsp_16, drsp_17,
    na.rm = TRUE
  ))
# Create "drsp_distress" - dep hope worth anx ang loss_interest overwhelm loss_control
df <- df %>%
  mutate(drsp_distress = rowMeans(select(., 
      drsp_1, drsp_2, drsp_3, drsp_4, drsp_16, drsp_17), na.rm = TRUE))

# create drsp_labilsocsx - mood_swing, rej_sens, conflicts,int_relat, angry
df <- df %>%
  mutate(drsp_labilsocsx = rowMeans(select(., drsp_5, drsp_6, drsp_8, drsp_7), na.rm = TRUE))

# create drsp_breastpain - breast_tend, breast_swell
df <- df %>%
  mutate(drsp_breastpain = rowMeans(select(., drsp_18, drsp_19), na.rm = TRUE))

# create drsp_eatsx - over_eat, sp_food_crav
df <- df %>%
  mutate(drsp_eatsx = rowMeans(select(., drsp_12, drsp_13), na.rm = TRUE))

# create drsp_lowarousal - tired, diff_conc, diff_sleep, high_sleep
df <- df %>%
  mutate(drsp_lowarousal = rowMeans(select(., drsp_11, drsp_10, drsp_14, drsp_15, drsp_9), na.rm = TRUE))

```


```{r remove-dupes}
df <- df %>%
  distinct(id, daterated, .keep_all = TRUE)

count_rows_ids(df)
```

## menstrualcycleR Processing

#Apply PACTS scaling and other menstrual cycle transformations using the menstrualcycleR package. This standardizes cycle phase coding across participants.

```{r run-menstrualcycler}
library(menstrualcycleR)

# Check required columns exist
required_cols <- c("id", "menses", "ovtoday")
date_col <- if("daterated" %in% names(df)) "daterated" else if("date" %in% names(df)) "date" else NULL

if(is.null(date_col)) {
  stop("ERROR: No date column found (daterated or date). Cannot run pacts_scaling.\n",
       "Required for menstrual cycle phase calculation.")
} else if(!all(required_cols %in% names(df))) {
  missing_cols <- setdiff(required_cols, names(df))
  stop("ERROR: Missing required columns for pacts_scaling: ", paste(missing_cols, collapse=", "), ".\n",
       "Required columns: id, date (or daterated), menses, ovtoday")
} else {
  # Run pacts_scaling 
  cat("\nRunning pacts_scaling from menstrualcycleR package...\n")
  
  if(date_col == "daterated") {
    # Need to rename for pacts_scaling
    df <- df %>% rename(date = daterated)
    df <- pacts_scaling(df, id=id, date=date, menses=menses, ovtoday=ovtoday, 
                        lower_cyclength_bound = 17, upper_cyclength_bound = 40)
    df <- df %>% rename(daterated = date)
  } else {
    df <- pacts_scaling(df, id=id, date=date, menses=menses, ovtoday=ovtoday, 
                        lower_cyclength_bound = 17, upper_cyclength_bound = 40)
  }
  
  cat("pacts_scaling complete.\n")
}


data_available_info <- cycledata_check(df, c("drsp_1"))

# View results
print(data_available_info$by_id)
print(data_available_info$overall)
print(data_available_info$data_symptom_plots$drsp_1)

#Save to output_folder: data_available_info$data_symptom_plots$drsp_1 

save_png(plot = data_available_info$data_symptom_plots$drsp_1, 
         filename = "drsp_1_data_availability.png",
         folder = data_prep_folder,
         height = 6)


```



```{r print variable names}

names(df)

```





##### END OF DATA PREP 

```{r save-prepped-dataset-box}

write.csv(df, paste0("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_", current_date, "_clean.csv"), row.names = FALSE)

# Save as .rds
saveRDS(df, paste0("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_", current_date, "_clean.rds"))
```


```{r save-prepped-dataset-data-folder}
# output a copy to the data folder
write.csv(df, paste0("data/HERA_", current_date, "_clean.csv"), row.names = FALSE)
# Save as .rds
saveRDS(df, paste0("data/HERA_", current_date, "_clean.rds"))

cat("\n✓ Clean dataset saved (no derived variables).\n")
cat("  Output: data/HERA_", current_date, "_clean.rds\n", sep="")
```

```{r placeholder-empty}

```

