---
title: "UKALC_ADHDCYCLE_1_Descriptives"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
    self_contained: true
---


```{r}
#Source R File with Setup, Package Loading, and Output Folder Creation

source("UKALC_00_setup.R")

# Create subfolder for descriptives outputs
descriptives_folder <- file.path(output_folder, "02_descriptives")
if (!dir.exists(descriptives_folder)) {
  dir.create(descriptives_folder, recursive = TRUE)
  cat("Created descriptives subfolder:", descriptives_folder, "\n")
}

# LOAD LATEST CLEAN DATA FILE 

# Get all files matching the pattern
files <- list.files(path = "data", 
                   pattern = "^analysis_ready_\\d{8}.*\\.rds$", 
                   full.names = TRUE)

# Sort files (dates in YYYYMMDD format will sort correctly alphabetically)
latest_file <- sort(files, decreasing = TRUE)[1]

# Load the dataset
df <- readRDS(latest_file)

# Print which file was loaded
cat("Loaded dataset from:", latest_file, "\n")

```


# Import Data
```{r}
# --- 1. Setup ---

# Add all your continuous outcome variable names here
continuous_vars <- c(
  "num_drinks_today", "bac_peak_percent", "crave_alc_am", "crave_alc_pm", "crave_nic_am", "crave_mar_am",
  "diff_stop", "diff_change", "immed_payoff", "res_urge", 
  "diff_long_rew", "diff_res_rew", "dep", "hopeless", 
  "worthless", "anxious", "mood_swing", "rej_sens", 
  "angry", "conflicts", "loss_interest", "diff_conc", 
  "tired", "over_eat", "sp_food_crav", "high_sleep", 
  "diff_sleep", "overwhelm", "loss_control", "breast_tend", 
  "breast_swell", "headache", "jm_pain", "loss_prod", 
  "int_relat", "crave_alc_pm", "crave_nic_pm", "crave_mar_pm"
)

outcome_labels <- c(
  # --- Substance Use & Craving ---
  "num_drinks_today" = "Number of Drinks",
  "bac_peak_percent" = "Peak BAC (%)",
  "drink_today"      = "Any Drinking Today",
  "fourplustoday"    = "Binge Drinking (4+)",
  "felt_drunk"       = "Felt Drunk",
  "crave_alc_am"     = "Alcohol Craving (AM)",
  "crave_alc_pm"     = "Alcohol Craving (PM)",
  "crave_nic_am"     = "Nicotine Craving (AM)",
  "crave_nic_pm"     = "Nicotine Craving (PM)",
  "crave_mar_am"     = "Marijuana Craving (AM)",
  "crave_mar_pm"     = "Marijuana Craving (PM)",

  # --- Impulsivity & Reward ---
  "diff_stop"        = "Difficulty Stopping",
  "diff_change"      = "Difficulty Changing",
  "immed_payoff"     = "Immediate Payoff Pref.",
  "res_urge"         = "Resistance to Urges",
  "diff_long_rew"    = "Diff. Waiting Long Reward",
  "diff_res_rew"     = "Diff. Resisting Reward",
  "bdefs_impulsiveaction" = "Impulsive Action (BDEFS)",
  "bdefs_delayaversion" = "Delay Aversion (BDEFS)",

  # --- Mood & Affect (DRSP/PMDD) ---
  "dep"              = "Depression",
  "hopeless"         = "Hopelessness",
  "worthless"        = "Worthlessness",
  "anxious"          = "Anxiety",
  "mood_swing"       = "Mood Swings",
  "rej_sens"         = "Rejection Sensitivity",
  "angry"            = "Anger/Irritability",
  "conflicts"        = "Interpersonal Conflicts",
  "loss_interest"    = "Loss of Interest",
  "overwhelm"        = "Feeling Overwhelmed",
  "loss_control"     = "Feeling Out of Control",
  "int_relat"        = "Relationship Conflicts",

  # --- DRSP Aggregate Scores ---
  "drsp_total"       = "DRSP Total",
  "drsp_distress"    = "Distress (DRSP)",
  "drsp_labilsocsx"  = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"  = "Breast Sx (DRSP)",
  "drsp_eatsx"       = "Eating Sx",
  "drsp_lowarousal"  = "Low Arousal Sx (DRSP)",

  # --- Physical & Cognitive Symptoms ---
  "diff_conc"        = "Difficulty Concentrating",
  "tired"            = "Fatigue/Lethargy",
  "high_sleep"       = "Hypersomnia",
  "diff_sleep"       = "Insomnia",
  "over_eat"         = "Overeating",
  "sp_food_crav"     = "Specific Food Cravings",
  "breast_tend"      = "Breast Tenderness",
  "breast_swell"     = "Breast Swelling",
  "headache"         = "Headache",
  "jm_pain"          = "Joint/Muscle Pain",
  "loss_prod"        = "Decreased Productivity"
)
```

# ------ SMM ELIGIBILITY at 5 OBS PER PHASE - MORNING/ALCOHOL OUTCOMES ------
```{r}
library(dplyr)
library(ggplot2)
library(tidyr) # For pivot_longer in the combined plot

# 1. DATA PREP ------------------------------------------------------------
# Save smm eligibility data
alcsmmelig <- cycledata_check(df, "drink_today")

# Merge back to df by id
df <- df %>%
  left_join(alcsmmelig$by_id %>% select(id, drink_today_follicular, drink_today_luteal), by = "id")

# Create a clean plotting dataset with an explicit "Low Data" flag
# This makes the ggplot logic much simpler
plot_data <- alcsmmelig$by_id %>%
  mutate(
    # Create a boolean flag: TRUE if EITHER phase is < 5
    low_data_flag = drink_today_follicular < 5 | drink_today_luteal < 5,
    # Convert to factor for nice legend labels
    status = ifelse(low_data_flag, "Excluded (<5 days)", "Included")
  )

# 2. INDIVIDUAL PLOTS -----------------------------------------------------

# Follicular Plot
p_drink_follicular <- ggplot(plot_data, aes(x = id, y = drink_today_follicular, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 5, label = "5-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Follicular Phase Drinking Days", y = "Days", x = "Participant ID") +
  theme_minimal()

print(p_drink_follicular)

# Luteal Plot (Fixed the object name to p_drink_luteal)
p_drink_luteal <- ggplot(plot_data, aes(x = id, y = drink_today_luteal, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 5, label = "5-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Luteal Phase Drinking Days", y = "Days", x = "Participant ID") +
  theme_minimal()

print(p_drink_luteal)


# 3. COMBINED PLOT --------------------------------------------------------

# To show both phases side-by-side per ID, we pivot to long format
p_combined_drink <- plot_data %>%
  pivot_longer(cols = c(drink_today_follicular, drink_today_luteal), 
               names_to = "phase", 
               values_to = "days") %>%
  mutate(phase = recode(phase, "drink_today_follicular" = "Follicular", "drink_today_luteal" = "Luteal")) %>%
  ggplot(aes(x = id, y = days, fill = status)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) + # 'dodge' puts bars side-by-side
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  labs(title = "Drinking Days by Phase (Red = <5 days in EITHER phase)", 
       y = "Number of Drinking Days", 
       x = "Participant ID") +
  facet_wrap(~phase, ncol = 1) + # Optional: Split panels if you have many IDs
  theme_minimal(base_size = 14)

print(p_combined_drink)

save_png(plot = p_combined_drink,
         filename = "combined_drink_data_availability.png",
         folder = descriptives_folder,
         width = 10,
         height = 8,
         dpi = 300)


# 4. COUNTS ---------------------------------------------------------------

# Count number of ids where either follicular or luteal is <5
# We can use the flag we created earlier
num_ids_below_threshold <- sum(plot_data$low_data_flag)
cat("Number of IDs with fewer than 5 drinking days in either phase:", num_ids_below_threshold, "\n")

# Add number that do have >=5 in both phases
num_ids_above_threshold <- sum(!plot_data$low_data_flag)
cat("Number of IDs with 5 or more drinking days in both phases:", num_ids_above_threshold, "\n")
```

# ------ SMM ELIGIBILITY at 5 OBS PER PHASE - EVENING/DRSP OUTCOMES ------

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)

# 1. DATA PREP ------------------------------------------------------------
# Save smm eligibility data for 'dep'
depsmmelig <- cycledata_check(df, "dep")

# Merge back to df by id
df <- df %>%
  left_join(depsmmelig$by_id %>% select(id, dep_follicular, dep_luteal), by = "id")

# Create a clean plotting dataset with an explicit "Low Data" flag
plot_data <- depsmmelig$by_id %>%
  mutate(
    # Create a boolean flag: TRUE if EITHER phase is < 5
    low_data_flag = dep_follicular < 5 | dep_luteal < 5,
    # Convert to factor for nice legend labels
    status = ifelse(low_data_flag, "Excluded (<5 days)", "Included")
  )

# 2. INDIVIDUAL PLOTS -----------------------------------------------------

# Follicular Plot
p_dep_follicular <- ggplot(plot_data, aes(x = id, y = dep_follicular, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 5, label = "5-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Follicular Phase Data (Dep)", y = "Days Present", x = "Participant ID") +
  theme_minimal()

print(p_dep_follicular)

# Luteal Plot
p_dep_luteal <- ggplot(plot_data, aes(x = id, y = dep_luteal, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 5, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 5, label = "5-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Luteal Phase Data (Dep)", y = "Days Present", x = "Participant ID") +
  theme_minimal()

print(p_dep_luteal)


# 3. COMBINED PLOT --------------------------------------------------------

# Pivot to long format to show both phases side-by-side
p_combined_dep <- plot_data %>%
  pivot_longer(cols = c(dep_follicular, dep_luteal), 
               names_to = "phase", 
               values_to = "days") %>%
  mutate(phase = recode(phase, "dep_follicular" = "Follicular", "dep_luteal" = "Luteal")) %>%
  ggplot(aes(x = id, y = days, fill = status)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) + 
  geom_hline(yintercept = 5, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Excluded (<5 days)" = "red", "Included" = "lightblue")) +
  labs(title = "Data Availability by Phase (Red = <5 days in EITHER phase)", 
       y = "Number of Days with Data", 
       x = "Participant ID") +
  facet_wrap(~phase, ncol = 1) + 
  theme_minimal(base_size = 14)

print(p_combined_dep)

save_png(plot = p_combined_dep,
         filename = "combined_dep_data_availability.png",
         folder = descriptives_folder,
         width = 10,
         height = 8,
         dpi = 300)

# 4. COUNTS ---------------------------------------------------------------

# Count number of ids where either follicular or luteal is <5
num_ids_below_threshold <- sum(plot_data$low_data_flag)
cat("Number of IDs with fewer than 5 days of 'dep' data in either phase:", num_ids_below_threshold, "\n")

# Add number that do have >=5 in both phases
num_ids_above_threshold <- sum(!plot_data$low_data_flag)
cat("Number of IDs with 5 or more 'dep' days in both phases:", num_ids_above_threshold, "\n")
```

# ------ OVERLAP IN AM/PM ELIGBILITY at 5 OBS PER PHASE ------

```{r}
# Create a venn diagram of (1) those with >=5 observations in both phases for drink_today, (2) those with >=5 observations in both phases for dep 


library(ggvenn)
# Step 1: Identify ids with >=5 non-missing observations for each variable in both phases
ids_with_drink_today <- alcsmmelig$by_id %>%
  filter(drink_today_follicular >= 5 & drink_today_luteal >= 5) %>%
  pull(id)

ids_with_dep <- depsmmelig$by_id %>%
  filter(dep_follicular >= 5 & dep_luteal >= 5) %>%
  pull(id)

# Step 2: Create a list of sets for the Venn diagram
sets <- list(
  DrinkToday = ids_with_drink_today,
  Dep = ids_with_dep
)

# Generate the Venn diagram
venn_plot <- ggvenn(
  sets, 
  show_percentage = FALSE,   # Show percentages in each intersection
  fill_color = c("#E69F00", "pink"),  # Colors for each circle
  stroke_size = 0.8,        # Border thickness
  set_name_size = 20,        # Set names size (Surveys, Hormones, CycleData)
  text_size = 20            # Size of numbers/percentages in intersections
) +
  ggtitle("Overlap of IDs with >=5 Observations in Both Phases for:") +
  theme(
    plot.title = element_text(size = 50, hjust = 0.5),  # Larger, centered title
    text = element_text(size = 20)                      # General text size adjustments
  )

# Display the plot
print(venn_plot)
# Save the plot with high resolution
save_png(plot = venn_plot, 
         filename = "venn_diagram_drink_dep.png",
         folder = descriptives_folder,
         height = 8)  # Keep taller for Venn diagram



```




```{r}
# ------ SMM ELIGIBILITY at 4 OBS PER PHASE - MORNING/ALCOHOL OUTCOMES ------
library(dplyr)
library(ggplot2)
library(tidyr) 

# 1. DATA PREP ------------------------------------------------------------
# Save smm eligibility data (saved as unique object '_4')
alcsmmelig_4 <- cycledata_check(df, "drink_today")

# Merge back to df by id - ensuring we don't overwrite main df columns if running multiple times
# (We select specific columns to avoid duplication issues if you run this repeatedly)
df <- df %>%
  left_join(alcsmmelig_4$by_id %>% select(id, drink_today_follicular, drink_today_luteal), by = "id", suffix = c("", "_4obs"))

# Create a clean plotting dataset 
plot_data_4 <- alcsmmelig_4$by_id %>%
  mutate(
    # Create a boolean flag: TRUE if EITHER phase is < 4
    low_data_flag = drink_today_follicular < 4 | drink_today_luteal < 4,
    status = ifelse(low_data_flag, "Excluded (<4 days)", "Included")
  )

# 2. INDIVIDUAL PLOTS -----------------------------------------------------

# Follicular Plot (Object renamed to end in _4)
p_drink_follicular_4 <- ggplot(plot_data_4, aes(x = id, y = drink_today_follicular, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 4, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 4, label = "4-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Follicular Phase Drinking Days (4 Obs Threshold)", y = "Days", x = "Participant ID") +
  theme_minimal()

print(p_drink_follicular_4)

# Luteal Plot
p_drink_luteal_4 <- ggplot(plot_data_4, aes(x = id, y = drink_today_luteal, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 4, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 4, label = "4-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Luteal Phase Drinking Days (4 Obs Threshold)", y = "Days", x = "Participant ID") +
  theme_minimal()

print(p_drink_luteal_4)


# 3. COMBINED PLOT --------------------------------------------------------

p_combined_drink_4 <- plot_data_4 %>%
  pivot_longer(cols = c(drink_today_follicular, drink_today_luteal), 
               names_to = "phase", 
               values_to = "days") %>%
  mutate(phase = recode(phase, "drink_today_follicular" = "Follicular", "drink_today_luteal" = "Luteal")) %>%
  ggplot(aes(x = id, y = days, fill = status)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) + 
  geom_hline(yintercept = 4, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  labs(title = "Drinking Days by Phase (Red = <4 days in EITHER phase)", 
       y = "Number of Drinking Days", 
       x = "Participant ID") +
  facet_wrap(~phase, ncol = 1) + 
  theme_minimal(base_size = 14)

print(p_combined_drink_4)

# SAVE WITH UNIQUE FILENAME
save_png(plot = p_combined_drink_4,
         filename = "combined_drink_data_availability_4obs.png", 
         folder = descriptives_folder,
         width = 10,
         height = 8,
         dpi = 300)


# 4. COUNTS ---------------------------------------------------------------

num_ids_below_threshold_4 <- sum(plot_data_4$low_data_flag)
cat("IDs with fewer than 4 drinking days in either phase:", num_ids_below_threshold_4, "\n")

num_ids_above_threshold_4 <- sum(!plot_data_4$low_data_flag)
cat("IDs with 4 or more drinking days in both phases:", num_ids_above_threshold_4, "\n")
```


```{r}
# ------ SMM ELIGIBILITY at 4 OBS PER PHASE - EVENING/DRSP OUTCOMES ------

library(dplyr)
library(ggplot2)
library(tidyr)

# 1. DATA PREP ------------------------------------------------------------
# Save smm eligibility data unique object
depsmmelig_4 <- cycledata_check(df, "dep")

# Merge back
df <- df %>%
  left_join(depsmmelig_4$by_id %>% select(id, dep_follicular, dep_luteal), by = "id", suffix = c("", "_4obs"))

# Create plotting dataset
plot_data_dep_4 <- depsmmelig_4$by_id %>%
  mutate(
    # Create a boolean flag: TRUE if EITHER phase is < 4
    low_data_flag = dep_follicular < 4 | dep_luteal < 4,
    status = ifelse(low_data_flag, "Excluded (<4 days)", "Included")
  )

# 2. INDIVIDUAL PLOTS -----------------------------------------------------

# Follicular Plot
p_dep_follicular_4 <- ggplot(plot_data_dep_4, aes(x = id, y = dep_follicular, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 4, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 4, label = "4-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Follicular Phase Data (Dep) - 4 Obs", y = "Days Present", x = "Participant ID") +
  theme_minimal()

print(p_dep_follicular_4)

# Luteal Plot
p_dep_luteal_4 <- ggplot(plot_data_dep_4, aes(x = id, y = dep_luteal, fill = status)) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 4, linetype = "dashed", color = "red") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  annotate("text", x = Inf, y = 4, label = "4-day threshold", hjust = 1.1, vjust = -0.5, color = "red", size = 4) +
  labs(title = "Luteal Phase Data (Dep) - 4 Obs", y = "Days Present", x = "Participant ID") +
  theme_minimal()

print(p_dep_luteal_4)


# 3. COMBINED PLOT --------------------------------------------------------

p_combined_dep_4 <- plot_data_dep_4 %>%
  pivot_longer(cols = c(dep_follicular, dep_luteal), 
               names_to = "phase", 
               values_to = "days") %>%
  mutate(phase = recode(phase, "dep_follicular" = "Follicular", "dep_luteal" = "Luteal")) %>%
  ggplot(aes(x = id, y = days, fill = status)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) + 
  geom_hline(yintercept = 4, linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("Excluded (<4 days)" = "red", "Included" = "lightblue")) +
  labs(title = "Data Availability by Phase (Red = <4 days in EITHER phase)", 
       y = "Number of Days with Data", 
       x = "Participant ID") +
  facet_wrap(~phase, ncol = 1) + 
  theme_minimal(base_size = 14)

print(p_combined_dep_4)

# SAVE WITH UNIQUE FILENAME
save_png(plot = p_combined_dep_4,
         filename = "combined_dep_data_availability_4obs.png",
         folder = descriptives_folder,
         width = 10,
         height = 8,
         dpi = 300)

# 4. COUNTS ---------------------------------------------------------------

num_ids_below_dep_4 <- sum(plot_data_dep_4$low_data_flag)
cat("IDs with fewer than 4 days of 'dep' data in either phase:", num_ids_below_dep_4, "\n")

num_ids_above_dep_4 <- sum(!plot_data_dep_4$low_data_flag)
cat("IDs with 4 or more 'dep' days in both phases:", num_ids_above_dep_4, "\n")
```


```{r}
# ------ OVERLAP IN AM/PM ELIGBILITY at 4 OBS PER PHASE ------

library(ggvenn)

# Step 1: Use the new unique objects (_4) to filter
ids_with_drink_today_4 <- alcsmmelig_4$by_id %>%
  filter(drink_today_follicular >= 4 & drink_today_luteal >= 4) %>%
  pull(id)

ids_with_dep_4 <- depsmmelig_4$by_id %>%
  filter(dep_follicular >= 4 & dep_luteal >= 4) %>%
  pull(id)

# Step 2: Create a list of sets
sets_4 <- list(
  DrinkToday = ids_with_drink_today_4,
  Dep = ids_with_dep_4
)

# Generate the Venn diagram
venn_plot_4 <- ggvenn(
  sets_4, 
  show_percentage = FALSE, 
  fill_color = c("#E69F00", "pink"), 
  stroke_size = 0.8,         
  set_name_size = 20,         
  text_size = 20             
) +
  ggtitle("Overlap of IDs with >=4 Observations in Both Phases") +
  theme(
    plot.title = element_text(size = 40, hjust = 0.5), # Slightly adjusted title size
    text = element_text(size = 20)                      
  )

print(venn_plot_4)

# SAVE WITH UNIQUE FILENAME
save_png(plot = venn_plot_4, 
         filename = "venn_diagram_drink_dep_4obs.png",
         folder = descriptives_folder,
         height = 8)
```






# Visualize >10 (non-SMM) data availability - BDEFS/DRSP outcomes
```{r venn-diagram-overlap-horm-surv-cyc, message=FALSE, warning=FALSE}
# Create a venn diagram (using ggvenn) to visualize the overlap of (1) ids with more than ten nonmissing observations of "dep", (2) ids with more than 10 nonmissing observations of e2, and (3) ids with more than 10 nonmissing observations of cyclic_time_impute_ov

# Step 1: Identify ids with more than 10 non-missing observations for each variable
ids_with_dep <- df %>%
  group_by(id) %>%
  summarize(nonmissing_dep = sum(!is.na(dep))) %>%
  filter(nonmissing_dep > 10) %>%
  pull(id)

ids_with_e2 <- df %>%
  group_by(id) %>%
  summarize(nonmissing_e2 = sum(!is.na(e2))) %>%
  filter(nonmissing_e2 > 10) %>%
  pull(id)

ids_with_cyclic_time_impute <- df %>%
  group_by(id) %>%
  summarize(nonmissing_cyclic_time_impute = sum(!is.na(cyclic_time_impute))) %>%
  filter(nonmissing_cyclic_time_impute > 10) %>%
  pull(id)

# Step 2: Create a list of sets for the Venn diagram
sets <- list(
  PMSurveys = ids_with_dep,
  Hormones = ids_with_e2,
  CycleData = ids_with_cyclic_time_impute
)

# Generate the Venn diagram
venn_plot <- ggvenn(
  sets, 
  show_percentage = FALSE,   # Show percentages in each intersection
  fill_color = c("pink", "#56B4E9", "#009E73"),  # Colors for each circle
  stroke_size = 0.8,        # Border thickness
  set_name_size = 20,        # Set names size (Surveys, Hormones, CycleData)
  text_size = 20            # Size of numbers/percentages in intersections
) +
  ggtitle("Overlap of IDs with More Than 10 Observations of:") +
  theme(
    plot.title = element_text(size = 50, hjust = 0.5),  # Larger, centered title
    text = element_text(size = 20)                      # General text size adjustments
  )

# Display the plot
print(venn_plot)

# Save the plot with high resolution
save_png(plot = venn_plot, 
         filename = "venn_diagram_drink_e2_scaledcycleday.png",
         folder = descriptives_folder,
         height = 8)  # Keep taller for Venn diagram


```
# Visualize data availability - Alcohol outcomes
```{r venn-diagram-overlap-horm-surv-cyc, message=FALSE, warning=FALSE}
# Create a venn diagram (using ggvenn) to visualize the overlap of (1) ids with more than ten nonmissing observations of "drink_today", (2) ids with more than 10 nonmissing observations of e2, and (3) ids with more than 10 nonmissing observations of cyclic_time_impute_ov

# Step 1: Identify ids with more than 10 non-missing observations for each variable
ids_with_drink_today <- df %>%
  group_by(id) %>%
  summarize(nonmissing_drink_today = sum(!is.na(drink_today))) %>%
  filter(nonmissing_drink_today > 10) %>%
  pull(id)

ids_with_e2 <- df %>%
  group_by(id) %>%
  summarize(nonmissing_e2 = sum(!is.na(e2))) %>%
  filter(nonmissing_e2 > 10) %>%
  pull(id)

ids_with_cyclic_time_impute <- df %>%
  group_by(id) %>%
  summarize(nonmissing_cyclic_time_impute = sum(!is.na(cyclic_time_impute))) %>%
  filter(nonmissing_cyclic_time_impute > 10) %>%
  pull(id)

# Step 2: Create a list of sets for the Venn diagram
sets <- list(
  AMSurveys = ids_with_drink_today,
  Hormones = ids_with_e2,
  CycleData = ids_with_cyclic_time_impute
)

# Generate the Venn diagram
venn_plot <- ggvenn(
  sets, 
  show_percentage = FALSE,   # Show percentages in each intersection
  fill_color = c("#E69F00", "#56B4E9", "#009E73"),  # Colors for each circle
  stroke_size = 0.8,        # Border thickness
  set_name_size = 20,        # Set names size (Surveys, Hormones, CycleData)
  text_size = 20            # Size of numbers/percentages in intersections
) +
  ggtitle("Overlap of IDs with More Than 10 Observations of:") +
  theme(
    plot.title = element_text(size = 50, hjust = 0.5),  # Larger, centered title
    text = element_text(size = 20)                      # General text size adjustments
  )

# Display the plot
print(venn_plot)

# Save the plot with high resolution
save_png(plot = venn_plot, 
         filename = "venn_diagram_drink_today_e2_scaledcycleday.png",
         folder = descriptives_folder,
         height = 8)  # Keep taller for Venn diagram


```


# Visualize Percentage of Days Drinking
```{r}
# 1. First, ensure your data exists and has values
# Check if df exists
if (!exists("df")) stop("Dataframe 'df' not found!")

# Check if the column is empty
if (all(is.na(df$drink_today))) stop("Column 'drink_today' is strictly N/A!")

# 2. Prepare data
drink_table <- table(df$drink_today, useNA = "no")

# Check if table is empty after removing NAs
if (length(drink_table) == 0) stop("Table is empty. Check if all data was NA.")

drink_percentages <- round(100 * prop.table(drink_table), 1)
drink_labels <- paste(names(drink_table), ": ", drink_table, " (", drink_percentages, "%)", sep = "")

# 3. PLOT TO SCREEN FIRST (Debug step)
# We run this WITHOUT png() to ensure the code works visually.
pie(drink_table, 
    labels = drink_labels, 
    col = c("skyblue", "lightcoral"), 
    main = "Percentage of Days Drinking")

# 4. Save to file using ragg device
# define output_folder if it doesn't exist (using current directory as fallback)
if (!exists("output_folder")) {
  output_folder <- getwd() 
  message("output_folder was missing. Saving to current directory: ", output_folder)
}

# Save PNG using ragg device
ragg::agg_png(filename = file.path(output_folder, "drink_today_pie_chart.png"), 
              width = 7, height = 5, units = "in", res = 300, bg = "white")
pie(drink_table, labels = drink_labels, col = c("skyblue", "lightcoral"), main = "Percentage of Days Drinking")
dev.off()

# Save SVG version with proper text sizing
# Set pointsize = 14 to match our standard base_size and prevent text overlap
svg(filename = file.path(output_folder, "drink_today_pie_chart.svg"), 
    width = 7, height = 5, bg = "white", pointsize = 14)
pie(drink_table, labels = drink_labels, col = c("skyblue", "lightcoral"), main = "Percentage of Days Drinking")
dev.off()

# Visualize percentage of days drinking by id
drink_by_id <- df %>%
  group_by(id) %>%
  summarize(
    total_days = n(),
    drinking_days = sum(drink_today == 1, na.rm = TRUE),
    percent_drinking = (drinking_days / total_days) * 100
  )
# Boxplot of percent drinking by id
p_drink_by_id <- ggplot(drink_by_id, aes(x = id, y = percent_drinking)) +
  geom_boxplot(fill = "lightgreen", alpha = 0.7) +
  labs(title = "Percentage of Days Drinking by Participant ID", x = "Participant ID", y = "Percentage of Days Drinking") +
  theme_minimal(base_size = 14)
print(p_drink_by_id)
save_png(plot = p_drink_by_id, 
         filename = "percent_drinking_by_id_boxplot.png",
         folder = descriptives_folder)

# hsitogram of days drinking by id
p_hist_drink_by_id <- ggplot(drink_by_id, aes(x = percent_drinking)) +
  geom_histogram(binwidth = 5, fill = "lightblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Percentage of Days Drinking by Participant ID", x = "Percentage of Days Drinking", y = "Count of Participants") +
  theme_minimal(base_size = 14)
print(p_hist_drink_by_id)
save_png(plot = p_hist_drink_by_id, 
         filename = "histogram_percent_drinking_by_id.png",
         folder = descriptives_folder)
```


```{r}
# Source the function and create variables

# Variables needed for this plotting chunk - only use variables that exist
drsp_vars <- c(
  "dep", "hopeless", "worthless", "anxious", "mood_swing", "rej_sens", 
  "angry", "conflicts", "loss_interest", "diff_conc", "tired", 
  "over_eat", "sp_food_crav", "high_sleep", "diff_sleep", "overwhelm", 
  "loss_control", "breast_tend", "breast_swell", "headache", 
  "jm_pain", "loss_prod", "int_relat"
)

bdefs_vars <- c(
  "diff_inhib", "diff_stop", "diff_change", "immed_payoff", 
  "res_urge", "diff_long_rew", "diff_res_rew"
)

hormlist <- c("e2", "p4", "lh")
hormwinsor <- c("e2_w", "p4_w", "lh_w")

daily_items_raw <- c(
  "num_drinks_today", "drink_today", "drink_today_bin", "fourplustoday", "fourplustoday_bin",
  "crave_alc_am", "crave_alc_pm", "bdefs_total", "bdefs_delayaversion", "bdefs_impulsiveaction", "drsp_total", "drsp_distress", "drsp_eatsx", "drsp_breastpain", "drsp_lowarousal")

# Combine and filter to only existing variables
alldailyvars_desired <- unique(c(daily_items_raw, drsp_vars, bdefs_vars, hormlist, hormwinsor))
alldailyvars <- intersect(alldailyvars_desired, names(df))


plot_day_of_week_effects <- function(df, outcome_vars) {
  # Set output directory path to descriptives subfolder
  output_folder_path <- descriptives_folder
  
  # Ensure day_of_week is an ordered factor
  df <- df %>%
    mutate(day_of_week = factor(day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")))
  
  # Define sine and cosine positions for each day (without labels)
  label_positions <- data.frame(
    day_of_week_sin = sin(2 * pi * (0:6) / 7),
    day_of_week_cos = cos(2 * pi * (0:6) / 7)
  )
  
  # Create output directory if it doesn’t exist
  if (!dir.exists(output_folder_path)) {
    dir.create(output_folder_path, recursive = TRUE)
  }
  
  # Helper function to plot smooth effect (sin/cos) for a given outcome
  plot_smooth_effect <- function(df, outcome_var, x_var, title_suffix, output_filename) {
    outcome_sym <- sym(outcome_var)
    x_sym <- sym(x_var)
    
    plot <- ggplot(df %>% filter(!is.na(!!x_sym) & !is.na(!!outcome_sym)), 
                   aes(x = !!x_sym, y = !!outcome_sym)) +
      geom_smooth(method = "loess") +
      labs(
        title = paste("Effects of Day of Week on", outcome_var, title_suffix),
        x = paste("Day of Week", title_suffix),
        y = outcome_var
      ) +
      theme_minimal(base_size = 14)
    
    # Save the plot
    save_png(plot = plot, 
             filename = output_filename,
             folder = output_folder_path)
  }
  
  # Loop over each outcome variable to generate the plots
  for (outcome_var in outcome_vars) {
    # Plot for day_of_week_sin
    plot_smooth_effect(df, outcome_var, "day_of_week_sin", "(sin)",
                       paste0("day_of_week_sin_", outcome_var, ".png"))
    
    # Plot for day_of_week_cos
    plot_smooth_effect(df, outcome_var, "day_of_week_cos", "(cos)",
                       paste0("day_of_week_cos_", outcome_var, ".png"))
    
    # Calculate weekly average or proportion for bar plot
    weekly_summary <- df %>%
      group_by(day_of_week) %>%
      summarize(mean_outcome = mean(!!sym(outcome_var), na.rm = TRUE)) %>%
      ungroup()
    
    # Plot the weekly average/proportion by day of the week (bar plot)
    plot_weekly <- ggplot(weekly_summary, aes(x = day_of_week, y = mean_outcome)) +
      geom_bar(stat = "identity", fill = "skyblue") +
      labs(
        title = paste("Weekly Average of", outcome_var, "by Day of the Week"),
        x = "Day of the Week",
        y = paste("Average", outcome_var)
      ) +
      theme_minimal(base_size = 14)
    
    # Save the bar plot for weekly summary by day of the week
    save_png(plot = plot_weekly, 
             filename = paste0("weekly_summary_", outcome_var, ".png"),
             folder = output_folder_path)
  }
  
  return("Plots generated and saved successfully.")
}

# Call the function
plot_day_of_week_effects(df = df, outcome_vars = alldailyvars)
```


```{r summarize-drinks-by-session}
df %>%
  group_by(session_today) %>%
  summarize(mean_drinks = mean(num_drinks_today, na.rm = TRUE),
            sd_drinks = sd(num_drinks_today, na.rm = TRUE),
            n = n())

# save to output folder
write.csv(
  df %>%
    group_by(session_today) %>%
    summarize(mean_drinks = mean(num_drinks_today, na.rm = TRUE),
              sd_drinks = sd(num_drinks_today, na.rm = TRUE),
              n = n()),
  file = file.path(output_folder, "summary_drinks_by_session.csv"),
  row.names = FALSE
)
```

# Print histogram of all e2 variables to visually inspect for outliers
```{r histograms}

names(df)
# Visualize distributions of e2, e2.zd, e2.szd, and log(e2 + 1)
# Histogram of raw e2 values

e2_histogram <- df %>%
  ggplot(aes(x = e2_w)) +
  geom_histogram(bins = 100, fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of e2 Values", x = "e2", y = "Count") +
  theme_light(base_size = 14)

print(e2_histogram)

save_png(plot = e2_histogram, 
         filename = "e2_histogram.png",
         folder = descriptives_folder)

# now do the same for person-standardized values
e2_person_std_histogram <- df %>%
  ggplot(aes(x = e2_w.zd)) +
  geom_histogram(bins = 100, fill = "darkgreen", alpha = 0.7) +
  labs(title = "Histogram of Person-Standardized e2 Values", x = "e2 Person-Standardized (d)", y = "Count") +
  theme_light(base_size = 14) 
print(e2_person_std_histogram)
save_png(plot = e2_person_std_histogram, 
         filename = "e2_person_standardized_histogram.png",
         folder = descriptives_folder)

# now do the same for sample-standardized values
e2_sample_std_histogram <- df %>%
  ggplot(aes(x = e2_w.szd)) +
  geom_histogram(bins = 100, fill = "purple", alpha = 0.7) +
  labs(title = "Histogram of Sample-Standardized e2 Values", x = "e2 Sample-Standardized (szd)", y = "Count") +
  theme_light(base_size = 14)
print(e2_sample_std_histogram)
save_png(plot = e2_sample_std_histogram, 
         filename = "e2_sample_standardized_histogram.png",
         folder = descriptives_folder)

# Test whether a log transform improves the distribution
e2_log_histogram <- df %>%
  mutate(E2_log = log(e2_w + 1)) %>% # Add 1
  ggplot(aes(x = E2_log)) +
  geom_histogram(bins = 100, fill = "orange", alpha = 0.7)+
  labs(title
       = "Histogram of Log-Transformed e2 Values", x = "Log(e2 + 1)", y = "Count") +
  theme_light(base_size = 14)
print(e2_log_histogram)
save_png(plot = e2_log_histogram,
         filename = "e2_log_transformed_histogram.png",
         folder = descriptives_folder)

# Stack the four histograms together for comparison
combined_e2_histograms <- ggarrange(
  e2_histogram + rremove("x.text") + rremove("x.title"),
  e2_person_std_histogram + rremove("x.text") + rremove("x.title"),
  e2_sample_std_histogram + rremove("x.text") + rremove("x.title"),
  e2_log_histogram,
  ncol = 1, nrow = 4,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE, legend = "right"
)
print(combined_e2_histograms)
save_png(plot = combined_e2_histograms,
         filename = "combined_e2_histograms.png",
         folder = descriptives_folder,
         height = 16)  # Keep tall height for stacked histograms

```
# Print histogram of all p4 variables to visually inspect for outliers
```{r histograms2}
# Visualize distributions of p4, p4.zd, p4.szd, and
# log(p4 + 1)
# Histogram of raw p4 values
p4_histogram <- df %>%
  ggplot(aes(x = p4_w)) +
  geom_histogram(bins = 100, fill = "blue", alpha = 0.7) +
  labs(title = "Histogram of p4 Values", x = "p4", y = "Count") +
  theme_light(base_size = 14)
print(p4_histogram)
save_png(plot = p4_histogram, 
         filename = "p4_histogram.png",
         folder = descriptives_folder)

# now do the same for person-standardized values
p4_person_std_histogram <- df %>%
  ggplot(aes(x = p4_w.zd)) +
  geom_histogram(bins = 100, fill = "darkgreen", alpha = 0.7) +
  labs(title = "Histogram of Person-Standardized p4 Values", x = "p4 Person-Standardized (d)", y = "Count") +
  theme_light(base_size = 14)
print(p4_person_std_histogram)
save_png(plot = p4_person_std_histogram,
         filename = "p4_person_standardized_histogram.png",
         folder = descriptives_folder)

# now do the same for sample-standardized values
p4_sample_std_histogram <- df %>%
  ggplot(aes(x = p4_w.szd)) +
  geom_histogram(bins = 100, fill = "purple", alpha = 0.7) +
  labs(title = "Histogram of Sample-Standardized p4 Values", x = "p4 Sample-Standardized (szd)", y = "Count") +
  theme_light(base_size = 14)
print(p4_sample_std_histogram)
save_png(plot = p4_sample_std_histogram,
         filename = "p4_sample_standardized_histogram.png",
         folder = descriptives_folder)

# Test whether a log transform improves the distribution
p4_log_histogram <- df %>%
  mutate(P4_log = log(p4_w + 1)) %>% # Add 1
  ggplot(aes(x = P4_log)) +
  geom_histogram(bins = 100, fill = "orange", alpha = 0.7) +
  labs(title  = "Histogram of Log-Transformed p4 Values", x = "Log(p4 + 1)", y = "Count") +
  theme_light(base_size = 14)
print(p4_log_histogram)
save_png(plot = p4_log_histogram,
         filename = "p4_log_transformed_histogram.png",
         folder = descriptives_folder)

# Stack the four histograms together for comparison
combined_p4_histograms <- ggarrange(
  p4_histogram + rremove("x.text") + rremove("x.title"),
  p4_person_std_histogram + rremove("x.text") + rremove("x.title"),
  p4_sample_std_histogram + rremove("x.text") + rremove("x.title"),
  p4_log_histogram,
  ncol = 1, nrow = 4,
  labels = c("A", "B", "C", "D"),
  common.legend = TRUE, legend = "right")
print(combined_p4_histograms)
save_png(plot = combined_p4_histograms,
         filename = "combined_p4_histograms.png",
         folder = descriptives_folder,
         height = 16)  # Keep tall height for stacked histograms

```


```{r}
# histogram of BAC values

hist(df$bac_peak_percent)

# boxplot of BAC values by id
p_bac <- ggplot(df, aes(x = as.factor(id), y = bac_peak_percent)) +
  geom_boxplot() +
  labs(title = "BAC Peak Percent by Participant ID", x = "Participant ID", y = "BAC Peak Percent") +
  theme_minimal(base_size = 14)

p_bac
save_png(plot = p_bac, 
         filename = "bac_by_id_boxplot.png",
         folder = descriptives_folder)

# winsorize peak bac at 99th percentile within person 
df <- df %>%
  group_by(id) %>%
  mutate(bac_peak_percent_w = ifelse(
    bac_peak_percent > quantile(bac_peak_percent, 0.99, na.rm = TRUE),
    quantile(bac_peak_percent, 0.99, na.rm = TRUE),
    bac_peak_percent
  )) %>%
  ungroup()
# boxplot of winsorized BAC values by id
p_bac_w <- ggplot(df, aes(x = as.factor(id), y = bac_peak_percent_w)) +
  geom_boxplot() +
  labs(title = "Winsorized BAC Peak Percent by Participant ID", x = "Participant ID", y = "Winsorized BAC Peak Percent") +
  theme_minimal(base_size = 14)
p_bac_w
save_png(plot = p_bac_w, 
         filename = "winsorized_bac_by_id_boxplot.png",
         folder = descriptives_folder)



```


```{r}
# -------------------------------------------------------------------------
# Violin Plots Loop
# -------------------------------------------------------------------------


# 2. Loop through the variables
# (Assumes 'continuous_vars' is already defined from your previous scripts)
for (outcome_name in continuous_vars) {
  
  message(paste0("Plotting violin for: ", outcome_name))
  
  # Create the plot using .data[[outcome_name]] for dynamic selection
  p <- ggplot(df, aes(x = "", y = .data[[outcome_name]])) +
    geom_violin(fill = "lightblue", color = "black", alpha = 0.7) +
    geom_boxplot(width = 0.1, position = position_dodge(0.9), outlier.color = "red", outlier.size = 2) +
    labs(
      title = paste0("Violin Plot of ", outcome_name),
      y = outcome_name,
      x = ""
    ) +
    theme_light(base_size = 14)
  
  # 3. Save the plot
  save_png(
    plot = p,
    filename = paste0("violin_", outcome_name, ".png"),
    folder = descriptives_folder
  )
  
  # Create a histogram for each and save that too
  p_hist <- ggplot(df, aes(x = .data[[outcome_name]])) +
    geom_histogram(bins = 30, fill = "lightblue", color = "black", alpha = 0.7) +
    labs(
      title = paste0("Histogram of ", outcome_name),
      x = outcome_name,
      y = "Count"
    ) +
    theme_light(base_size = 14)
  save_png(
    plot = p_hist,
    filename = paste0("histogram_", outcome_name, ".png"),
    folder = descriptives_folder
  )
}

```




# ICCs, Between- and Within-Person Associations of Study Variables
```{r f1_corr, echo=TRUE, message=FALSE, warning=FALSE}
# ===================================================================
# 1. Load Required Libraries
# ===================================================================
library(tidyverse)
library(rmcorr)
library(performance)
library(lme4)
library(ggplot2)

# ===================================================================
# 2. Define the Heatmap Generation Function (Final Version)
# ===================================================================
generate_correlation_heatmap <- function(
    data, outcome_vars, id_var, final_labels = NULL, # <-- New argument
    sig.level = 0.05,
    highlight_diag = TRUE,
    show_legend = TRUE,
    label_size = 4.6,
    diag_label_size = 6,
    subtitle_text = "Upper: Within-person (rmcorr); Lower: Between-person (Spearman); Diagonal: ICC"
) {
  # (The first part of the function is the same as the last version)
  n_vars <- length(outcome_vars)
  results_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars,
                           dimnames = list(outcome_vars, outcome_vars))
  pval_matrix <- matrix(NA, nrow = n_vars, ncol = n_vars,
                        dimnames = list(outcome_vars, outcome_vars))
  
  summary_expressions <- purrr::map(outcome_vars, ~ rlang::quo(mean(!!rlang::sym(.x), na.rm = TRUE))) %>%
    rlang::set_names(outcome_vars)
    
  person_means <- data %>%
    group_by(!!rlang::sym(id_var)) %>%
    summarise(!!!summary_expressions, .groups = "drop")

  between_corr <- suppressWarnings(
    cor(person_means[, outcome_vars], method = "spearman", use = "pairwise.complete.obs")
  )

  for (i in 1:(n_vars - 1)) for (j in (i + 1):n_vars) {
    x <- person_means[[outcome_vars[i]]]
    y <- person_means[[outcome_vars[j]]]
    if (sum(!is.na(x) & !is.na(y)) > 2) {
      test <- suppressWarnings(cor.test(x, y, method = "spearman"))
      results_matrix[i, j] <- between_corr[i, j]
      pval_matrix[i, j] <- test$p.value
    }
  }

  for (i in 1:n_vars) {
    var <- outcome_vars[i]
    fml <- as.formula(paste0("`", var, "` ~ 1 + (1|`", id_var, "`)"))
    if (sum(!is.na(data[[var]])) > 1) {
      model_fit <- try(lmer(fml, data = data), silent = TRUE)
      if (!inherits(model_fit, "try-error")) {
        icc_val <- performance::icc(model_fit)$ICC_adjusted
        results_matrix[i, i] <- icc_val
        pval_matrix[i, i] <- 0
      }
    }
  }

  for (i in 2:n_vars) for (j in 1:(i - 1)) {
    var1 <- outcome_vars[i]; var2 <- outcome_vars[j]
    rmcorr_data <- data %>% select(all_of(c(id_var, var1, var2))) %>% drop_na()
    if (nrow(rmcorr_data) > 2) {
      rmcorr_result <- rmcorr::rmcorr(participant = id_var, measure1 = var1, measure2 = var2, dataset = rmcorr_data)
      results_matrix[i, j] <- rmcorr_result$r
      pval_matrix[i, j] <- rmcorr_result$p
    }
  }

  within_corr_matrix <- results_matrix
  within_corr_matrix[upper.tri(within_corr_matrix, diag = TRUE)] <- 0
  dist_matrix <- as.dist(1 - abs(within_corr_matrix))
  hclust_obj <- hclust(dist_matrix, method = "ward.D2")
  corr_order <- rownames(within_corr_matrix)[hclust_obj$order]

  results_matrix_ordered <- results_matrix[corr_order, corr_order]
  pval_matrix_ordered    <- pval_matrix[corr_order, corr_order]

  plot_df <- as.data.frame(as.table(results_matrix_ordered))
  names(plot_df) <- c("Var1", "Var2", "value")
  plot_df$p.value <- as.vector(pval_matrix_ordered)
  plot_df$Var1 <- factor(plot_df$Var1, levels = rev(corr_order))
  plot_df$Var2 <- factor(plot_df$Var2, levels = corr_order)

  no_leading_zero <- function(x) {
    ifelse(is.na(x), NA_character_, sub("^(-?)0\\.", "\\1.", sprintf("%.2f", x)))
  }

  plot_df <- plot_df %>%
    mutate(
      is_diag = as.character(Var1) == as.character(Var2),
      is_sig  = p.value < sig.level,
      label   = no_leading_zero(value)
    )

  diag_df    <- plot_df %>% filter(is_diag)
  sig_df     <- plot_df %>% filter(!is_diag & is_sig)
  nonsig_df  <- plot_df %>% filter(!is_diag & !is_sig)

  p <- ggplot() +
    geom_tile(data = plot_df %>% filter(!is_diag),
              aes(x = Var2, y = Var1, fill = value), color = "gray70", linewidth = 0.5) +
    geom_tile(data = diag_df, aes(x = Var2, y = Var1),
              fill = "gray92", color = "gray70", linewidth = 0.5) +
    scale_fill_gradient2(low = "#6D9EC1", mid = "white", high = "#E46726",
                       midpoint = 0, limits = c(-1, 1), name = "Correlation") +
    labs(title = "Multilevel Correlation Matrix of Daily Outcomes", subtitle = subtitle_text) +
    theme_light(base_size = 14) +
    theme(
      plot.title    = element_text(hjust = 0.5, face = "bold", size = 20),
      plot.subtitle = element_text(hjust = 0.5, face = "italic", size = 12, margin = margin(b = 6)),
      axis.text.x   = element_text(angle = 45, vjust = 1, hjust = 1),
      axis.text.y   = element_text(face = "bold"),
      axis.title    = element_blank(),
      legend.position = if (show_legend) "right" else "none",
      panel.grid.major = element_blank()
    )

  p <- p +
    geom_text(data = sig_df, aes(x = Var2, y = Var1, label = label),
              fontface = "bold", color = "black", size = label_size, na.rm = TRUE) +
    geom_text(data = nonsig_df, aes(x = Var2, y = Var1, label = label),
              color = "gray40", size = label_size, na.rm = TRUE)

  if (highlight_diag && nrow(diag_df) > 0) {
    p <- p +
      geom_text(data = diag_df, aes(x = Var2, y = Var1, label = label),
                fontface = "bold", color = "#4B0082", size = diag_label_size, na.rm = TRUE)
  }
  
  # === NEW CODE TO RELABEL AXES AT THE END ===
  if (!is.null(final_labels)) {
    p <- p +
      scale_x_discrete(labels = final_labels) +
      scale_y_discrete(labels = final_labels)
  }
  # ===========================================

  return(p)
}

# ===================================================================
# 3. Define Parameters and Run the Plot
# ===================================================================
# Technical names mapped to the desired "pretty" labels
corrplotlist <- c(
  # --- Substance Use & Craving ---
  "num_drinks_today" = "Number of Drinks",
  "bac_peak_percent" = "Peak BAC (%)",
  #"drink_today"      = "Any Drinking Today",
  #"fourplustoday"    = "Binge Drinking (4+)",
  #"felt_drunk"       = "Felt Drunk",
  #"crave_alc_am"     = "Alcohol Craving (AM)",
  "crave_alc_pm"     = "Alcohol Craving (PM)",
  #"crave_nic_am"     = "Nicotine Craving (AM)",
  #"crave_nic_pm"     = "Nicotine Craving (PM)",
  #"crave_mar_am"     = "Marijuana Craving (AM)",
  #"crave_mar_pm"     = "Marijuana Craving (PM)",

  # --- Impulsivity & Reward ---
  #"diff_stop"        = "Difficulty Stopping",
  #"diff_change"      = "Difficulty Changing",
  #"immed_payoff"     = "Immediate Payoff Pref.",
  #"res_urge"         = "Resistance to Urges",
  #"diff_long_rew"    = "Diff. Waiting Long Reward",
  #"diff_res_rew"     = "Diff. Resisting Reward",
  "bdefs_impulsiveaction" = "Impulsive Action (BDEFS)",
  "bdefs_delayaversion" = "Delay Aversion (BDEFS)",

  # --- Mood & Affect (DRSP/PMDD) ---
  #"dep"              = "Depression",
  #"hopeless"         = "Hopelessness",
  #"worthless"        = "Worthlessness",
  #"anxious"          = "Anxiety",
  #"mood_swing"       = "Mood Swings",
  #"rej_sens"         = "Rejection Sensitivity",
  #"angry"            = "Anger/Irritability",
  #"conflicts"        = "Interpersonal Conflicts",
  #"loss_interest"    = "Loss of Interest",
  #"overwhelm"        = "Feeling Overwhelmed",
  #"loss_control"     = "Feeling Out of Control",
  #"int_relat"        = "Relationship Impairment",
  
    # --- DRSP Aggregate Scores ---
  "drsp_total"       = "DRSP Total",
  "drsp_distress"    = "Distress (DRSP)",
  "drsp_labilsocsx"  = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"  = "Breast Sx (DRSP)",
  "drsp_eatsx"       = "Eating Sx",
  "drsp_lowarousal"  = "Low Arousal Sx (DRSP)",

  # --- Physical & Cognitive Symptoms ---
  #"diff_conc"        = "Difficulty Concentrating",
  #"tired"            = "Fatigue/Lethargy",
  #"high_sleep"       = "Hypersomnia (High Sleep)",
  "diff_sleep"       = "Insomnia/Sleep Diff.",
  #"over_eat"         = "Overeating",
  #"sp_food_crav"     = "Specific Food Cravings",
  #"breast_tend"      = "Breast Tenderness",
  #"breast_swell"     = "Breast Swelling",
  #"headache"         = "Headache",
  "jm_pain"          = "Joint/Muscle Pain",
  #"loss_prod"        = "Decreased Productivity",
  
  # --- Hormones  ---
  "e2_w"             = "Estradiol (E2)",
  "p4_w"             = "Progesterone (P4)",
  "lh_w"             = "Luteinizing Hormone (LH)"
)

# This assumes `df` exists in your environment
if (exists("df") && exists("corrplotlist")) {

  # === NEW: Create safe labels for calculation and a lookup table for plotting ===
  # 1. Create safe labels (e.g., "Hyperactive.Impulsive.Sx") for calculations
  safe_labels <- make.names(corrplotlist)
  
  # 2. Create a lookup table to map safe labels back to pretty labels
  label_lookup <- setNames(corrplotlist, safe_labels)
  # ===============================================================================

  # Prepare the data by renaming columns to the SAFE labels
  original_vars <- names(corrplotlist)
  plot_data <- df[, c("id", original_vars)]
  names(plot_data)[match(original_vars, names(plot_data))] <- safe_labels # Use safe_labels here

  # Call the function with the safe labels for calculation AND the lookup table for the final plot
  correlation_heatmap <- generate_correlation_heatmap(
    data = plot_data,
    outcome_vars = safe_labels, # Use safe_labels here
    id_var = "id",
    final_labels = label_lookup # Pass the lookup table for final relabeling
  )

  print(correlation_heatmap)

  # Define the final filename, including today's date
dated_output_file <- file.path(output_folder, 
                               paste0("Fig1_corr_", 
                                      format(Sys.Date(), "%Y%m%d"), 
                                      ".png"))

# Save the plot directly to the final, dated file
save_png(plot = correlation_heatmap,
         filename = paste0("Fig1_corr_", 
                          format(Sys.Date(), "%Y%m%d"), 
                          ".png"),
         folder = descriptives_folder,
         width = 11,
         height = 8.5)

} else {
  
  stop("Error: 'df' or 'corrplotlist' not found in the environment.")
}
```

```{r} 

# Load necessary libraries
library(tidyverse)
library(tibble)
library(Hmisc) # Required for rcorr()

# 1. Define Variable Lists
# ---------------------------------------------------------
# Daily/EMA variables (Key = Raw Name, Value = Pretty Label)
corrplotlist <- c(
  "num_drinks_today"      = "Number of Drinks",
  "bac_peak_percent"      = "Peak BAC (%)",
  "crave_alc_pm"          = "Alcohol Craving (PM)",
  "bdefs_impulsiveaction" = "Impulsive Action (BDEFS)",
  "bdefs_delayaversion"   = "Delay Aversion (BDEFS)",
  "drsp_total"            = "DRSP Total",
  "drsp_distress"         = "Distress (DRSP)",
  "drsp_labilsocsx"       = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"       = "Breast Sx (DRSP)",
  "drsp_eatsx"            = "Eating Sx",
  "drsp_lowarousal"       = "Low Arousal Sx (DRSP)",
  "diff_sleep"            = "Insomnia/Sleep Diff.",
  "jm_pain"               = "Joint/Muscle Pain",
  "e2_w"                  = "Estradiol (E2)",
  "p4_w"                  = "Progesterone (P4)",
  "lh_w"                  = "Luteinizing Hormone (LH)"
)

# Trait variables (Raw Names)
trait_vars_raw <- c("age", "audit_score", "weight_kg", "height_cm", 
                    "tlfb30d_daysdrank", "tlfb30d_cumuldrinks", 
                    "tlfb30d_daysbinge", "tlfb30d_daysdrunk", 
                    "pdhq_daysdrinkperwk", "pdhq_numtypicaldrinks", 
                    "pdhq_hrstypdrinking")

# Create a mapping for trait vars too (Raw Name -> Raw Name) 
# so they can be processed by the same logic
trait_list <- setNames(trait_vars_raw, trait_vars_raw) 

# Combine into one MASTER list (Raw Name -> Pretty Label)
all_var_map <- c(corrplotlist, trait_list)


# 2. Aggregation (Create the Data)
# ---------------------------------------------------------
# A. Summarize Traits (First non-missing value)
vars_to_summarize <- setdiff(trait_vars_raw, "id")

traits <- df %>%
  group_by(id) %>%
  summarize(across(all_of(vars_to_summarize), ~ first(na.omit(.x))), .groups = "drop")

# B. Summarize Daily Vars (Mean)
daily_raw_names <- names(corrplotlist)

trait_means <- df %>%
  group_by(id) %>%
  summarize(across(all_of(daily_raw_names), ~ mean(.x, na.rm = TRUE)))

# C. Join
combined_trait_data <- trait_means %>%
  inner_join(traits, by = "id") %>%
  select(-id) # Remove ID before correlation


# 3. Apply Your Requested Formatting (Safe Labels & Lookup)
# ---------------------------------------------------------

# Identify which columns we actually have in the data
current_cols <- names(combined_trait_data)

# Subset the master map to only variables present in the data
active_map <- all_var_map[current_cols]

# === THE REQUESTED FORMATTING LOGIC ===
# 1. Create safe labels (e.g., "Number.of.Drinks") derived from the Pretty Labels
safe_labels <- make.names(active_map)

# 2. Create a lookup table to map safe labels back to pretty labels
label_lookup <- setNames(active_map, safe_labels)

# 3. Rename the dataframe columns to the SAFE labels
names(combined_trait_data) <- safe_labels
# ======================================


# 4. Calculate Correlation AND P-Values
# ---------------------------------------------------------
# rcorr requires a matrix
mat_data <- as.matrix(combined_trait_data)

# Use rcorr to get both correlation (r) and p-values (P)
cor_res <- rcorr(mat_data, type = "spearman")

# Extract matrices
r_mat <- cor_res$r
p_mat <- cor_res$P

# 5. Prepare Data for Plotting
# ---------------------------------------------------------
# Flatten correlations
r_long <- r_mat %>%
  as.data.frame() %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "r_value")

# Flatten p-values
p_long <- p_mat %>%
  as.data.frame() %>%
  rownames_to_column("Var1") %>%
  pivot_longer(-Var1, names_to = "Var2", values_to = "p_value")

# Join them
plot_data <- left_join(r_long, p_long, by = c("Var1", "Var2")) %>%
  mutate(
    # Apply Lookup for Pretty Labels
    Label1 = label_lookup[Var1],
    Label2 = label_lookup[Var2],
    
    # Set Factors for ordering based on the matrix structure
    Label1 = factor(Label1, levels = label_lookup[rownames(r_mat)]),
    Label2 = factor(Label2, levels = label_lookup[colnames(r_mat)]),
    
    # Define significance flag
    is_sig = !is.na(p_value) & p_value < 0.05
  )

# 6. Plot
# ---------------------------------------------------------
p_trait_corr <- ggplot(plot_data, aes(Label2, Label1, fill = r_value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
                       midpoint = 0, limit = c(-1,1), space = "Lab", 
                       name="Spearman\nCorrelation") +
  
  # --- CONDITIONAL BOLDING ---
  geom_text(aes(
      label = round(r_value, 2), 
      # Bold if significant (p < .05), Plain otherwise
      fontface = ifelse(is_sig, "bold", "plain"),
      # Optional: Dim non-significant numbers slightly?
      alpha = ifelse(is_sig, 1, 0.7) 
    ), 
    size = 3,
    color = "black" 
  ) +
  # ---------------------------

  theme_minimal(base_size = 14) + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
        axis.title = element_blank()) +
  coord_fixed() +
  labs(title = "Between-Person Spearman Correlations",
       subtitle = "Bold values indicate p < .05")

print(p_trait_corr)

save_png(plot = p_trait_corr, 
         filename = "between_person_trait_correlation_heatmap_sig.png",
         folder = descriptives_folder,
         width = 14, 
         height = 12)

```

