---
title: "UKALC_GAMMs_cyclic"
author: "Tory Eisenlohr-Moul, PhD"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
    self_contained: true
---




# =============================================================================
# GENERALIZED ADDITIVE MIXED MODELS (GAMMs) FOR CYCLICAL ANALYSIS
# =============================================================================
# This script runs GAMMs to model cyclical patterns in alcohol use and other
# outcomes across the menstrual cycle. It processes both binary and continuous
# outcome variables using menses-centered and ovulation-centered time frames.
#
# WHAT IT DOES:
# 1. Sources setup file for packages and output folder configuration
# 2. Loads preprocessed hormone/daily survey data
# 3. Runs GAMMs for binary outcomes (e.g., drink_today, felt_drunk)
# 4. Runs GAMMs for continuous outcomes (e.g., num_drinks, cravings)
# 5. Creates visualizations with rolling averages and model predictions
# 6. Saves all outputs (models, summaries, plots) to dated output folder
#
# OUTPUT STRUCTURE:
# - All outputs saved to: output_folder/[outcome_name]_gamm/
# - Each outcome gets: models (.rds), summaries (.txt), and plots (.png)
#
# MODELS:
# - Binary outcomes: Logistic regression with cyclic splines
# - Continuous outcomes: Gaussian regression on log-transformed data
# - Both include: random intercepts, cyclic trends, random cyclic slopes
# - Covariates: 
#   * day_of_week: controls for weekday effects (e.g., weekend drinking)
#   * session_today: controls for laboratory visit effects
#
# NOTE: Core GAMM structure (cyclic splines, random effects) is preserved.
#       session_today was added to continuous-outcome models for consistency
#       with the binary-outcome models.
# =============================================================================

```{r setup-packages}
# Source R File with Setup, Package Loading, and Output Folder Creation
# This loads all required packages and creates the dated output folder
source("UKALC_00_setup.R")

# Create subfolder for GAMM outputs (keep root-level output_folder unchanged)
gamms_folder <- file.path(output_folder, "03_gamms")
if (!dir.exists(gamms_folder)) {
  dir.create(gamms_folder, recursive = TRUE)
  cat("Created GAMMs subfolder:", gamms_folder, "\n")
}

```

# =============================================================================
# DATA IMPORT
# =============================================================================
# Load the preprocessed data file containing hormone measurements and daily
# survey responses. This dataset should include:
# - Participant ID (id)
# - Cyclic time variables (cyclic_time, cyclic_time_ov)
# - Binary outcome variables (drink_today, fourplustoday, felt_drunk)
# - Continuous outcome variables (num_drinks, cravings, symptoms)
# - Hormone measurements (e2, p4, lh)
# - Covariates (day_of_week, session_today)
# =============================================================================

```{r import-data}
# -------------------------------------------------------------------------
# Configure Data File Path
# -------------------------------------------------------------------------
# LOAD LATEST CLEAN DATA FILE 

# Load the dataset
df <- readRDS("data/analysis_ready_withsessiondow.rds")

message(paste0("Data loaded successfully: ", nrow(df), " rows, ", ncol(df), " columns"))
```

# =============================================================================
# CUSTOM FUNCTIONS CHECK
# =============================================================================
# This script uses custom plotting functions for rolling averages and data
# availability visualization. If these functions are not available, those
# specific plots will be skipped while GAMMs still run.
#
# REQUIRED CUSTOM FUNCTIONS:
# - cycle_plot(): Creates rolling average plots centered on menses or ovulation
# - cycledata_check(): Visualizes data availability across the cycle
#
# If you have these functions in a separate file:
# source("path/to/custom_functions.R")
# =============================================================================

```{r check-custom-functions}
# Check if custom functions are available
has_cycle_plot <- exists("cycle_plot") && is.function(cycle_plot)
has_cycledata_check <- exists("cycledata_check") && is.function(cycledata_check)

if (!has_cycle_plot) {
  message("WARNING: cycle_plot() function not found. Rolling average plots will be skipped.")
}

if (!has_cycledata_check) {
  message("WARNING: cycledata_check() function not found. Data availability plots will be skipped.")
}
```
'


```{r}
names(df)
```



# =============================================================================
# BINARY OUTCOMES: GAMMs for Dichotomous Variables
# =============================================================================
# This section analyzes binary outcomes (0/1) using logistic GAMMs.
# Each outcome is modeled with both menses-centered and ovulation-centered
# time frames to capture different aspects of cyclical patterns.
#
# BINARY OUTCOMES ANALYZED:
# - drink_today: Whether participant drank alcohol (yes/no)
# - fourplustoday: Whether participant had 4+ drinks (binge drinking)
# - felt_drunk: Whether participant reported feeling drunk
#
# MODEL SPECIFICATION:
# - Family: Binomial with logit link
# - Fixed effects: day_of_week, session_today
# - Cyclic smooth: s(cyclic_time, bs='cc') - captures average cycle pattern
# - Random intercept: s(id, bs='re') - accounts for baseline differences
# - Random cyclic slope: s(cyclic_time, id, bs=c('re','cc')) - individual variation
#
# OUTPUTS FOR EACH VARIABLE:
# - Rolling average plots (menses and ovulation centered)
# - Data availability check plot
# - Model objects (.rds files)
# - Model summaries (.txt files)
# - Prediction plots with confidence bands
# =============================================================================

```{r binary-outcomes-gamms}
# -------------------------------------------------------------------------
# 1. SETUP: Define Binary Outcome Variables
# -------------------------------------------------------------------------

# List of binary outcome variable names to analyze
binary_vars <- c("drink_today") #, "felt_drunk") 

# Helper function to convert p-values to significance stars
# *** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1
get_stars <- function(p) {
  stats::symnum(p, corr = FALSE, na = FALSE, 
                cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
                symbols = c("***", "**", "*", ".", " "))
}

# -------------------------------------------------------------------------
# 2. LOOP THROUGH EACH BINARY OUTCOME
# -------------------------------------------------------------------------

for (outcome_name in binary_vars) {
  
  message(paste0("\n===== Processing Binary Outcome: ", outcome_name, " ====="))
  
  # Create outcome-specific subdirectory in the gamms_folder
  # This keeps all outputs for each variable organized together
  save_dir <- file.path(gamms_folder, paste0(outcome_name, "_gamm"))
  
  if (!dir.exists(save_dir)) {
    dir.create(save_dir, recursive = TRUE)
    message(paste0("Created directory: ", save_dir))
  }
  
  # -----------------------------------------------------------------------
  # 3. RAW DATA VISUALIZATION: Rolling Averages
  # -----------------------------------------------------------------------
  # These plots show the observed data patterns before model fitting.
  # Rolling averages smooth out day-to-day noise to reveal trends.
  
  # --- Menses-Centered Rolling Average Plot ---
  # Time axis: ovulation (-1) to menses (0) to ovulation (+1)
  if (has_cycle_plot) {
    message("  Creating menses-centered rolling average plot...")
    tryCatch({
      rollplot_menses <- cycle_plot(
        df,
        outcome_name,
        centering = "menses",
        include_impute = TRUE,
        y_scale = "person-centered_roll",
        se = TRUE
      )
      
      save_png(
        plot = rollplot_menses$plot,
        filename = "rollplot_menses.png",
        folder = save_dir
      )
    }, error = function(e) {
      message(paste0("  Error creating menses-centered plot: ", e$message))
    })
  } else {
    message("  Skipping menses-centered rolling average plot (cycle_plot not available)")
  }
  
  # --- Ovulation-Centered Rolling Average Plot ---
  # Time axis: menses (-1) to ovulation (0) to menses (+1)
  if (has_cycle_plot) {
    message("  Creating ovulation-centered rolling average plot...")
    tryCatch({
      rollplot_ov <- cycle_plot(
        df,
        outcome_name,
        centering = "ovulation",
        include_impute = TRUE,
        y_scale = "person-centered_roll", 
        se = TRUE
      )
      
      save_png(
        plot = rollplot_ov$plot,
        filename = "rollplot_ovulation.png",
        folder = save_dir
      )
    }, error = function(e) {
      message(paste0("  Error creating ovulation-centered plot: ", e$message))
    })
  } else {
    message("  Skipping ovulation-centered rolling average plot (cycle_plot not available)")
  }
  
  # --- Data Availability Check ---
  # Visualizes which participants have data at which cycle times
  if (has_cycledata_check) {
    message("  Creating data availability plot...")
    tryCatch({
      checkdata <- cycledata_check(df, symptom_columns = c(outcome_name))
      
      save_png(
        plot = checkdata$data_symptom_plots,
        filename = paste0("data_availability_", outcome_name, ".png"),
        folder = save_dir
      )
    }, error = function(e) {
      message(paste0("  Error creating data availability plot: ", e$message))
    })
  } else {
    message("  Skipping data availability plot (cycledata_check not available)")
  }
  
  # -----------------------------------------------------------------------
  # 4. GAMM: MENSES-CENTERED MODEL
  # -----------------------------------------------------------------------
  # Fits a logistic GAMM with menses onset as the reference point (time = 0).
  # This model estimates the probability of the outcome across the cycle.
  
  message("  Fitting menses-centered GAMM...")
  
  # Prepare data: keep only complete cases for required variables
  selected_vars <- c("cyclic_time", "day_of_week", "session_today", outcome_name, "id")
  dat_menses <- df[complete.cases(df[selected_vars]), ]
  dat_menses$id <- as.factor(dat_menses$id)
  
  # Construct model formula
  # - day_of_week: controls for weekday effects
  # - session_today: controls for laboratory visit effects
  # - s(cyclic_time, bs='cc'): cyclic smooth for average effect
  # - s(id, bs='re'): random intercept for each person
  # - s(cyclic_time, id, bs=c('re','cc')): random cyclic slope
  f_menses <- as.formula(paste(outcome_name, "~ day_of_week + session_today + s(cyclic_time, bs = 'cc') + s(id, bs = 're') + s(cyclic_time, id, bs = c('re', 'cc'))"))
  
  # Fit the model
  gamm_menses <- mgcv::gam(
    formula = f_menses,
    knots = list(cyclic_time = c(-1, 1)),
    data = dat_menses, 
    method = 'REML',
    family = binomial(link = "logit")
  )
  
  # Save model object and text summary
  saveRDS(gamm_menses, file = file.path(save_dir, "model_menses.rds"))
  
  sink(file.path(save_dir, "summary_menses.txt"))
  summ_menses <- summary(gamm_menses)
  print(summ_menses)
  sink()
  
  # --- Extract Statistics for Plot Annotation ---
  edf_fixed <- round(summ_menses$s.table[1, "edf"], 2)
  stars_fixed <- get_stars(summ_menses$s.table[1, "p-value"])
  
  edf_rand <- round(summ_menses$s.table[3, "edf"], 2)
  stars_rand <- get_stars(summ_menses$s.table[3, "p-value"])
  
  stat_label_menses <- paste0("Average Effect: edf=", edf_fixed, stars_fixed, "\n",
                              "Individual Differences: edf=", edf_rand, stars_rand)
  
  # --- Generate Predictions for Plotting ---
  message("  Generating predictions for menses-centered model...")
  
  # Set reference values for covariates (held constant during prediction)
  ref_day_menses <- dat_menses$day_of_week[1]
  ref_session_menses <- dat_menses$session_today[1]

  # Create prediction grid
  grid_list <- list(
    cyclic_time = seq(-1, 1, by = 0.05),
    id = 0,
    day_of_week = ref_day_menses,
    session_today = ref_session_menses
  )
  plotdat_menses <- do.call(expand.grid, grid_list)
  
  # Get predictions with confidence intervals
  pred_menses <- marginaleffects::predictions(
    gamm_menses,
    newdata = plotdat_menses,
    type = "response"
  )
  
  plotdat_menses$estimate <- pred_menses$estimate
  plotdat_menses$conf.low <- pred_menses$conf.low
  plotdat_menses$conf.high <- pred_menses$conf.high
  
  # --- Create Prediction Plot ---
  message("  Creating menses-centered prediction plot...")
  gammplot_menses <- ggplot(plotdat_menses, aes(x = cyclic_time, y = estimate)) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.50),
      labels = c("Ovulation", "50%L", "Menses Onset", "50%F", "Ovulation")
    ) +
    labs(x = "", y = paste("Probability of", outcome_name, "(Menses-Centered)"),
         caption = paste("Predictions holding day_of_week at:", ref_day_menses, 
                        "and session_today at:", ref_session_menses)) +
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) +
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) +
    annotate("text", x = -1, y = Inf, label = stat_label_menses, 
             hjust = 0, vjust = 1.1, size = 5, fontface = "italic") +
    theme_minimal(base_size = 14) +
    theme(text = element_text(size = 16),
          axis.text = element_text(size = 14))
  
  save_png(
    plot = gammplot_menses,
    filename = "gamm_plot_menses.png",
    folder = save_dir
  )
  
  # -----------------------------------------------------------------------
  # 5. GAMM: OVULATION-CENTERED MODEL
  # -----------------------------------------------------------------------
  # Fits a logistic GAMM with ovulation as the reference point (time = 0).
  
  message("  Fitting ovulation-centered GAMM...")
  
  # Prepare data: keep only complete cases for required variables
  selected_vars_ov <- c("cyclic_time_ov", "day_of_week", "session_today", outcome_name, "id")
  dat_ov <- df[complete.cases(df[selected_vars_ov]), ]
  dat_ov$id <- as.factor(dat_ov$id)
  
  # Construct model formula
  f_ov <- as.formula(paste(outcome_name, "~ day_of_week + session_today + s(cyclic_time_ov, bs = 'cc') + s(id, bs = 're') + s(cyclic_time_ov, id, bs = c('re', 'cc'))"))
  
  # Fit the model
  gamm_ov <- mgcv::gam(
    formula = f_ov,
    knots = list(cyclic_time_ov = c(-1, 1)),
    data = dat_ov, 
    method = 'REML',
    family = binomial(link = "logit")
  )
  
  # Save model object and text summary
  saveRDS(gamm_ov, file = file.path(save_dir, "model_ovulation.rds"))
  
  sink(file.path(save_dir, "summary_ovulation.txt"))
  summ_ov <- summary(gamm_ov)
  print(summ_ov)
  sink()
  
  # --- Extract Statistics for Plot Annotation ---
  edf_fixed_ov <- round(summ_ov$s.table[1, "edf"], 2)
  stars_fixed_ov <- get_stars(summ_ov$s.table[1, "p-value"])
  
  edf_rand_ov <- round(summ_ov$s.table[3, "edf"], 2)
  stars_rand_ov <- get_stars(summ_ov$s.table[3, "p-value"])
  
  stat_label_ov <- paste0("Average Effect: edf=", edf_fixed_ov, stars_fixed_ov, "\n",
                          "Individual Differences: edf=", edf_rand_ov, stars_rand_ov)
  
  # --- Generate Predictions for Plotting ---
  message("  Generating predictions for ovulation-centered model...")
  
  # Set reference values for covariates
  ref_day_ov <- dat_ov$day_of_week[1]
  ref_session_ov <- dat_ov$session_today[1]
  
  # Create prediction grid
  grid_list_ov <- list(
    cyclic_time_ov = seq(-1, 1, by = 0.05), 
    id = 0, 
    day_of_week = ref_day_ov,
    session_today = ref_session_ov
  )
  plotdat_ov <- do.call(expand.grid, grid_list_ov)
  
  # Get predictions with confidence intervals
  pred_ov <- marginaleffects::predictions(
    gamm_ov,
    newdata = plotdat_ov,
    type = "response"
  )
  
  plotdat_ov$estimate <- pred_ov$estimate
  plotdat_ov$conf.low <- pred_ov$conf.low
  plotdat_ov$conf.high <- pred_ov$conf.high
  
  # --- Create Prediction Plot ---
  message("  Creating ovulation-centered prediction plot...")
  gammplot_ov <- ggplot(plotdat_ov, aes(x = cyclic_time_ov, y = estimate)) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.50),
      labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses")
    ) +
    labs(x = "", y = paste("Probability of", outcome_name, "(Ovulation-Centered)"),
         caption = paste("Predictions holding day_of_week at:", ref_day_ov,
                        "and session_today at:", ref_session_ov)) +
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) +
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) +
    annotate("text", x = -1, y = Inf, label = stat_label_ov, 
             hjust = 0, vjust = 1.1, size = 5, fontface = "italic") +
    theme_minimal(base_size = 14) +
    theme(text = element_text(size = 16),
          axis.text = element_text(size = 14))
  
  save_png(
    plot = gammplot_ov,
    filename = "gamm_plot_ovulation.png",
    folder = save_dir
  )
  
  message(paste0("===== Completed: ", outcome_name, " =====\n"))
}
```



# =============================================================================
# CONTINUOUS OUTCOMES: GAMMs for Quantitative Variables
# =============================================================================
# This section analyzes continuous outcomes using Gaussian GAMMs on
# log-transformed data. Log transformation handles skewness and zeros.
#
# CONTINUOUS OUTCOMES ANALYZED:
# - Alcohol: num_drinks_today, crave_alc_am/pm, crave_nic_am/pm, crave_mar_am/pm
# - Impulsivity: diff_stop, diff_change, immed_payoff, res_urge, diff_long_rew, diff_res_rew
# - Mood: dep, hopeless, worthless, anxious, mood_swing, rej_sens, angry, conflicts
# - Cognition: loss_interest, diff_conc, tired
# - Appetite: over_eat, sp_food_crav
# - Sleep: high_sleep, diff_sleep
# - Other symptoms: overwhelm, loss_control, breast_tend, breast_swell, headache, jm_pain
# - Functioning: loss_prod, int_relat
#
# MODEL SPECIFICATION:
# - Family: Gaussian (suitable for log-transformed continuous data)
# - Data transformation: log(x + 1) to handle zeros
# - Fixed effects: day_of_week, session_today (for consistency with binary models)
# - Cyclic smooth: s(cyclic_time, bs='cc') - captures average cycle pattern
# - Random intercept: s(id, bs='re') - accounts for baseline differences
# - Random cyclic slope: s(cyclic_time, id, bs=c('re','cc')) - individual variation
#
# OUTPUTS FOR EACH VARIABLE:
# - Rolling average plots on original scale (menses and ovulation centered)
# - Model objects fitted on log scale (.rds files)
# - Model summaries (.txt files)
# - Prediction plots back-transformed to original scale
#
# NOTE: Models are fitted on log scale but predictions are back-transformed
# for interpretability using exp(y) - 1
# =============================================================================

  
```{r continuous-outcomes-gamms}
# -------------------------------------------------------------------------
# 1. SETUP: Define Continuous Outcome Variables
# -------------------------------------------------------------------------

# List of continuous outcome variable names to analyze
continuous_vars <- c(
  # Alcohol and substance use
  "num_drinks_today",
#, "bac_peak_percent", "crave_alc_am", "crave_alc_pm", "crave_nic_am", "crave_mar_am",
  # Impulsivity and reward processing
#"bdefs_delayaversion", "bdefs_impulsiveaction", "bdefs_total",
  #"diff_stop", "diff_change", "immed_payoff", "res_urge", 
  #"diff_long_rew", "diff_res_rew", 
  # Mood and emotional symptoms
  "dep", "hopeless", "worthless", "anxious", "mood_swing", "rej_sens", 
  "angry", "conflicts",
  # DRSP aggregate scores
  "drsp_total", "drsp_distress", "drsp_labilsocsx", "drsp_breastpain", 
  "drsp_eatsx", "drsp_lowarousal", 
  # Cognitive and physical symptoms
  "loss_interest", "diff_conc", "tired", 
  # Appetite and eating
  "over_eat", "sp_food_crav", 
  # Sleep
  "high_sleep", "diff_sleep", 
  # Other physical and psychological symptoms
  "overwhelm", "loss_control", "breast_tend", "breast_swell", 
  "headache", "jm_pain", 
  # Functioning
  "loss_prod", "int_relat") #, 
  # Additional substance cravings (evening assessments)
  #"crave_alc_pm", "crave_nic_pm", "crave_mar_pm")

# Helper function to convert p-values to significance stars (same as binary section)
get_stars <- function(p) {
  stats::symnum(p, corr = FALSE, na = FALSE, 
                cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1), 
                symbols = c("***", "**", "*", ".", " "))
}

# -------------------------------------------------------------------------
# 2. LOOP THROUGH EACH CONTINUOUS OUTCOME
# -------------------------------------------------------------------------

for (outcome_name in continuous_vars) {
  
  message(paste0("\n===== Processing Continuous Outcome: ", outcome_name, " ====="))
  
  # --- Check if Variable Exists and Has Data ---
  # Skip this outcome if the variable doesn't exist or has no valid data
  if (!outcome_name %in% names(df)) {
    message(paste0("  WARNING: Variable '", outcome_name, "' not found in dataframe. Skipping."))
    next
  }
  
  outcome_data <- df[[outcome_name]]
  if (length(outcome_data) == 0 || all(is.na(outcome_data))) {
    message(paste0("  WARNING: Variable '", outcome_name, "' has no valid data. Skipping."))
    next
  }
  
  # --- Create Log-Transformed Variable ---
  # Log transformation: log(x + 1) handles zeros and reduces skewness
  # The +1 ensures log(0) = 0 rather than -Inf
  log_outcome_name <- paste0(outcome_name, "_log")
  df[[log_outcome_name]] <- log(df[[outcome_name]] + 1)
  
  # --- Create Output Directory ---
  save_dir <- file.path(gamms_folder, paste0(outcome_name, "_gamm"))
  
  if (!dir.exists(save_dir)) {
    dir.create(save_dir, recursive = TRUE)
    message(paste0("Created directory: ", save_dir))
  }
  
  # -----------------------------------------------------------------------
  # 3. RAW DATA VISUALIZATION: Rolling Averages (Original Scale)
  # -----------------------------------------------------------------------
  # These plots show the observed data patterns before model fitting on the
  # original (non-logged) scale for interpretability.
  
  # --- Menses-Centered Rolling Average Plot ---
  if (has_cycle_plot) {
    message("  Creating menses-centered rolling average plot...")
    tryCatch({
      rollplot_menses <- cycle_plot(
        df,
        outcome_name,
        centering = "menses",
        include_impute = TRUE,
        y_scale = "person-centered_roll", 
        se = TRUE
      )
      
      save_png(
        plot = rollplot_menses$plot,
        filename = "rollplot_menses_raw.png",
        folder = save_dir
      )
    }, error = function(e) {
      message(paste0("  Error creating menses-centered plot: ", e$message))
    })
  } else {
    message("  Skipping menses-centered rolling average plot (cycle_plot not available)")
  }
  
  # --- Ovulation-Centered Rolling Average Plot ---
  if (has_cycle_plot) {
    message("  Creating ovulation-centered rolling average plot...")
    tryCatch({
      rollplot_ov <- cycle_plot(
        df,
        outcome_name,
        centering = "ovulation",
        include_impute = TRUE,
        y_scale = "person-centered_roll", 
        se = TRUE
      )
      
      save_png(
        plot = rollplot_ov$plot,
        filename = "rollplot_ovulation_raw.png",
        folder = save_dir
      )
    }, error = function(e) {
      message(paste0("  Error creating ovulation-centered plot: ", e$message))
    })
  } else {
    message("  Skipping ovulation-centered rolling average plot (cycle_plot not available)")
  }
  
  # -----------------------------------------------------------------------
  # 4. GAMM: MENSES-CENTERED MODEL (Log Scale)
  # -----------------------------------------------------------------------
  # Fits a Gaussian GAMM on log-transformed data with menses onset as
  # the reference point. Predictions will be back-transformed for plotting.
  
  message("  Fitting menses-centered GAMM (on log scale)...")
  
  # Prepare data: filter based on LOG variable to ensure proper sample matching
  selected_vars <- c("cyclic_time", "day_of_week", "session_today", log_outcome_name, "id")
  dat_menses <- df[complete.cases(df[selected_vars]), ]
  dat_menses$id <- as.factor(dat_menses$id) 
  
  # Construct model formula using logged outcome
  # - day_of_week: controls for weekday effects
  # - session_today: controls for laboratory visit effects
  # - s(cyclic_time, bs='cc'): cyclic smooth for average effect
  # - s(id, bs='re'): random intercept for each person
  # - s(cyclic_time, id, bs=c('re','cc')): random cyclic slope
  f_menses <- as.formula(paste(log_outcome_name, "~ day_of_week + session_today + s(cyclic_time, bs = 'cc') + s(id, bs = 're') + s(cyclic_time, id, bs = c('re', 'cc'))"))
  
  # Fit the model
  gamm_menses <- mgcv::gam(
    formula = f_menses,
    knots = list(cyclic_time = c(-1, 1)),
    data = dat_menses, 
    method = 'REML',
    family = gaussian()
  )
  
  # Save model object and text summary
  saveRDS(gamm_menses, file = file.path(save_dir, "model_menses_log.rds"))
  
  sink(file.path(save_dir, "summary_menses_log.txt"))
  summ_menses <- summary(gamm_menses)
  print(summ_menses)
  sink()
  
  # --- Extract Statistics for Plot Annotation ---
  edf_fixed <- round(summ_menses$s.table[1, "edf"], 2)
  stars_fixed <- get_stars(summ_menses$s.table[1, "p-value"])
  
  edf_rand <- round(summ_menses$s.table[3, "edf"], 2)
  stars_rand <- get_stars(summ_menses$s.table[3, "p-value"])
  
  stat_label_menses <- paste0("Average Effect: edf=", edf_fixed, stars_fixed, "\n",
                              "Individual Differences: edf=", edf_rand, stars_rand)
  
  # --- Generate Predictions and Back-Transform to Original Scale ---
  message("  Generating predictions for menses-centered model...")
  
  # Set reference values for covariates
  ref_day_menses <- dat_menses$day_of_week[1]
  ref_session_menses <- dat_menses$session_today[1]
  
  # Create prediction grid
  grid_list <- list(
    cyclic_time = seq(-1, 1, by = 0.05), 
    id = 0,
    day_of_week = ref_day_menses,
    session_today = ref_session_menses
  )
  plotdat_menses <- do.call(expand.grid, grid_list)
  
  # Get predictions on log scale
  pred_menses <- marginaleffects::predictions(
    gamm_menses,
    newdata = plotdat_menses,
    type = "response"
  )
  
  # Back-transform from log scale: exp(y) - 1
  plotdat_menses$estimate <- exp(pred_menses$estimate) - 1
  plotdat_menses$conf.low <- exp(pred_menses$conf.low) - 1
  plotdat_menses$conf.high <- exp(pred_menses$conf.high) - 1
  
  # --- Create Prediction Plot (Original Scale) ---
  message("  Creating menses-centered prediction plot...")
  gammplot_menses <- ggplot(plotdat_menses, aes(x = cyclic_time, y = estimate)) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.50),
      labels = c("Ovulation", "50%L", "Menses Onset", "50%F", "Ovulation")
    ) +
    labs(x = "", y = paste(outcome_name, "(Original Scale)"),
         caption = paste("Predictions holding day_of_week at:", ref_day_menses, 
                        "and session_today at:", ref_session_menses)) +
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) +
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) +
    annotate("text", x = -1, y = Inf, label = stat_label_menses, 
             hjust = 0, vjust = 1.1, size = 5, fontface = "italic") +
    theme_minimal(base_size = 14) +
    theme(text = element_text(size = 16),
          axis.text = element_text(size = 14))
  
  save_png(
    plot = gammplot_menses,
    filename = "gamm_plot_menses_backtrans.png",
    folder = save_dir
  )
  
  # -----------------------------------------------------------------------
  # 5. GAMM: OVULATION-CENTERED MODEL (Log Scale)
  # -----------------------------------------------------------------------
  # Fits a Gaussian GAMM on log-transformed data with ovulation as
  # the reference point. Predictions will be back-transformed for plotting.
  
  message("  Fitting ovulation-centered GAMM (on log scale)...")
  
  # Prepare data: keep only complete cases for required variables
  selected_vars_ov <- c("cyclic_time_ov", "day_of_week", "session_today", log_outcome_name, "id")
  dat_ov <- df[complete.cases(df[selected_vars_ov]), ]
  dat_ov$id <- as.factor(dat_ov$id)
  
  # Construct model formula
  f_ov <- as.formula(paste(log_outcome_name, "~ day_of_week + session_today + s(cyclic_time_ov, bs = 'cc') + s(id, bs = 're') + s(cyclic_time_ov, id, bs = c('re', 'cc'))"))
  
  # Fit the model
  gamm_ov <- mgcv::gam(
    formula = f_ov,
    knots = list(cyclic_time_ov = c(-1, 1)),
    data = dat_ov, 
    method = 'REML',
    family = gaussian()
  )
  
  # Save model object and text summary
  saveRDS(gamm_ov, file = file.path(save_dir, "model_ovulation_log.rds"))
  
  sink(file.path(save_dir, "summary_ovulation_log.txt"))
  summ_ov <- summary(gamm_ov)
  print(summ_ov)
  sink()
  
  # --- Extract Statistics for Plot Annotation ---
  edf_fixed_ov <- round(summ_ov$s.table[1, "edf"], 2)
  stars_fixed_ov <- get_stars(summ_ov$s.table[1, "p-value"])
  
  edf_rand_ov <- round(summ_ov$s.table[3, "edf"], 2)
  stars_rand_ov <- get_stars(summ_ov$s.table[3, "p-value"])
  
  stat_label_ov <- paste0("Average Effect: edf=", edf_fixed_ov, stars_fixed_ov, "\n",
                          "Individual Differences: edf=", edf_rand_ov, stars_rand_ov)
  
  # --- Generate Predictions and Back-Transform to Original Scale ---
  message("  Generating predictions for ovulation-centered model...")
  
  # Set reference values for covariates
  ref_day_ov <- dat_ov$day_of_week[1]
  ref_session_ov <- dat_ov$session_today[1]
  
  # Create prediction grid
  grid_list_ov <- list(
    cyclic_time_ov = seq(-1, 1, by = 0.05), 
    id = 0, 
    day_of_week = ref_day_ov,
    session_today = ref_session_ov
  )
  plotdat_ov <- do.call(expand.grid, grid_list_ov)
  
  # Get predictions on log scale
  pred_ov <- marginaleffects::predictions(
    gamm_ov,
    newdata = plotdat_ov,
    type = "response"
  )
  
  # Back-transform from log scale: exp(y) - 1
  plotdat_ov$estimate <- exp(pred_ov$estimate) - 1
  plotdat_ov$conf.low <- exp(pred_ov$conf.low) - 1
  plotdat_ov$conf.high <- exp(pred_ov$conf.high) - 1
  
  # --- Create Prediction Plot (Original Scale) ---
  message("  Creating ovulation-centered prediction plot...")
  gammplot_ov <- ggplot(plotdat_ov, aes(x = cyclic_time_ov, y = estimate)) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.50),
      labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses")
    ) +
    labs(x = "", y = paste(outcome_name, "(Original Scale)"),
         caption = paste("Predictions holding day_of_week at:", ref_day_ov, 
                        "and session_today at:", ref_session_ov)) +
    geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) +
    geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) +
    geom_line(linewidth = 1) +
    geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) +
    annotate("text", x = -1, y = Inf, label = stat_label_ov, 
             hjust = 0, vjust = 1.1, size = 5, fontface = "italic") +
    theme_minimal(base_size = 14) +
    theme(text = element_text(size = 16),
          axis.text = element_text(size = 14))
  
  save_png(
    plot = gammplot_ov,
    filename = "gamm_plot_ovulation_backtrans.png",
    folder = save_dir
  )
  
  message(paste0("===== Completed: ", outcome_name, " =====\n"))
}

```


# =============================================================================
# EXPLORATORY ANALYSIS: Day of Week Effects
# =============================================================================
# These plots explore whether drinking patterns vary by day of the week
# across the menstrual cycle. This helps identify potential confounds
# (e.g., weekend drinking) that need to be controlled in the GAMMs.
#
# WHAT IT DOES:
# 1. Creates person-centered drinking variables
# 2. Aggregates by cyclic time and day of week
# 3. Plots smoothed trends separately for each day of week
#
# OUTPUTS:
# - drink_today_by_day_of_week.png: Binary drinking indicator
# - num_drinks_by_day_of_week.png: Number of drinks consumed
# =============================================================================

```{r day-of-week-plots}
library(tidyverse)

# -------------------------------------------------------------------------
# Plot 1: Binary Drinking Indicator by Day of Week
# -------------------------------------------------------------------------

message("Creating day-of-week plot for binary drinking...")

# Prepare data: person-center the binary drink_today variable
plot_data <- df %>%
  # Calculate each person's average drinking frequency
  group_by(id) %>% 
  mutate(
    person_mean = mean(drink_today, na.rm = TRUE),
    drink_centered = drink_today - person_mean
  ) %>%
  ungroup() %>%
  # Aggregate to get mean levels for each time point and day of week
  group_by(cyclic_time, day_of_week) %>%
  summarize(
    mean_drink = mean(drink_centered, na.rm = TRUE),
    .groups = "drop"
  )

# Create smoothed plot showing trends by day of week
p_drink_today_dow <- ggplot(plot_data, aes(x = cyclic_time, y = mean_drink, color = day_of_week, fill = day_of_week)) +
  geom_point(alpha = 0.5) +  # Show actual mean levels
  # Add loess smoothing to reveal trends
  geom_smooth(method = "loess", se = FALSE, alpha = 0.2) + 
  theme_minimal(base_size = 14) +
  theme(text = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(
    x = "Cycle Time",
    y = "Mean Person-Centered Drink Levels",
    color = "Day of Week",
    fill = "Day of Week",
    title = "Smoothed Trends of Person-Centered Drinking Across Cycle"
  )

print(p_drink_today_dow)

save_png(
  plot = p_drink_today_dow,
  filename = "drink_today_by_day_of_week.png"
)

# -------------------------------------------------------------------------
# Plot 2: Number of Drinks by Day of Week
# -------------------------------------------------------------------------

message("Creating day-of-week plot for number of drinks...")

# Prepare data: person-center the num_drinks variable
plot_data <- df %>%
  group_by(id) %>% 
  mutate(
    person_mean = mean(num_drinks, na.rm = TRUE),
    drink_centered = num_drinks - person_mean
  ) %>%
  ungroup() %>%
  group_by(cyclic_time, day_of_week) %>%
  summarize(
    mean_drink = mean(drink_centered, na.rm = TRUE),
    .groups = "drop"
  )

# Create smoothed plot
p_num_drinks_dow <- ggplot(plot_data, aes(x = cyclic_time, y = mean_drink, color = day_of_week, fill = day_of_week)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "loess", se = FALSE, alpha = 0.2) + 
  theme_minimal(base_size = 14) +
  theme(text = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(
    x = "Cycle Time",
    y = "Mean Person-Centered Number of Drinks",
    color = "Day of Week",
    fill = "Day of Week",
    title = "Smoothed Trends of Person-Centered Drinks Across Cycle"
  )

print(p_num_drinks_dow)

save_png(
  plot = p_num_drinks_dow,
  filename = "num_drinks_by_day_of_week.png"
)



# -------------------------------------------------------------------------
# Plot 2: Peak BAC Across the Cycle by Day of Week
# -------------------------------------------------------------------------

message("Creating day-of-week plot for peak BAC...")

# Prepare data: person-center the BAC variable
plot_data <- df %>%
  group_by(id) %>% 
  mutate(
    person_mean = mean(bac_peak_percent, na.rm = TRUE),
    bac_centered = bac_peak_percent - person_mean
  ) %>%
  ungroup() %>%
  group_by(cyclic_time, day_of_week) %>%
  summarize(
    mean_bac = mean(bac_centered, na.rm = TRUE),
    .groups = "drop"
  )

# Create smoothed plot
p_bac_dow <- ggplot(plot_data, aes(x = cyclic_time, y = mean_bac, color = day_of_week, fill = day_of_week)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "loess", se = FALSE, alpha = 0.2) + 
  theme_minimal(base_size = 14) +
  theme(text = element_text(size = 16),
        axis.text = element_text(size = 14)) +
  labs(
    x = "Cycle Time",
    y = "Mean Person-Centered Peak BAC",
    color = "Day of Week",
    fill = "Day of Week",
    title = "Smoothed Trends of Person-Centered Peak BAC Across Cycle"
  )

print(p_bac_dow)

save_png(
  plot = p_bac_dow,
  filename = "peak_bac_by_day_of_week.png"
)



```


# =============================================================================
# HORMONE ANALYSIS: Combined Hormone Trajectories
# =============================================================================
# This section models hormone levels (E2, P4, LH) across the menstrual cycle
# and creates a combined visualization. Hormones are analyzed on an
# ovulation-centered time frame as this is biologically most relevant.
#
# HORMONES ANALYZED:
# - E2 (Estradiol): Primary estrogen, peaks at ovulation
# - P4 (Progesterone): Rises after ovulation, peaks in luteal phase
# - LH (Luteinizing Hormone): Triggers ovulation, sharp peak at mid-cycle
#
# MODEL SPECIFICATION:
# - Family: Gaussian on log-transformed data
# - Data transformation: log(x + 1) to handle zeros
# - Fixed effects: None (population-level cyclic trend only)
# - Cyclic smooth: s(cyclic_time_ov, bs='cc', k=10) - captures average pattern
# - Random intercept: s(id, bs='re') - accounts for baseline differences
# - Random cyclic slope: s(id, cyclic_time_ov, bs=c('re','cc'), k=c(NA,10))
#
# VISUALIZATION:
# - All three hormones plotted on same axes using standardized (z-scored) values
# - This allows comparison of relative patterns despite different measurement scales
# - Predictions are back-transformed to original scale before standardization
#
# OUTPUTS:
# - Combined plot showing all three hormones (hormones_combined/combined_hormone_plot.png)
# =============================================================================

```{r hormone-analysis}
# -------------------------------------------------------------------------
# Setup
# -------------------------------------------------------------------------

# Define hormone variable names
hormone_vars <- c("e2", "p4", "lh")

# Use ovulation-centered time (most relevant for hormone patterns)
time_var <- "cyclic_time_ov"

# Create output subdirectory for combined hormone plots
save_dir <- file.path(gamms_folder, "hormones_combined")
if (!dir.exists(save_dir)) {
  dir.create(save_dir, recursive = TRUE)
  message(paste0("Created directory: ", save_dir))
}

# List to store prediction data for all hormones
all_plotdat <- list() 

# -------------------------------------------------------------------------
# Loop Through Each Hormone: Fit Models and Generate Predictions
# -------------------------------------------------------------------------

for (h_name in hormone_vars) {
  
  message(paste0("\n===== Processing Hormone: ", h_name, " ====="))
  
  # --- Create Log-Transformed Variable ---
  log_name <- paste0(h_name, "_log")
  df[[log_name]] <- log(df[[h_name]] + 1)
  
  # --- Prepare Data ---
  selected_vars <- c(time_var, log_name, "id")
  datSX <- df[complete.cases(df[selected_vars]), ]
  datSX$id <- as.factor(datSX$id)
  
  # --- Fit GAMM ---
  # - Random intercept for each person
  # - Cyclic smooth term (k=10 basis functions)
  # - Random cyclic slope allowing individual variation
  message(paste0("  Fitting GAMM for ", h_name, "..."))
  
  f_hormone <- as.formula(paste(
    log_name, 
    "~ s(id, bs = 're') + s(", time_var, ", bs = 'cc', k = 10) + s(id, ", time_var, ", bs = c('re','cc'), k = c(NA, 10))"
  ))
  
  knots_list <- list(c(-1, 1))
  names(knots_list) <- time_var
  
  gamm_model <- mgcv::gam(
    formula = f_hormone,
    data = datSX,
    knots = knots_list,
    method = 'REML'
  )
  
  # --- Generate Predictions ---
  message(paste0("  Generating predictions for ", h_name, "..."))
  
  # Create prediction grid across cycle
  # Use setNames to dynamically name the cyclic time variable
  grid_list <- setNames(list(seq(-1, 1, by = 0.05), 0), c(time_var, "id"))
  plotdat <- do.call(expand.grid, grid_list)
  
  # Get predictions on log-transformed outcome scale
  pred <- marginaleffects::predictions(
    gamm_model,
    newdata = plotdat,
    type = "response"
  )
  
  plotdat$estimate <- pred$estimate
  plotdat$conf.low <- pred$conf.low
  plotdat$conf.high <- pred$conf.high
  
  # --- Back-Transform from Log Scale ---
  plotdat$estimate <- exp(plotdat$estimate) - 1
  plotdat$conf.low <- exp(plotdat$conf.low) - 1
  plotdat$conf.high <- exp(plotdat$conf.high) - 1
  
  # --- Standardize (Z-Score) for Combined Plotting ---
  # Puts E2, P4, and LH on the same relative y-axis
  # Note: Checks for zero SD to prevent Inf/NaN values
  mean_est <- mean(plotdat$estimate)
  sd_est <- sd(plotdat$estimate)
  
  plotdat <- plotdat %>%
    mutate(
      estimate_ss = if_else(is.na(sd_est) | sd_est == 0, NA_real_, (estimate - mean_est) / sd_est),
      conf.low_ss = if_else(is.na(sd_est) | sd_est == 0, NA_real_, (conf.low - mean_est) / sd_est),
      conf.high_ss = if_else(is.na(sd_est) | sd_est == 0, NA_real_, (conf.high - mean_est) / sd_est),
      hormone = h_name
    )
  
  # Store results
  all_plotdat[[h_name]] <- plotdat
  
  message(paste0("===== Completed: ", h_name, " ====="))
}

# -------------------------------------------------------------------------
# Combine Data and Create Visualization
# -------------------------------------------------------------------------

# Merge all hormone data into single dataframe
combined_df <- do.call(rbind, all_plotdat)

# Define color scheme for hormones
hormone_colors <- c("e2" = "navy", "p4" = "maroon", "lh" = "forestgreen")

message("\nGenerating combined hormone plot...")

# Create combined plot with all three hormones
combined_plot <- ggplot(combined_df, aes(x = .data[[time_var]], group = hormone)) +
  
  # Shaded background regions for menses and ovulation
  geom_rect(aes(xmin = -1, xmax = -0.92, ymin = -Inf, ymax = Inf), fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
  geom_rect(aes(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf), fill = "grey87", alpha = 0.5, inherit.aes = FALSE) +
  geom_rect(aes(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf), fill = "grey70", alpha = 0.5, inherit.aes = FALSE) +
  
  # Lines and confidence bands for each hormone
  geom_line(aes(y = estimate_ss, color = hormone), linewidth = 1.5) +
  geom_ribbon(aes(ymin = conf.low_ss, ymax = conf.high_ss, fill = hormone), alpha = 0.1) +
  
  # Scales and colors
  scale_x_continuous(
    limits = c(-1, 1),
    breaks = seq(-1, 1, by = 0.50),
    labels = c("Menses Onset", "50% Fol", "Ovulation", "50% Lut", "Menses Onset")
  ) +
  scale_color_manual(values = hormone_colors, name = "Hormone") +
  scale_fill_manual(values = hormone_colors, name = "Hormone") +
  
  # Labels and theme
  labs(
    x = "Cycle Phase", 
    y = "Sample-Standardized Hormone Levels (Z-Score)",
    title = "Combined Hormone Trajectories (Ovulation-Centered)"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    text = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.position = "top",
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# Save combined plot
save_png(
  plot = combined_plot,
  filename = "combined_hormone_plot.png",
  folder = save_dir
)

message(paste0("\nHormone analysis complete! Check ", save_dir, "/"))
```

```{r}

```




