---
title: "HERA Individual GAM Reports"
author: "Tory Eisenlohr-Moul, PhD"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

> **Script Purpose:** Automated, participant-centric analysis of symptom cyclicity using cyclic GAMs.
> 1) **Setup & Preparation** — load data, define outcomes/layouts, precompute transforms.
> 2) **Automated Reporting** — one **landscape** PDF per participant with compact cover and paged plots.
> 3) **CSV Audit** — per-ID × outcome × time-var stats for triage/QC.

---

# Libraries
```{r}
library(dplyr)
library(ggplot2)
library(mgcv)
library(patchwork)
library(gridExtra)
library(grid)
library(tibble)
library(tidyr)
library(knitr)
library(broom)
library(scales)
library(glue)

options(scipen = 999)
set.seed(12345)
```

# Data
```{r}

df <- readRDS('/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_20260414_clean.rds')
```

# TEST CHUNK - KEEP COMMENTED OUT UNLESS TESTING NEW CODE

```{r}
#set.seed(12345)
#test_ids <- sample(unique(df$id), 5)
#df <- df %>% filter(id %in% test_ids)
```

# Section 1: Set output folder & time vars
```{r}
# --- Path to the main output directory ---
gam_output_base <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/06_Single_GAMs/"

# --- Define all Time Variables to iterate over (must be scaled to [-1, 1]) ---
time_vars_to_run <- c("cyclic_time_impute")

  #"cyclic_time", "cyclic_time_ov", "cyclic_time_impute", "cyclic_time_imp_ov")

# --- Table Formatting ---
table_cols_per_page <- 6

# --- Plot Layout (grid per page) ---
plots_per_col <- 3
plots_per_row <- 3

# --- Minimum observations per cycle-quarter to run GAMs ---
# The cycle is divided into 4 equal quarters over [-1, 1].
# If any quarter has fewer than this many observations, the GAM is skipped
# for that specific outcome × time-var combination.
min_obs_per_quarter <- 2
```

# Section 2: Create output folders & check time vars
```{r}
# --- Create output folder with current date ---
current_date <- format(Sys.Date(), "%Y%m%d")
output_folder <- file.path(gam_output_base, paste0("GAM_Reports_", current_date))
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)

# --- Robustness: Verify time variables exist ---
time_vars <- intersect(time_vars_to_run, names(df))
if (length(time_vars) < length(time_vars_to_run)) {
  missing_vars <- setdiff(time_vars_to_run, time_vars)
  warning("The following time_vars were not found and will be skipped: ", paste(missing_vars, collapse = ", "))
}
```

# Section 3: Define outcomes
```{r}
# --- Define Outcomes to Model ---
outcomes <- c(
  "drsp_NA"          = "DRSP Negative Affect Mean",
  "drsp_distress"    = "Distress (DRSP)",
  "drsp_labilsocsx"  = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"  = "Breast Sx (DRSP)",
  "drsp_eatsx"       = "Eating Sx",
  "drsp_lowarousal"  = "Low Arousal Sx (DRSP)",
  "drsp_1"           = "Depression",
  "drsp_2"           = "Hopelessness",
  "drsp_3"           = "Worthlessness",
  "drsp_4"           = "Anxiety",
  "drsp_5"           = "Mood Swings",
  "drsp_6"           = "Rejection Sensitivity",
  "drsp_7"           = "Anger/Irritability",
  "drsp_8"           = "Interpersonal Conflicts",
  "drsp_9"           = "Loss of Interest",
  "drsp_16"          = "Feeling Overwhelmed",
  "drsp_17"          = "Feeling Out of Control",
  "drsp_10"          = "Difficulty Concentrating",
  "drsp_11"          = "Fatigue/Lethargy",
  "drsp_14"          = "Hypersomnia",
  "drsp_15"          = "Insomnia",
  "drsp_12"          = "Overeating",
  "drsp_13"          = "Specific Food Cravings",
  "drsp_18"          = "Breast Tenderness",
  "drsp_19"          = "Breast Swelling",
  "drsp_20"          = "Headache",
  "drsp_21"          = "Joint/Muscle Pain"
)

# --- Physical vs Affect symptom classification (raw DRSP items only) ---
physical_symptoms <- c("drsp_12", "drsp_13", "drsp_14", "drsp_15",
                       "drsp_18", "drsp_19", "drsp_20", "drsp_21")

# --- Pattern constants for variable matching ---
DRSP_VAR_PATTERN <- "^drsp_\\d+$"

# --- Max possible counts per category (for density-plot Y-axis scaling) ---
all_raw_drsp <- names(outcomes)[grepl(DRSP_VAR_PATTERN, names(outcomes))]
n_physical_max <- length(intersect(all_raw_drsp, physical_symptoms))
n_affect_max   <- length(setdiff(all_raw_drsp, physical_symptoms))

# --- Define colors for plots (each outcome stays consistent) ---
## Use a colorblind friendly palette
drsp_color_map <- c(
  # DRSP mood/affect individual items
  "drsp_1"  = "#0072B2",  # blue
  "drsp_2"  = "#E69F00",  # orange
  "drsp_3"  = "#009E73",  # green
  "drsp_4"  = "#D55E00",  # vermillion
  "drsp_5"  = "#CC79A7",  # reddish purple
  "drsp_6"  = "#56B4E9",  # sky blue
  "drsp_7"  = "#F0E442",  # yellow
  "drsp_8"  = "#000000",  # black
  "drsp_9"  = "#999999",  # gray
  "drsp_10" = "#004488",  # dark blue
  "drsp_11" = "#BB5566",  # muted red
  "drsp_12" = "#228833",  # dark green
  "drsp_13" = "#EE6677",  # soft red
  "drsp_14" = "#AA3377",  # magenta
  "drsp_15" = "#66CCEE",  # light cyan
  "drsp_16" = "#332288",  # deep purple
  "drsp_17" = "#117733",  # dark green 2
  # DRSP physical items
  "drsp_18" = "#44AA99",  # teal
  "drsp_19" = "#88CCEE",  # light blue
  "drsp_20" = "#DDCC77",  # sand
  "drsp_21" = "#CC6677",  # rose
  # DRSP composite outcomes (Okabe-Ito colorblind-friendly palette)
  "drsp_NA"         = "#0072B2",  # blue
  "drsp_distress"   = "#E69F00",  # orange
  "drsp_labilsocsx" = "#009E73",  # green
  "drsp_breastpain" = "#D55E00",  # vermillion
  "drsp_eatsx"      = "#CC79A7",  # reddish purple
  "drsp_lowarousal" = "#56B4E9"   # sky blue
)

# Safety check: warn (do not stop) if any outcome is missing a color
missing_colors <- setdiff(names(outcomes), names(drsp_color_map))
if (length(missing_colors) > 0) {
  warning("Missing colors for: ", paste(missing_colors, collapse = ", "),
          ". Falling back to #0072B2 for these outcomes.")
}

# --- Per-outcome y-axis limits and reference-line settings ---
# y_axis_mode "fixed" uses a hard y_min/y_max (Likert 1-6 scale).
# y_axis_mode "data"  uses data-driven limits; no fixed reference line.
outcome_plot_settings <- tribble(
  ~variable,          ~y_axis_mode, ~y_min, ~y_max, ~ref_line,
  # DRSP composite outcomes — default 1–4.5 scale; expands to 5 or 6 if predictions exceed 4.5
  "drsp_NA",          "fixed",       1,      4.5,   NA,
  "drsp_distress",    "fixed",       1,      4.5,   NA,
  "drsp_labilsocsx",  "fixed",       1,      4.5,   NA,
  "drsp_breastpain",  "fixed",       1,      4.5,   NA,
  "drsp_eatsx",       "fixed",       1,      4.5,   NA,
  "drsp_lowarousal",  "fixed",       1,      4.5,   NA,
  # DRSP individual items — default 1–4.5 scale; expands to 5 or 6 if predictions exceed 4.5
  "drsp_1",           "fixed",       1,      4.5,    4,
  "drsp_2",           "fixed",       1,      4.5,    4,
  "drsp_3",           "fixed",       1,      4.5,    4,
  "drsp_4",           "fixed",       1,      4.5,    4,
  "drsp_5",           "fixed",       1,      4.5,    4,
  "drsp_6",           "fixed",       1,      4.5,    4,
  "drsp_7",           "fixed",       1,      4.5,    4,
  "drsp_8",           "fixed",       1,      4.5,    4,
  "drsp_9",           "fixed",       1,      4.5,    4,
  "drsp_16",          "fixed",       1,      4.5,    4,
  "drsp_17",          "fixed",       1,      4.5,    4,
  "drsp_10",          "fixed",       1,      4.5,    4,
  "drsp_11",          "fixed",       1,      4.5,    4,
  "drsp_14",          "fixed",       1,      4.5,    4,
  "drsp_15",          "fixed",       1,      4.5,    4,
  "drsp_12",          "fixed",       1,      4.5,    4,
  "drsp_13",          "fixed",       1,      4.5,    4,
  "drsp_18",          "fixed",       1,      4.5,    4,
  "drsp_19",          "fixed",       1,      4.5,    4,
  "drsp_20",          "fixed",       1,      4.5,    4,
  "drsp_21",          "fixed",       1,      4.5,    4
)

# --- Normalization scale ranges for normalized overlay pages ---
# Fixed-scale outcomes use their configured y_min/y_max.
# Data-mode outcomes fall back to per-curve GAM min/max at plot time.
norm_scale_ranges <- outcome_plot_settings %>%
  dplyr::mutate(
    norm_min = dplyr::if_else(!is.na(y_min), y_min, NA_real_),
    norm_max = dplyr::if_else(!is.na(y_max), y_max, NA_real_)
  ) %>%
  dplyr::select(variable, norm_min, norm_max)

# --- Define Categories for organization ---
outcome_categories <- tribble(
  ~category,                ~variable,
  "DRSP Composites",        "drsp_NA",
  "DRSP Composites",        "drsp_distress",
  "DRSP Composites",        "drsp_labilsocsx",
  "DRSP Composites",        "drsp_breastpain",
  "DRSP Composites",        "drsp_eatsx",
  "DRSP Composites",        "drsp_lowarousal",
  "Mood & Physical (DRSP)", "drsp_1",
  "Mood & Physical (DRSP)", "drsp_2",
  "Mood & Physical (DRSP)", "drsp_3",
  "Mood & Physical (DRSP)", "drsp_4",
  "Mood & Physical (DRSP)", "drsp_5",
  "Mood & Physical (DRSP)", "drsp_6",
  "Mood & Physical (DRSP)", "drsp_7",
  "Mood & Physical (DRSP)", "drsp_8",
  "Mood & Physical (DRSP)", "drsp_9",
  "Mood & Physical (DRSP)", "drsp_16",
  "Mood & Physical (DRSP)", "drsp_17",
  "Mood & Physical (DRSP)", "drsp_10",
  "Mood & Physical (DRSP)", "drsp_11",
  "Mood & Physical (DRSP)", "drsp_14",
  "Mood & Physical (DRSP)", "drsp_15",
  "Mood & Physical (DRSP)", "drsp_12",
  "Mood & Physical (DRSP)", "drsp_13",
  "Mood & Physical (DRSP)", "drsp_18",
  "Mood & Physical (DRSP)", "drsp_19",
  "Mood & Physical (DRSP)", "drsp_20",
  "Mood & Physical (DRSP)", "drsp_21"
) %>%
  right_join(enframe(outcomes, name = "variable", value = "outcome_label"), by = "variable")

# --- Define custom page layouts ---
page_definitions <- c(
  list(
    Composites_Page = outcome_categories %>%
      filter(category == "DRSP Composites") %>%
      pull(outcome_label)
  ),
  split(
    outcome_categories %>%
      filter(category == "Mood & Physical (DRSP)") %>%
      pull(outcome_label),
    ceiling(seq_along(
      outcome_categories %>%
        filter(category == "Mood & Physical (DRSP)") %>%
        pull(outcome_label)
    ) / 9)
  )
)
names(page_definitions)[2:length(page_definitions)] <-
  paste0("DRSP_Page_", seq_len(length(page_definitions) - 1))

# --- Pre-calculate log-transformations ---
for (var in names(outcomes)) {
  if (var %in% names(df)) {
    log_var_name <- paste0(var, "_Log")
    df[[log_var_name]] <- log(df[[var]] + 1)
  }
}
```

# Section 4: Per ID outcomes (fit functions + helpers)
```{r}
# --- Prediction grids reused per time var ---
pred_grid_list <- setNames(lapply(time_vars, function(tv) {
  pg_df <- data.frame(seq(-1, 1, by = 0.05))
  names(pg_df) <- tv
  pg_df
}), time_vars)

# --- Axis labels + reference lines per time var ---
axis_meta <- lapply(time_vars, function(tv) {
  if (grepl("ov", tv, ignore.case = TRUE)) {
    list(labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses"),
         menses = c(-1, 1), ov = 0)
  } else {
    list(labels = c("Ovulation", "50%L", "Menses", "50%F", "Ovulation"),
         menses = 0, ov = c(-1, 1))
  }
})
names(axis_meta) <- time_vars

# --- Shared helper: count observations in each cycle quarter ---
# Accepts a numeric vector of cycle-time values (NAs allowed; they are dropped).
# Returns a named integer vector of length 4 (Q1–Q4) with counts per quarter.
# Uses left-closed intervals: [-1,-0.5), [-0.5,0), [0,0.5), [0.5,1].
quarter_bin_counts <- function(vals) {
  vals <- vals[!is.na(vals)]
  qbin <- cut(vals,
              breaks = seq(-1, 1, length.out = 5),
              labels = paste0("Q", 1:4),
              include.lowest = TRUE, right = FALSE)
  counts <- as.integer(table(factor(qbin, levels = paste0("Q", 1:4))))
  setNames(counts, paste0("Q", 1:4))
}

# --- Check cycle-quarter coverage for a given time variable ---
# Wrapper around quarter_bin_counts() that extracts the cycle column from a
# data frame and returns a list with quarter_counts, min_per_quarter, and
# passes (TRUE when every quarter has >= min_obs_per_quarter observations).
check_quarter_coverage <- function(data, cycle_var) {
  counts <- quarter_bin_counts(data[[cycle_var]])
  list(
    quarter_counts  = counts,
    min_per_quarter = min(counts),
    passes          = all(counts >= min_obs_per_quarter)
  )
}

# --- Fit + predict one outcome for one time var (bam + predict) ---
fit_outcome <- function(dat, outcome_var, tv, outcomes_map) {
  log_out <- paste0(outcome_var, "_Log")
  keep <- c(log_out, tv)
  if (!all(keep %in% names(dat))) return(NULL)
  dat <- dat[complete.cases(dat[, keep]), keep, drop = FALSE]
  if (nrow(dat) < 10 || var(dat[[log_out]], na.rm = TRUE) < 1e-6) return(NULL)

  # --- Quarter-based coverage check ---
  qcov <- check_quarter_coverage(data = dat, cycle_var = tv)

  message(glue::glue(
    "[{outcome_var}] N={nrow(dat)}, Quarter counts: ",
    "{paste(names(qcov$quarter_counts), qcov$quarter_counts, sep='=', collapse=', ')}, ",
    "Min per quarter: {qcov$min_per_quarter}, Passes: {qcov$passes}"
  ))

  if (!qcov$passes) {
    message(glue::glue(
      "Skipping GAM for {outcome_var}: quarter coverage requirement not met ",
      "(need >={min_obs_per_quarter} per quarter, min={qcov$min_per_quarter})"
    ))
    return(NULL)
  }

  k <- min(6L, nrow(dat) - 1L)
  if (k < 3L) return(NULL)

  fml <- as.formula(paste0(log_out, " ~ s(", tv, ", bs = 'cc', k = ", k, ")"))
  knots <- setNames(list(c(-1, 1)), tv)

  mf <- try(mgcv::bam(fml, data = dat, knots = knots, method = "fREML",
                      discrete = TRUE, select = TRUE),
            silent = TRUE)
  if (inherits(mf, "try-error")) return(NULL)

  s  <- summary(mf)
  p  <- suppressWarnings(as.numeric(s$s.pv[1]))
  st <- s$s.table
  edf <- suppressWarnings(as.numeric(st[1, "edf"]))
  Fval <- suppressWarnings(as.numeric(st[1, "F"]))

  pg <- pred_grid_list[[tv]]
  pr <- predict(mf, newdata = pg, se.fit = TRUE, type = "link")  # link scale for log(y+1)
  est <- exp(pr$fit) - 1
  conf.low  <- exp(pr$fit - 1.96 * pr$se.fit) - 1
  conf.high <- exp(pr$fit + 1.96 * pr$se.fit) - 1

  # Record cycle time at peak (max) and trough (min) of predicted curve
  peak_time   <- pg[[tv]][which.max(est)]
  trough_time <- pg[[tv]][which.min(est)]

  # Count observations in each positional half for reporting.
  n_neg    <- sum(dat[[tv]] <  0, na.rm = TRUE)
  n_nonneg <- sum(dat[[tv]] >= 0, na.rm = TRUE)

  pred <- cbind(pg, estimate = est, conf.low = conf.low, conf.high = conf.high)
  pred$outcome_var   <- outcome_var
  pred$outcome_label <- outcomes_map[[outcome_var]]
  pred$p.value <- p
  pred$raw_range <- diff(range(est, na.rm = TRUE))
  pred$n <- nrow(dat)
  pred$edf <- edf

  list(
    pred = pred,
    sig  = tibble::tibble(
      variable = outcome_var,
      outcome_label = outcomes_map[[outcome_var]],
      time_var = tv,
      p.value = p,
      edf = edf,
      F = Fval,
      n = nrow(dat),
      n_neg = n_neg,
      n_nonneg = n_nonneg,
      range = diff(range(est, na.rm = TRUE)),
      peak_time = peak_time,
      trough_time = trough_time,
      coverage_mode = "quarter_check"
    )
  )
}

# --- Plot one outcome curve with annotations ---
# y_limits = NULL means: use per-outcome defaults from outcome_plot_settings
#   (fixed outcomes → c(y_min, y_max); data-mode outcomes → data-driven, no clamp)
# y_limits = c(lo, hi) means: use those limits as-is (e.g. from shared-y section)
create_annotated_plot <- function(plot_df, tv, axis_meta_list, y_limits = NULL,
                                  raw_data = NULL) {
  if (is.null(plot_df) || nrow(plot_df) == 0) return(NULL)

  # Track whether y_limits were provided by the caller (vs. computed internally)
  y_limits_from_caller <- !is.null(y_limits)

  max_pt <- dplyr::slice_max(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  min_pt <- dplyr::slice_min(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  rng   <- dplyr::first(plot_df$raw_range)
  p_val <- dplyr::first(plot_df$p.value)
  n_obs <- dplyr::first(plot_df$n)
  edf_val <- dplyr::first(plot_df$edf)

  outcome_var <- dplyr::first(plot_df$outcome_var)
  line_col <- drsp_color_map[[outcome_var]]
  if (is.null(line_col) || is.na(line_col)) line_col <- "#0072B2"

  # Look up per-outcome y-axis and reference-line settings
  ps        <- outcome_plot_settings[outcome_plot_settings$variable == outcome_var, ]
  mode      <- if (nrow(ps) > 0) ps$y_axis_mode[1] else "fixed"
  ref_line  <- if (nrow(ps) > 0) ps$ref_line[1]    else 4
  y_fix_min <- if (nrow(ps) > 0) ps$y_min[1]       else 1
  y_fix_max <- if (nrow(ps) > 0) ps$y_max[1]       else 4.5

  # Resolve y-axis limits when not explicitly supplied by the caller
  if (is.null(y_limits)) {
    if (mode == "fixed" && !is.na(y_fix_min) && !is.na(y_fix_max)) {
      y_limits <- c(y_fix_min, y_fix_max)
    }
    # data mode: leave y_limits = NULL → ggplot auto-scales
  }

  # Prepare raw data overlay using the same inclusion rule as fit_outcome:
  # require non-missing log-transformed outcome and time var, then plot on original scale.
  raw_pts <- NULL
  log_out <- paste0(outcome_var, "_Log")
  if (!is.null(raw_data) && outcome_var %in% names(raw_data) &&
      log_out %in% names(raw_data) && tv %in% names(raw_data)) {
    raw_pts <- raw_data[complete.cases(raw_data[, c(log_out, tv)]), c(outcome_var, tv), drop = FALSE]
    if (nrow(raw_pts) == 0) raw_pts <- NULL
  }

  # Dynamic Y-axis expansion for fixed-scale outcomes (where y_limits is non-NULL).
  # Data-mode outcomes leave y_limits = NULL and rely on ggplot auto-scaling instead.
  #
  # Two expansion paths:
  #   1. Caller-provided y_limits (GAM-based global ymax for this ID):
  #      Only expand further if raw data points for THIS outcome exceed the limit.
  #   2. Internally-computed y_limits (no caller override):
  #      Expand based on both GAM predictions and raw data.
  if (!is.null(y_limits)) {
    raw_max <- if (!is.null(raw_pts) && nrow(raw_pts) > 0 && outcome_var %in% names(raw_pts)) {
      vals <- raw_pts[[outcome_var]]
      if (any(is.finite(vals))) max(vals[is.finite(vals)]) else NA_real_
    } else {
      NA_real_
    }
    if (y_limits_from_caller) {
      # Caller already applied GAM-based global expansion; only expand further for raw data
      expanded_ymax <- expand_ymax(y_limits[2], raw_max)
    } else {
      # No caller limits: expand for both predictions and raw data
      pred_max <- if (any(is.finite(plot_df$estimate))) max(plot_df$estimate, na.rm = TRUE) else NA_real_
      max_val <- if (!is.na(raw_max) && !is.na(pred_max)) max(raw_max, pred_max) else
                  if (!is.na(raw_max)) raw_max else pred_max
      expanded_ymax <- expand_ymax(y_limits[2], max_val)
    }
    if (!is.na(expanded_ymax) && expanded_ymax > y_limits[2]) {
      y_limits[2] <- expanded_ymax
      y_fix_max <- expanded_ymax
    }
  }

  # Sync y_fix_max to y_limits[2] so axis breaks always cover the full displayed range
  if (mode == "fixed" && !is.null(y_limits)) {
    y_fix_max <- max(y_fix_max, y_limits[2])
  }

  cpass_flag <- !is.null(rng) && is.finite(rng) && rng >= 1.5
  stars <- dplyr::case_when(
    is.finite(p_val) & p_val < 0.001 ~ "***",
    is.finite(p_val) & p_val < 0.01  ~ "**",
    is.finite(p_val) & p_val < 0.05  ~ "*",
    TRUE ~ ""
  )
  cpass_label <- if (cpass_flag) "\u25B2 Range \u22651.5 | " else ""
  edf_str <- if (!is.null(edf_val) && !is.na(edf_val)) sprintf(", EDF = %.2f", edf_val) else ""
  ann <- if (!is.null(n_obs) && !is.na(n_obs)) {
    sprintf("%s%s Range = %.2f%s, N = %d", cpass_label, stars, rng, edf_str, as.integer(n_obs))
  } else {
    sprintf("%s%s Range = %.2f%s", cpass_label, stars, rng, edf_str)
  }
  meta  <- axis_meta_list[[tv]]

  p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = .data[[tv]], y = estimate))

  # Pink shading over the expanded y-axis region (above 4.5) — rendered first so it sits behind data
  if (!is.null(y_limits) && y_limits[2] > 4.5) {
    p <- p + ggplot2::annotate(
      "rect",
      xmin = -Inf, xmax = Inf,
      ymin = 4.5,  ymax = y_limits[2],
      fill = "pink", alpha = 0.25
    )
  }

  # Reference line only for outcomes that define one (Likert-scale outcomes)
  if (!is.na(ref_line)) {
    p <- p + ggplot2::geom_hline(yintercept = ref_line, color = "red",
                                 linewidth = 0.4, alpha = 0.4)
  }

  # Overlay raw data points behind the GAM curve (slightly prominent for visibility)
  if (!is.null(raw_pts)) {
    p <- p + ggplot2::geom_point(
      data = raw_pts,
      ggplot2::aes(x = .data[[tv]], y = .data[[outcome_var]]),
      color = line_col, alpha = 0.45, size = 1.8, shape = 16,
      inherit.aes = FALSE
    )
  }

  p <- p +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red") +
    ggplot2::geom_vline(xintercept = meta$ov,     color = "forestgreen") +
    ggplot2::geom_line(linewidth = 1, color = line_col) +
    ggplot2::geom_point(data = max_pt, size = 2.5, color = line_col) +
    ggplot2::geom_point(data = min_pt, size = 2.5, color = line_col) +
    ggplot2::annotate("text", x = -1, y = Inf, label = ann,
                      hjust = -0.05, vjust = 1.5, size = 2.5, fontface = "bold") +
    ggplot2::labs(title = dplyr::first(plot_df$outcome_label), x = "Cycle Phase", y = NULL) +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      plot.title = ggplot2::element_text(size = 9, face = "bold", hjust = 0.5),
      plot.margin = ggplot2::margin(t = 10, r = 5, b = 5, l = 5, unit = "pt"),
      axis.text = ggplot2::element_text(size = 7)
    )

  # For fixed-scale outcomes, show integer breaks on the y-axis up to y_fix_max
  if (mode == "fixed" && !is.na(y_fix_min) && !is.na(y_fix_max)) {
    p <- p + ggplot2::scale_y_continuous(
      breaks = seq(y_fix_min, y_fix_max, by = 1)
    )
  }

  # Always disable clipping so top annotations (y = Inf) are never cut off.
  # When y_limits is defined, also lock the y-axis range.
  if (!is.null(y_limits)) {
    p <- p + ggplot2::coord_cartesian(ylim = y_limits, clip = "off")
  } else {
    p <- p + ggplot2::coord_cartesian(clip = "off")
  }

  # Colored effect-size borders for DRSP individual items
  is_drsp <- grepl(DRSP_VAR_PATTERN, outcome_var)

  if (is_drsp) {
    border_col <- dplyr::case_when(
      is.finite(rng) && rng >= 1.5 ~ "red",
      is.finite(rng) && rng >= 1.0 ~ "salmon",
      is.finite(rng) && rng >= 0.5 ~ "pink",
      TRUE                         ~ NA_character_
    )

    if (!is.na(border_col)) {
      p <- p + ggplot2::theme(
        panel.border = ggplot2::element_rect(color = border_col, linewidth = 1.5, fill = NA)
      )
    }
  }

  p
}

# --- Dynamic Y-axis expansion for all fixed-scale outcomes ---
# All fixed-scale outcomes default to y_max = 4.5 but expand to 5 or 6 if observed values exceed it.
# GAM predictions exceeding 4.5 drive global (all-outcome) expansion for the person.
# Raw data points exceeding 4.5 drive per-outcome expansion only.
expand_ymax <- function(y_max_current, max_observed, absolute_max = 6) {
  if (is.na(max_observed) || !is.finite(max_observed) || max_observed <= y_max_current) {
    return(y_max_current)
  }
  min(ceiling(max_observed), absolute_max)
}
```

# Section 5: Per ID outcome reports (main PDF loop)
```{r}
report_folder <- file.path(output_folder, "Participant_Reports_PDF_byID")
dir.create(report_folder, showWarnings = FALSE, recursive = TRUE)

ids_to_run    <- unique(df$id)   # set to subset for testing, e.g., c("203","109")
skip_existing <- TRUE            # set FALSE to overwrite

# collector for a single CSV at the end
all_ids_signif <- list()

# Per-ID global ymax lookup: stores the expanded y-axis max for each ID
# (computed across ALL outcomes and time vars). Used in both person-reports
# (Section 5) and outcome-reports (Section 7).
id_global_ymax_map <- c()

# fixed layout controls
max_plots_per_page <- plots_per_col * plots_per_row
table_theme <- gridExtra::ttheme_default(
  base_size = 7,
  core    = list(fg_params = list(hjust = 0, x = 0.05)),
  colhead = list(fg_params = list(hjust = 0, x = 0.05)),
  padding = grid::unit(c(3, 2), "mm")
)

for (current_id in ids_to_run) {
  message(sprintf("--- Processing ID: %s ---", current_id))
  datSX_id <- subset(df, id == current_id)

  # ---- Open a single LANDSCAPE PDF per participant ----
  pdf_filename <- file.path(report_folder, sprintf("Report_ID_%s.pdf", current_id))
  if (isTRUE(skip_existing) && file.exists(pdf_filename)) {
    message("    -> File exists; skipping. Set skip_existing=FALSE to overwrite.")
    next
  }
  grDevices::pdf(pdf_filename, width = 11, height = 8.5)   # LANDSCAPE

  # ---- Page 0: Raw data coverage (Joint/Muscle Pain + DRSP Negative Affect) ----
  # Shows raw data points across the cycle variable so the reader can quickly
  # judge whether there is enough data across phases for interpretation.
  # This page is always produced, even if GAMs cannot be run.
  if (length(time_vars) > 0) {
    tv0 <- time_vars[1]
    meta0 <- axis_meta[[tv0]]

    # Returns TRUE when any quarter has fewer than min_obs_per_quarter observations.
    is_quarter_sparse <- function(qcounts) {
      any(qcounts < min_obs_per_quarter)
    }

    make_raw_panel <- function(dat, tv, meta, var, y_lab, color) {
      if (!var %in% names(dat)) {
        p <- ggplot2::ggplot() +
          ggplot2::annotate("text", x = 0, y = 0, label = paste0(var, " not in data"),
                            size = 5, color = "grey50") +
          ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = "n = 0 daily observations") +
          ggplot2::theme_minimal(base_size = 12)
        return(p)
      }
      vals <- dat[[var]]
      tv_vals <- dat[[tv]]
      valid <- !is.na(vals) & !is.na(tv_vals)
      n_obs <- sum(valid)
      has_data <- n_obs > 0
      qcounts <- quarter_bin_counts(tv_vals[valid])
      sub_label <- paste0(
        "n = ", n_obs, " daily observations",
        "  (Q1: ", qcounts["Q1"], ", Q2: ", qcounts["Q2"],
        ", Q3: ", qcounts["Q3"], ", Q4: ", qcounts["Q4"], ")"
      )
      p <- ggplot2::ggplot(dat, ggplot2::aes(x = .data[[tv]], y = .data[[var]])) +
        ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::scale_x_continuous(
          limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
          labels = meta$labels
        ) +
        ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = sub_label) +
        ggplot2::theme_minimal(base_size = 12)
      if (has_data) {
        p <- p + ggplot2::geom_point(color = color, alpha = 0.45, size = 1.8)
      } else {
        p <- p + ggplot2::annotate("text", x = 0, y = 0, label = "No data",
                                   size = 5, color = "grey50")
      }
      p
    }

    p_joint  <- make_raw_panel(datSX_id, tv0, meta0, "drsp_21",
                               "Joint/Muscle Pain (DRSP 21)", "#CC6677")
    p_drsp_na <- make_raw_panel(datSX_id, tv0, meta0, "drsp_NA",
                                "DRSP Negative Affect Mean", "#D55E00")

    tv0_vals <- datSX_id[[tv0]]
    tv0_vals <- tv0_vals[!is.na(tv0_vals)]
    tv0_qcounts <- quarter_bin_counts(tv0_vals)
    tv0_sparse  <- is_quarter_sparse(tv0_qcounts)

    # Subtitle is informational only; the skip decision is made per time variable.
    coverage_subtitle <- paste0("Time variable: ", tv0)
    if (tv0_sparse) {
      coverage_subtitle <- paste0(
        coverage_subtitle,
        "\n\u26A0 Fewer than ", min_obs_per_quarter,
        " observations in one or more cycle quarters for ", tv0
      )
    }

    p_cover <- p_joint / p_drsp_na +
      patchwork::plot_annotation(
        title    = paste0("Raw Data Coverage \u2014 ID: ", current_id),
        subtitle = coverage_subtitle,
        theme    = ggplot2::theme(
          plot.title    = ggplot2::element_text(face = "bold", size = 12),
          plot.subtitle = ggplot2::element_text(
            size = if (tv0_sparse) 10 else 9,
            color = if (tv0_sparse) "red" else "grey30"
          )
        )
      )
    print(p_cover)
  } else {
    grid::grid.newpage()
    grid::grid.text(
      label = paste0(
        "Raw Data Coverage \u2014 ID: ", current_id,
        "\n\nNo eligible time variables were available for GAM fitting.",
        "\nA coverage-only placeholder report was generated."
      ),
      x = 0.5, y = 0.55,
      gp = grid::gpar(fontsize = 12, fontface = "bold")
    )
    grDevices::dev.off()
    next
  } # end Page 0

  # ---- Pass 1: fit & collect (coverage gate applied per time variable) ----
  preds_by_tv <- list()
  sig_by_tv   <- list()

  for (tv in time_vars) {
    message(sprintf("  ... time variable: %s", tv))

    # Per-tv coverage gate — require ≥min_obs_per_quarter in each cycle quarter.
    tv_vals_all <- datSX_id[[tv]]
    tv_vals_all <- tv_vals_all[!is.na(tv_vals_all)]
    tv_qcounts  <- quarter_bin_counts(tv_vals_all)
    if (is_quarter_sparse(tv_qcounts)) {
      message(sprintf("    -> Skipping tv=%s: insufficient quarter coverage (Q1=%d, Q2=%d, Q3=%d, Q4=%d, min=%d)",
                      tv, tv_qcounts["Q1"], tv_qcounts["Q2"], tv_qcounts["Q3"], tv_qcounts["Q4"],
                      min_obs_per_quarter))
      next
    }

    all_preds <- list()
    sig_list  <- list()

    for (outcome_var in names(outcomes)) {
      res <- fit_outcome(datSX_id, outcome_var, tv, outcomes)
      if (is.null(res)) next
      all_preds[[length(all_preds) + 1]] <- res$pred
      sig_list[[length(sig_list) + 1]]  <- res$sig
    }

    if (length(all_preds) > 0) {
      preds_by_tv[[tv]] <- dplyr::bind_rows(all_preds)
      sig_by_tv[[tv]]   <- dplyr::bind_rows(sig_list)
    }
  }

  if (!length(preds_by_tv)) {
    message("    -> No successful models for this ID; skipping.")
    grDevices::dev.off()
    next
  }

  # ---- Compute global ymax for this ID across ALL outcomes and time vars ----
  # Only GAM predictions drive the global expansion: if any outcome's predicted
  # curve exceeds 4.5, ALL plots for this person expand to the same ymax (5 or 6).
  # Raw data points above 4.5 expand only the individual outcome plot where they
  # occur (handled inside create_annotated_plot).
  global_max_val <- 4.5
  # Check GAM predictions only (raw-data expansion is per-outcome, not global)
  for (tv_name in names(preds_by_tv)) {
    est_vals <- preds_by_tv[[tv_name]]$estimate
    finite_est <- est_vals[is.finite(est_vals)]
    if (length(finite_est) > 0) global_max_val <- max(global_max_val, max(finite_est))
  }
  id_global_ymax <- expand_ymax(4.5, global_max_val)
  id_global_ylimits <- c(1, id_global_ymax)
  id_global_ymax_map[as.character(current_id)] <- id_global_ymax

  # Build per-ID significance object for CSV
  sig_all <- dplyr::bind_rows(sig_by_tv)
  sig_all$id <- current_id
  all_ids_signif[[as.character(current_id)]] <- sig_all

  # ---- COVER: Compact counts by time_var ----
  sig_all_clean <- sig_all %>%
    dplyr::mutate(
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    )

  cover_counts <- sig_all_clean %>%
    dplyr::group_by(time_var) %>%
    dplyr::summarise(
      tested = dplyr::n(),
      n_sig  = sum(stars != "", na.rm = TRUE),
      n_cpass = sum(is.finite(range) & range >= 1.5 & stars != "", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::unite("count", n_sig, tested, sep = " / ") %>%
    dplyr::rename(`Sig / Tested` = count, `Range ≥1.5 sig` = n_cpass)

  cover_theme <- gridExtra::ttheme_default(
    base_size = 9,
    core = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"),
      padding  = grid::unit(c(2, 2), "mm")
    ),
    colhead = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
    )
  )

  title_grob    <- grid::textGrob(paste("Compact Summary — ID:", current_id),
                                  gp = grid::gpar(fontsize = 16, fontface = "bold"))
  subtitle_grob <- grid::textGrob("Rows show: # significant / # tested (p < .05) for each time variable",
                                  gp = grid::gpar(fontsize = 9))
  cover_grob    <- gridExtra::tableGrob(cover_counts, rows = NULL, theme = cover_theme)
  gridExtra::grid.arrange(title_grob, subtitle_grob, cover_grob,
                          ncol = 1, heights = grid::unit(c(0.12, 0.08, 0.80), "npc"))

  # ---- OPTIONAL: compact stars matrix for outcomes with any significance ----
  compact_stars <- sig_all_clean %>%
    dplyr::group_by(variable) %>%
    dplyr::mutate(any_sig = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(any_sig) %>%
    dplyr::mutate(Outcome = dplyr::coalesce(outcome_label, variable)) %>%
    dplyr::mutate(
      Outcome = ifelse(nchar(Outcome) > 22, variable, Outcome),
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    ) %>%
    dplyr::select(Outcome, time_var, stars) %>%
    tidyr::pivot_wider(names_from = time_var, values_from = stars, values_fill = "") %>%
    dplyr::arrange(Outcome)

  if (nrow(compact_stars) > 0) {
    compact_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(
        fg_params = list(hjust = 0, x = 0.02),
        padding = grid::unit(c(1.5, 2), "mm")
      ),
      colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
    )
    header <- grid::textGrob("Significant Outcomes Only (★ = p < .05)",
                             gp = grid::gpar(fontsize = 12, fontface = "bold"))
    tbl <- gridExtra::tableGrob(compact_stars, rows = NULL, theme = compact_theme)
    gridExtra::grid.arrange(header, tbl, ncol = 1, heights = grid::unit(c(0.10, 0.90), "npc"))
  }

  # ---- PAGE 2: Significant EDFs ranked by Range (across all time vars) ----
  sig_ranked <- sig_all_clean %>%
    dplyr::filter(is.finite(p.value) & p.value < 0.05) %>%
    dplyr::mutate(
      Outcome = dplyr::coalesce(outcome_label, variable),
      Outcome = ifelse(nchar(Outcome) > 40, variable, Outcome)
    ) %>%
    dplyr::transmute(
      Outcome,
      TimeVar  = time_var,
      EDF      = round(edf, 2),
      Range    = round(range, 2),
      F        = round(F, 2),
      N        = n
    ) %>%
    dplyr::arrange(dplyr::desc(Range), Outcome, TimeVar)

  if (nrow(sig_ranked) == 0) {
    header <- grid::textGrob(
      paste("Significant Smooths — ID:", current_id, "(none at p < .05)"),
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    sub <- grid::textGrob("No significant cyclic effects detected across any time variable.",
                          gp = grid::gpar(fontsize = 10, fontface = "italic"))
    gridExtra::grid.arrange(header, sub, ncol = 1,
                            heights = grid::unit(c(0.2, 0.8), "npc"))
  } else {
    ranked_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(
        fg_params = list(hjust = 0, x = 0.02),
        padding = grid::unit(c(1.2, 2), "mm")
      ),
      colhead = list(
        fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
      )
    )
    title_grob <- grid::textGrob(
      paste("Significant Smooths (p < .05), Ranked by Range — ID:", current_id),
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    note_grob <- grid::textGrob(
      "Columns: EDF = effective degrees of freedom; Range = max(pred) − min(pred) on original scale.",
      gp = grid::gpar(fontsize = 9, fontface = "italic")
    )
    ranked_tbl <- gridExtra::tableGrob(sig_ranked, rows = NULL, theme = ranked_theme)
    gridExtra::grid.arrange(
      title_grob, note_grob, ranked_tbl,
      ncol = 1,
      heights = grid::unit(c(0.12, 0.06, 0.82), "npc")
    )
  }

  # ---- Helper: generate an overlay page for qualifying GAM curves ----
  # @param var_pattern    regex to select variables; used if var_list is NULL
  # @param var_list       character vector of variable names; overrides var_pattern if provided
  # @param range_thresh   minimum range for inclusion alongside p < .05
  # @param title_prefix   text before " — ID: ..." in plot title
  # @param subtitle_label description of items shown
  # @param ylim_strategy  "fixed" uses fixed_ylim arg; "data" derives from conf intervals
  # @param fixed_ylim     c(lo, hi) used when ylim_strategy == "fixed"
  # @param show_hline     whether to show the horizontal reference line at hline_y
  # @param hline_y        y-intercept for the reference line
  # @param legend_ncol    number of columns in the legend
  plot_overlay_page <- function(sig_tv, pred_tv, tv, meta,
                                range_thresh, title_prefix, subtitle_label,
                                var_pattern = NULL, var_list = NULL,
                                ylim_strategy = "fixed", fixed_ylim = c(1, 6),
                                show_hline = TRUE, hline_y = 4,
                                legend_ncol = 3,
                                show_range_in_labels = FALSE,
                                normalize_01 = FALSE,
                                ref_lines_per_var = NULL) {
    if (!is.null(var_list)) {
      qualifying <- sig_tv %>%
        dplyr::filter(variable %in% var_list)
    } else if (!is.null(var_pattern)) {
      qualifying <- sig_tv %>%
        dplyr::filter(grepl(var_pattern, variable))
    } else {
      return(invisible(NULL))
    }
    qualifying <- qualifying %>%
      dplyr::filter(
        (is.finite(p.value) & p.value < 0.05) |
        (is.finite(range) & range >= range_thresh)
      )
    if (nrow(qualifying) == 0) return(invisible(NULL))

    ov_vars <- unique(qualifying$variable)
    ov_data <- pred_tv %>% dplyr::filter(outcome_var %in% ov_vars)
    if (nrow(ov_data) == 0) return(invisible(NULL))

    n_curves <- length(ov_vars)

    ov_colors <- drsp_color_map[ov_vars]
    ov_colors[is.na(ov_colors)] <- "#0072B2"
    names(ov_colors) <- ov_vars

    range_lookup <- qualifying %>%
      dplyr::distinct(variable, .keep_all = TRUE) %>%
      dplyr::select(variable, range)

    if (show_range_in_labels) {
      ov_labels <- sapply(ov_vars, function(v) {
        r <- range_lookup$range[range_lookup$variable == v]
        if (length(r) > 0 && is.finite(r)) {
          paste0(outcomes[v], " (", sprintf("%.2f", r), ")")
        } else {
          outcomes[v]
        }
      })
    } else {
      ov_labels <- outcomes[ov_vars]
    }

    if (normalize_01) {
      ov_data <- ov_data %>%
        dplyr::left_join(
          norm_scale_ranges,
          by = c("outcome_var" = "variable")
        ) %>%
        dplyr::group_by(outcome_var) %>%
        dplyr::mutate(
          scale_min = dplyr::coalesce(norm_min, min(estimate, na.rm = TRUE)),
          scale_max = dplyr::coalesce(norm_max, max(estimate, na.rm = TRUE)),
          scale_range = scale_max - scale_min,
          estimate  = ifelse(scale_range > 0, (estimate  - scale_min) / scale_range, 0.5),
          conf.low  = ifelse(scale_range > 0, (conf.low  - scale_min) / scale_range, 0.5),
          conf.high = ifelse(scale_range > 0, (conf.high - scale_min) / scale_range, 0.5)
        ) %>%
        dplyr::select(-norm_min, -norm_max, -scale_min, -scale_max, -scale_range) %>%
        dplyr::ungroup()
      ylim_strategy <- "fixed"
      fixed_ylim <- c(0, 1)
    }

    if (ylim_strategy == "data") {
      y_min <- min(ov_data$conf.low, na.rm = TRUE)
      y_max <- max(ov_data$conf.high, na.rm = TRUE)
      y_pad <- (y_max - y_min) * 0.05
      plot_ylim <- c(max(0, y_min - y_pad), y_max + y_pad)
    } else {
      plot_ylim <- fixed_ylim
      # Auto-expand y_max from 4.5 to 5 or 6 if data exceeds it (only for non-normalized axes)
      if (!normalize_01 && any(is.finite(ov_data$estimate))) {
        max_est <- max(ov_data$estimate, na.rm = TRUE)
        expanded_ymax <- expand_ymax(plot_ylim[2], max_est)
        if (expanded_ymax > plot_ylim[2]) {
          plot_ylim[2] <- expanded_ymax
        }
      }
    }

    range_label <- format(range_thresh, nsmall = 1)
    y_axis_label <- if (normalize_01) "Normalized Score (0\u20131)" else "Predicted Score"

    if (is.infinite(range_thresh)) {
      filter_label <- "p < .05"
    } else {
      filter_label <- paste0("p < .05 or range \u2265 ", range_label)
    }

    p <- ggplot2::ggplot(ov_data,
           ggplot2::aes(x = .data[[tv]], y = estimate, color = outcome_var))

    # Pink shading over the expanded y-axis region (above 4.5) — rendered first
    if (!normalize_01 && plot_ylim[2] > 4.5) {
      p <- p + ggplot2::annotate(
        "rect",
        xmin = -Inf, xmax = Inf,
        ymin = 4.5,  ymax = plot_ylim[2],
        fill = "pink", alpha = 0.25
      )
    }

    if (show_hline) {
      p <- p + ggplot2::geom_hline(yintercept = hline_y, color = "red",
                                   linewidth = 0.4, alpha = 0.4)
    }

    if (!is.null(ref_lines_per_var)) {
      for (rv in names(ref_lines_per_var)) {
        if (rv %in% ov_vars) {
          ref_col <- ov_colors[[rv]]
          if (is.null(ref_col) || is.na(ref_col)) ref_col <- "#0072B2"
          p <- p + ggplot2::geom_hline(
            yintercept = ref_lines_per_var[[rv]],
            color = ref_col, linewidth = 0.5, linetype = "solid", alpha = 0.7
          )
        }
      }
    }

    p <- p +
      ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                          linewidth = 0.6, linetype = "dashed") +
      ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                          linewidth = 0.6, linetype = "dashed") +
      ggplot2::geom_line(linewidth = 0.9) +
      ggplot2::scale_color_manual(
        values = ov_colors,
        breaks = ov_vars,
        labels = ov_labels
      ) +
      ggplot2::scale_x_continuous(
        limits = c(-1, 1),
        breaks = seq(-1, 1, by = 0.5),
        labels = meta$labels
      ) +
      ggplot2::coord_cartesian(ylim = plot_ylim) +
      ggplot2::labs(
        title = paste0(title_prefix, " \u2014 ID: ", current_id),
        subtitle = paste0(
          subtitle_label, " with ", filter_label,
          "  |  N curves = ", n_curves, "  |  Time variable: ", tv
        ),
        x = "Cycle Phase",
        y = y_axis_label,
        color = "Outcome"
      ) +
      ggplot2::theme_minimal(base_size = 10) +
      ggplot2::theme(
        legend.position = "bottom",
        legend.text = ggplot2::element_text(size = 6),
        plot.title = ggplot2::element_text(face = "bold", size = 12),
        plot.subtitle = ggplot2::element_text(size = 8),
        plot.margin = ggplot2::margin(t = 8, r = 8, b = 8, l = 8, unit = "pt")
      ) +
      ggplot2::guides(
        color = ggplot2::guide_legend(ncol = legend_ncol)
      )

    print(p)
  }

  # ---- Overlaid GAM pages by outcome category ----
  for (tv in names(preds_by_tv)) {

    # (1) Individual DRSP items: sig OR range >= 0.5, Y scale uses per-ID global ymax
    plot_overlay_page(
      sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
      tv = tv, meta = axis_meta[[tv]],
      var_list       = all_raw_drsp,
      range_thresh   = 0.5,
      title_prefix   = "Overlaid GAM Curves (Individual DRSP Items)",
      subtitle_label = "Individual DRSP items",
      ylim_strategy  = "fixed", fixed_ylim = id_global_ylimits,
      show_hline = TRUE, hline_y = 4,
      legend_ncol = 4,
      show_range_in_labels = TRUE
    )

    # (2) DRSP composite outcomes: sig only, Y scale uses per-ID global ymax
    composite_vars <- names(outcomes)[!grepl(DRSP_VAR_PATTERN, names(outcomes))]
    plot_overlay_page(
      sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
      tv = tv, meta = axis_meta[[tv]],
      var_list       = composite_vars,
      range_thresh   = Inf,
      title_prefix   = "Overlaid GAM Curves (DRSP Composites)",
      subtitle_label = "DRSP composite outcomes",
      ylim_strategy  = "fixed", fixed_ylim = id_global_ylimits,
      show_hline = FALSE,
      legend_ncol = 3,
      show_range_in_labels = TRUE
    )
  }

  # ---- Sections per time variable: compact table + pages of plots ----
  for (tv in names(preds_by_tv)) {

    summary_df <- sig_by_tv[[tv]] %>%
      dplyr::left_join(dplyr::select(outcome_categories, variable, category), by = "variable") %>%
      dplyr::mutate(
        stars = dplyr::case_when(
          is.finite(p.value) & p.value < 0.001 ~ "***",
          is.finite(p.value) & p.value < 0.01  ~ "**",
          is.finite(p.value) & p.value < 0.05  ~ "*",
          TRUE ~ ""
        ),
        Outcome = dplyr::coalesce(outcome_label, variable)
      ) %>%
      dplyr::select(Category = category, Outcome, N = n, EDF = edf, F = F, Range = range, Sig = stars) %>%
      dplyr::arrange(Category, Outcome)

    tv_title <- grid::textGrob(paste("Significance — ID:", current_id, "| Time Variable:", tv),
                               gp = grid::gpar(fontsize = 14, fontface = "bold"))
    tv_table <- gridExtra::tableGrob(summary_df, rows = NULL, theme = table_theme)
    gridExtra::grid.arrange(tv_title, tv_table, ncol = 1,
                            heights = grid::unit(c(0.12, 0.88), "npc"))

    # Plots organized by outcome category (using page_definitions)
    combined_plot_data <- preds_by_tv[[tv]]

    for (page_name in names(page_definitions)) {
      page_outcome_labels <- page_definitions[[page_name]]
      page_data <- dplyr::filter(combined_plot_data, outcome_label %in% page_outcome_labels)
      if (nrow(page_data) == 0) next

      plot_list <- lapply(page_outcome_labels, function(lab) {
        plot_df <- dplyr::filter(page_data, outcome_label == lab)
        if (nrow(plot_df) == 0) return(NULL)
        create_annotated_plot(plot_df, tv, axis_meta, y_limits = id_global_ylimits, raw_data = datSX_id)
      })
      plot_list <- Filter(Negate(is.null), plot_list)

      if (length(plot_list) > 0) {
        if (length(plot_list) < max_plots_per_page) {
          plot_list <- c(plot_list,
                         rep(list(patchwork::plot_spacer()), max_plots_per_page - length(plot_list)))
        }
        page_title_label <- gsub("_", " ", page_name)
        page_title <- paste(page_title_label, "\u2014 ID:", current_id, "| Time Variable:", tv)
        final_page <- patchwork::wrap_plots(plotlist = plot_list, ncol = plots_per_col) +
          patchwork::plot_annotation(
            title = page_title,
            theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 12, face = "bold"))
          )
        print(final_page)
      }
    }
  }

  grDevices::dev.off()
  message(sprintf("    --- \u2705 Saved Report: %s ---", pdf_filename))
}

# ---- Write long CSV for evaluation across IDs ----
if (length(all_ids_signif)) {
  eval_csv <- file.path(output_folder, paste0("participant_significance_summary_", current_date, ".csv"))
  write.csv(dplyr::bind_rows(all_ids_signif), eval_csv, row.names = FALSE)
  message(sprintf("Wrote summary CSV: %s", eval_csv))
}
```

# Section 5b: Overall Sample Summary Report
```{r overall-summary-report}
# =========================================================
# OVERALL SAMPLE SUMMARY REPORT
# Generates a landscape PDF summarizing the distributions and
# descriptives of cyclical symptom change across the full sample.
# =========================================================

if (length(all_ids_signif) > 0) {
  gam_summary_all <- dplyr::bind_rows(all_ids_signif)

  overall_pdf <- file.path(output_folder, paste0("Overall_Sample_Summary_", current_date, ".pdf"))
  grDevices::pdf(overall_pdf, width = 11, height = 8.5)

  # --- Page numbering helper ---
  # NOTE: Update total_report_pages if pages are added/removed from this report.
  total_report_pages <- 9L
  page_counter <- 0L
  draw_page_number <- function() {
    page_counter <<- page_counter + 1L
    grid::grid.text(
      paste0("Page ", page_counter, " of ", total_report_pages),
      x = 0.98, y = 0.02,
      just = c("right", "bottom"),
      gp = grid::gpar(fontsize = 8, col = "grey50")
    )
  }

  # ---- COVER PAGE (p.1): How to Read This Report ----
  cover_title <- grid::textGrob(
    "How to Read This Report",
    gp = grid::gpar(fontsize = 24, fontface = "bold")
  )
  cover_purpose <- grid::textGrob(
    paste0(
      "This report summarizes cyclical symptom patterns across the menstrual cycle\n",
      "for all participants in your sample. Each participant\u2019s daily symptom ratings\n",
      "were modeled using Generalized Additive Models (GAMs) to capture smooth,\n",
      "non-linear patterns of change across the cycle."
    ),
    gp = grid::gpar(fontsize = 11, lineheight = 1.3), just = "left", x = 0.05
  )
  cover_defs <- grid::textGrob(
    paste0(
      "KEY DEFINITIONS\n\n",
      "\u2022 Range: The difference between the highest and lowest predicted values from\n",
      "  each person\u2019s GAM curve \u2014 how much a symptom fluctuates across the cycle.\n\n",
      "\u2022 Significant GAM (p < .05): The cyclical pattern is statistically\n",
      "  distinguishable from a flat line \u2014 evidence symptoms truly vary across\n",
      "  the cycle.\n\n",
      "\u2022 Range \u2265 1.5: A swing of 1.5+ points on the 1\u20136 DRSP scale, equivalent to\n",
      "  the C-PASS 30% change criterion used to benchmark PMDD cyclical change.\n",
      "  This is a reference value for comparison with the existing C-PASS standard,\n",
      "  not a validated clinical cutpoint.\n\n",
      "\u2022 Affective outcomes: Mood/emotional symptoms (depression, anxiety,\n",
      "  irritability, etc.) and composites (Neg. Affect, Distress, Lability,\n",
      "  Low Arousal).\n\n",
      "\u2022 Physical outcomes: Somatic symptoms (breast tenderness, headache,\n",
      "  overeating, insomnia, etc.) and composites (Breast Sx, Eating Sx).\n\n",
      "\u2022 GAM-valid: Participants with enough data coverage across the menstrual\n",
      "  cycle (\u22652 obs per quarter) to fit a reliable model."
    ),
    gp = grid::gpar(fontsize = 10, lineheight = 1.2), just = "left", x = 0.05
  )
  cover_toc <- grid::textGrob(
    paste0(
      "REPORT CONTENTS\n\n",
      "  Page  1 \u2014 How to Read This Report (this page)\n",
      "  Page  2 \u2014 Key Metrics Summary with Interpretation Guide\n",
      "  Page  3 \u2014 Per-Outcome Descriptives Table (color-coded)\n",
      "  Page  4 \u2014 Distribution of Ranges by Outcome (histograms)\n",
      "  Page  5 \u2014 Significance & Range Rates by Outcome (grouped bars)\n",
      "  Page  6 \u2014 Per-Person Range Distributions (dot + box plot)\n",
      "  Page  7 \u2014 Significance/Range Heatmap (Outcome \u00d7 Participant)\n",
      "  Page  8 \u2014 Affective vs. Physical Symptom Comparison\n",
      "  Page  9 \u2014 Symptom Co-Occurrence Matrix"
    ),
    gp = grid::gpar(fontsize = 10, lineheight = 1.2), just = "left", x = 0.05
  )
  gridExtra::grid.arrange(
    cover_title, cover_purpose, cover_defs, cover_toc,
    ncol = 1,
    heights = grid::unit(c(0.06, 0.12, 0.55, 0.27), "npc"),
    padding = grid::unit(0.5, "lines")
  )
  draw_page_number()

  # --- Classify outcomes for affective calculations ---
  # Affective raw DRSP items: all individual DRSP items EXCEPT physical symptoms
  affective_raw_drsp <- setdiff(all_raw_drsp, physical_symptoms)

  # Affective composites: composites capturing mood/affect constructs
  affective_composites <- c("drsp_NA", "drsp_distress", "drsp_labilsocsx", "drsp_lowarousal")

  # All affective outcomes (raw + composite)
  all_affective_outcomes <- c(affective_raw_drsp, affective_composites)

  # Physical composites and all physical outcomes
  phys_composites <- c("drsp_breastpain", "drsp_eatsx")
  all_physical_outcomes <- c(physical_symptoms, phys_composites)

  # --- Per-ID summary: best result per outcome (one row per id × variable) ---
  # Use finite-safe aggregation to avoid Inf/-Inf when all values are NA.
  id_outcome_summary <- gam_summary_all %>%
    dplyr::group_by(id, variable) %>%
    dplyr::summarise(
      best_p      = ifelse(any(is.finite(p.value)), min(p.value[is.finite(p.value)]), NA_real_),
      best_range  = ifelse(any(is.finite(range)),   max(range[is.finite(range)]),     NA_real_),
      best_n      = ifelse(any(is.finite(n)),        max(n[is.finite(n)]),             NA_real_),
      sig         = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE),
      .groups     = "drop"
    )

  all_ids <- unique(gam_summary_all$id)
  n_gam_valid <- length(all_ids)
  n_total_sample <- length(unique(df$id))

  # --- KEY METRIC 1: % of GAM-valid sample with at least one symptom/composite range >= 1.5 ---
  ids_any_range15 <- id_outcome_summary %>%
    dplyr::filter(is.finite(best_range) & best_range >= 1.5) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_any_range15 <- length(ids_any_range15) / n_gam_valid * 100

  # --- KEY METRIC 2: % of GAM-valid sample with at least one affective symptom/composite range >= 1.5 ---
  ids_affective_range15 <- id_outcome_summary %>%
    dplyr::filter(variable %in% all_affective_outcomes,
                  is.finite(best_range) & best_range >= 1.5) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_affective_range15 <- length(ids_affective_range15) / n_gam_valid * 100

  # --- KEY METRIC 3: % of GAM-valid sample with at least one significant GAM (any outcome) ---
  ids_any_sig <- id_outcome_summary %>%
    dplyr::filter(sig) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_any_sig <- length(ids_any_sig) / n_gam_valid * 100

  # --- KEY METRIC 4: % of GAM-valid sample with at least one significant affective GAM ---
  ids_affective_sig <- id_outcome_summary %>%
    dplyr::filter(variable %in% all_affective_outcomes, sig) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_affective_sig <- length(ids_affective_sig) / n_gam_valid * 100

  # ---- PAGE 1: Overall Descriptive Summary ----
  # Per-outcome descriptive statistics across all participants
  # Percentages use n_gam_valid (participants with valid GAM data) as denominator
  outcome_desc <- id_outcome_summary %>%
    dplyr::group_by(variable) %>%
    dplyr::summarise(
      N           = dplyr::n(),
      mean_range  = mean(best_range, na.rm = TRUE),
      sd_range    = sd(best_range, na.rm = TRUE),
      median_range = median(best_range, na.rm = TRUE),
      q25_range   = quantile(best_range, 0.25, na.rm = TRUE),
      q75_range   = quantile(best_range, 0.75, na.rm = TRUE),
      min_range   = min(best_range, na.rm = TRUE),
      max_range   = max(best_range, na.rm = TRUE),
      n_sig       = sum(sig, na.rm = TRUE),
      pct_sig     = sum(sig, na.rm = TRUE) / n_gam_valid * 100,
      n_range15   = sum(is.finite(best_range) & best_range >= 1.5, na.rm = TRUE),
      pct_range15 = sum(is.finite(best_range) & best_range >= 1.5, na.rm = TRUE) / n_gam_valid * 100,
      n_sig_or_range15   = sum(sig | (is.finite(best_range) & best_range >= 1.5), na.rm = TRUE),
      pct_sig_or_range15 = sum(sig | (is.finite(best_range) & best_range >= 1.5), na.rm = TRUE) / n_gam_valid * 100,
      .groups     = "drop"
    ) %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(label = dplyr::coalesce(label, variable)) %>%
    dplyr::arrange(dplyr::desc(median_range))

  # Key metrics summary table
  key_metrics <- data.frame(
    Metric = c(
      "Total Participants in Data",
      "Participants with Valid GAM Data",
      "",
      "N with \u2265 1 significant GAM (any outcome)",
      "% with \u2265 1 significant GAM (any outcome)",
      "N with \u2265 1 significant affective GAM",
      "% with \u2265 1 significant affective GAM",
      "",
      "N with \u2265 1 symptom/composite range \u2265 1.5",
      "% with \u2265 1 symptom/composite range \u2265 1.5",
      "N with \u2265 1 affective symptom/composite range \u2265 1.5",
      "% with \u2265 1 affective symptom/composite range \u2265 1.5"
    ),
    Value = c(
      as.character(n_total_sample),
      as.character(n_gam_valid),
      "",
      as.character(length(ids_any_sig)),
      sprintf("%.1f%%", pct_any_sig),
      as.character(length(ids_affective_sig)),
      sprintf("%.1f%%", pct_affective_sig),
      "",
      as.character(length(ids_any_range15)),
      sprintf("%.1f%%", pct_any_range15),
      as.character(length(ids_affective_range15)),
      sprintf("%.1f%%", pct_affective_range15)
    ),
    stringsAsFactors = FALSE
  )

  key_theme <- gridExtra::ttheme_default(
    base_size = 11,
    core    = list(fg_params = list(hjust = 0, x = 0.05)),
    colhead = list(fg_params = list(hjust = 0, x = 0.05, fontface = "bold"))
  )

  title_grob <- grid::textGrob(
    "Overall Sample Summary \u2014 Cyclical Symptom Change",
    gp = grid::gpar(fontsize = 18, fontface = "bold")
  )
  subtitle_grob <- grid::textGrob(
    paste0("N = ", n_total_sample, " total | N = ", n_gam_valid,
           " with valid GAM data (percentages use GAM-valid N as denominator)"),
    gp = grid::gpar(fontsize = 11, fontface = "italic")
  )
  key_tbl <- gridExtra::tableGrob(key_metrics, rows = NULL, theme = key_theme)
  note_grob <- grid::textGrob(
    paste0("Affective outcomes include: raw DRSP emotional/affective items (excl. physical) ",
           "and composites (Neg. Affect, Distress, Social/Affect Lability, Low Arousal). ",
           "All percentages use GAM-valid N = ", n_gam_valid, " as denominator."),
    gp = grid::gpar(fontsize = 8, fontface = "italic")
  )

  gridExtra::grid.arrange(
    title_grob, subtitle_grob, key_tbl,
    grid::textGrob(
      paste0(
        "INTERPRETING THESE METRICS\n\n",
        "\u2022 Significant GAM: The person\u2019s symptom shows a reliable pattern of change across the\n",
        "  menstrual cycle (p < .05). Higher % = more people show cyclical symptoms.\n\n",
        "\u2022 Range \u2265 1.5: The symptom swings by at least 1.5 points on the 1\u20136 DRSP scale across\n",
        "  the cycle \u2014 equivalent to the C-PASS 30% change criterion for PMDD. Higher % = more\n",
        "  people with large cyclical swings.\n\n",
        "\u2022 Affective vs. Any: Affective metrics focus on mood/emotional symptoms only (excluding\n",
        "  physical symptoms). This shows whether cyclicity is primarily mood-driven.\n\n",
        "\u2022 Denominator: All percentages use GAM-valid N (participants with enough data to model)\n",
        "  rather than total N, giving a more accurate picture of cyclicity rates."
      ),
      gp = grid::gpar(fontsize = 9), just = "left", x = 0.05
    ),
    note_grob,
    ncol = 1,
    heights = grid::unit(c(0.07, 0.04, 0.45, 0.37, 0.04), "npc")
  )
  draw_page_number()

  # ---- PAGE 3 (p.3): Per-Outcome Range Descriptives Table (color-coded) ----
  display_desc <- outcome_desc %>%
    dplyr::transmute(
      Outcome       = label,
      N             = N,
      `Median Range` = sprintf("%.2f", median_range),
      `Mean Range`  = sprintf("%.2f", mean_range),
      `SD Range`    = sprintf("%.2f", sd_range),
      `IQR`         = sprintf("%.2f-%.2f", q25_range, q75_range),
      `Min-Max`     = sprintf("%.2f-%.2f", min_range, max_range),
      `N Sig`       = n_sig,
      `% Sig`       = sprintf("%.1f%%", pct_sig),
      `N ≥1.5`  = n_range15,
      `% ≥1.5`  = sprintf("%.1f%%", pct_range15),
      `N Either`    = n_sig_or_range15,
      `% Either`    = sprintf("%.1f%%", pct_sig_or_range15)
    )

  # --- Conditional color-coding for key columns ---
  # Color thresholds and shading for table highlighting
  HIGHLIGHT_BLUE_DARK  <- "#BBDEFB"   # >= 50%
  HIGHLIGHT_BLUE_LIGHT <- "#E3F2FD"   # >= 25%
  HIGHLIGHT_ORG_DARK   <- "#FFE0B2"   # >= 50%
  HIGHLIGHT_ORG_MED    <- "#FFDDC1"   # median range >= 1.5
  HIGHLIGHT_ORG_LIGHT  <- "#FFF3E0"   # >= 25% or median range >= 1.0
  HIGHLIGHT_GRN_DARK   <- "#C8E6C9"   # >= 50%
  HIGHLIGHT_GRN_LIGHT  <- "#E8F5E9"   # >= 25%

  n_desc_rows <- nrow(display_desc)
  n_desc_cols <- ncol(display_desc)
  fill_matrix <- matrix("white", nrow = n_desc_rows, ncol = n_desc_cols)
  # Median Range (col 3): orange shading for high values
  fill_matrix[, 3] <- ifelse(outcome_desc$median_range >= 1.5, HIGHLIGHT_ORG_MED,
                              ifelse(outcome_desc$median_range >= 1.0, HIGHLIGHT_ORG_LIGHT, "white"))
  # % Sig (col 9): blue shading for high rates
  fill_matrix[, 9] <- ifelse(outcome_desc$pct_sig >= 50, HIGHLIGHT_BLUE_DARK,
                              ifelse(outcome_desc$pct_sig >= 25, HIGHLIGHT_BLUE_LIGHT, "white"))
  # % >=1.5 (col 11): orange shading for high rates
  fill_matrix[, 11] <- ifelse(outcome_desc$pct_range15 >= 50, HIGHLIGHT_ORG_DARK,
                               ifelse(outcome_desc$pct_range15 >= 25, HIGHLIGHT_ORG_LIGHT, "white"))
  # % Either (col 13): green shading for high rates
  fill_matrix[, 13] <- ifelse(outcome_desc$pct_sig_or_range15 >= 50, HIGHLIGHT_GRN_DARK,
                               ifelse(outcome_desc$pct_sig_or_range15 >= 25, HIGHLIGHT_GRN_LIGHT, "white"))

  desc_theme <- gridExtra::ttheme_default(
    base_size = 7,
    core    = list(fg_params = list(hjust = 0, x = 0.02),
                   bg_params = list(fill = fill_matrix)),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")),
    padding = grid::unit(c(2, 1.5), "mm")
  )

  title2 <- grid::textGrob(
    "Per-Outcome Range Descriptives (sorted by median range)",
    gp = grid::gpar(fontsize = 14, fontface = "bold")
  )
  note2 <- grid::textGrob(
    paste0("Range = max(predicted) \u2013 min(predicted) on original scale; Sig = p < .05; ",
           "% Sig and % \u22651.5 use GAM-valid N = ", n_gam_valid, " as denominator"),
    gp = grid::gpar(fontsize = 9, fontface = "italic")
  )
  footnote2 <- grid::textGrob(
    paste0(
      "COLUMN GUIDE: N = participants with valid GAM data for this outcome; ",
      "Median Range = typical amount of cyclical variation;\n",
      "% Sig = % with statistically significant cyclicity (p < .05); ",
      "% \u22651.5 = % with range meeting C-PASS benchmark (\u22651.5-point swing);\n",
      "% Either = % meeting either criterion. ",
      "Shading: blue = high % Sig; orange = high Median Range or % \u22651.5; green = high % Either."
    ),
    gp = grid::gpar(fontsize = 7, fontface = "italic")
  )
  desc_tbl <- gridExtra::tableGrob(display_desc, rows = NULL, theme = desc_theme)
  gridExtra::grid.arrange(
    title2, note2, desc_tbl, footnote2,
    ncol = 1,
    heights = grid::unit(c(0.06, 0.04, 0.78, 0.09), "npc")
  )
  draw_page_number()

  # ---- PAGE 3: Distribution of Ranges (histogram faceted by outcome) ----
  # Show range distributions for all outcomes, faceted
  range_plot_data <- id_outcome_summary %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(label = dplyr::coalesce(label, variable))

  p_range_dist <- ggplot2::ggplot(range_plot_data,
                                   ggplot2::aes(x = best_range, fill = sig)) +
    ggplot2::geom_histogram(bins = 20, alpha = 0.7, position = "identity") +
    ggplot2::geom_vline(xintercept = 1.5, color = "red", linetype = "dashed",
                        linewidth = 0.6) +
    ggplot2::facet_wrap(~ label, scales = "free_y", ncol = 4) +
    ggplot2::scale_fill_manual(
      values = c("TRUE" = "#0072B2", "FALSE" = "#999999"),
      labels = c("TRUE" = "Significant", "FALSE" = "Not Significant"),
      name = NULL
    ) +
    ggplot2::labs(
      title = "Distribution of GAM Ranges Across Participants",
      subtitle = "Red dashed line = range \u2265 1.5 threshold",
      x = "Range (max \u2013 min predicted)",
      y = "Count",
      caption = paste0(
        "Each panel shows one outcome. Blue bars = significant cyclicity (p < .05); grey = not significant.\n",
        "Outcomes where most bars fall right of the red line have widespread variation meeting the C-PASS benchmark."
      )
    ) +
    ggplot2::theme_minimal(base_size = 8) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 12),
      plot.caption = ggplot2::element_text(size = 7, hjust = 0, face = "italic"),
      strip.text = ggplot2::element_text(size = 6),
      legend.position = "bottom"
    )
  print(p_range_dist)
  draw_page_number()

  # ---- PAGE 4 (was 6): % Significant and % Range >= 1.5 by outcome (faceted bars) ----
  # Precompute ordering by % Range >= 1.5 before pivoting so all metrics
  # share a consistent outcome order (rather than ordering by the mean of both).
  pct_data <- outcome_desc %>%
    dplyr::mutate(label = factor(label, levels = label[order(pct_range15)])) %>%
    dplyr::select(label, pct_sig, pct_range15, pct_sig_or_range15) %>%
    tidyr::pivot_longer(cols = c(pct_sig, pct_range15, pct_sig_or_range15),
                        names_to = "metric", values_to = "pct") %>%
    dplyr::mutate(
      metric = dplyr::case_when(
        metric == "pct_sig"             ~ "% Significant (p < .05)",
        metric == "pct_range15"         ~ "% Range \u2265 1.5",
        metric == "pct_sig_or_range15"  ~ "% Sig OR Range \u2265 1.5",
        TRUE ~ metric
      ),
      metric = factor(metric, levels = c("% Significant (p < .05)",
                                          "% Range \u2265 1.5",
                                          "% Sig OR Range \u2265 1.5")),
      pct_label = sprintf("%.1f%%", pct)
    )

  p_pct_bars <- ggplot2::ggplot(pct_data,
                                 ggplot2::aes(x = label, y = pct, fill = metric)) +
    ggplot2::geom_col(alpha = 0.85, width = 0.7) +
    ggplot2::geom_text(
      ggplot2::aes(label = pct_label),
      hjust = -0.1, size = 2.5
    ) +
    ggplot2::coord_flip() +
    ggplot2::facet_wrap(~ metric, ncol = 3) +
    ggplot2::scale_y_continuous(limits = c(0, 105),
                                expand = ggplot2::expansion(mult = c(0, 0.05))) +
    ggplot2::scale_fill_manual(values = c("% Significant (p < .05)" = "#0072B2",
                                          "% Range \u2265 1.5" = "#D55E00",
                                          "% Sig OR Range \u2265 1.5" = "#009E73")) +
    ggplot2::labs(
      title = "Percentage of GAM-Valid Sample by Outcome",
      subtitle = paste0("Denominator: GAM-valid N = ", n_gam_valid),
      x = NULL,
      y = "Percentage",
      caption = paste0(
        "Blue = statistically significant cyclicity (p < .05); orange = range meeting C-PASS benchmark (\u22651.5);\n",
        "green = either criterion met. Outcomes ordered by % Range \u2265 1.5 (bottom to top)."
      )
    ) +
    ggplot2::theme_minimal(base_size = 9) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 13),
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      legend.position = "none",
      strip.text = ggplot2::element_text(face = "bold", size = 9),
      panel.spacing = grid::unit(1, "lines")
    )
  print(p_pct_bars)
  draw_page_number()

  # ---- PAGE 5 (was 8): Dot/Strip Plot with Box-and-Whisker ----
  p_dot_strip <- ggplot2::ggplot(
    range_plot_data,
    ggplot2::aes(x = reorder(label, best_range, FUN = median, na.rm = TRUE),
                 y = best_range)
  ) +
    ggplot2::geom_boxplot(width = 0.5, fill = "#E8E8E8", outlier.shape = NA, alpha = 0.7) +
    ggplot2::geom_jitter(
      ggplot2::aes(color = sig),
      width = 0.2, alpha = 0.5, size = 1.5
    ) +
    ggplot2::geom_hline(yintercept = 1.5, color = "red", linetype = "dashed", linewidth = 0.6) +
    ggplot2::coord_flip() +
    ggplot2::scale_color_manual(
      values = c("TRUE" = "#0072B2", "FALSE" = "#999999"),
      labels = c("TRUE" = "Significant", "FALSE" = "Not Significant"),
      name = NULL
    ) +
    ggplot2::labs(
      title = "Individual Participant Ranges by Outcome",
      subtitle = paste0(
        "Each dot = one participant; box shows median \u00b1 IQR; ",
        "red dashed line = range \u2265 1.5 threshold"
      ),
      caption = paste0(
        "This plot reveals the full distribution of cyclical variation for each outcome.\n",
        "Blue dots = significant cyclicity; grey dots = not significant.\n",
        "Outcomes are ordered by median range (bottom to top)."
      ),
      x = NULL,
      y = "Range (max \u2013 min predicted)"
    ) +
    ggplot2::theme_minimal(base_size = 9) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 13),
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      legend.position = "bottom"
    )
  print(p_dot_strip)
  draw_page_number()

  # ---- PAGE 6 (was 9): Significance/Range Heatmap (Outcome x Participant) ----
  heatmap_data <- id_outcome_summary %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(
      label = dplyr::coalesce(label, variable),
      category = dplyr::case_when(
        sig & is.finite(best_range) & best_range >= 1.5 ~ "Sig + Range \u2265 1.5",
        sig                                              ~ "Significant Only",
        is.finite(best_range) & best_range >= 1.5        ~ "Range \u2265 1.5 Only",
        TRUE                                              ~ "Neither"
      ),
      category = factor(category, levels = c("Neither", "Significant Only",
                                              "Range \u2265 1.5 Only", "Sig + Range \u2265 1.5"))
    )

  # Order outcomes: physical on bottom, affective on top, within each group
  # sorted by % meeting either criterion (ascending so top = highest %)
  outcome_order_df <- heatmap_data %>%
    dplyr::group_by(label, variable) %>%
    dplyr::summarise(pct_either = mean(category != "Neither", na.rm = TRUE), .groups = "drop") %>%
    dplyr::mutate(
      domain = dplyr::case_when(
        variable %in% all_affective_outcomes ~ "Affective",
        variable %in% all_physical_outcomes  ~ "Physical",
        TRUE ~ "Other"
      ),
      # Physical at bottom (1), then Affective on top (2)
      domain_rank = dplyr::case_when(
        domain == "Physical"  ~ 1L,
        domain == "Affective" ~ 2L,
        TRUE                  ~ 0L
      )
    ) %>%
    dplyr::arrange(domain_rank, pct_either)
  outcome_order <- outcome_order_df$label

  # Order participants by total number of outcomes meeting either criterion
  id_order <- heatmap_data %>%
    dplyr::group_by(id) %>%
    dplyr::summarise(n_either = sum(category != "Neither", na.rm = TRUE), .groups = "drop") %>%
    dplyr::arrange(n_either) %>%
    dplyr::pull(id)

  heatmap_data <- heatmap_data %>%
    dplyr::mutate(
      label = factor(label, levels = outcome_order),
      id = factor(id, levels = id_order)
    )

  p_heatmap <- ggplot2::ggplot(heatmap_data,
                                ggplot2::aes(x = id, y = label, fill = category)) +
    ggplot2::geom_tile(color = "white", linewidth = 0.2) +
    ggplot2::scale_fill_manual(
      values = c("Neither" = "#E0E0E0",
                 "Significant Only" = "#90CAF9",
                 "Range \u2265 1.5 Only" = "#FFCC80",
                 "Sig + Range \u2265 1.5" = "#E65100"),
      name = NULL
    ) +
    ggplot2::labs(
      title = "Cyclical Symptom Patterns: Outcome \u00d7 Participant Heatmap",
      subtitle = paste0("Participants (columns) ordered by total outcomes affected; ",
                        "outcomes (rows) grouped by domain: affective (top), physical (bottom)"),
      caption = paste0(
        "Each cell shows one participant\u2019s result for one outcome.\n",
        "Dark orange = both statistically significant AND range meeting C-PASS benchmark; ",
        "blue = significant only;\n",
        "light orange = range meeting C-PASS benchmark but not significant; grey = neither."
      ),
      x = "Participants",
      y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 9) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 13),
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      axis.text.x = ggplot2::element_blank(),
      axis.ticks.x = ggplot2::element_blank(),
      legend.position = "bottom",
      panel.grid = ggplot2::element_blank()
    )
  print(p_heatmap)
  draw_page_number()

  # ---- PAGE 7 (was 10): Affective vs. Physical Symptom Comparison ----

  comparison_data <- id_outcome_summary %>%
    dplyr::mutate(
      domain = dplyr::case_when(
        variable %in% all_affective_outcomes ~ "Affective",
        variable %in% all_physical_outcomes  ~ "Physical",
        TRUE ~ NA_character_
      )
    ) %>%
    dplyr::filter(!is.na(domain))

  # Per-domain summary (each row = one outcome x participant pair)
  domain_summary <- comparison_data %>%
    dplyr::group_by(domain) %>%
    dplyr::summarise(
      n_pairs      = dplyr::n(),
      mean_range   = mean(best_range, na.rm = TRUE),
      median_range = median(best_range, na.rm = TRUE),
      pct_sig      = mean(sig, na.rm = TRUE) * 100,
      pct_range15  = mean(is.finite(best_range) & best_range >= 1.5, na.rm = TRUE) * 100,
      pct_either   = mean(sig | (is.finite(best_range) & best_range >= 1.5), na.rm = TRUE) * 100,
      .groups      = "drop"
    )

  # Grouped bar chart for key percentage metrics
  domain_long <- domain_summary %>%
    tidyr::pivot_longer(
      cols = c(pct_sig, pct_range15, pct_either),
      names_to = "metric", values_to = "pct"
    ) %>%
    dplyr::mutate(
      metric = dplyr::case_when(
        metric == "pct_sig"     ~ "% Significant",
        metric == "pct_range15" ~ "% Range \u2265 1.5",
        metric == "pct_either"  ~ "% Either",
        TRUE ~ metric
      ),
      metric = factor(metric, levels = c("% Significant", "% Range \u2265 1.5", "% Either")),
      pct_label = sprintf("%.1f%%", pct)
    )

  p_domain <- ggplot2::ggplot(domain_long,
                               ggplot2::aes(x = metric, y = pct, fill = domain)) +
    ggplot2::geom_col(position = "dodge", alpha = 0.85, width = 0.6) +
    ggplot2::geom_text(
      ggplot2::aes(label = pct_label),
      position = ggplot2::position_dodge(width = 0.6),
      vjust = -0.3, size = 4
    ) +
    ggplot2::scale_fill_manual(
      values = c("Affective" = "#0072B2", "Physical" = "#D55E00"),
      name = NULL
    ) +
    ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
    ggplot2::labs(
      title = "Affective vs. Physical Symptom Cyclicity",
      subtitle = "Comparing cyclicity rates across outcome \u00d7 participant pairs",
      x = NULL,
      y = "Percentage"
    ) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 14),
      legend.position = "bottom"
    )

  # Caption grob for the layout
  domain_caption <- grid::textGrob(
    paste0(
      "Affective = mood/emotional symptoms and composites (Neg. Affect, Distress, Lability, Low Arousal). ",
      "Physical = somatic symptoms and composites (Breast Sx, Eating Sx).\n",
      "% of outcome \u00d7 participant pairs meeting each criterion. ",
      "Range \u2265 1.5 reference line corresponds to the C-PASS 30% change benchmark."
    ),
    gp = grid::gpar(fontsize = 9, fontface = "italic"),
    just = "left", x = 0.02
  )
  gridExtra::grid.arrange(
    p_domain,
    domain_caption,
    ncol = 1,
    heights = grid::unit(c(0.88, 0.12), "npc")
  )
  draw_page_number()

  # ---- PAGE 8 (was 11): Symptom Co-Occurrence Matrix ----
  # For each pair of outcomes, what % of GAM-valid participants meet
  # "sig OR range >= 1.5" in both outcomes?
  cooccur_data <- id_outcome_summary %>%
    dplyr::mutate(meets = sig | (is.finite(best_range) & best_range >= 1.5)) %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(label = dplyr::coalesce(label, variable))

  # Build binary matrix: rows = participants, cols = outcomes (TRUE/FALSE)
  meets_wide <- cooccur_data %>%
    dplyr::select(id, label, meets) %>%
    tidyr::pivot_wider(names_from = label, values_from = meets,
                       values_fill = FALSE, values_fn = list(meets = any))

  meets_mat <- as.matrix(meets_wide[, -1]) * 1L  # convert logical to integer
  rownames(meets_mat) <- meets_wide$id

  # Cross-product gives count of participants meeting criterion for both
  cooccur_counts <- crossprod(meets_mat)
  cooccur_pct <- cooccur_counts / n_gam_valid * 100

  # Reshape for ggplot
  cooccur_df <- as.data.frame(as.table(cooccur_pct))
  names(cooccur_df) <- c("Outcome1", "Outcome2", "pct_both")

  # Order outcomes by their individual % meeting criterion
  outcome_pct_order <- cooccur_data %>%
    dplyr::group_by(label) %>%
    dplyr::summarise(pct = mean(meets, na.rm = TRUE) * 100, .groups = "drop") %>%
    dplyr::arrange(pct) %>%
    dplyr::pull(label)

  cooccur_df <- cooccur_df %>%
    dplyr::mutate(
      Outcome1 = factor(Outcome1, levels = outcome_pct_order),
      Outcome2 = factor(Outcome2, levels = rev(outcome_pct_order))
    )

  p_cooccur <- ggplot2::ggplot(cooccur_df,
                                ggplot2::aes(x = Outcome1, y = Outcome2, fill = pct_both)) +
    ggplot2::geom_tile(color = "white", linewidth = 0.3) +
    ggplot2::geom_text(
      ggplot2::aes(label = ifelse(pct_both >= 1, sprintf("%.0f", pct_both), "")),
      size = 2.0
    ) +
    ggplot2::scale_fill_gradient2(
      low = "white", mid = "#BBDEFB", high = "#0D47A1",
      midpoint = 25, limits = c(0, max(cooccur_df$pct_both, na.rm = TRUE)),
      name = "% of Sample"
    ) +
    ggplot2::labs(
      title = "Symptom Co-Occurrence: % of Sample Meeting Criterion for Both Outcomes",
      subtitle = "Criterion: significant (p < .05) OR range \u2265 1.5",
      caption = paste0(
        "Each cell shows the % of GAM-valid participants (N = ", n_gam_valid,
        ") who meet the criterion for BOTH outcomes.\n",
        "Darker blue = more co-occurrence. Diagonal = % meeting criterion for that single outcome.\n",
        "High co-occurrence suggests symptoms that tend to be cyclical in the same people."
      ),
      x = NULL, y = NULL
    ) +
    ggplot2::theme_minimal(base_size = 7) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 12),
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1, size = 6),
      axis.text.y = ggplot2::element_text(size = 6),
      panel.grid = ggplot2::element_blank(),
      legend.position = "right"
    )
  print(p_cooccur)
  draw_page_number()

  grDevices::dev.off()
  message(sprintf("Wrote overall summary report: %s", overall_pdf))
} else {
  message("No significance data available; skipping overall summary report.")
}
```

---

# Section 6: Per outcome exploration (analysis functions)

```{r exploration-analysis-functions}
# =========================================================
# EXPLORATION / THRESHOLD ANALYSIS FUNCTIONS
# These power the per-outcome exploration PDFs below.
# =========================================================

# --- Basic cleaning: add numeric significance flag, coerce types ---
clean_gam_summary <- function(df) {
  df %>%
    dplyr::mutate(
      significant     = ifelse(p.value < 0.05, "Significant", "Not Significant"),
      significant_num = ifelse(significant == "Significant", 1, 0),
      significant_f   = factor(significant_num,
                               levels = c(0, 1),
                               labels = c("Not Significant", "Significant")),
      range    = as.numeric(range),
      n        = as.numeric(n),
      variable = as.character(variable),
      id       = as.character(id)
    ) %>%
    dplyr::filter(!is.na(range), !is.na(significant_num),
                  !is.na(variable), !is.na(id))
}

# --- Filter to one outcome variable, one row per ID ---
# Aggregates across potential multiple time_var rows per id/variable.
# peak_time, trough_time, and best_time_var come from the row with the
# smallest p.value (best significance), ensuring internal consistency.
get_outcome_df <- function(df, target_var) {
  df %>%
    dplyr::filter(variable == target_var) %>%
    dplyr::group_by(id, variable) %>%
    dplyr::summarise(
      p.value         = min(p.value, na.rm = TRUE),
      range           = max(range,   na.rm = TRUE),
      n               = max(n,       na.rm = TRUE),
      significant_num = ifelse(any(significant_num == 1, na.rm = TRUE), 1, 0),
      peak_time       = peak_time[which.min(p.value)][1],
      trough_time     = trough_time[which.min(p.value)][1],
      best_time_var   = time_var[which.min(p.value)][1],
      .groups         = "drop"
    ) %>%
    dplyr::mutate(
      significant_f = factor(
        significant_num,
        levels = c(0, 1),
        labels = c("Not Significant", "Significant")
      )
    )
}

# --- Descriptive summary (overall + by significance) ---
summarize_cyclicity <- function(df_outcome) {
  overall <- df_outcome %>%
    dplyr::summarise(
      n_participants   = dplyr::n(),
      n_significant    = sum(significant_num == 1, na.rm = TRUE),
      prop_significant = mean(significant_num == 1, na.rm = TRUE)
    )
  by_sig <- df_outcome %>%
    dplyr::group_by(significant_f) %>%
    dplyr::summarise(
      n            = dplyr::n(),
      mean_range   = mean(range, na.rm = TRUE),
      sd_range     = sd(range, na.rm = TRUE),
      median_range = median(range, na.rm = TRUE),
      q25_range    = quantile(range, .25, na.rm = TRUE),
      q75_range    = quantile(range, .75, na.rm = TRUE),
      min_range    = min(range, na.rm = TRUE),
      max_range    = max(range, na.rm = TRUE),
      mean_n       = mean(n, na.rm = TRUE),
      .groups      = "drop"
    )
  list(overall = overall, by_significance = by_sig)
}

# --- Proportion significant at various range thresholds ---
threshold_table <- function(df_outcome,
                            thresholds = c(0.10, 0.25, 0.36, 0.50, 0.75, 1.00)) {
  do.call(rbind, lapply(thresholds, function(th) {
    tmp <- df_outcome %>% dplyr::filter(range >= th)
    tibble::tibble(
      threshold        = th,
      n_at_or_above    = nrow(tmp),
      prop_of_sample   = ifelse(nrow(df_outcome) > 0,
                                nrow(tmp) / nrow(df_outcome), NA_real_),
      n_significant    = sum(tmp$significant_num == 1, na.rm = TRUE),
      prop_significant = ifelse(nrow(tmp) > 0,
                                mean(tmp$significant_num == 1, na.rm = TRUE),
                                NA_real_)
    )
  }))
}

# --- Logistic regression: significance ~ range (& + n) ---
fit_logistic_models <- function(df_outcome) {
  model_range   <- glm(significant_num ~ range,     family = binomial, data = df_outcome)
  model_range_n <- glm(significant_num ~ range + n, family = binomial, data = df_outcome)
  list(
    model_range   = model_range,
    model_range_n = model_range_n,
    tidy_range    = broom::tidy(model_range,   exponentiate = TRUE),
    tidy_range_n  = broom::tidy(model_range_n, exponentiate = TRUE)
  )
}

# --- Predicted probability of significance across range values ---
make_prediction_df <- function(df_outcome, model_range, model_range_n = NULL,
                               n_value = NULL) {
  xseq <- seq(min(df_outcome$range, na.rm = TRUE),
              max(df_outcome$range, na.rm = TRUE),
              length.out = 300)
  pred1 <- tibble::tibble(range = xseq) %>%
    dplyr::mutate(pred_prob = predict(model_range, newdata = ., type = "response"),
                  model     = "Range Only (no covariates)")
  if (!is.null(model_range_n)) {
    if (is.null(n_value)) n_value <- median(df_outcome$n, na.rm = TRUE)
    pred2 <- tibble::tibble(range = xseq, n = n_value) %>%
      dplyr::mutate(pred_prob = predict(model_range_n, newdata = ., type = "response"),
                    model     = paste0("Range + Obs. Count (n held at median = ",
                                       round(n_value, 0), ")")) %>%
      dplyr::select(range, pred_prob, model)
    dplyr::bind_rows(pred1, pred2)
  } else {
    pred1
  }
}

# --- Histogram of range, colored by significance ---
plot_range_histogram <- function(df_outcome) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, fill = significant_f)) +
    ggplot2::geom_histogram(position = "identity", alpha = 0.65, bins = 30) +
    ggplot2::labs(
      title = paste0("Histogram of ", unique(df_outcome$variable), " Range by Significance"),
      x = "Cyclical range", y = "Count", fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- Logistic curve with jittered observed points ---
plot_significance_probability <- function(df_outcome, pred_df,
                                          thresholds = c(0.25, 0.36, 0.50),
                                          jitter_height = 0.04) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, y = significant_num)) +
    ggplot2::geom_jitter(ggplot2::aes(color = significant_f),
                         width = 0, height = jitter_height,
                         alpha = 0.45, size = 2) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2, inherit.aes = FALSE) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(-0.02, 1.02), breaks = seq(0, 1, by = 0.1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title = paste0("Probability of Significant Cyclicity: ", unique(df_outcome$variable)),
      subtitle = paste0(
        "Each dot = one participant (jittered vertically). ",
        "Lines = logistic model-predicted probability of significance."
      ),
      x = "Cyclical range",
      y = "Predicted probability of significant GAM",
      color = NULL, linetype = NULL,
      caption = paste0(
        "'Range Only' = logistic model predicting significance from cyclical range alone.\n",
        "'Range + Obs. Count' = model also adjusting for number of observations ",
        "(held at its median for the prediction curve).\n",
        "Dashed vertical lines mark range thresholds."
      )
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      plot.subtitle = ggplot2::element_text(size = 10)
    )
}

# --- Binned empirical proportions ---
make_empirical_bins <- function(df_outcome, n_bins = 10) {
  df_outcome %>%
    dplyr::mutate(range_bin = dplyr::ntile(range, n_bins)) %>%
    dplyr::group_by(range_bin) %>%
    dplyr::summarise(
      mean_range       = mean(range, na.rm = TRUE),
      prop_significant = mean(significant_num, na.rm = TRUE),
      n                = dplyr::n(),
      .groups          = "drop"
    )
}

# --- Empirical bins + model overlay ---
plot_empirical_and_model <- function(df_outcome, pred_df, empirical_df,
                                     thresholds = c(0.25, 0.36, 0.50)) {
  ggplot2::ggplot() +
    ggplot2::geom_point(data = empirical_df,
                        ggplot2::aes(x = mean_range, y = prop_significant, size = n),
                        alpha = 0.8) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(0, 1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title    = paste0("Empirical and Modeled Probability: ", unique(df_outcome$variable)),
      subtitle = paste0(
        "Dots = observed proportion significant in each range bin (size = bin count).\n",
        "Lines = logistic model predictions. Dashed verticals = range thresholds."
      ),
      x        = "Cyclical range",
      y        = "Probability of significant GAM",
      size     = "Bin n",
      linetype = NULL,
      caption  = paste0(
        "This plot shows how the probability of achieving a significant GAM increases with cyclical range.\n",
        "Each dot represents a bin of participants grouped by similar range values; larger dots = more participants.\n",
        "The lines show logistic regression model predictions of significance probability as a function of range."
      )
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.caption = ggplot2::element_text(size = 8, hjust = 0, face = "italic"),
      plot.subtitle = ggplot2::element_text(size = 10)
    )
}

# --- All-in-one analysis wrapper for one outcome variable ---
analyze_one_outcome <- function(df, target_var,
                                thresholds = c(0.25, 0.36, 0.50),
                                n_bins = 10) {
  df_clean   <- clean_gam_summary(df)
  df_outcome <- get_outcome_df(df_clean, target_var)
  if (nrow(df_outcome) == 0) stop(paste("No rows found for variable:", target_var))

  desc      <- summarize_cyclicity(df_outcome)
  thresh_tab <- threshold_table(df_outcome, thresholds = thresholds)
  mods      <- fit_logistic_models(df_outcome)
  pred_df   <- make_prediction_df(df_outcome  = df_outcome,
                                  model_range   = mods$model_range,
                                  model_range_n = mods$model_range_n)
  emp_df    <- make_empirical_bins(df_outcome, n_bins = n_bins)

  p_hist <- plot_range_histogram(df_outcome)
  p_prob <- plot_significance_probability(df_outcome, pred_df, thresholds = thresholds)
  p_emp  <- plot_empirical_and_model(df_outcome, pred_df, emp_df, thresholds = thresholds)

  list(
    data                        = df_outcome,
    descriptives_overall        = desc$overall,
    descriptives_by_significance = desc$by_significance,
    threshold_table             = thresh_tab,
    model_range                 = mods$model_range,
    model_range_n               = mods$model_range_n,
    model_range_tidy_OR         = mods$tidy_range,
    model_range_n_tidy_OR       = mods$tidy_range_n,
    predictions                 = pred_df,
    empirical_bins              = emp_df,
    plot_histogram              = p_hist,
    plot_probability            = p_prob,
    plot_empirical              = p_emp
  )
}

# --- Extract DRSP variable names from the summary data ---
get_drsp_variables <- function(df) {
  df %>%
    clean_gam_summary() %>%
    dplyr::distinct(variable) %>%
    dplyr::pull(variable) %>%
    .[grepl("^drsp", ., ignore.case = TRUE)]
}

# --- Combined overlay GAM plot for a group of participants ---
# Overlays individual GAM fitted curves on a single plot.
# fit_results_list: named list (keyed by id) of fit_outcome() results
# ids: character vector of participant IDs to include
# tv: time variable name
# axis_meta_list: axis metadata (labels, reference lines)
# outcome_label: display label for the outcome
# title_prefix: e.g., "Significant GAMs" or "Non-Significant GAMs"
plot_overlay_gam <- function(fit_results_list, ids, tv, axis_meta_list,
                             outcome_label, title_prefix) {
  meta <- axis_meta_list[[tv]]

  overlay_data <- do.call(rbind, lapply(ids, function(qid) {
    res <- fit_results_list[[as.character(qid)]]
    if (is.null(res)) return(NULL)
    pred_df <- res$pred
    pred_df$id <- qid
    pred_df
  }))

  if (is.null(overlay_data) || nrow(overlay_data) == 0) return(NULL)

  n_people <- length(unique(overlay_data$id))

  # Compute y-axis limits: default 1–4.5, expand to 5 or 6 if predictions exceed 4.5
  estimate_vals <- overlay_data$estimate
  finite_estimate_vals <- estimate_vals[is.finite(estimate_vals)]
  max_pred <- if (length(finite_estimate_vals) > 0) max(finite_estimate_vals) else NA_real_
  ymax_final <- expand_ymax(4.5, max_pred)

  p <- ggplot2::ggplot(overlay_data,
                       ggplot2::aes(x = .data[[tv]], y = estimate, group = id))

  # Pink shading over expanded region (above 4.5) — rendered first
  if (ymax_final > 4.5) {
    p <- p + ggplot2::annotate(
      "rect",
      xmin = -Inf, xmax = Inf,
      ymin = 4.5,  ymax = ymax_final,
      fill = "pink", alpha = 0.25
    )
  }

  p <- p +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                        linewidth = 0.8) +
    ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                        linewidth = 0.8) +
    ggplot2::geom_line(alpha = 0.4, linewidth = 0.6, color = "#0072B2") +
    ggplot2::scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.5),
      labels = meta$labels
    ) +
    ggplot2::scale_y_continuous(breaks = seq(1, ceiling(ymax_final), by = 1)) +
    ggplot2::coord_cartesian(ylim = c(1, ymax_final)) +
    ggplot2::labs(
      title    = paste0(title_prefix, ": ", outcome_label),
      subtitle = paste0("N = ", n_people, " participants | Time variable: ", tv),
      x        = "Cycle Phase",
      y        = outcome_label
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(size = 16, face = "bold"),
      plot.subtitle = ggplot2::element_text(size = 12)
    )

  p
}

# --- Combined overlay GAM plot coloring two groups of participants differently ---
# group1_ids: IDs drawn in group1_color (e.g., significant participants)
# group2_ids: IDs drawn in group2_color (e.g., non-significant participants)
# group1_label / group2_label: legend labels for each group
plot_overlay_gam_twogroup <- function(fit_results_list, group1_ids, group2_ids,
                                      tv, axis_meta_list,
                                      outcome_label, title_prefix,
                                      group1_label = "Group 1",
                                      group2_label = "Group 2",
                                      group1_color = "#0072B2",
                                      group2_color = "#999999") {
  meta <- axis_meta_list[[tv]]

  build_group_data <- function(ids, grp) {
    do.call(rbind, lapply(ids, function(qid) {
      res <- fit_results_list[[as.character(qid)]]
      if (is.null(res)) return(NULL)
      pred_df <- res$pred
      pred_df$id    <- qid
      pred_df$group <- grp
      pred_df
    }))
  }

  d1 <- build_group_data(group1_ids, "group1")
  d2 <- build_group_data(group2_ids, "group2")
  overlay_data <- dplyr::bind_rows(d1, d2)

  if (is.null(overlay_data) || nrow(overlay_data) == 0) return(NULL)

  n1 <- length(unique(overlay_data$id[overlay_data$group == "group1"]))
  n2 <- length(unique(overlay_data$id[overlay_data$group == "group2"]))

  # Compute y-axis limits: default 1–4.5, expand to 5 or 6 if predictions exceed 4.5
  finite_ests <- overlay_data$estimate[is.finite(overlay_data$estimate)]
  max_pred <- if (length(finite_ests) > 0) max(finite_ests) else NA_real_
  ymax_final <- expand_ymax(4.5, max_pred)

  p <- ggplot2::ggplot(
    overlay_data,
    ggplot2::aes(x = .data[[tv]], y = estimate, group = id, color = group)
  )

  # Pink shading over expanded region (above 4.5) — rendered first
  if (ymax_final > 4.5) {
    p <- p + ggplot2::annotate(
      "rect",
      xmin = -Inf, xmax = Inf,
      ymin = 4.5,  ymax = ymax_final,
      fill = "pink", alpha = 0.25
    )
  }

  p <- p +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red",    linewidth = 0.8) +
    ggplot2::geom_vline(xintercept = meta$ov,     color = "forestgreen", linewidth = 0.8) +
    ggplot2::geom_line(alpha = 0.5, linewidth = 0.6) +
    ggplot2::scale_color_manual(
      values = c(group1 = group1_color, group2 = group2_color),
      labels = c(group1 = paste0(group1_label, " (N=", n1, ")"),
                 group2 = paste0(group2_label, " (N=", n2, ")"))
    ) +
    ggplot2::scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.5),
      labels = meta$labels
    ) +
    ggplot2::scale_y_continuous(breaks = seq(1, ceiling(ymax_final), by = 1)) +
    ggplot2::coord_cartesian(ylim = c(1, ymax_final)) +
    ggplot2::labs(
      title    = paste0(title_prefix, ": ", outcome_label),
      subtitle = paste0("Time variable: ", tv, "  |  N total = ", n1 + n2),
      x        = "Cycle Phase",
      y        = outcome_label,
      color    = NULL
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(size = 16, face = "bold"),
      plot.subtitle = ggplot2::element_text(size = 12),
      legend.position = "bottom"
    )

  p
}

# --- Peak and trough timing distribution plot for qualifying participants ---
plot_peak_trough_timing <- function(df_sig, tv, axis_meta_list, outcome_label) {
  meta <- axis_meta_list[[tv]]

  tl_peak   <- df_sig %>%
    dplyr::filter(!is.na(peak_time)) %>%
    dplyr::transmute(time = peak_time, type = "Peak (Max)")
  tl_trough <- df_sig %>%
    dplyr::filter(!is.na(trough_time)) %>%
    dplyr::transmute(time = trough_time, type = "Trough (Min)")

  if (nrow(tl_peak) < 2 || nrow(tl_trough) < 2) return(NULL)

  tl_all <- dplyr::bind_rows(tl_peak, tl_trough)

  ggplot2::ggplot() +
    ggplot2::geom_density(data = tl_peak,
                          ggplot2::aes(x = time,
                                       y = ggplot2::after_stat(ndensity),
                                       fill = "Peak (Max)", color = "Peak (Max)"),
                          alpha = 0.3, adjust = 1.2) +
    ggplot2::geom_density(data = tl_trough,
                          ggplot2::aes(x = time,
                                       y = -ggplot2::after_stat(ndensity),
                                       fill = "Trough (Min)", color = "Trough (Min)"),
                          alpha = 0.3, adjust = 1.2) +
    ggplot2::geom_rug(data = tl_all, ggplot2::aes(x = time, color = type),
                      alpha = 0.5, length = grid::unit(0.03, "npc")) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.5, color = "grey40") +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::scale_y_continuous(limits = c(-1, 1),
                                breaks = seq(-1, 1, by = 0.25),
                                labels = abs) +
    ggplot2::labs(
      title = paste("Peak & Trough Timing:", outcome_label),
      subtitle = paste0("Among participants with significant GAMs (p < .05) ",
                        "or range \u2265 1.5; N = ",
                        nrow(df_sig)),
      x = "Cycle Phase",
      y = "Peaks \u2191 / Troughs \u2193 (normalized density)",
      fill = NULL, color = NULL
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::scale_fill_manual(values = c("Peak (Max)" = "#D55E00",
                                          "Trough (Min)" = "#0072B2")) +
    ggplot2::scale_color_manual(values = c("Peak (Max)" = "#D55E00",
                                           "Trough (Min)" = "#0072B2"))
}
```

# Section 7: Per outcome reports (exploration PDF loop)

```{r generate-explore-pdfs, results='asis'}
# =========================================================
# GENERATE PER-OUTCOME EXPLORATION PDF REPORTS
# =========================================================
# For each DRSP outcome, generates a landscape PDF containing:
#   1. Summary page: descriptive stats + threshold table + peak/trough timing
#   2. All GAM curves colored by significance
#   3. All GAM curves colored by sig. or range >= 1.5
#   4. Significance probability plot (logistic curve)
#   5. Histogram of range by significance
#   6. Empirical + modeled probability plot
#   Per time variable:
#   7. Peak & trough timing density (significant or range >= 1.5 participants)
#   8+. Individual GAM trajectory pages (all IDs, sorted by descending range)
# =========================================================

if (length(all_ids_signif) > 0) {
  gam_summary_df <- dplyr::bind_rows(all_ids_signif)

  # Create output sub-folder for exploration PDFs
  explore_folder <- file.path(output_folder, "Outcome_Exploration_PDFs")
  dir.create(explore_folder, showWarnings = FALSE, recursive = TRUE)

  drsp_vars <- get_drsp_variables(gam_summary_df)
  all_outcome_vars <- drsp_vars
  cat("Found", length(drsp_vars), "DRSP variables to process\n")

  # Thresholds for the exploration
  explore_thresholds <- c(0.10, 0.20, 0.25, 0.36, 0.50, 0.75, 1.00)

  # Plot layout for individual GAM pages
  plots_per_col_exp <- 3
  plots_per_row_exp <- 3
  max_plots_per_page_exp <- plots_per_col_exp * plots_per_row_exp

  # Table themes for summary pages
  summary_table_theme <- gridExtra::ttheme_default(
    base_size = 11,
    core    = list(fg_params = list(hjust = 0, x = 0.05)),
    colhead = list(fg_params = list(hjust = 0, x = 0.05, fontface = "bold"))
  )
  threshold_table_theme <- gridExtra::ttheme_default(
    base_size = 10,
    core    = list(fg_params = list(hjust = 0.5, x = 0.5)),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
  )

  for (target_var in all_outcome_vars) {
    cat(sprintf("\n=== Processing: %s ===\n", target_var))

    # 1. Run exploration analysis
    result <- tryCatch(
      analyze_one_outcome(gam_summary_df, target_var,
                          thresholds = explore_thresholds,
                          n_bins = 10),
      error = function(e) {
        message("  Skipping ", target_var, ": ", e$message)
        NULL
      }
    )
    if (is.null(result)) next

    # 2. Identify all IDs sorted by range descending, plus sig/nonsig splits for overlay pages
    sig_ids <- result$data %>%
      dplyr::filter(significant_num == 1) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    nonsig_ids <- result$data %>%
      dplyr::filter(significant_num == 0) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    # IDs where GAM is significant OR range >= 1.5 (for two-group colored overlay)
    sig_or_range_ids <- result$data %>%
      dplyr::filter(significant_num == 1 | (is.finite(range) & range >= 1.5)) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    neither_ids <- result$data %>%
      dplyr::filter(significant_num == 0 & !(is.finite(range) & range >= 1.5)) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    qualifying_ids <- result$data %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    # 3. Get outcome label
    outcome_label <- if (target_var %in% names(outcomes)) {
      outcomes[[target_var]]
    } else {
      target_var
    }

    cat(sprintf("  N participants: %d | N significant: %d | N qualifying IDs: %d\n",
                result$descriptives_overall$n_participants,
                result$descriptives_overall$n_significant,
                length(qualifying_ids)))

    # 4. Pre-compute GAM fit results for all qualifying IDs (all time variables)
    all_fit_results_by_tv <- list()
    for (tv in time_vars) {
      tv_fits <- list()
      for (qid in qualifying_ids) {
        dat_id <- subset(df, as.character(id) == qid)
        if (nrow(dat_id) == 0) next
        res <- tryCatch(
          fit_outcome(dat_id, target_var, tv, outcomes),
          error = function(e) NULL
        )
        if (!is.null(res)) tv_fits[[as.character(qid)]] <- res
      }
      all_fit_results_by_tv[[tv]] <- tv_fits
    }

    # 5. Open PDF
    pdf_path <- file.path(explore_folder, paste0("Explore_", target_var, ".pdf"))
    grDevices::pdf(pdf_path, width = 11, height = 8.5)

    # ---- PAGE 1: Summary + Threshold Table ----
    title_grob <- grid::textGrob(
      paste("GAM Cyclicity Exploration:", outcome_label),
      gp = grid::gpar(fontsize = 18, fontface = "bold")
    )
    subtitle_grob <- grid::textGrob(
      paste("Variable:", target_var),
      gp = grid::gpar(fontsize = 12, fontface = "italic")
    )

    n_cpass <- sum(result$data$significant_num == 1 &
                    is.finite(result$data$range) &
                    result$data$range >= 1.5, na.rm = TRUE)

    n_total <- result$descriptives_overall$n_participants
    pct_cpass <- if (n_total > 0) n_cpass / n_total * 100 else NA_real_

    desc_df <- data.frame(
      Metric = c("Total Participants",
                 "N Significant (p < .05)",
                 "Proportion Significant",
                 "N Range \u22651.5 (p < .05 & range \u2265 1.5)",
                 "% Range \u22651.5 (of total sample)",
                 "Median Range",
                 "Mean Observations per ID"),
      Value = c(
        as.character(n_total),
        as.character(result$descriptives_overall$n_significant),
        sprintf("%.1f%%", result$descriptives_overall$prop_significant * 100),
        as.character(n_cpass),
        ifelse(is.na(pct_cpass), "N/A", sprintf("%.1f%%", pct_cpass)),
        sprintf("%.2f", median(result$data$range, na.rm = TRUE)),
        sprintf("%.1f", mean(result$data$n, na.rm = TRUE))
      ),
      stringsAsFactors = FALSE
    )

    sig_peak_data <- result$data %>%
      dplyr::filter(
        (significant_num == 1) | (is.finite(range) & range >= 1.5),
        !is.na(peak_time), !is.na(trough_time)
      )

    if (nrow(sig_peak_data) > 0) {
      timing_by_tv <- sig_peak_data %>%
        dplyr::group_by(best_time_var) %>%
        dplyr::summarise(
          med_peak   = median(peak_time, na.rm = TRUE),
          med_trough = median(trough_time, na.rm = TRUE),
          n_qual     = dplyr::n(),
          .groups    = "drop"
        )
      for (i in seq_len(nrow(timing_by_tv))) {
        tv_label <- timing_by_tv$best_time_var[i]
        suffix <- if (nrow(timing_by_tv) > 1) paste0(" [", tv_label, "]") else ""
        desc_df <- rbind(desc_df, data.frame(
          Metric = c(paste0("Median Peak Time (sig. or range\u22651.5)", suffix),
                      paste0("Median Trough Time (sig. or range\u22651.5)", suffix)),
          Value = c(
            sprintf("%.2f (N=%d)", timing_by_tv$med_peak[i], timing_by_tv$n_qual[i]),
            sprintf("%.2f (N=%d)", timing_by_tv$med_trough[i], timing_by_tv$n_qual[i])
          ),
          stringsAsFactors = FALSE
        ))
      }
    }

    desc_tbl <- gridExtra::tableGrob(desc_df, rows = NULL, theme = summary_table_theme)

    thresh_display <- result$threshold_table %>%
      dplyr::mutate(
        threshold        = sprintf("%.2f", threshold),
        n_at_or_above    = as.character(n_at_or_above),
        prop_of_sample   = sprintf("%.1f%%", prop_of_sample * 100),
        n_significant    = as.character(n_significant),
        prop_significant = ifelse(is.na(prop_significant), "N/A",
                                  sprintf("%.1f%%", prop_significant * 100))
      )
    names(thresh_display) <- c("Range Threshold", "N >= Threshold",
                               "% of Sample", "N Significant", "% Significant")

    thresh_tbl <- gridExtra::tableGrob(thresh_display, rows = NULL,
                                       theme = threshold_table_theme)
    thresh_header <- grid::textGrob(
      "Significance Rates at Range Thresholds",
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    note_grob <- if (length(qualifying_ids) > 0) {
      grid::textGrob(
        sprintf("Individual GAM pages follow for %d IDs (sorted by descending range)",
                length(qualifying_ids)),
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    } else {
      grid::textGrob(
        "No qualifying IDs for individual GAM pages",
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    }

    gridExtra::grid.arrange(
      title_grob, subtitle_grob, desc_tbl, thresh_header, thresh_tbl, note_grob,
      ncol = 1,
      heights = grid::unit(c(0.10, 0.05, 0.25, 0.08, 0.38, 0.08), "npc")
    )

    # ---- PAGE 2: All curves — significant (vermillion) vs. non-significant (sky blue) ----
    for (tv in time_vars) {
      p_twocol_sig <- plot_overlay_gam_twogroup(
        fit_results_list = all_fit_results_by_tv[[tv]],
        group1_ids       = sig_ids,
        group2_ids       = nonsig_ids,
        tv               = tv,
        axis_meta_list   = axis_meta,
        outcome_label    = outcome_label,
        title_prefix     = "All GAM Curves: Significant vs. Non-Significant",
        group1_label     = "Significant (p < .05)",
        group2_label     = "Non-Significant (p \u2265 .05)",
        group1_color     = "#D55E00",
        group2_color     = "#56B4E9"
      )
      if (!is.null(p_twocol_sig)) print(p_twocol_sig)
    }

    # ---- PAGE 3: All curves — significant OR range >= 1.5 (reddish purple) vs. neither (bluish green) ----
    for (tv in time_vars) {
      p_twocol_range <- plot_overlay_gam_twogroup(
        fit_results_list = all_fit_results_by_tv[[tv]],
        group1_ids       = sig_or_range_ids,
        group2_ids       = neither_ids,
        tv               = tv,
        axis_meta_list   = axis_meta,
        outcome_label    = outcome_label,
        title_prefix     = "All GAM Curves: Sig. or Range \u22651.5 vs. Neither",
        group1_label     = "Sig. or Range \u22651.5",
        group2_label     = "Neither",
        group1_color     = "#CC79A7",
        group2_color     = "#009E73"
      )
      if (!is.null(p_twocol_range)) print(p_twocol_range)
    }

    # ---- PAGE 4: Significance Probability Plot ----
    print(result$plot_probability)

    # ---- PAGE 5: Histogram ----
    print(result$plot_histogram)

    # ---- PAGE 6: Empirical + Model ----
    print(result$plot_empirical)

    # ---- PAGE 7+: Peak & Trough Timing Summary (per time variable) ----
    for (tv in time_vars) {
      timing_data <- result$data %>%
        dplyr::filter(
          (significant_num == 1) | (is.finite(range) & range >= 1.5),
          best_time_var == tv,
          !is.na(peak_time), !is.na(trough_time)
        )

      if (nrow(timing_data) > 0) {
        p_timing <- plot_peak_trough_timing(timing_data, tv, axis_meta,
                                            outcome_label)
        if (!is.null(p_timing)) print(p_timing)
      }
    }

    # ---- Individual GAM trajectory plots for all IDs (per time variable) ----
    # All IDs sorted by descending range
    for (tv in time_vars) {
      if (length(qualifying_ids) == 0) next

      all_fit_results <- all_fit_results_by_tv[[tv]]

      id_plots <- list()
      for (qid in qualifying_ids) {
        res <- all_fit_results[[as.character(qid)]]
        if (is.null(res)) next

        datSX_qid <- subset(df, as.character(id) == qid)
        # For outcome-specific reports, compute y-axis from this outcome's
        # data only — do NOT use the per-ID global ymax from person reports.
        # Only expand above 4.5 if this outcome's predictions or raw data
        # for this participant actually exceed 4.5.
        outcome_max <- 4.5
        if (!is.null(res$pred$estimate)) {
          pred_estimates <- res$pred$estimate[is.finite(res$pred$estimate)]
          if (length(pred_estimates) > 0) {
            outcome_max <- max(outcome_max, max(pred_estimates))
          }
        }
        # Also check raw data points for this specific outcome
        if (nrow(datSX_qid) > 0 && target_var %in% names(datSX_qid)) {
          raw_vals <- datSX_qid[[target_var]]
          raw_vals <- raw_vals[is.finite(raw_vals)]
          if (length(raw_vals) > 0) {
            outcome_max <- max(outcome_max, max(raw_vals))
          }
        }
        qid_ymax <- expand_ymax(4.5, outcome_max)
        p <- create_annotated_plot(res$pred, tv, axis_meta,
                                   y_limits = c(1, qid_ymax), raw_data = datSX_qid)
        if (!is.null(p)) {
          id_range <- round(res$sig$range, 2)
          id_n     <- res$sig$n
          id_pval  <- res$sig$p.value
          id_stars <- dplyr::case_when(
            is.finite(id_pval) & id_pval < 0.001 ~ "***",
            is.finite(id_pval) & id_pval < 0.01  ~ "**",
            is.finite(id_pval) & id_pval < 0.05  ~ "*",
            TRUE ~ ""
          )
          p <- p + ggplot2::labs(
            title = sprintf("ID %s %s (range=%.2f, N=%d)", qid, id_stars, id_range, id_n)
          )
          # Red title when either positional half has fewer than 5 observations
          if (isTRUE(res$sig$n_neg < 5) || isTRUE(res$sig$n_nonneg < 5)) {
            p <- p + ggplot2::theme(
              plot.title = ggplot2::element_text(size = 10, face = "bold",
                                                 hjust = 0.5, color = "red")
            )
          }
          id_plots[[length(id_plots) + 1]] <- p
        }
      }

      if (length(id_plots) > 0) {
        pages <- split(id_plots,
                       ceiling(seq_along(id_plots) / max_plots_per_page_exp))

        for (page_plots in pages) {
          if (length(page_plots) < max_plots_per_page_exp) {
            page_plots <- c(page_plots,
                           rep(list(patchwork::plot_spacer()),
                               max_plots_per_page_exp - length(page_plots)))
          }
          final_page <- patchwork::wrap_plots(plotlist = page_plots,
                                              ncol = plots_per_col_exp) +
            patchwork::plot_annotation(
              title    = paste("Individual GAMs:", outcome_label),
              subtitle = paste("Sorted by descending range | Time variable:", tv),
              theme    = ggplot2::theme(
                plot.title    = ggplot2::element_text(size = 12, face = "bold"),
                plot.subtitle = ggplot2::element_text(size = 9)
              )
            )
          print(final_page)
        }
      }
    }

    grDevices::dev.off()
    message(sprintf("  Saved: %s", pdf_path))
  }

  cat("\n=== All exploration PDFs generated ===\n")
  cat("Output folder:", explore_folder, "\n")
} else {
  message("No significance data available; skipping per-outcome exploration PDFs.")
}
```
