title: "UKALC_07_smms_binary"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
    self_contained: true
---

# =============================================================================
# SMOOTHING MIXTURE MODELS (SMMs) - BINARY OUTCOMES ONLY
# =============================================================================
# This script is a specialized version of UKALC_06_smms.Rmd focused exclusively
# on binary (0/1) drinking outcomes using Bernoulli distribution with logit link.
#
# WHAT IT DOES:
# 1. Sources setup file and loads preprocessed data
# 2. Configures SMM analyses for binary alcohol outcomes (drink_today, fourplustoday, felt_drunk)
# 3. Runs run_smm_bernoulli() for each outcome/centering/g combination
# 4. Generates probability-scale trajectory plots and classification results
# 5. Compares models via BIC to identify optimal group numbers
# 6. Saves all outputs (models, plots, classifications, BIC tables)
#
# DIFFERENCE FROM MAIN SMM SCRIPT:
# - Uses ONLY run_smm_bernoulli.R (simplified Bernoulli-only implementation)
# - Main SMM script (UKALC_06_smms.Rmd) uses run_smm_cyclic.R (handles both types)
# - This script kept for legacy compatibility and focused binary analysis
#
# OUTPUT STRUCTURE:
# - Base directory: output_folder/04b_smms_binary/SMMs_Binary_Only/
# - Organized by: date (YYYYMMDD) → centering (menses/ovulation) → outcome/
# - Per outcome/g: classification CSVs, trajectory plots (mean/roll/centered), BIC
#
# KEY FEATURES:
# - Bernoulli distribution with logit link for binary outcomes
# - Random intercepts only (no random slopes due to convergence issues with sparse data)
# - Predictions on probability scale (0-1) for interpretability
# - EM algorithm with RSS-based group assignment using Pearson residuals
#
# BINARY OUTCOMES ANALYZED:
# - drink_today: Any alcohol consumption (0=no, 1=yes)
# - fourplustoday: Heavy drinking episode (0=<4 drinks, 1=4+ drinks)
# - felt_drunk: Subjective intoxication (0=no, 1=yes)
#
# SEE ALSO:
# - scripts/run_smm_bernoulli.R for Bernoulli-specific implementation
# - UKALC_06_smms.Rmd for unified continuous+binary SMM analysis
# - docs/BINARY_OUTCOME_TESTING_GUIDE.md for testing procedures
# =============================================================================

```{r setup-binary-smm}
# Source R File with Setup, Package Loading, and Output Folder Creation
# 
# This chunk loads the UKALC_00_setup.R script which:
# - Loads all required packages (tidyverse, lme4, mgcv, gamm4, etc.)
# - Sets global options (no scientific notation, 3 decimal places)
# - Creates date-stamped output folder for all analysis products
# - Sources graphics configuration for consistent plotting
#
# Then creates a subfolder specifically for binary SMM outputs

source("UKALC_00_setup.R")

# Create subfolder for Binary SMM outputs (keep output_folder as root-level directory)
smms_binary_folder <- file.path(output_folder, "04b_smms_binary")
if (!dir.exists(smms_binary_folder)) {
  dir.create(smms_binary_folder, recursive = TRUE)
  cat("Created Binary SMMs subfolder:", smms_binary_folder, "\n")
}

# Note: Do not overwrite output_folder; use smms_binary_folder explicitly for Binary SMM-specific files

```


# --- 1. SETUP & PARAMETERS ---

```{r}
## WHAT THIS SECTION DOES:
# Loads required libraries and sets random seed for reproducible analyses.
#
## WHY IT'S NECESSARY:
# - Libraries provide statistical modeling (gamm4, mgcv), data manipulation (dplyr, tidyr),
#   and visualization (ggplot2) functionality
# - Random seed ensures k-means initialization produces consistent group assignments
#   across analysis runs
#
## IMPORTANT NOTES:
# - Most packages already loaded by UKALC_00_setup.R
# - slider package: Used for rolling averages (though not needed for binary outcomes)
# - marginaleffects: Used for model predictions and contrasts

# Load required libraries (most already loaded via UKALC_00_setup.R)
library(dplyr)
library(mgcv)
library(gamm4)
library(ggplot2)
library(glue)
library(tidyr)
library(purrr)
library(lubridate)
library(zoo)
library(marginaleffects)
  
# Set random seed for reproducibility
set.seed(123)

```


# Load the data file
```{r}
## WHAT THIS SECTION DOES:
# Loads the most recent cleaned/preprocessed UKALC dataset from the data/ folder.
#
## DATA SOURCE:
# - Pattern: data/ukalc_YYYYMMDD*.rds
# - Files are date-stamped (YYYYMMDD format sorts chronologically)
# - Latest file = most recent data processing run
#
## PROCESSING:
# - Lists all files matching pattern
# - Sorts in descending order (newest first)
# - Loads first (most recent) file
#
## OUTPUT:
# - df: Main analysis dataframe with all participants and variables

# LOAD LATEST CLEAN DATA FILE 

# Get all files matching the pattern
files <- list.files(path = "data", 
                   pattern = "^analysis_ready_\\d{8}.*\\.rds$", 
                   full.names = TRUE)

# Sort files (dates in YYYYMMDD format will sort correctly alphabetically)
latest_file <- sort(files, decreasing = TRUE)[1]

# Load the dataset
df <- readRDS(latest_file)

# Print which file was loaded
cat("Loaded dataset from:", latest_file, "\n")


# Optional: Create sensitivity analysis subset (currently commented out)
# Uncomment to restrict analyses to participants with confirmed ovulatory cycles
# create df_sens that selects only those for whom Ov_Status = "Ovulatory"
#df_sens <- df %>%
  #filter(Ov_Status == "Ovulatory")

#df <- df_sens
```

# Define the base folder for all outputs
```{r}
## WHAT THIS SECTION DOES:
# Creates the output folder for all Binary SMM analysis products.
#
## FOLDER STRUCTURE:
# - Base: smms_binary_folder (created in setup chunk)
# - Subfolder: SMMs_Binary_Only
# - Date-stamped subfolders created automatically by run_smm_bernoulli()
#
## WHY IT'S NECESSARY:
# - Organizes outputs separate from continuous/mixed SMM analyses
# - Prevents overwriting of other analysis products
# - Facilitates comparison between binary-only and full SMM approaches

smm_output_folder <- file.path(smms_binary_folder, "SMMs_OvandBord_Binary_Only")
if (!dir.exists(smm_output_folder)) {
  dir.create(smm_output_folder, recursive = TRUE)
}
```

# Choose your time variables
```{r}
## WHAT THIS SECTION DOES:
# Defines cyclic time variables for menses-centered and ovulation-centered analyses.
#
## TIME VARIABLES:
# - cyclic_time_impute: Menses-centered time (-1 to +1, where 0 = menses onset)
# - cyclic_time_imp_ov: Ovulation-centered time (-1 to +1, where 0 = ovulation)
#
## WHY IT'S NECESSARY:
# - Different outcomes show clearer patterns with different centering
# - Allows comparison of trajectory patterns relative to both cycle events
# - Variables created in UKALC_01_data_prep.Rmd with forward-backward fill imputation

menses_time_variable <- "cyclic_time_impute"
ovulation_time_variable <- "cyclic_time_imp_ov"
```


# Define binary outcomes
```{r}
## WHAT THIS SECTION DOES:
# Defines the complete list of BINARY outcomes to analyze with SMMs.
#
## FORMAT:
# - Names: Technical variable names in the dataframe
# - Values: Human-readable labels for plot titles
#
## IMPORTANT NOTES:
# - All outcomes MUST be binary (0/1) variables
# - run_smm_bernoulli() validates binary format and will error if not 0/1
# - These are the only outcomes analyzed in this script (no continuous variables)
#
## BINARY DRINKING OUTCOMES:
# - drink_today_bin: Any alcohol consumption (yes/no)
# - fourplustoday_bin: Binge drinking, defined as 4+ drinks in one day
# - felt_drunk_today_bin: Self-reported subjective intoxication

binary_outcomes <- c(
  # --- Binary Substance Use Outcomes ---
  "drink_today_bin"      = "Any Drinking Today",
  "fourplustoday_bin"    = "Binge Drinking (4+)",
  "felt_drunk_today_bin" = "Felt Drunk"
)

```



# --- 2. AUTOMATED ANALYSIS LOOP ---

# Source helper functions
```{r}
## WHAT THIS SECTION DOES:
# Loads helper functions for SMM analysis pipeline.
#
## FUNCTIONS LOADED:
# - preprocess_outcome.R: Log transformation and person-centering (NOT USED for binary)
# - subset_by_phase_cyclic.R: Filters to participants with sufficient observations per phase
# - run_smm_bernoulli.R: Binary-only SMM fitting with Bernoulli distribution
# - model_plot_modx_gam_cyclic.R: Generates GAM plots with moderator effects
# - compare_bic_cyclic.R: Creates BIC comparison plots across group sizes
#
## WHY IT'S NECESSARY:
# - Modular functions enable reuse across different outcome types
# - Centralized functions ensure consistent analysis approach
# - Easier debugging and maintenance than inline code

# Source function files
source("scripts/preprocess_outcome.R")
source("scripts/subset_by_phase_cyclic.R")
source("scripts/run_smm_bernoulli.R")
source("scripts/model_plot_modx_gam_cyclic.R")
source("scripts/compare_bic_cyclic.R")

```

```{r}
## WHAT THIS SECTION DOES:
# Runs Smoothing Mixture Models (SMMs) for all binary outcomes with both
# menses-centered and ovulation-centered time variables.
#
## ANALYSIS PIPELINE:
# For each outcome × centering combination:
# 1. Preprocess outcome (though binary outcomes skip log/centering transformations)
# 2. Subset to participants with ≥4 observations per phase (sufficient data)
# 3. Fit SMM models with g=2, 3, and 4 groups
# 4. Generate BIC comparison plots
# 5. Generate GAM plots showing group trajectories
# 6. Save all results (classifications, plots, BIC tables)
#
## COVARIATES:
# - day_of_week: Fixed effect for day-of-week patterns (Friday = reference level)
# - session_today: Fixed effect for lab session presence (0/1)
#
## ERROR HANDLING:
# - Skips outcomes with insufficient data (<15 total observations after subsetting)
# - Singular matrix errors handled gracefully within run_smm_bernoulli()
# - Convergence issues logged but do not halt pipeline
#
## OUTPUTS:
# - all_smm_results: List of SMM model objects and classifications
# - all_bic_results: List of BIC comparison data
# - all_gam_results: List of GAM smoothing results

# --- AUTOMATED ANALYSIS LOOP ---
analysis_combinations <- expand.grid(
  outcome = names(binary_outcomes),
  centering = c("menses", "ovulation"),
  stringsAsFactors = FALSE
)

all_smm_results <- list()
all_bic_results <- list()
all_gam_results <- list()

# Set reference levels for covariates (before analysis loop)
# Check if covariates exist in the data
if ("day_of_week" %in% names(df)) {
  # Set Friday as reference for day_of_week
  df$day_of_week <- relevel(factor(df$day_of_week), ref = "Friday")
} else {
  warning("Column 'day_of_week' not found in data. Covariates will not be used.")
}

if ("session_today" %in% names(df)) {
  # Ensure session_today is numeric 0/1 (0 is automatically the reference)
  df$session_today <- as.numeric(df$session_today)
} else {
  warning("Column 'session_today' not found in data. Covariates will not be used.")
}

for (i in 1:nrow(analysis_combinations)) {
  current_outcome <- analysis_combinations$outcome[i]
  current_centering <- analysis_combinations$centering[i]
  current_plot_label <- binary_outcomes[current_outcome]
  current_time_var <- ifelse(current_centering == "menses", menses_time_variable, ovulation_time_variable)

  message(paste0("\n🚀 Starting analysis for: '", current_plot_label, "' (", current_centering, "-centered)"))

  # For binary outcomes, preprocess_outcome is not strictly needed (no log transform),
  # but we call it for consistency with the pipeline structure
  preprocessed_data <- preprocess_outcome(data = df, outcome = current_outcome)

  subset_result <- subset_by_phase_cyclic(
    data = preprocessed_data,
    outcome = current_outcome,
    time_var = current_time_var,
    min_obs = 4  # Require at least 4 observations per phase
  )
  data_for_analysis <- subset_result$data_subset

  if (nrow(data_for_analysis) < 10) { # Require at least 10 total observations to proceed
      message(paste0("⚠️ Skipping '", current_plot_label, "' due to insufficient data after subsetting."))
      next
  }

  smm_results <- run_smm_bernoulli(
    data = data_for_analysis,
    outcome = current_outcome,
    time_var = current_time_var,
    plot_label = current_plot_label,
    g = 2:4,
    centering = current_centering,
    save_dir = smm_output_folder,
    covariates = c("day_of_week", "session_today")  # EXPLICIT
  )

  list_name <- paste0(current_outcome, "_", current_centering)
  all_smm_results[[list_name]] <- smm_results

  # Generate BIC comparison plot
  if (!is.null(smm_results) && length(smm_results$all_results) > 0) {
    bic_result <- compare_bic_cyclic(
      data = data_for_analysis,
      outcome = current_outcome,
      time_var = current_time_var,
      centering = current_centering,
      smm_results = smm_results,
      save_dir = smm_output_folder
    )
    all_bic_results[[list_name]] <- bic_result

    gams <- model_plot_modx_gam_cyclic(
  data = data_for_analysis,
  outcome = current_outcome,
  time_var = current_time_var,
  smm_result = smm_results,
  centering = current_centering,
  save_dir = smm_output_folder,
  covariates = c("day_of_week", "session_today")  # EXPLICIT - must match run_smm_bernoulli
)
    all_gam_results[[list_name]] <- gams
  }

  message(paste0("✅ Finished analysis for: '", current_plot_label, "' (", current_centering, "-centered)"))
}

message("\n🏁 ALL BINARY OUTCOMES PROCESSED! 🏁")
```


# --- 3. SAVE ALL RESULTS ---

```{r}
## WHAT THIS SECTION DOES:
# Saves all Smoothing Mixture Model (SMM) analysis results to RDS files for:
# - all_smm_results: SMM model objects, classifications, and posterior probabilities
# - all_bic_results: BIC comparison data for model selection
# - all_gam_results: GAM (Generalized Additive Model) smoothing results
#
## WHY IT'S NECESSARY:
# - Enables reloading results in future R sessions without re-running analyses
# - SMM analyses can take considerable time; saving results improves workflow efficiency
# - Allows iterative labeling and exploration without recalculating models
#
## IMPORTANT NOTES:
# - Files saved to smm_output_folder (defined in Section 1)
# - Each RDS file preserves complete R objects with all metadata
# - To reload: use readRDS() as shown in Section 7

message("\n💾 Saving results objects...")

saveRDS(all_smm_results, file = file.path(smm_output_folder, "all_smm_results.rds"))
message("✅ Saved: all_smm_results.rds")

saveRDS(all_bic_results, file = file.path(smm_output_folder, "all_bic_results.rds"))
message("✅ Saved: all_bic_results.rds")

saveRDS(all_gam_results, file = file.path(smm_output_folder, "all_gam_results.rds"))
message("✅ Saved: all_gam_results.rds")

message(glue("\n📁 All results saved to {smm_output_folder}"))
```


# --- 4. SELECT CONFIGURATION FOR EACH OUTCOME (OPTIONAL) ---

```{r}
## WHAT THIS SECTION DOES:
# Selects optimal cycle centering (menses vs ovulation) AND group size (g) for each 
# outcome based on:
# - BIC values from Section 2 analysis loop
# - Visual inspection of trajectory plots
# - Domain knowledge about binary drinking patterns
#
## HOW IT INTEGRATES WITH PREVIOUS SECTIONS:
# - Section 2 runs SMM analyses for ALL outcomes with BOTH centering methods and g=2:4
# - This section lets you choose specific centering + g combinations for each outcome
# - Selected configurations are used for labeling in Section 5
#
## USER DECISIONS NEEDED:
# - Review BIC comparison plots (saved by Section 2)
# - Adjust outcome_config_selections dataframe as needed
# - You can now select MULTIPLE versions per outcome (e.g., multiple g values or centering methods)
# - Default: All binary outcomes with both menses and ovulation centering with g=2:4
#
## WHY SELECT CONFIGURATIONS?
# Different outcomes may fit better with different centering:
# - Drinking patterns may align better with menses-centered (premenstrual drinking)
# - Or ovulation-centered (mid-cycle drinking patterns)
# 
# You can select multiple versions per outcome. For example:
# - drink_today_bin: 2-group menses-centered, 3-group ovulation-centered, 4-group ovulation-centered
# - fourplustoday_bin: 3-group menses-centered, 4-group menses-centered
# 
# Select the best centering + g combinations based on BIC values and visual inspection.

# Set to TRUE to enable configuration selection (recommended)
# Set to FALSE to label all centering methods and group sizes for all outcomes
use_selected_configs <- TRUE

if (use_selected_configs) {
  # Only include outcomes that actually have SMM results available in `all_smm_results`
  # Extract outcome names by removing the centering suffix (_menses or _ovulation)
  available_outcomes <- unique(sub("_(menses|ovulation)$", "", names(all_smm_results)))
  selected_outcomes <- intersect(names(binary_outcomes), available_outcomes)
  
  # Inform the user if any outcomes were excluded because they had no SMM results
  excluded_outcomes <- setdiff(names(binary_outcomes), selected_outcomes)
  if (length(excluded_outcomes) > 0L) {
    message(
      "⚠️  Excluding outcomes with no SMM results from configuration selection: ",
      paste(excluded_outcomes, collapse = ", ")
    )
  }
  
  # ===== EASIER CONFIGURATION APPROACH (Recommended for multiple outcomes) =====
  # 
  # Instead of manually building dataframes, use this compact list-based approach:
  # - Specify configurations per outcome using named lists
  # - Much easier to read and maintain with many outcomes
  # - Each outcome can have different centering methods and g values
  
  # Define your configuration as a named list
  # Format: outcome_name = list(menses = c(g_values), ovulation = c(g_values))
  smm_config <- list(
    # Binary drinking outcomes - typically test both menses and ovulation centering
    drink_today_bin      = list(menses = 2:4, ovulation = 2:4),
    fourplustoday_bin    = list(menses = 2:4, ovulation = 2:4),
    felt_drunk_today_bin = list(menses = 2:4, ovulation = 2:4)
  )
  
  # Example customizations:
  # - Select only specific g values: drink_today_bin = list(menses = c(2, 4))  # Skip g=3
  # - Mix centering methods: drink_today_bin = list(menses = 2, ovulation = c(3, 4))
  # - Single version only: fourplustoday_bin = list(menses = 3)
  
  # Convert the compact list format to the dataframe format
  outcome_config_selections <- do.call(rbind, lapply(names(smm_config), function(outcome) {
    config <- smm_config[[outcome]]
    rows <- list()
    for (centering_method in names(config)) {
      g_values <- config[[centering_method]]
      for (g_val in g_values) {
        rows[[length(rows) + 1]] <- data.frame(
          outcome = outcome,
          centering = centering_method,
          g = g_val,
          stringsAsFactors = FALSE
        )
      }
    }
    do.call(rbind, rows)
  }))
  
  # Filter to only outcomes with available results
  outcome_config_selections <- subset(
    outcome_config_selections,
    outcome %in% selected_outcomes
  )
  
  # Validate selections exist in results
  message("\n📋 Configuration selections:")
  print(outcome_config_selections)
  
  valid_selections <- TRUE
  for (config_idx in 1:nrow(outcome_config_selections)) {
    outcome <- outcome_config_selections$outcome[config_idx]
    centering <- outcome_config_selections$centering[config_idx]
    g <- outcome_config_selections$g[config_idx]
    list_name <- paste0(outcome, "_", centering)
    
    if (!(list_name %in% names(all_smm_results))) {
      message(glue("⚠️  Configuration not found: {list_name}"))
      valid_selections <- FALSE
    } else if (!(as.character(g) %in% names(all_smm_results[[list_name]]$all_results))) {
      message(glue("⚠️  Group size g={g} not found for: {list_name}"))
      valid_selections <- FALSE
    }
  }
  
  if (valid_selections) {
    message("✅ All selected configurations found in results!")
  } else {
    message("⚠️  Some configurations not found. Available:")
    print(names(all_smm_results))
  }
} else {
  outcome_config_selections <- NULL
  message("ℹ️  Configuration selection disabled - will process all centering methods")
}
```


# --- 5. INTERACTIVE GROUP LABELING (OPTIONAL BUT RECOMMENDED) ---

```{r}
## WHAT IT DOES:
# - Loads helper functions for interactive group labeling
# - Builds list of labeling targets based on configuration selections
# - Sets up batch processing for assigning meaningful names to trajectory groups
#
## WHY IT'S NECESSARY:
# - SMM assigns arbitrary group numbers (1, 2, 3) with no inherent meaning
# - Meaningful labels (e.g., "Luteal-Peak", "Stable-Low") improve interpretability
# - Consistent labeling ensures clarity in manuscripts and presentations
#
## IMPORTANT NOTES:
# - Labeling functions: defined in scripts/label_smm_groups.R
# - Common label patterns for binary outcomes:
#   * Timing-based: "Perimenstrual-High", "Luteal-Peak", "Follicular-Low"
#   * Pattern-based: "Stable-Low", "Cyclic-Pattern", "Stable-High"
#   * Probability-based: "High-Risk", "Low-Risk", "Variable"
# - Labels are saved to CSV files alongside plots for reproducibility
# - With the new configuration format, you can now select specific g values per outcome

# Load labeling functions
if (!file.exists("scripts/label_smm_groups.R")) {
  stop("Required file not found: scripts/label_smm_groups.R")
}
source("scripts/label_smm_groups.R")

# Safety check: if user requested selected configs, ensure selections are available
if (isTRUE(use_selected_configs) &&
    (!exists("outcome_config_selections") || is.null(outcome_config_selections))) {
  stop(
    "use_selected_configs is TRUE, but 'outcome_config_selections' is not available.\n",
    "Please run the configuration selection step before this labeling chunk."
  )
}

# Define which analyses to label
labeling_targets <- list()

if (use_selected_configs && !is.null(outcome_config_selections)) {
  # Use selected configurations - now includes specific g values per outcome
  message("\n🏷️  Setting up labeling for selected configurations...")
  
  for (config_idx in 1:nrow(outcome_config_selections)) {
    outcome <- outcome_config_selections$outcome[config_idx]
    centering <- outcome_config_selections$centering[config_idx]
    g <- outcome_config_selections$g[config_idx]
    
    # Add labeling target with the specific g value from configuration
    labeling_targets[[length(labeling_targets) + 1]] <- list(
      outcome = outcome,
      centering = centering,
      g = g
    )
  }
} else {
  # Manual configuration - example for key binary outcomes
  message("\n🏷️  Add your own labeling targets...")
  
  # Example: Label drinking with menses-centered, 3 groups
  # labeling_targets[[1]] <- list(outcome = "drink_today_bin", centering = "menses", g = 3)
  
  # Example: Label binge drinking with ovulation-centered, 3 groups
  # labeling_targets[[2]] <- list(outcome = "fourplustoday_bin", centering = "ovulation", g = 3)
}

message(glue("📝 Total labeling targets: {length(labeling_targets)}"))
```

```{r eval=FALSE}
# INTERACTIVE LABELING
# 
# TO RUN THIS INTERACTIVELY:
# - Either change `eval=FALSE` to `eval=TRUE` in the chunk header above, OR
# - Run this chunk manually (Ctrl+Enter / Cmd+Enter in RStudio)
# 
# When run, this code will prompt you to view plots and enter labels for each group.
#
# DEPENDENCIES:
# This chunk requires objects created in earlier sections:
# - labeling_targets: list of labeling configurations (from Section 5)
# - all_smm_results: SMM model results (from Section 2)
# - smm_output_folder: output directory (from Section 1)
#
# If you enable this chunk without running the preceding setup and model-fitting
# sections, these objects will be missing and the code will fail.

# Safety check to ensure required objects exist before proceeding
required_objects <- c("labeling_targets", "all_smm_results", "smm_output_folder")
missing_objects <- required_objects[!vapply(required_objects, exists, logical(1))]
if (length(missing_objects) > 0) {
  stop(
    "This interactive labeling chunk depends on objects created in earlier chunks:\n",
    "  - labeling_targets (Section 5)\n",
    "  - all_smm_results (Section 2)\n",
    "  - smm_output_folder (Section 1)\n\n",
    "The following objects are missing: ",
    paste(missing_objects, collapse = ", "), "\n",
    "Please run the SMM setup and model-fitting chunks before enabling this chunk."
  )
}

if (length(labeling_targets) > 0) {
  date_folder <- format(Sys.Date(), "%Y%m%d")
  all_group_labels <- list()
  
  for (target in labeling_targets) {
    outcome <- target$outcome
    centering <- target$centering
    g <- target$g
    
    message(glue("\n{strrep('=', 80)}"))
    message(glue("🏷️  LABELING: {outcome} ({centering}-centered, g={g})"))
    message(glue("{strrep('=', 80)}\n"))
    
    # Check if this analysis exists
    list_name <- paste0(outcome, "_", centering)
    if (!list_name %in% names(all_smm_results)) {
      message(glue("⚠️  Skipping - no results found for {list_name}"))
      next
    }
    
    # Check if this group size exists
    if (!as.character(g) %in% names(all_smm_results[[list_name]]$all_results)) {
      message(glue("⚠️  Skipping - g={g} not available for {list_name}"))
      next
    }
    
    # Interactive labeling
    labels <- label_smm_groups(
      base_save_dir = smm_output_folder,
      outcome = outcome,
      centering = centering,
      g = g,
      date_folder = date_folder,
      interactive = TRUE
    )
    
    # Save labels to CSV
    save_group_labels(
      labels_df = labels,
      base_save_dir = smm_output_folder,
      outcome = outcome,
      centering = centering,
      g = g,
      date_folder = date_folder
    )
    
    # Store in list
    label_key <- glue("{outcome}_{centering}_g{g}")
    all_group_labels[[label_key]] <- labels
    
    message(glue("✅ Labels saved for {label_key}\n"))
  }
  
  message("\n🎉 ALL LABELING COMPLETE!")
  
  # Save all labels object
  saveRDS(all_group_labels, file = file.path(smm_output_folder, "all_group_labels.rds"))
  message(glue("💾 Saved: all_group_labels.rds"))
}
```


# --- 6. APPLY LABELS TO CLASSIFICATION DATA ---

```{r}
## WHAT THIS SECTION DOES:
# Applies human-readable labels to SMM classification dataframes by:
# - Merging group labels (from Section 5) with classification results (from Section 2)
# - Creating labeled dataframes ready for downstream analyses
# - Saving labeled classifications for use in subsequent R sessions
#
## INPUT DATA:
# - all_group_labels: Label mappings created in interactive labeling (Section 5)
#   Structure: list of dataframes with smm_group and group_label columns
#
## PROCESSING:
# - Parses label keys to extract outcome, centering, and group size
# - Retrieves corresponding classification dataframe from all_smm_results
# - Applies labels via apply_group_labels() helper function
# - Preserves all original columns (id, smm_group, posterior probabilities)
#
## OUTPUT:
# - all_labeled_classifications.rds: List of labeled classification dataframes
#   saved to smm_output_folder for merging with main analysis dataset

# Check if labels exist (defined here for chunk independence)
has_labels <- exists("all_group_labels") && length(all_group_labels) > 0

if (has_labels) {
  message("\n📊 Applying labels to classification dataframes...")
  
  all_labeled_classifications <- list()
  
  for (label_key in names(all_group_labels)) {
    # Parse the key: expected pattern "outcome_centering_g#"
    # Use regex to safely handle outcome names that may contain underscores or "g"
    key_match <- regexec("^(.*)_g([0-9]+)$", label_key)
    match_vec <- regmatches(label_key, key_match)[[1]]
    # If the key does not match the expected pattern, skip to avoid mis-parsing
    if (length(match_vec) != 3L) next
    
    outcome_centering <- match_vec[2]
    g <- as.numeric(match_vec[3])
    
    # Check if results exist
    if (!outcome_centering %in% names(all_smm_results)) next
    if (!as.character(g) %in% names(all_smm_results[[outcome_centering]]$all_results)) next
    
    # Get classification dataframe
    class_df <- all_smm_results[[outcome_centering]]$all_results[[as.character(g)]]$class
    
    # Apply labels
    class_labeled <- apply_group_labels(class_df, all_group_labels[[label_key]])
    
    # Store
    all_labeled_classifications[[label_key]] <- class_labeled
    
    message(glue("✅ Applied labels to {label_key}"))
  }
  
  # Save labeled classifications
  saveRDS(all_labeled_classifications, 
          file = file.path(smm_output_folder, "all_labeled_classifications.rds"))
  message(glue("\n💾 Saved: all_labeled_classifications.rds"))
}
```


# --- 7. LOAD SAVED RESULTS (FOR FUTURE SESSIONS) ---

```{r}
## WHAT THIS SECTION PROVIDES:
# Template code for reloading saved SMM results in future R sessions without
# re-running the computationally intensive analyses from Section 2.
#
## WHEN TO USE:
# - Starting a new R session after completing SMM analyses
# - Returning to labeling or exploration after closing RStudio
# - Sharing results with collaborators who don't need to refit models
#
## WHAT CAN BE LOADED:
# - all_smm_results.rds: Complete SMM model objects and classifications
# - all_bic_results.rds: BIC comparison data for model selection
# - all_gam_results.rds: GAM smoothing results
# - all_group_labels.rds: Interactive label mappings (if previously created)
# - all_labeled_classifications.rds: Labeled classification dataframes
# - Individual label CSV files: Specific labels for outcome/centering/g combinations
#
## IMPORTANT NOTES:
# - All lines below are commented by default
# - Uncomment only the lines you need for your current session
# - Ensure file paths match your output folder structure

## If you need to reload results from a previous session:

# Load all results
# all_smm_results <- readRDS(file.path(smm_output_folder, "all_smm_results.rds"))
# all_bic_results <- readRDS(file.path(smm_output_folder, "all_bic_results.rds"))
# all_gam_results <- readRDS(file.path(smm_output_folder, "all_gam_results.rds"))

# Load labels (if you've already created them)
# all_group_labels <- readRDS(file.path(smm_output_folder, "all_group_labels.rds"))
# all_labeled_classifications <- readRDS(file.path(smm_output_folder, "all_labeled_classifications.rds"))

# Or load specific labels from CSV:
# labels_drink_g3 <- load_group_labels(
#   base_save_dir = smm_output_folder,
#   outcome = "drink_today_bin",
#   centering = "menses",
#   g = 3,
#   date_folder = "20260107"
# )
```


# --- 8. ANALYSIS SUMMARY ---

```{r}
## WHAT THIS SECTION DOES:
# Provides a comprehensive summary of the completed Binary SMM analyses including:
# - Count of outcomes successfully analyzed
# - List of available results with centering methods and group sizes
# - Configurations that have been labeled (if Section 5 was run)
# - Locations of all saved output files
#
## USE CASE:
# Quick reference to understand what was accomplished in this session and
# what files are available for downstream analyses or future sessions.

# Check if labels exist (defined here for chunk independence)
# Note: Redefined here to allow Section 8 to run independently of Section 6
has_labels <- exists("all_group_labels") && length(all_group_labels) > 0

message("\n📊 ========== BINARY SMM ANALYSIS SUMMARY ==========")
message(glue("Total outcomes analyzed: {length(all_smm_results)}"))
message("\n✅ Available results:")
for (name in names(all_smm_results)) {
  if (!is.null(all_smm_results[[name]]$all_results)) {
    group_sizes <- names(all_smm_results[[name]]$all_results)
    message(glue("  - {name}: g = {paste(group_sizes, collapse=', ')}"))
  }
}

if (has_labels) {
  message("\n🏷️  Labeled configurations:")
  for (label_key in names(all_group_labels)) {
    message(glue("  - {label_key}"))
  }
}

message("\n📁 Saved files:")
message(glue("  - all_smm_results.rds"))
message(glue("  - all_bic_results.rds"))
message(glue("  - all_gam_results.rds"))
if (has_labels) {
  message(glue("  - all_group_labels.rds"))
  message(glue("  - all_labeled_classifications.rds"))
}

message(glue("\n📂 Output folder: {smm_output_folder}"))
message("\n🎉 Binary SMM Analysis complete!")
```


```{r}
# Restore original dataframe (df_orig) to df to reset any modifications
# made during the SMM analyses. This prepares for potential subsequent analyses
# or code chunks that expect the unmodified input data.
if (exists("df_orig")) {
  df <- df_orig
} else {
  warning("df_orig does not exist; df remains unchanged")
}

```

