---
title: "HERA Individual GAM Reports"
author: "Tory Eisenlohr-Moul, PhD"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

> **Script Purpose:** Automated, participant-centric analysis of symptom cyclicity using cyclic GAMs.
> 1) **Setup & Preparation** — load data, define outcomes/layouts, precompute transforms.
> 2) **Automated Reporting** — one **landscape** PDF per participant with compact cover and paged plots.
> 3) **CSV Audit** — per-ID × outcome × time-var stats for triage/QC.

---

### 1. Setup & Configuration

#### 1.1: Load Libraries
```{r}
#install.packages(c("dplyr","ggplot2","mgcv","patchwork","gridExtra","grid","tibble","tidyr","knitr"))

library(dplyr)
library(ggplot2)
library(mgcv)
library(patchwork)
library(gridExtra)
library(grid)
library(tibble)
library(tidyr)
library(knitr)
library(broom)
library(scales)

options(scipen = 999)
set.seed(12345)
```



```{r}
# Import Data

df <- readRDS('/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_20260319_clean.rds')

```


```{r}
# --- Path to the main output directory ---
gam_output_base <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/03_dataedits_output/06_Single_GAMs/"

# --- Define all Time Variables to iterate over (must be scaled to [-1, 1]) ---
time_vars_to_run <- c("cyclic_time_impute")
  
  #"cyclic_time", "cyclic_time_ov", "cyclic_time_impute", "cyclic_time_imp_ov")

# --- Table Formatting ---
table_cols_per_page <- 6

# --- Plot Layout (grid per page) ---
plots_per_col <- 3
plots_per_row <- 3 

```

```{r}
# --- Create output folder with current date ---
current_date <- format(Sys.Date(), "%Y%m%d")
output_folder <- file.path(gam_output_base, paste0("GAM_Reports_", current_date))
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)

# --- Load Scaled Dataset ---
#df

# --- Robustness: Verify time variables exist ---
time_vars <- intersect(time_vars_to_run, names(df))
if (length(time_vars) < length(time_vars_to_run)) {
  missing_vars <- setdiff(time_vars_to_run, time_vars)
  warning("The following time_vars were not found and will be skipped: ", paste(missing_vars, collapse = ", "))
}

```

```{r}
# --- Define Outcomes to Model ---
outcomes <- c(
  #"numdrinkstoday" = "Number of Drinks",
  "drsp_NA"          = "DRSP Negative Affect Mean",
  "drsp_NAmax"       = "DRSP Negative Affect Max",
  "drsp_total"       = "DRSP Total",
  "drsp_distress"    = "Distress (DRSP)",
  "drsp_labilsocsx"  = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"  = "Breast Sx (DRSP)",
  "drsp_eatsx"       = "Eating Sx",
  "drsp_lowarousal"  = "Low Arousal Sx (DRSP)",
  "drsp_1"              = "Depression",
  "drsp_2"         = "Hopelessness",
  "drsp_3"        = "Worthlessness",
  "drsp_4"          = "Anxiety",
  "drsp_5"       = "Mood Swings",
  "drsp_6"         = "Rejection Sensitivity",
  "drsp_7"            = "Anger/Irritability",
  "drsp_8"        = "Interpersonal Conflicts",
  "drsp_9"    = "Loss of Interest",
  "drsp_16"        = "Feeling Overwhelmed",
  "drsp_17"     = "Feeling Out of Control",
  "drsp_10"        = "Difficulty Concentrating",
  "drsp_11"            = "Fatigue/Lethargy",
  "drsp_14"       = "Hypersomnia",
  "drsp_15"       = "Insomnia",
  "drsp_12"         = "Overeating",
  "drsp_13"     = "Specific Food Cravings",
  "drsp_18"      = "Breast Tenderness",
  "drsp_19"     = "Breast Swelling",
  "drsp_20"         = "Headache",
  "drsp_21"          = "Joint/Muscle Pain"
)

# --- Physical vs Affect symptom classification (raw DRSP items only) ---
physical_symptoms <- c("drsp_12", "drsp_13", "drsp_14", "drsp_15",
                       "drsp_18", "drsp_19", "drsp_20", "drsp_21")

# --- Pre-calculate log-transformations ---
for (var in names(outcomes)) {
  if (var %in% names(df)) {
    log_var_name <- paste0(var, "_Log")
    df[[log_var_name]] <- log(df[[var]] + 1)
  }
}

```

```{r}
# --- Prediction grids reused per time var ---
pred_grid_list <- setNames(lapply(time_vars, function(tv) {
  df <- data.frame(seq(-1, 1, by = 0.05))
  names(df) <- tv
  df
}), time_vars)

# --- Axis labels + reference lines per time var ---
axis_meta <- lapply(time_vars, function(tv) {
  if (grepl("ov", tv, ignore.case = TRUE)) {
    list(labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses"),
         menses = c(-1, 1), ov = 0)
  } else {
    list(labels = c("Ovulation", "50%L", "Menses", "50%F", "Ovulation"),
         menses = 0, ov = c(-1, 1))
  }
})
names(axis_meta) <- time_vars

# --- Fit + predict one outcome for one time var (bam + predict) ---
fit_outcome <- function(dat, outcome_var, tv, outcomes_map, k_max = 10L) {
  log_out <- paste0(outcome_var, "_Log")
  keep <- c(log_out, tv)
  if (!all(keep %in% names(dat))) return(NULL)
  dat <- dat[complete.cases(dat[, keep]), keep, drop = FALSE]
  if (nrow(dat) < 10 || var(dat[[log_out]], na.rm = TRUE) < 1e-6) return(NULL)

  k <- min(k_max, nrow(dat) - 1L); if (k < 3L) return(NULL)

  fml <- as.formula(paste0(log_out, " ~ s(", tv, ", bs = 'cc', k = ", k, ")"))
  knots <- setNames(list(c(-1, 1)), tv)

  mf <- try(mgcv::bam(fml, data = dat, knots = knots, method = "fREML",
                      discrete = TRUE, select = TRUE),
            silent = TRUE)
  if (inherits(mf, "try-error")) return(NULL)

  s  <- summary(mf)
  p  <- suppressWarnings(as.numeric(s$s.pv[1]))
  st <- s$s.table
  edf <- suppressWarnings(as.numeric(st[1, "edf"]))
  Fval <- suppressWarnings(as.numeric(st[1, "F"]))

  pg <- pred_grid_list[[tv]]
  pr <- predict(mf, newdata = pg, se.fit = TRUE, type = "link")  # link scale for log(y+1)
  est <- exp(pr$fit) - 1
  conf.low  <- exp(pr$fit - 1.96 * pr$se.fit) - 1
  conf.high <- exp(pr$fit + 1.96 * pr$se.fit) - 1

  # Record cycle time at peak (max) and trough (min) of predicted curve
  # pg and est are aligned: pg is the prediction grid, est is the back-transformed prediction
  peak_time   <- pg[[tv]][which.max(est)]
  trough_time <- pg[[tv]][which.min(est)]

  pred <- cbind(pg, estimate = est, conf.low = conf.low, conf.high = conf.high)
  pred$outcome_label <- outcomes_map[[outcome_var]]
  pred$p.value <- p
  pred$raw_range <- diff(range(est, na.rm = TRUE))
  pred$n <- nrow(dat)

  list(
    pred = pred,
    sig  = tibble::tibble(
      variable = outcome_var,
      outcome_label = outcomes_map[[outcome_var]],
      time_var = tv,
      p.value = p,
      edf = edf,
      F = Fval,
      n = nrow(dat),
      range = diff(range(est, na.rm = TRUE)),
      peak_time = peak_time,
      trough_time = trough_time
    )
  )
}

# --- Plot one outcome curve with annotations ---
create_annotated_plot <- function(plot_df, tv, axis_meta_list) {
  if (is.null(plot_df) || nrow(plot_df) == 0) return(NULL)
  max_pt <- dplyr::slice_max(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  min_pt <- dplyr::slice_min(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  rng   <- dplyr::first(plot_df$raw_range)
  p_val <- dplyr::first(plot_df$p.value)
  n_obs <- dplyr::first(plot_df$n)
  cpass_flag <- !is.null(rng) && is.finite(rng) && rng >= 1.5
  stars <- dplyr::case_when(
    is.finite(p_val) & p_val < 0.001 ~ "***",
    is.finite(p_val) & p_val < 0.01  ~ "**",
    is.finite(p_val) & p_val < 0.05  ~ "*",
    TRUE ~ ""
  )
  cpass_label <- if (cpass_flag) "\u25B2 CPASS | " else ""
  ann <- if (!is.null(n_obs) && !is.na(n_obs)) {
    sprintf("%s%s Range = %.2f, N = %d", cpass_label, stars, rng, as.integer(n_obs))
  } else {
    sprintf("%s%s Range = %.2f", cpass_label, stars, rng)
  }
  meta  <- axis_meta_list[[tv]]

  p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = .data[[tv]], y = estimate)) +
    ggplot2::geom_hline(yintercept = 4, color = "red", linewidth = 0.4, alpha = 0.4) +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red") +
    ggplot2::geom_vline(xintercept = meta$ov,     color = "forestgreen") +
    ggplot2::geom_line(linewidth = 1, color = "#0072B2") +
    ggplot2::geom_ribbon(ggplot2::aes(ymin = conf.low, ymax = conf.high),
                         alpha = 0.2, fill = "#0072B2") +
    ggplot2::geom_point(data = max_pt, size = 2.5) +
    ggplot2::geom_point(data = min_pt, size = 2.5) +
    ggplot2::annotate("text", x = -1, y = Inf, label = ann,
                      hjust = -0.1, vjust = 1.5, size = 3, fontface = "bold") +
    ggplot2::labs(title = dplyr::first(plot_df$outcome_label), x = "Cycle Phase", y = NULL) +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::coord_cartesian(ylim = c(1, 6)) +
    ggplot2::theme_minimal() +
    ggplot2::theme(plot.title = ggplot2::element_text(size = 10, face = "bold", hjust = 0.5))

  # Highlight plots that meet CPASS range threshold with a gold border
  if (cpass_flag) {
    p <- p + ggplot2::theme(
      panel.border = ggplot2::element_rect(color = "goldenrod2", linewidth = 1.5, fill = NA)
    )
  }
  p
}

```

```{r}
report_folder <- file.path(output_folder, "Participant_Reports_PDF_byID")
dir.create(report_folder, showWarnings = FALSE, recursive = TRUE)

ids_to_run    <- unique(df$id)   # set to subset for testing, e.g., c("203","109")
skip_existing <- TRUE                         # set FALSE to overwrite

# collector for a single CSV at the end
all_ids_signif <- list()

# fixed layout controls
max_plots_per_page <- plots_per_col * plots_per_row
table_theme <- gridExtra::ttheme_default(
  base_size = 9,
  core   = list(fg_params = list(hjust = 0, x = 0.05)),
  colhead = list(fg_params = list(hjust = 0, x = 0.05))
)

for (current_id in ids_to_run) {
  message(sprintf("--- Processing ID: %s ---", current_id))
  datSX_id <- subset(df, id == current_id)

  # Pass 1: fit & collect
  preds_by_tv <- list()
  sig_by_tv   <- list()

  for (tv in time_vars) {
    message(sprintf("  ... time variable: %s", tv))
    all_preds <- list()
    sig_list  <- list()

    for (outcome_var in names(outcomes)) {
      res <- fit_outcome(datSX_id, outcome_var, tv, outcomes)
      if (is.null(res)) next
      all_preds[[length(all_preds) + 1]] <- res$pred
      sig_list[[length(sig_list) + 1]]  <- res$sig
    }

    if (length(all_preds) > 0) {
      preds_by_tv[[tv]] <- dplyr::bind_rows(all_preds)
      sig_by_tv[[tv]]   <- dplyr::bind_rows(sig_list)
    }
  }

  if (!length(preds_by_tv)) {
    message("    -> No successful models for this ID; skipping.")
    next
  }

  # Build per-ID significance object for CSV
  sig_all <- dplyr::bind_rows(sig_by_tv)
  sig_all$id <- current_id
  all_ids_signif[[as.character(current_id)]] <- sig_all

  # ---- Open a single LANDSCAPE PDF per participant ----
  pdf_filename <- file.path(report_folder, sprintf("Report_ID_%s.pdf", current_id))
  if (isTRUE(skip_existing) && file.exists(pdf_filename)) {
    message("    -> File exists; skipping. Set skip_existing=FALSE to overwrite.")
    next
  }
  grDevices::pdf(pdf_filename, width = 11, height = 8.5)   # LANDSCAPE

  # ---- COVER: Compact counts by time_var ----
  sig_all_clean <- sig_all %>%
    dplyr::mutate(
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    )

  cover_counts <- sig_all_clean %>%
    dplyr::group_by(time_var) %>%
    dplyr::summarise(
      tested = dplyr::n(),
      n_sig  = sum(stars != "", na.rm = TRUE),
      n_cpass = sum(is.finite(range) & range >= 1.5 & stars != "", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::unite("count", n_sig, tested, sep = " / ") %>%
    dplyr::rename(`Sig / Tested` = count, `CPASS (range >= 1.5)` = n_cpass)

  cover_theme <- gridExtra::ttheme_default(
    base_size = 9,
    core = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"),
      padding  = grid::unit(c(2, 2), "mm")
    ),
    colhead = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
    )
  )

  title_grob    <- grid::textGrob(paste("Compact Summary — ID:", current_id),
                                  gp = grid::gpar(fontsize = 16, fontface = "bold"))
  subtitle_grob <- grid::textGrob("Rows show: # significant / # tested (p < .05) for each time variable",
                                  gp = grid::gpar(fontsize = 9))
  cover_grob    <- gridExtra::tableGrob(cover_counts, rows = NULL, theme = cover_theme)
  gridExtra::grid.arrange(title_grob, subtitle_grob, cover_grob,
                          ncol = 1, heights = grid::unit(c(0.12, 0.08, 0.80), "npc"))

  # ---- OPTIONAL: compact stars matrix for outcomes with any significance ----
  compact_stars <- sig_all_clean %>%
    dplyr::group_by(variable) %>%
    dplyr::mutate(any_sig = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(any_sig) %>%
    dplyr::mutate(Outcome = dplyr::coalesce(outcome_label, variable)) %>%
    dplyr::mutate(
      Outcome = ifelse(nchar(Outcome) > 22, variable, Outcome),
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    ) %>%
    dplyr::select(Outcome, time_var, stars) %>%
    tidyr::pivot_wider(names_from = time_var, values_from = stars, values_fill = "") %>%
    dplyr::arrange(Outcome)

  if (nrow(compact_stars) > 0) {
    compact_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(
        fg_params = list(hjust = 0, x = 0.02),
        padding = grid::unit(c(1.5, 2), "mm")
      ),
      colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
    )
    header <- grid::textGrob("Significant Outcomes Only (★ = p < .05)",
                             gp = grid::gpar(fontsize = 12, fontface = "bold"))
    tbl <- gridExtra::tableGrob(compact_stars, rows = NULL, theme = compact_theme)
    gridExtra::grid.arrange(header, tbl, ncol = 1, heights = grid::unit(c(0.10, 0.90), "npc"))
  }
  
  # ---- PAGE 2: Significant EDFs ranked by Range (across all time vars) ----
sig_ranked <- sig_all_clean %>%
  dplyr::filter(is.finite(p.value) & p.value < 0.05) %>%     # keep significant smooths
  dplyr::mutate(
    Outcome = dplyr::coalesce(outcome_label, variable),
    Outcome = ifelse(nchar(Outcome) > 40, variable, Outcome) # shorten long names
  ) %>%
  dplyr::transmute(
    Outcome,
    TimeVar  = time_var,
    EDF      = round(edf, 2),
    Range    = round(range, 2),
    F        = round(F, 2),
    N        = n
  ) %>%
  dplyr::arrange(dplyr::desc(Range), Outcome, TimeVar)

if (nrow(sig_ranked) == 0) {
  # show an empty-state page
  header <- grid::textGrob(
    paste("Significant Smooths — ID:", current_id, "(none at p < .05)"),
    gp = grid::gpar(fontsize = 14, fontface = "bold")
  )
  sub <- grid::textGrob("No significant cyclic effects detected across any time variable.",
                        gp = grid::gpar(fontsize = 10, fontface = "italic"))
  gridExtra::grid.arrange(header, sub, ncol = 1,
                          heights = grid::unit(c(0.2, 0.8), "npc"))
} else {
  # compact table theme to fit on a single landscape page
  ranked_theme <- gridExtra::ttheme_default(
    base_size = 8,
    core = list(
      fg_params = list(hjust = 0, x = 0.02),
      padding = grid::unit(c(1.2, 2), "mm")
    ),
    colhead = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
    )
  )
  title_grob <- grid::textGrob(
    paste("Significant Smooths (p < .05), Ranked by Range — ID:", current_id),
    gp = grid::gpar(fontsize = 14, fontface = "bold")
  )
  note_grob <- grid::textGrob(
    "Columns: EDF = effective degrees of freedom; Range = max(pred) − min(pred) on original scale.",
    gp = grid::gpar(fontsize = 9, fontface = "italic")
  )
  ranked_tbl <- gridExtra::tableGrob(sig_ranked, rows = NULL, theme = ranked_theme)

  gridExtra::grid.arrange(
    title_grob, note_grob, ranked_tbl,
    ncol = 1,
    heights = grid::unit(c(0.12, 0.06, 0.82), "npc")
  )
}

  # ---- Person-Specific Peak & Trough Timing Distribution ----
  # Separate density plots for Physical vs Affect symptoms (raw items only)
  for (tv in names(sig_by_tv)) {
    sig_tv <- sig_by_tv[[tv]]
    if (nrow(sig_tv) == 0) next

    # Keep only raw DRSP items (drsp_1, drsp_2, …), exclude composites
    sig_tv_raw <- sig_tv %>% dplyr::filter(grepl("^drsp_\\d+$", variable))
    if (nrow(sig_tv_raw) == 0) next

    meta <- axis_meta[[tv]]

    # Helper to build a density plot for one symptom category
    make_timing_density <- function(sig_subset, category_label) {
      tl <- dplyr::bind_rows(
        sig_subset %>% dplyr::filter(!is.na(peak_time)) %>%
          dplyr::transmute(time = peak_time, type = "Peak (Max)", outcome = outcome_label),
        sig_subset %>% dplyr::filter(!is.na(trough_time)) %>%
          dplyr::transmute(time = trough_time, type = "Trough (Min)", outcome = outcome_label)
      )
      if (nrow(tl) < 4) return(NULL)  # need >= 2 per group for density estimation

      ggplot2::ggplot(tl, ggplot2::aes(x = time, fill = type, color = type)) +
        ggplot2::geom_density(alpha = 0.3, adjust = 1.2) +
        ggplot2::geom_rug(alpha = 0.5, length = grid::unit(0.03, "npc")) +
        ggplot2::geom_vline(xintercept = meta$menses, color = "red", linewidth = 0.8,
                            linetype = "dashed") +
        ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen", linewidth = 0.8,
                            linetype = "dashed") +
        ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                    labels = meta$labels) +
        ggplot2::labs(
          title = paste0(category_label, " Symptoms \u2014 ID: ", current_id),
          subtitle = paste0("N outcomes = ", nrow(sig_subset), " | Time variable: ", tv),
          x = "Cycle Phase", y = "Density", fill = NULL, color = NULL
        ) +
        ggplot2::theme_minimal(base_size = 12) +
        ggplot2::scale_fill_manual(values = c("Peak (Max)" = "#D55E00",
                                              "Trough (Min)" = "#0072B2")) +
        ggplot2::scale_color_manual(values = c("Peak (Max)" = "#D55E00",
                                               "Trough (Min)" = "#0072B2"))
    }

    p_physical <- make_timing_density(
      sig_tv_raw %>% dplyr::filter(variable %in% physical_symptoms), "Physical"
    )
    p_affect <- make_timing_density(
      sig_tv_raw %>% dplyr::filter(!variable %in% physical_symptoms), "Affect"
    )

    # Stack the two plots on one page; print whichever are available
    plots <- Filter(Negate(is.null), list(p_physical, p_affect))
    if (length(plots) == 2) {
      gridExtra::grid.arrange(grobs = plots, ncol = 1)
    } else if (length(plots) == 1) {
      print(plots[[1]])
    }
  }

  # ---- Sections per time variable: compact table + pages of plots ----
  for (tv in names(preds_by_tv)) {

    # NA-safe compact significance table (optionally filter to significant only)
    summary_df <- sig_by_tv[[tv]] %>%
      dplyr::mutate(
        stars = dplyr::case_when(
          is.finite(p.value) & p.value < 0.001 ~ "***",
          is.finite(p.value) & p.value < 0.01  ~ "**",
          is.finite(p.value) & p.value < 0.05  ~ "*",
          TRUE ~ ""
        ),
        Outcome = dplyr::coalesce(outcome_label, variable)
      ) %>%
      dplyr::select(Outcome, N = n, EDF = edf, F = F, Range = range, Sig = stars) %>%
      dplyr::arrange(Outcome)
    # Uncomment to show significant rows only:
    # summary_df <- dplyr::filter(summary_df, Sig != "")

    tv_title <- grid::textGrob(paste("Significance — ID:", current_id, "| Time Variable:", tv),
                               gp = grid::gpar(fontsize = 14, fontface = "bold"))
    tv_table <- gridExtra::tableGrob(summary_df, rows = NULL, theme = table_theme)
    gridExtra::grid.arrange(tv_title, tv_table, ncol = 1,
                            heights = grid::unit(c(0.12, 0.88), "npc"))

    # Plots paginated across all outcomes
    combined_plot_data <- preds_by_tv[[tv]]
    all_outcome_labels <- unique(combined_plot_data$outcome_label)
    outcome_pages <- split(all_outcome_labels, ceiling(seq_along(all_outcome_labels) / max_plots_per_page))

    for (page_outcomes in outcome_pages) {
      page_data <- dplyr::filter(combined_plot_data, outcome_label %in% page_outcomes)
      if (nrow(page_data) == 0) next

      plot_list <- lapply(page_outcomes, function(lab) {
        plot_df <- dplyr::filter(page_data, outcome_label == lab)
        create_annotated_plot(plot_df, tv, axis_meta)
      })
      plot_list <- Filter(Negate(is.null), plot_list)

      if (length(plot_list) > 0) {
        if (length(plot_list) < max_plots_per_page) {
          plot_list <- c(plot_list,
                         rep(list(patchwork::plot_spacer()), max_plots_per_page - length(plot_list)))
        }
        page_title <- paste("Symptom Trajectories — ID:", current_id, "| Time Variable:", tv)
        final_page <- patchwork::wrap_plots(plotlist = plot_list, ncol = plots_per_col) +
          patchwork::plot_annotation(
            title = page_title,
            theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 14, face = "bold"))
          )
        print(final_page)
      }
    }
  }

  grDevices::dev.off()
  message(sprintf("    --- ✅ Saved Report: %s ---", pdf_filename))
}

# ---- Write long CSV for evaluation across IDs ----
if (length(all_ids_signif)) {
  eval_csv <- file.path(output_folder, paste0("participant_significance_summary_", current_date, ".csv"))
  write.csv(dplyr::bind_rows(all_ids_signif), eval_csv, row.names = FALSE)
  message(sprintf("Wrote summary CSV: %s", eval_csv))
}

```

---

### 4. Per-Outcome Exploration PDFs

> **Purpose:** For each DRSP outcome, generates a landscape PDF containing summary statistics, threshold analysis, logistic significance probability curves, peak/trough timing distribution, and individual GAM trajectory pages — significant GAMs first, then non-significant sorted by range. All individual GAM plots share a fixed Y axis within each outcome for cross-person comparison.

#### 4.1: Exploration Analysis Functions

```{r exploration-analysis-functions}
# =========================================================
# EXPLORATION / THRESHOLD ANALYSIS FUNCTIONS
# These power the per-outcome exploration PDFs below.
# =========================================================

# --- Basic cleaning: add numeric significance flag, coerce types ---
clean_gam_summary <- function(df) {
  df %>%
    dplyr::mutate(
      significant     = ifelse(p.value < 0.05, "Significant", "Not Significant"),
      significant_num = ifelse(significant == "Significant", 1, 0),
      significant_f   = factor(significant_num,
                               levels = c(0, 1),
                               labels = c("Not Significant", "Significant")),
      range    = as.numeric(range),
      n        = as.numeric(n),
      variable = as.character(variable),
      id       = as.character(id)
    ) %>%
    dplyr::filter(!is.na(range), !is.na(significant_num),
                  !is.na(variable), !is.na(id))
}

# --- Filter to one outcome variable, one row per ID ---
# Aggregates across potential multiple time_var rows per id/variable.
# peak_time, trough_time, and best_time_var come from the row with the
# smallest p.value (best significance), ensuring internal consistency.
get_outcome_df <- function(df, target_var) {
  df %>%
    dplyr::filter(variable == target_var) %>%
    dplyr::group_by(id, variable) %>%
    dplyr::summarise(
      p.value         = min(p.value, na.rm = TRUE),
      range           = max(range,   na.rm = TRUE),
      n               = max(n,       na.rm = TRUE),
      significant_num = ifelse(any(significant_num == 1, na.rm = TRUE), 1, 0),
      # Take timing from the most-significant (smallest p) time_var row
      peak_time       = peak_time[which.min(p.value)][1],
      trough_time     = trough_time[which.min(p.value)][1],
      best_time_var   = time_var[which.min(p.value)][1],
      .groups         = "drop"
    ) %>%
    dplyr::mutate(
      significant_f = factor(
        significant_num,
        levels = c(0, 1),
        labels = c("Not Significant", "Significant")
      )
    )
}

# --- Descriptive summary (overall + by significance) ---
summarize_cyclicity <- function(df_outcome) {
  overall <- df_outcome %>%
    dplyr::summarise(
      n_participants   = dplyr::n(),
      n_significant    = sum(significant_num == 1, na.rm = TRUE),
      prop_significant = mean(significant_num == 1, na.rm = TRUE)
    )
  by_sig <- df_outcome %>%
    dplyr::group_by(significant_f) %>%
    dplyr::summarise(
      n            = dplyr::n(),
      mean_range   = mean(range, na.rm = TRUE),
      sd_range     = sd(range, na.rm = TRUE),
      median_range = median(range, na.rm = TRUE),
      q25_range    = quantile(range, .25, na.rm = TRUE),
      q75_range    = quantile(range, .75, na.rm = TRUE),
      min_range    = min(range, na.rm = TRUE),
      max_range    = max(range, na.rm = TRUE),
      mean_n       = mean(n, na.rm = TRUE),
      .groups      = "drop"
    )
  list(overall = overall, by_significance = by_sig)
}

# --- Proportion significant at various range thresholds ---
threshold_table <- function(df_outcome,
                            thresholds = c(0.10, 0.25, 0.36, 0.50, 0.75, 1.00)) {
  do.call(rbind, lapply(thresholds, function(th) {
    tmp <- df_outcome %>% dplyr::filter(range >= th)
    tibble::tibble(
      threshold        = th,
      n_at_or_above    = nrow(tmp),
      prop_of_sample   = ifelse(nrow(df_outcome) > 0,
                                nrow(tmp) / nrow(df_outcome), NA_real_),
      n_significant    = sum(tmp$significant_num == 1, na.rm = TRUE),
      prop_significant = ifelse(nrow(tmp) > 0,
                                mean(tmp$significant_num == 1, na.rm = TRUE),
                                NA_real_)
    )
  }))
}

# --- Logistic regression: significance ~ range (& + n) ---
fit_logistic_models <- function(df_outcome) {
  model_range   <- glm(significant_num ~ range,     family = binomial, data = df_outcome)
  model_range_n <- glm(significant_num ~ range + n, family = binomial, data = df_outcome)
  list(
    model_range   = model_range,
    model_range_n = model_range_n,
    tidy_range    = broom::tidy(model_range,   exponentiate = TRUE),
    tidy_range_n  = broom::tidy(model_range_n, exponentiate = TRUE)
  )
}

# --- Predicted probability of significance across range values ---
make_prediction_df <- function(df_outcome, model_range, model_range_n = NULL,
                               n_value = NULL) {
  xseq <- seq(min(df_outcome$range, na.rm = TRUE),
              max(df_outcome$range, na.rm = TRUE),
              length.out = 300)
  pred1 <- tibble::tibble(range = xseq) %>%
    dplyr::mutate(pred_prob = predict(model_range, newdata = ., type = "response"),
                  model     = "Range only")
  if (!is.null(model_range_n)) {
    if (is.null(n_value)) n_value <- median(df_outcome$n, na.rm = TRUE)
    pred2 <- tibble::tibble(range = xseq, n = n_value) %>%
      dplyr::mutate(pred_prob = predict(model_range_n, newdata = ., type = "response"),
                    model     = paste0("Range + n (n fixed at ", round(n_value, 1), ")")) %>%
      dplyr::select(range, pred_prob, model)
    dplyr::bind_rows(pred1, pred2)
  } else {
    pred1
  }
}

# --- Histogram of range, colored by significance ---
plot_range_histogram <- function(df_outcome) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, fill = significant_f)) +
    ggplot2::geom_histogram(position = "identity", alpha = 0.65, bins = 30) +
    ggplot2::labs(
      title = paste0("Histogram of ", unique(df_outcome$variable), " Range by Significance"),
      x = "Cyclical range", y = "Count", fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- Logistic curve with jittered observed points ---
plot_significance_probability <- function(df_outcome, pred_df,
                                          thresholds = c(0.25, 0.36, 0.50),
                                          jitter_height = 0.04) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, y = significant_num)) +
    ggplot2::geom_jitter(ggplot2::aes(color = significant_f),
                         width = 0, height = jitter_height,
                         alpha = 0.45, size = 2) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2, inherit.aes = FALSE) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(-0.02, 1.02), breaks = seq(0, 1, by = 0.1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title = paste0("Probability of Significant Cyclicity: ", unique(df_outcome$variable)),
      x = "Cyclical range",
      y = "Predicted probability of significant GAM",
      color = NULL, linetype = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- Binned empirical proportions ---
make_empirical_bins <- function(df_outcome, n_bins = 10) {
  df_outcome %>%
    dplyr::mutate(range_bin = dplyr::ntile(range, n_bins)) %>%
    dplyr::group_by(range_bin) %>%
    dplyr::summarise(
      mean_range       = mean(range, na.rm = TRUE),
      prop_significant = mean(significant_num, na.rm = TRUE),
      n                = dplyr::n(),
      .groups          = "drop"
    )
}

# --- Empirical bins + model overlay ---
plot_empirical_and_model <- function(df_outcome, pred_df, empirical_df,
                                     thresholds = c(0.25, 0.36, 0.50)) {
  ggplot2::ggplot() +
    ggplot2::geom_point(data = empirical_df,
                        ggplot2::aes(x = mean_range, y = prop_significant, size = n),
                        alpha = 0.8) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(0, 1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title    = paste0("Empirical and Modeled Probability: ", unique(df_outcome$variable)),
      x        = "Cyclical range",
      y        = "Probability of significant GAM",
      size     = "Bin n",
      linetype = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- All-in-one analysis wrapper for one outcome variable ---
analyze_one_outcome <- function(df, target_var,
                                thresholds = c(0.25, 0.36, 0.50),
                                n_bins = 10) {
  df_clean   <- clean_gam_summary(df)
  df_outcome <- get_outcome_df(df_clean, target_var)
  if (nrow(df_outcome) == 0) stop(paste("No rows found for variable:", target_var))

  desc      <- summarize_cyclicity(df_outcome)
  thresh_tab <- threshold_table(df_outcome, thresholds = thresholds)
  mods      <- fit_logistic_models(df_outcome)
  pred_df   <- make_prediction_df(df_outcome  = df_outcome,
                                  model_range   = mods$model_range,
                                  model_range_n = mods$model_range_n)
  emp_df    <- make_empirical_bins(df_outcome, n_bins = n_bins)

  p_hist <- plot_range_histogram(df_outcome)
  p_prob <- plot_significance_probability(df_outcome, pred_df, thresholds = thresholds)
  p_emp  <- plot_empirical_and_model(df_outcome, pred_df, emp_df, thresholds = thresholds)

  list(
    data                        = df_outcome,
    descriptives_overall        = desc$overall,
    descriptives_by_significance = desc$by_significance,
    threshold_table             = thresh_tab,
    model_range                 = mods$model_range,
    model_range_n               = mods$model_range_n,
    model_range_tidy_OR         = mods$tidy_range,
    model_range_n_tidy_OR       = mods$tidy_range_n,
    predictions                 = pred_df,
    empirical_bins              = emp_df,
    plot_histogram              = p_hist,
    plot_probability            = p_prob,
    plot_empirical              = p_emp
  )
}

# --- Extract DRSP variable names from the summary data ---
get_drsp_variables <- function(df) {
  df %>%
    clean_gam_summary() %>%
    dplyr::distinct(variable) %>%
    dplyr::pull(variable) %>%
    .[grepl("^drsp", ., ignore.case = TRUE)]
}

# --- Peak and trough timing distribution plot for significant participants ---
plot_peak_trough_timing <- function(df_sig, tv, axis_meta_list, outcome_label) {
  meta <- axis_meta_list[[tv]]

  timing_long <- dplyr::bind_rows(
    df_sig %>% dplyr::transmute(time = peak_time, type = "Peak (Max)"),
    df_sig %>% dplyr::transmute(time = trough_time, type = "Trough (Min)")
  )

  if (nrow(timing_long) < 4) return(NULL)  # need >= 2 per group for density estimation

  ggplot2::ggplot(timing_long, ggplot2::aes(x = time, fill = type, color = type)) +
    ggplot2::geom_density(alpha = 0.3, adjust = 1.2) +
    ggplot2::geom_rug(alpha = 0.5, length = grid::unit(0.03, "npc")) +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::labs(
      title = paste("Peak & Trough Timing:", outcome_label),
      subtitle = paste0("Among participants with significant GAMs (p < .05); N = ",
                        nrow(df_sig)),
      x = "Cycle Phase", y = "Density", fill = NULL, color = NULL
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::scale_fill_manual(values = c("Peak (Max)" = "#D55E00",
                                          "Trough (Min)" = "#0072B2")) +
    ggplot2::scale_color_manual(values = c("Peak (Max)" = "#D55E00",
                                           "Trough (Min)" = "#0072B2"))
}
```

#### 4.2: Generate Per-Outcome Exploration PDFs

```{r generate-explore-pdfs, results='asis'}
# =========================================================
# GENERATE PER-OUTCOME EXPLORATION PDF REPORTS
# =========================================================
# For each DRSP outcome, generates a landscape PDF containing:
#   1. Summary page: descriptive stats + threshold table + peak/trough timing
#   2. Significance probability plot (logistic curve)
#   3. Histogram of range by significance
#   4. Empirical + modeled probability plot
#   Per time variable:
#   5. Peak & trough timing density (significant participants)
#   6+. Individual GAM trajectory pages (fixed Y axes across people):
#       - Significant with range > 0.33 first, then
#       - Significant with range <= 0.33, then
#       - Not significant with range > 0.33
# =========================================================

# --- Build the exploration summary from in-memory results ---
if (length(all_ids_signif) > 0) {
  gam_summary_df <- dplyr::bind_rows(all_ids_signif)

  # Create output sub-folder for exploration PDFs
  explore_folder <- file.path(output_folder, "Outcome_Exploration_PDFs")
  dir.create(explore_folder, showWarnings = FALSE, recursive = TRUE)

  drsp_vars <- get_drsp_variables(gam_summary_df)
  cat("Found", length(drsp_vars), "DRSP variables to process\n")

  # Thresholds for the exploration
  explore_thresholds <- c(0.10, 0.20, 0.25, 0.36, 0.50, 0.75, 1.00)

  # Plot layout for individual GAM pages
  plots_per_col_exp <- 3
  plots_per_row_exp <- 3
  max_plots_per_page_exp <- plots_per_col_exp * plots_per_row_exp

  # Table themes for summary pages
  summary_table_theme <- gridExtra::ttheme_default(
    base_size = 11,
    core    = list(fg_params = list(hjust = 0, x = 0.05)),
    colhead = list(fg_params = list(hjust = 0, x = 0.05, fontface = "bold"))
  )
  threshold_table_theme <- gridExtra::ttheme_default(
    base_size = 10,
    core    = list(fg_params = list(hjust = 0.5, x = 0.5)),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
  )

  for (target_var in drsp_vars) {
    cat(sprintf("\n=== Processing: %s ===\n", target_var))

    # 1. Run exploration analysis
    result <- tryCatch(
      analyze_one_outcome(gam_summary_df, target_var,
                          thresholds = explore_thresholds,
                          n_bins = 10),
      error = function(e) {
        message("  Skipping ", target_var, ": ", e$message)
        NULL
      }
    )
    if (is.null(result)) next

    # 2. Identify qualifying IDs in three groups, each sorted by range descending:
    #    (a) Significant & range > 0.33
    #    (b) Significant & range <= 0.33
    #    (c) Not significant & range > 0.33
    sig_above_cutoff <- result$data %>%
      dplyr::filter(significant_num == 1 & range > 0.33) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    sig_below_cutoff <- result$data %>%
      dplyr::filter(significant_num == 1 & range <= 0.33) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    nonsig_above_cutoff <- result$data %>%
      dplyr::filter(significant_num == 0 & range > 0.33) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    qualifying_ids <- c(sig_above_cutoff, sig_below_cutoff, nonsig_above_cutoff)

    # 3. Get outcome label
    outcome_label <- if (target_var %in% names(outcomes)) {
      outcomes[[target_var]]
    } else {
      target_var
    }

    cat(sprintf("  N participants: %d | N significant: %d | N qualifying IDs: %d\n",
                result$descriptives_overall$n_participants,
                result$descriptives_overall$n_significant,
                length(qualifying_ids)))

    # 4. Open PDF
    pdf_path <- file.path(explore_folder, paste0("Explore_", target_var, ".pdf"))
    grDevices::pdf(pdf_path, width = 11, height = 8.5)

    # ---- PAGE 1: Summary + Threshold Table ----
    title_grob <- grid::textGrob(
      paste("GAM Cyclicity Exploration:", outcome_label),
      gp = grid::gpar(fontsize = 18, fontface = "bold")
    )
    subtitle_grob <- grid::textGrob(
      paste("Variable:", target_var),
      gp = grid::gpar(fontsize = 12, fontface = "italic")
    )

    n_cpass <- sum(result$data$significant_num == 1 &
                    is.finite(result$data$range) &
                    result$data$range >= 1.5, na.rm = TRUE)

    desc_df <- data.frame(
      Metric = c("Total Participants",
                 "N Significant (p < .05)",
                 "Proportion Significant",
                 "N CPASS-level (p < .05 & range \u2265 1.5)",
                 "Median Range",
                 "Mean Observations per ID"),
      Value = c(
        as.character(result$descriptives_overall$n_participants),
        as.character(result$descriptives_overall$n_significant),
        sprintf("%.1f%%", result$descriptives_overall$prop_significant * 100),
        as.character(n_cpass),
        sprintf("%.2f", median(result$data$range, na.rm = TRUE)),
        sprintf("%.1f", mean(result$data$n, na.rm = TRUE))
      ),
      stringsAsFactors = FALSE
    )

    # Add peak/trough timing summary for significant participants, per time_var
    sig_peak_data <- result$data %>%
      dplyr::filter(significant_num == 1, !is.na(peak_time), !is.na(trough_time))

    if (nrow(sig_peak_data) > 0) {
      timing_by_tv <- sig_peak_data %>%
        dplyr::group_by(best_time_var) %>%
        dplyr::summarise(
          med_peak   = median(peak_time, na.rm = TRUE),
          med_trough = median(trough_time, na.rm = TRUE),
          n_sig      = dplyr::n(),
          .groups    = "drop"
        )
      for (i in seq_len(nrow(timing_by_tv))) {
        tv_label <- timing_by_tv$best_time_var[i]
        suffix <- if (nrow(timing_by_tv) > 1) paste0(" [", tv_label, "]") else ""
        desc_df <- rbind(desc_df, data.frame(
          Metric = c(paste0("Median Peak Time (sig.)", suffix),
                      paste0("Median Trough Time (sig.)", suffix)),
          Value = c(
            sprintf("%.2f (N=%d)", timing_by_tv$med_peak[i], timing_by_tv$n_sig[i]),
            sprintf("%.2f (N=%d)", timing_by_tv$med_trough[i], timing_by_tv$n_sig[i])
          ),
          stringsAsFactors = FALSE
        ))
      }
    }

    desc_tbl <- gridExtra::tableGrob(desc_df, rows = NULL, theme = summary_table_theme)

    thresh_display <- result$threshold_table %>%
      dplyr::mutate(
        threshold        = sprintf("%.2f", threshold),
        n_at_or_above    = as.character(n_at_or_above),
        prop_of_sample   = sprintf("%.1f%%", prop_of_sample * 100),
        n_significant    = as.character(n_significant),
        prop_significant = ifelse(is.na(prop_significant), "N/A",
                                  sprintf("%.1f%%", prop_significant * 100))
      )
    names(thresh_display) <- c("Range Threshold", "N >= Threshold",
                               "% of Sample", "N Significant", "% Significant")

    thresh_tbl <- gridExtra::tableGrob(thresh_display, rows = NULL,
                                       theme = threshold_table_theme)
    thresh_header <- grid::textGrob(
      "Significance Rates at Range Thresholds",
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    note_grob <- if (length(qualifying_ids) > 0) {
      grid::textGrob(
        sprintf("Individual GAM pages follow for %d IDs (sig range>0.33, sig range\u22640.33, then non-sig range>0.33)",
                length(qualifying_ids)),
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    } else {
      grid::textGrob(
        "No qualifying IDs for individual GAM pages",
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    }

    gridExtra::grid.arrange(
      title_grob, subtitle_grob, desc_tbl, thresh_header, thresh_tbl, note_grob,
      ncol = 1,
      # Heights: title, subtitle, desc table, thresh header, thresh table, note
      heights = grid::unit(c(0.10, 0.05, 0.25, 0.08, 0.38, 0.08), "npc")
    )

    # ---- PAGE 2: Significance Probability Plot ----
    print(result$plot_probability)

    # ---- PAGE 3: Histogram ----
    print(result$plot_histogram)

    # ---- PAGE 4: Empirical + Model ----
    print(result$plot_empirical)

    # ---- PAGE 5+: Peak & Trough Timing Summary (per time variable) ----
    for (tv in time_vars) {
      sig_timing_data <- result$data %>%
        dplyr::filter(significant_num == 1, best_time_var == tv,
                      !is.na(peak_time), !is.na(trough_time))

      if (nrow(sig_timing_data) > 0) {
        p_timing <- plot_peak_trough_timing(sig_timing_data, tv, axis_meta,
                                            outcome_label)
        if (!is.null(p_timing)) print(p_timing)
      }
    }

    # ---- Individual GAM trajectory plots for all IDs (per time variable) ----
    # Pre-compute predictions for global y-axis; groups: sig>0.33, sig<=0.33, nonsig>0.33
    for (tv in time_vars) {
      if (length(qualifying_ids) == 0) next

      # Pre-compute all predictions for global y-axis range
      all_fit_results <- list()
      for (qid in qualifying_ids) {
        dat_id <- subset(df, as.character(id) == qid)
        if (nrow(dat_id) == 0) next

        res <- tryCatch(
          fit_outcome(dat_id, target_var, tv, outcomes),
          error = function(e) NULL
        )
        if (!is.null(res)) all_fit_results[[as.character(qid)]] <- res
      }

      groups <- list(
        list(ids = sig_above_cutoff,
             label = "Significant, Range > 0.33"),
        list(ids = sig_below_cutoff,
             label = "Significant, Range \u2264 0.33"),
        list(ids = nonsig_above_cutoff,
             label = "Not Significant, Range > 0.33")
      )

      for (grp in groups) {
        if (length(grp$ids) == 0) next

        id_plots <- list()
        for (qid in grp$ids) {
          res <- all_fit_results[[as.character(qid)]]
          if (is.null(res)) next

          p <- create_annotated_plot(res$pred, tv, axis_meta)
          if (!is.null(p)) {
            id_range <- round(res$sig$range, 2)
            id_n     <- res$sig$n
            id_pval  <- res$sig$p.value
            id_stars <- dplyr::case_when(
              is.finite(id_pval) & id_pval < 0.001 ~ "***",
              is.finite(id_pval) & id_pval < 0.01  ~ "**",
              is.finite(id_pval) & id_pval < 0.05  ~ "*",
              TRUE ~ ""
            )
            p <- p + ggplot2::labs(
              title = sprintf("ID %s %s (range=%.2f, N=%d)", qid, id_stars, id_range, id_n)
            )
            id_plots[[length(id_plots) + 1]] <- p
          }
        }

        if (length(id_plots) > 0) {
          pages <- split(id_plots,
                         ceiling(seq_along(id_plots) / max_plots_per_page_exp))

          for (page_plots in pages) {
            if (length(page_plots) < max_plots_per_page_exp) {
              page_plots <- c(page_plots,
                             rep(list(patchwork::plot_spacer()),
                                 max_plots_per_page_exp - length(page_plots)))
            }
            final_page <- patchwork::wrap_plots(plotlist = page_plots,
                                                ncol = plots_per_col_exp) +
              patchwork::plot_annotation(
                title    = paste(grp$label, ":", outcome_label),
                subtitle = paste("Time variable:", tv),
                theme    = ggplot2::theme(
                  plot.title    = ggplot2::element_text(size = 14, face = "bold"),
                  plot.subtitle = ggplot2::element_text(size = 11)
                )
              )
            print(final_page)
          }
        }
      }
    }

    grDevices::dev.off()
    message(sprintf("  Saved: %s", pdf_path))
  }

  cat("\n=== All exploration PDFs generated ===\n")
  cat("Output folder:", explore_folder, "\n")
} else {
  message("No significance data available; skipping per-outcome exploration PDFs.")
}
```
