title: "UKALC_primary"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document: 
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
---

# UKALC Primary Data Preparation Pipeline

## Overview

Prepares CLEAN datasets for R01AA027990: "Estradiol Effects on Behavioral and Reward Sensitivity to Alcohol across the Menstrual Cycle" (PIs: Mark Fillmore & Michelle Martel, University of Kentucky).

**Pipeline:** Import → Clean → Merge surveys (AM/PM) → Integrate cycle data → Merge hormone assays → Generate diagnostics → Save clean output

**Key Outputs:** `ukalc_YYYYMMDD_clean.rds` (clean dataset with NO derived variables), diagnostic plots

**IMPORTANT CHANGE:** This file now outputs clean data ONLY (no rolling averages, person-standardization, or other derived variables). Derived variables are created in UKALC_03_processing_shared.Rmd AFTER merging with ADHD data to ensure both datasets receive identical processing.


```{r setup}
# Load packages, set options, create output folder
source("UKALC_00_setup.R")

# Create subfolder for data prep outputs
data_prep_folder <- file.path(output_folder, "01_data_preparation")
if (!dir.exists(data_prep_folder)) {
  dir.create(data_prep_folder, recursive = TRUE)
  cat("Created subfolder:", data_prep_folder, "\n")
}
```

```{r source-custom-fxs, quiet=TRUE, message=FALSE, warning=FALSE}
# Load helper functions from scripts/
source("scripts/plot_day_of_week.R")
source("scripts/print_variable_names.R")
source("scripts/count_rows_ids.R")
source("scripts/standardize_tubenumber.R")

cat("Custom functions sourced.\n")
```

## Data Import

```{r import-survey-data}
# Import AM and PM surveys from REDCap (includes historical Qualtrics data)
# Handles UTF-8 BOM, standardizes column names, removes blank rows

# Define Box folder path (UPDATE THIS for your machine)
data_path <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_DAILYSURVEYS/02_data_prep_workspace"

# Helper function: Load CSV, clean column names, standardize ID
load_and_clean_survey <- function(filename) {
  
  # Read CSV with UTF-8 encoding (handles BOM)
  df <- read_csv(
    file.path(data_path, filename), 
    show_col_types = FALSE,
    locale = locale(encoding = "UTF-8"),
    name_repair = "unique_quiet"
  )
  
  # Strip UTF-8 BOM from column names (\ufeffid → id)
  names(df) <- sub("^\\ufeff", "", names(df))
  
  # Standardize all column names to snake_case (MoodSwing → mood_swing)
  df <- df %>% janitor::clean_names()
  
  # Remove empty columns
  df <- df %>%
    select(-matches("^x\\d+$")) %>%  # Remove auto-named columns like "x1", "x2"
    select(where(~ !all(is.na(.))))   # Remove columns with only NA values
  
  # Ensure 'id' column exists after cleaning
  # janitor::clean_names() may have renamed it, so check common variants
  if (!"id" %in% names(df)) {
    # Check for common id column names after cleaning
    id_candidates <- c("subject_id", "participant_id", "record_id", "user_id", "patient_id")
    id_col <- intersect(id_candidates, names(df))
    if (length(id_col) > 0) {
      df <- df %>% rename(id = !!sym(id_col[1]))
      cat(sprintf("  Renamed '%s' to 'id'\n", id_col[1]))
    } else {
      stop(sprintf("Cannot find 'id' column in %s after cleaning. Available columns: %s", 
                   filename, paste(names(df), collapse=", ")))
    }
  }
  
  # Keep only digits in 'id' column (handles various ID formats like "UKALC_101" -> "101")
  df <- df %>%
    mutate(id = str_extract(as.character(id), "\\d+"))
  
  # Remove completely empty rows (where id is NA/empty AND all other non-id columns are NA or whitespace)
  # Count rows before filtering
  n_rows_before <- nrow(df)
  
  df <- df %>%
    filter(!is.na(id) & id != "")
  
  # Report removed rows
  n_removed <- n_rows_before - nrow(df)
  if (n_removed > 0) {
    cat(sprintf("  Removed %d rows with missing/empty id from %s\n", n_removed, filename))
  }
  
  # Standardize hormone column names (E2, P4) - inlined from rename_id_E2_P4.R
  # Standardize estradiol/estrogen-related column names to start with "e2"
  # Only rename columns where the hormone term is the leading identifier (e.g., "estradiol_ng_ml"),
  # avoiding changes to names where it appears later (e.g., "phase2_estrogen_level").
  df <- df %>%
    rename_with(
      ~ gsub("^(estrogen|estradiol|e2)(_?)", "e2\\2", ., ignore.case = TRUE),
      matches("^(estrogen|estradiol|e2)($|_)", ignore.case = TRUE)
    )
  
  # Standardize progesterone-related column names to start with "p4"
  # Only rename columns where the hormone term is the leading identifier.
  df <- df %>%
    rename_with(
      ~ gsub("^(progesterone|p4)(_?)", "p4\\2", ., ignore.case = TRUE),
      matches("^(progesterone|p4)($|_)", ignore.case = TRUE)
    )
  
  # Convert id to factor
  df <- df %>% mutate(id = as.factor(id))
  
  return(df)
}

# ---- 2. Load & Process Datasets ----

am_rc   <- load_and_clean_survey("AM_REDCap_csv.csv")
pm_rc   <- load_and_clean_survey("PM_REDCap_csv.csv")

# ---- 3. Diagnostics ----

# Print quick summary of unique IDs per dataset
datasets <- list(AM_REDCap = am_rc, PM_REDCap = pm_rc)

cat("\n--- Processing Complete ---\n")
for (name in names(datasets)) {
  n_ids <- n_distinct(datasets[[name]]$id)
  n_rows <- nrow(datasets[[name]])
  cat(sprintf("%s: %d rows, %d unique IDs\n", name, n_rows, n_ids))
}

```

## Clean AM Survey Data

```{r clean-am-rc, warning=FALSE}
# Clean morning surveys: alcohol use, cravings, stress, medications
# Parses dates, standardizes craving names (crav_* → crave_*_am), validates 0-10 scale
# Imputes num_drinks=0 when craving reported but drinks missing

# ---- 1. Process Dates ----
# NOTE: janitor::clean_names() has already been applied in load_and_clean_survey()
# so date_time, tube_number, date_rated are already lowercase

# Parse date_time column to create daterated, handling blank/NA values safely
am_rc <- am_rc %>%
  mutate(
    # Only parse non-empty, non-NA date_time values
    daterated = if_else(
      !is.na(date_time) & date_time != "" & !str_detect(date_time, "^\\s*$"),
      as.Date(suppressWarnings(lubridate::mdy_hm(date_time))),
      as.Date(NA_character_)
    )
  ) %>%
  select(-date_time)

# ---- 1a. Remove Pre-existing Cleaned Columns (if any) ----
# Some CSV files may contain pre-cleaned versions of alcohol timing columns
# (e.g., drink_time_start_clean, drink_dur_hrs) from external processing.
# Remove these to avoid duplicates, as we will generate our own cleaned versions.
pre_cleaned_cols <- c("drink_time_start_clean", "drink_dur_hrs")
existing_pre_cleaned <- intersect(pre_cleaned_cols, names(am_rc))

if (length(existing_pre_cleaned) > 0) {
  cat(sprintf("\n=== Removing Pre-existing Cleaned Columns ===\n"))
  cat(sprintf("Found %d pre-cleaned columns in CSV: %s\n", 
              length(existing_pre_cleaned), 
              paste(existing_pre_cleaned, collapse=", ")))
  cat("These will be removed and regenerated using standardized text parsing functions.\n")
  
  am_rc <- am_rc %>%
    select(-any_of(pre_cleaned_cols))
}

# ---- 1b. Standardize Craving Column Names ----
# NOTE: janitor::clean_names() has converted crav_alc_AM -> crav_alc_am already
# Now we just need to change "crav_" prefix to "crave_"
am_rc <- am_rc %>%
  rename_with(~ gsub("^crav_", "crave_", .), starts_with("crav_"))

# Validate transformation: Check if any old-format craving columns remain
remaining_crav <- names(am_rc)[grepl("^crav_", names(am_rc), ignore.case = TRUE)]
if (length(remaining_crav) > 0) {
  warning(sprintf("AM: Found %d craving columns that were not transformed: %s", 
                  length(remaining_crav), paste(remaining_crav, collapse = ", ")))
}

# ---- 1c. Track Missing Data BEFORE Filtering ----
# Document missing data patterns before removing any rows
n_total_rows <- nrow(am_rc)
n_missing_id <- sum(is.na(am_rc$id))
n_missing_daterated <- sum(is.na(am_rc$daterated))
n_missing_both <- sum(is.na(am_rc$id) & is.na(am_rc$daterated))

cat("\n=== AM Survey Missing Data Report (Pre-filtering) ===\n")
cat(sprintf("Total rows: %d\n", n_total_rows))
cat(sprintf("Rows with missing id: %d (%.1f%%)\n", n_missing_id, 100 * n_missing_id / n_total_rows))
cat(sprintf("Rows with missing daterated: %d (%.1f%%)\n", n_missing_daterated, 100 * n_missing_daterated / n_total_rows))
cat(sprintf("Rows with both missing: %d (%.1f%%)\n", n_missing_both, 100 * n_missing_both / n_total_rows))

# ---- 1d. Filter Rows with Missing Keys ----
# Drop observations where id or daterated are missing (cannot be matched in merge)
am_rc <- am_rc %>%
  filter(!is.na(id), !is.na(daterated))

cat(sprintf("Rows after removing missing id/daterated: %d (removed %d rows)\n\n", 
            nrow(am_rc), n_total_rows - nrow(am_rc)))

# ---- 2. Clean Alcohol Variables ----
# Convert drank_alcohol to factor (1/0)
am_rc <- am_rc %>%
  mutate(drank_alcohol = as.factor(drank_alcohol))

# Ensure num_drinks is numeric
am_rc <- am_rc %>% 
  mutate(num_drinks = as.numeric(num_drinks))

# ---- 2a. Text Parsing Functions for Alcohol Duration and Start Time ----
# These functions clean free-text responses from participants about drinking timing
# Adapted from archive/old_scripts/UKALC_textparse.Rmd

#' Clean Time String to HH:MM Format
#'
#' Converts free-text time entries (e.g., "7pm", "around 9:30", "10") to 24-hour HH:MM format.
#' Handles common variations including AM/PM indicators, words like "noon"/"midnight",
#' and missing/invalid entries.
#'
#' @param x Character string representing a time (e.g., "7pm", "21:30", "noon")
#' @return Character string in HH:MM format (24-hour), or NA if unparseable
#'
#' @examples
#' clean_time_single("7pm")        # Returns "19:00"
#' clean_time_single("around 9:30") # Returns "21:30" (assumes PM if no indicator)
#' clean_time_single("noon")       # Returns "12:00"
#' clean_time_single("midnight")   # Returns "00:00"
#' clean_time_single("no drink")   # Returns NA
clean_time_single <- function(x) {
  if (is.na(x) || x == "") return(NA_character_)
  x <- str_to_lower(str_trim(x))
  
  # Handle explicit "no drinking" responses
  if (x %in% c("0", "nan", "none", "na") || str_detect(x, "no drink|didn't|consume")) {
    return(NA_character_)
  }
  
  # Handle special keywords
  if (str_detect(x, "noon")) return("12:00")
  if (str_detect(x, "midnight")) return("00:00")
  
  # Remove ranges and qualifiers, keep first time mentioned
  x <- str_trim(str_split(x, "[-/]| and | for ")[[1]][1])
  x <- str_remove_all(x, "around|at|approx|start|time|about")
  x <- str_trim(x)
  
  # Extract time pattern: digits with optional minutes and AM/PM
  pattern <- "(\\d{1,2})(?:[:.](\\d{2}))?\\s*([ap]\\.?m?\\.?)?"
  match <- str_match(x, pattern)
  
  if (is.na(match[1,1])) return(NA_character_)
  
  h <- as.integer(match[1,2])
  m <- as.integer(match[1,3])
  if (is.na(m)) m <- 0
  ampm <- match[1,4]
  
  # Convert to 24-hour format
  if (!is.na(ampm) && ampm != "") {
    ampm <- str_remove_all(ampm, "\\.")
    if (str_detect(ampm, "p") && h < 12) h <- h + 12
    if (str_detect(ampm, "a") && h == 12) h <- 0
  } else {
    # If no AM/PM indicator, assume PM for drinking hours (evening bias)
    if (h <= 11 && h != 0) h <- h + 12
    if (h == 24) h <- 0
  }
  
  # Validate time bounds
  if (h > 23 || m > 59) return(NA_character_)
  
  return(sprintf("%02d:%02d", h, m))
}

#' Clean Duration String to Numeric Hours
#'
#' Converts free-text duration entries (e.g., "2-3 hours", "1h 30m", "90 minutes") 
#' to numeric hours. Handles ranges (takes midpoint), explicit units, and common
#' text variations.
#'
#' @param x Character string representing duration (e.g., "2 hours", "1:30", "90 min")
#' @return Numeric value in hours, or NA if unparseable
#'
#' @examples
#' clean_duration_single("2 hours")    # Returns 2.0
#' clean_duration_single("1h 30m")     # Returns 1.5
#' clean_duration_single("2-3")        # Returns 2.5 (midpoint)
#' clean_duration_single("90 minutes") # Returns 1.5
#' clean_duration_single("1:30")       # Returns 1.5
#' clean_duration_single("no drink")   # Returns 0.0
clean_duration_single <- function(x) {
  if (is.na(x) || x == "") return(NA_real_)
  x <- str_to_lower(str_trim(x))
  
  # Handle explicit "no drinking" responses - return 0
  if (x %in% c("0", "nan", "none", "na", "no") || 
      str_detect(x, "no drink|didn't|consume|zero")) {
    return(0.0)
  }
  
  # Reject timestamps mistakenly entered as duration (e.g., "4pm")
  if (str_detect(x, "[0-9] ?[ap]\\.?m")) return(NA_real_)
  
  # Standardize text to numeric
  x <- x %>%
    str_replace_all("one", "1") %>% str_replace_all("two", "2") %>%
    str_replace_all("three", "3") %>% str_replace_all("four", "4") %>%
    str_replace_all("five", "5") %>% str_replace_all("half", "0.5") %>%
    str_replace_all("an hour", "1 hour") %>% str_replace_all("a minute", "1 minute") %>%
    str_replace_all("1/2", "0.5") %>%
    str_replace_all("hrs|hr|hours|hour", "h") %>%
    str_replace_all("mins|min|minutes|minute", "m")
  
  # Handle Range Pattern (e.g., "3-4" or "2 to 3") - take midpoint
  range_match <- str_match(x, "^(\\d+(\\.\\d+)?)\\s*[-to]+\\s*(\\d+(\\.\\d+)?)$")
  if (!is.na(range_match[1,1])) {
    return((as.numeric(range_match[1,2]) + as.numeric(range_match[1,4])) / 2)
  }
  
  # Handle Explicit Unit Pattern (e.g., "1h 30m")
  h_val <- 0; m_val <- 0; found_unit <- FALSE
  h_match <- str_match(x, "(\\d+(\\.\\d+)?)\\s*h")
  if (!is.na(h_match[1,1])) { h_val <- as.numeric(h_match[1,2]); found_unit <- TRUE }
  m_match <- str_match(x, "(\\d+(\\.\\d+)?)\\s*m")
  if (!is.na(m_match[1,1])) { m_val <- as.numeric(m_match[1,2]); found_unit <- TRUE }
  if (found_unit) return(h_val + (m_val / 60))
  
  # Handle Colon Pattern (e.g., "1:30" as 1 hour 30 minutes)
  colon_match <- str_match(x, "^(\\d{1,2}):(\\d{2})$")
  if (!is.na(colon_match[1,1])) {
    h <- as.numeric(colon_match[1,2]); m <- as.numeric(colon_match[1,3])
    if (h <= 12) return(h + (m / 60))
  }
  
  # Handle Bare Number (assume hours if < 15, minutes if >= 15)
  num_clean <- str_remove_all(x, "[^0-9\\.]")
  if (num_clean != "" && !is.na(as.numeric(num_clean))) {
    val <- as.numeric(num_clean)
    if (val >= 15) return(val / 60) else return(val)
  }
  
  return(NA_real_)
}

# ---- 2b. Apply Text Parsing to Alcohol Variables ----
# Clean time_start_alcohol and dur_alcohol using the parsing functions
cat("\n=== Cleaning Alcohol Duration and Start Time ===\n")

# Count non-missing values before cleaning
n_start_before <- sum(!is.na(am_rc$time_start_alcohol))
n_dur_before <- sum(!is.na(am_rc$dur_alcohol))

# Apply cleaning functions
am_rc <- am_rc %>%
  mutate(
    time_start_alcohol_clean = map_chr(time_start_alcohol, clean_time_single),
    dur_alcohol_hours = map_dbl(dur_alcohol, clean_duration_single)
  )

# Count non-missing values after cleaning
n_start_after <- sum(!is.na(am_rc$time_start_alcohol_clean))
n_dur_after <- sum(!is.na(am_rc$dur_alcohol_hours))

cat(sprintf("time_start_alcohol: %d non-missing → %d cleaned (%.1f%% success rate)\n",
            n_start_before, n_start_after, 
            100 * n_start_after / max(n_start_before, 1)))
cat(sprintf("dur_alcohol: %d non-missing → %d cleaned (%.1f%% success rate)\n",
            n_dur_before, n_dur_after,
            100 * n_dur_after / max(n_dur_before, 1)))

# Show example conversions for validation
if (n_start_before > 0) {
  cat("\nExample time conversions:\n")
  examples_time <- am_rc %>%
    filter(!is.na(time_start_alcohol)) %>%
    select(time_start_alcohol, time_start_alcohol_clean) %>%
    distinct() %>%
    head(5)
  print(examples_time)
}

if (n_dur_before > 0) {
  cat("\nExample duration conversions:\n")
  examples_dur <- am_rc %>%
    filter(!is.na(dur_alcohol)) %>%
    select(dur_alcohol, dur_alcohol_hours) %>%
    distinct() %>%
    head(5)
  print(examples_dur)
}

# ---- Quality Check: Validate Parsed Values ----
# Check for unreasonable duration values that may indicate parsing errors
extreme_durations <- am_rc %>%
  filter(!is.na(dur_alcohol_hours)) %>%
  filter(dur_alcohol_hours > 12 | dur_alcohol_hours < 0) %>%
  select(id, daterated, dur_alcohol, dur_alcohol_hours)

if (nrow(extreme_durations) > 0) {
  cat("\n⚠️  WARNING: Found", nrow(extreme_durations), "drinking episodes with unusual duration (>12 hours or <0):\n")
  print(extreme_durations)
  cat("Review these values manually to ensure they were parsed correctly.\n")
}

# Check for extreme number of drinks that may be data entry errors
extreme_drinks <- am_rc %>%
  filter(!is.na(num_drinks_today)) %>%
  filter(num_drinks_today > 20) %>%
  select(id, daterated, num_drinks_today)

if (nrow(extreme_drinks) > 0) {
  cat("\n⚠️  WARNING: Found", nrow(extreme_drinks), "drinking episodes with >20 drinks:\n")
  print(extreme_drinks)
  cat("These values may be data entry errors. Review and consider capping if implausible.\n")
}

# ---- 3. Clean Craving Variables ----
# Craving columns have been standardized to: crave_alc_am, crave_nic_am, crave_mar_am, etc.
# (renamed from crav_*_AM to crave_*_am in section 1b above)
# NOTE: The _am suffix is already present from section 1b

# Record NA rates before transformation
crave_cols_pre <- names(am_rc)[grepl("^crave_.*_am$", names(am_rc), ignore.case = TRUE)]
if (length(crave_cols_pre) > 0) {
  crave_pre <- am_rc %>% 
    summarise(across(matches("^crave_.*_am$"), ~ mean(is.na(.))))
  cat("AM Craving Variables NA proportions (before cleaning):\n")
  print(crave_pre)
}

# Clean craving variables - extract leading digits if text format present
# Handles formats like: "5 Moderate Urge", "10", 10
am_rc <- am_rc %>%
  mutate(across(matches("^crave_.*_am$"), ~ suppressWarnings(as.integer(stringr::str_extract(as.character(.), "^\\d+")))))

# Record NA rates after transformation
if (length(crave_cols_pre) > 0) {
  crave_post <- am_rc %>% 
    summarise(across(matches("^crave_.*_am$"), ~ mean(is.na(.))))
  cat("\nAM Craving Variables NA proportions (after cleaning):\n")
  print(crave_post)
  
  # Report if cleaning introduced NAs
  for(col in crave_cols_pre) {
    na_before <- crave_pre[[col]]
    na_after <- crave_post[[col]]
    if(na_after > na_before) {
      cat(sprintf("WARNING: %s had %.1f%% NAs increase from %.1f%% to %.1f%% due to text extraction\n", 
                  col, 100*(na_after - na_before), 100*na_before, 100*na_after))
    }
  }
}

# Validate craving values are in expected range (0-10)
crave_cols <- names(am_rc)[grepl("^crave_.*_am$", names(am_rc), ignore.case = TRUE)]
if (length(crave_cols) > 0) {
  for (col in crave_cols) {
    out_of_range <- am_rc[[col]][!is.na(am_rc[[col]]) & (am_rc[[col]] < 0 | am_rc[[col]] > 10)]
    if (length(out_of_range) > 0) {
      warning(sprintf("AM RC: %s has %d values outside expected 0-10 range", col, length(out_of_range)))
    }
  }
}

# ---- 4. Imputation Logic with Documentation ----
# If crave_alc_am is not missing but num_drinks is missing, impute num_drinks = 0
# RATIONALE: Craving question comes after drinking question in survey flow.
# If participant answered craving, they saw drinking question but didn't report drinks.
# This likely means no drinking occurred (rather than technical skip).
if("crave_alc_am" %in% names(am_rc) && "num_drinks" %in% names(am_rc)) {
  n_imputed <- sum(!is.na(am_rc$crave_alc_am) & is.na(am_rc$num_drinks))
  if(n_imputed > 0) {
    cat(sprintf("\n=== Imputation: Setting num_drinks = 0 for %d rows where crave_alc_am exists but num_drinks is NA ===\n", n_imputed))
    am_rc$num_drinks[!is.na(am_rc$crave_alc_am) & is.na(am_rc$num_drinks)] <- 0
  }
}

# ---- 5. Select Final Columns and Standardize Names ----
# Keep all relevant AM variables
# NOTE: Craving variables already have _am suffix from section 1b (crave_alc_am, crave_nic_am, etc.)

# Standardize TubeNumber using helper function for consistency with hormone data
am_rc <- standardize_tubenumber(am_rc)

am_rc <- am_rc %>% 
  select(id, daterated, 
         # Tube number for hormone data matching
         any_of("TubeNumber"),
         # Alcohol variables (raw and cleaned)
         num_drinks, drank_alcohol, felt_drunk, 
         time_start_alcohol, time_start_alcohol_clean,
         dur_alcohol, dur_alcohol_hours,
         # Craving variables (all substances) - with _am suffix
         crave_alc_am, crave_nic_am, crave_mar_am, crave_coc_am, crave_benz_am, 
         crave_hall_am, crave_opiate_am, crave_oth_am,
         # Stress variables
         had_stress, stress_work_school, stress_social, stress_health, 
         stress_financial, stress_logistics, stress_study_related,
         # Medication/drug use variables
         took_meds, drug_pain_nsaid, drug_cold_allergy, drug_antibiotic, 
         drug_psych_sleep, drug_gi, drug_steroid, drug_spironolactone, drug_supplement,
         # Other substance use
         used_nicotine, used_marijuana, used_cocaine, used_benzos, 
         used_hallucinogens, used_opiates)

# ---- 6. Visualization and Diagnostics ----
# Visualize Missing Data
cat("\n=== AM Survey Missing Data Pattern (Final) ===\n")
#vis_miss(am_rc) 

# Print summary
cat("\n=== AM Survey Cleaning Complete ===\n")
cat("Variables: All in snake_case format\n")
cat("Craving variables: crave_alc_am, crave_nic_am, crave_mar_am, etc. (validated 0-10 range, _am suffix already present)\n")
cat("Alcohol variables: num_drinks (numeric), drank_alcohol (factor), felt_drunk\n")
cat("Alcohol timing variables: time_start_alcohol_clean (HH:MM), dur_alcohol_hours (numeric hours)\n")
cat("Stress variables: had_stress, stress_work_school, stress_social, etc.\n")
cat("Medication variables: took_meds, drug_pain_nsaid, drug_cold_allergy, etc.\n")
cat("Substance use: used_nicotine, used_marijuana, used_cocaine, etc.\n\n")

count_rows_ids(am_rc)
```

```{r create-merged-amdf}
# Fill missing dates in AM data: creates row for each day between first/last survey
# Ensures craving variables have "_am" suffix for merge

# Standardize dates and remove empty rows
amdf <- am_rc %>%
  mutate(
    daterated = as_date(daterated),
    id = as.character(id)
  ) %>%
  filter(!is.na(id))

# Fill missing dates (creates row for each day between first and last survey)
amdf_complete <- amdf %>%
  group_by(id) %>%
  filter(any(!is.na(daterated))) %>% 
  complete(daterated = seq(min(daterated, na.rm = TRUE), 
                           max(daterated, na.rm = TRUE), 
                           by = "day")) %>%
  ungroup()

# Verify craving variables have "_am" suffix (safety check)
amdf_complete <- amdf_complete %>%
  rename_with(~ ifelse(grepl("_(am|pm)$", .), ., paste0(., "_am")), starts_with("crave_")) %>%
  distinct()

cat("Final AM DataFrame Rows:", nrow(amdf_complete), "\n")
count_rows_ids(amdf_complete)

vis_miss(amdf_complete)
```

## Clean PM Survey Data

```{r clean-pm-rc, warning=FALSE}
# Clean evening surveys: mood (DRSP), impulse control (bdefs), cravings
# Parses dates (handles "before noon" → previous day), standardizes names (crav_* → crave_*_pm)
# Recodes DRSP (1-6 scale) and validates cravings (0-10 scale)
# Note: Temporary datepart/timepart/b4noon kept for duplicate removal, dropped before merge

# ---- 0. Define Variable Lists ----
# These variable lists are needed for recoding DRSP and bdefs scales below
# Following refactoring guidelines: each chunk defines its own variable lists

drsp_vars <- c(
  "dep", "hopeless", "worthless", "anxious", "mood_swing", "rej_sens", 
  "angry", "conflicts", "loss_interest", "diff_conc", "tired", 
  "over_eat", "sp_food_crav", "high_sleep", "diff_sleep", "overwhelm", 
  "loss_control", "breast_tend", "breast_swell", "headache", 
  "jm_pain", "loss_prod", "int_relat"
)

bdefs_vars <- c(
  "diff_inhib", "diff_stop", "diff_change", "immed_payoff", 
  "res_urge", "diff_long_rew", "diff_res_rew"
)

# ---- 1. Process Dates ----
# NOTE: janitor::clean_names() has already been applied in load_and_clean_survey()
# so date_time is already lowercase and ready to use

# Parse date_time column to create daterated
# The raw data has been fixed, so we can now parse dates uniformly
# Handles multiple date formats flexibly using lubridate::parse_date_time
pm_rc <- pm_rc %>%
  mutate(
    # Only parse non-empty, non-NA date_time values
    daterated = if_else(
      ! is.na(date_time) & date_time != "" & ! str_detect(date_time, "^\\s*$"),
      as.Date(coalesce(
        # Try mdy_hms first (handles formats with seconds like "8/12/2021 7:22:00PM")
        suppressWarnings(lubridate::mdy_hms(date_time)),
        # Fall back to mdy_hm (handles "8/30/21 20:14" format without seconds)
        suppressWarnings(lubridate::mdy_hm(date_time))
      )),
      as.Date(NA_character_)
    )
  )


# Save "before noon" completions as day prior
# Handle blank date_time values that would cause str_split to fail
pm_rc <- pm_rc %>%
  mutate(
    date_time_clean = if_else(is.na(date_time) | date_time == "", "___MISSING_DATE___ ___MISSING_TIME___", date_time)
  )

pm_rc[c('datepart', 'timepart')] <- str_split_fixed(pm_rc$date_time_clean, ' ', 2)
pm_rc <- pm_rc %>% select(-date_time, -date_time_clean)

# Parse timepart, handling NA values
pm_rc <- pm_rc %>%
  mutate(
    timepart = suppressWarnings(lubridate::hm(timepart)),
    b4noon = if_else(!is.na(timepart) & timepart < lubridate::hm("12:00"), 1, 0, missing = 0)
  )

pm_rc <- pm_rc %>%
  mutate(daterated = case_when(
    b4noon == 1 & !is.na(daterated) ~ daterated - 1,
    TRUE ~ daterated
  ))

# ---- 1a. Standardize Craving Column Names ----
# NOTE: janitor::clean_names() has converted crav_alc_PM -> crav_alc_pm already
# Now we just need to change "crav_" prefix to "crave_"
pm_rc <- pm_rc %>%
  rename_with(~ gsub("^crav_", "crave_", .), starts_with("crav_"))

# Validate transformation: Check if any old-format craving columns remain
remaining_crav <- names(pm_rc)[grepl("^crav_", names(pm_rc), ignore.case = TRUE)]
if (length(remaining_crav) > 0) {
  warning(sprintf("PM: Found %d craving columns that were not transformed: %s", 
                  length(remaining_crav), paste(remaining_crav, collapse = ", ")))
}

# ---- 2. Remove Rows with Blank Data ----
# Identify rows where all non-key fields are NA or empty/whitespace
# Key fields: id, daterated, datepart, timepart, b4noon
# This removes the "blank" rows seen in the PM export

n_total_rows <- nrow(pm_rc)
n_missing_id <- sum(is.na(pm_rc$id))
n_missing_daterated <- sum(is.na(pm_rc$daterated))

cat("\n=== PM Survey Missing Data Report (Pre-filtering) ===\n")
cat(sprintf("Total rows: %d\n", n_total_rows))
cat(sprintf("Rows with missing id: %d (%.1f%%)\n", n_missing_id, 100 * n_missing_id / n_total_rows))
cat(sprintf("Rows with missing daterated: %d (%.1f%%)\n", n_missing_daterated, 100 * n_missing_daterated / n_total_rows))

# Define key columns to exclude from "all NA" check
key_cols <- c("id", "daterated", "datepart", "timepart", "b4noon")
data_cols <- setdiff(names(pm_rc), key_cols)

# Remove rows where:
# 1. id is missing (can't be matched), OR
# 2. All data columns are NA/empty (blank survey row)
# Note: We keep rows with missing daterated if they have actual data,
# as these may have date parsing issues but still contain valid survey responses
pm_rc <- pm_rc %>%
  filter(
    !is.na(id) &
    # At least one data column must have a non-NA value
    rowSums(!is.na(select(., all_of(data_cols)))) > 0
  )

n_removed <- n_total_rows - nrow(pm_rc)
cat(sprintf("Rows after removing missing id and blank data rows: %d (removed %d rows)\n", 
            nrow(pm_rc), n_removed))

# Check for rows with missing daterated that were kept
n_kept_missing_daterated <- sum(is.na(pm_rc$daterated))
if (n_kept_missing_daterated > 0) {
  affected_ids <- unique(pm_rc$id[is.na(pm_rc$daterated)])
  cat(sprintf("⚠ WARNING: %d rows have missing daterated but were kept because they contain data\n", 
              n_kept_missing_daterated))
  cat("  These rows may not merge correctly. Investigate date_time values for parsing issues.\n")
  if (length(affected_ids) <= 10) {
    cat("  Affected IDs:", paste(affected_ids, collapse = ", "), "\n")
  } else {
    cat("  Affected IDs (first 10):", paste(head(affected_ids, 10), collapse = ", "), 
        sprintf("... and %d more\n", length(affected_ids) - 10))
  }
  cat("  Next steps:\n")
  cat("  1. Check original date_time format in raw data for these IDs\n")
  cat("  2. Verify date_time values are in 'M/D/YY H:MM' format (e.g., '8/30/21 20:14')\n")
  cat("  3. Look for blank, malformed, or non-standard date strings\n")
}
cat("\n")

# ---- 3. Variable Names Already Cleaned ----
# NOTE: janitor::clean_names() in load_and_clean_survey() has already:
# - Converted Dep -> dep, MoodSwing -> mood_swing, RejSens -> rej_sens
# - Converted Diff_Inhib -> diff_inhib, Diff_Stop -> diff_stop, etc.
# So we don't need manual name mappings anymore!

cat("Variable names already standardized by janitor::clean_names():\n")
cat("  - DRSP: dep, hopeless, mood_swing, rej_sens, etc.\n")
cat("  - bdefs: diff_inhib, diff_stop, diff_change, etc.\n")
cat("  - Craving: crave_alc_pm, crave_nic_pm, etc.\n\n")

# ---- 4. Clean Craving Variables ----
# Craving columns should now be: crave_alc_pm, crave_nic_pm, etc.
# (renamed from crav_* in section 1a above)

# Record NA rates before transformation
crave_cols_pre <- names(pm_rc)[grepl("^crave_.*_pm$", names(pm_rc), ignore.case = TRUE)]
if (length(crave_cols_pre) > 0) {
  crave_pre <- pm_rc %>% 
    summarise(across(matches("^crave_.*_pm$"), ~ mean(is.na(.))))
  cat("\nPM Craving Variables NA proportions (before cleaning):\n")
  print(crave_pre)
}

# Clean craving variables - extract leading digits if text format present
# This handles both:
#   - Numeric values: 0, 1, 2, ..., 10 (kept as-is)
#   - Text labels: "5 Moderate Urge" (extracts the leading "5" → 5)
# The str_extract gets the first digits, as.integer converts to numeric
pm_rc <- pm_rc %>%
  mutate(across(matches("^crave_.*_pm$"), ~ suppressWarnings(as.integer(stringr::str_extract(as.character(.), "^\\d+")))))

# Record NA rates after transformation
if (length(crave_cols_pre) > 0) {
  crave_post <- pm_rc %>% 
    summarise(across(matches("^crave_.*_pm$"), ~ mean(is.na(.))))
  cat("\nPM Craving Variables NA proportions (after cleaning):\n")
  print(crave_post)
}

# Validate craving values are in expected range (0-10)
crave_cols <- names(pm_rc)[grepl("^crave_.*_pm$", names(pm_rc), ignore.case = TRUE)]
if (length(crave_cols) > 0) {
  for (col in crave_cols) {
    out_of_range <- pm_rc[[col]][!is.na(pm_rc[[col]]) & (pm_rc[[col]] < 0 | pm_rc[[col]] > 10)]
    if (length(out_of_range) > 0) {
      warning(sprintf("PM RC: %s has %d values outside expected 0-10 range", col, length(out_of_range)))
    }
  }
}

# ---- 5. Recode DRSP Variables (Only if Character/String) ----
# DRSP uses 6-point scale: Not at All(1) to Extreme(6)
# The example data has numeric values (1-6), so we should preserve those
# Only recode if values are character/string labels

drsp_cols_present <- intersect(drsp_vars, names(pm_rc))
if(length(drsp_cols_present) > 0) {
  # Check data type: numeric values (normal case) or character strings (rare case)
  # Data typically comes in as numeric 1-6, but handle strings just in case
  drsp_is_char <- sapply(pm_rc[drsp_cols_present], function(x) is.character(x) | is.factor(x))
  drsp_char_cols <- names(drsp_is_char)[drsp_is_char]
  
  if(length(drsp_char_cols) > 0) { # Only if strings detected
    cat(sprintf("\nRecoding %d DRSP character variables to numeric scale (1-6)\n", length(drsp_char_cols)))
    
    # Define the recoding as a named vector
    recode_levels <- c(
      "Not at All" = 1, 
      "Minimal" = 2, 
      "Mild" = 3, 
      "Moderate" = 4, 
      "Severe" = 5, 
      "Extreme" = 6
    )
    
    # Identify any unexpected DRSP response levels before recoding
    unexpected_levels <- lapply(pm_rc[drsp_char_cols], function(x) {
      vals <- unique(as.character(x))
      vals[!(is.na(vals) | vals %in% names(recode_levels))]
    })
    cols_with_unexpected <- names(unexpected_levels)[vapply(unexpected_levels, length, integer(1L)) > 0L]
    if (length(cols_with_unexpected) > 0) {
      cat("\nWarning: Unexpected DRSP levels detected in the following variables (will be set to NA):\n")
      for (nm in cols_with_unexpected) {
        cat(sprintf("  - %s: %s\n", nm, paste(unexpected_levels[[nm]], collapse = ", ")))
      }
    }
    
    pm_rc <- pm_rc %>%
      mutate(across(all_of(drsp_char_cols), ~ as.numeric(recode(as.character(.), !!!recode_levels, .default = NA_real_))))
  } else { # No character columns - data is numeric (normal case)
    cat(sprintf("\nDRSP variables (%d found) are already numeric; skipping recoding\n", length(drsp_cols_present)))
  }
  
  # Ensure all DRSP columns are numeric
  pm_rc <- pm_rc %>%
    mutate(across(all_of(drsp_cols_present), ~ as.numeric(.)))
}

# view id, daterated, and diff_inhib columns for checking
# pm_rc %>% select(id, daterated, diff_inhib) %>% View()

# Set bdefs columns to numeric
  
pm_rc <- pm_rc %>%
  mutate(across(all_of(bdefs_vars), ~ as.numeric(.)))


# ---- 6a. Standardize TubeNumber (if present) ----
# Use helper function to standardize tube_number to TubeNumber
# This ensures consistency with hormone data and AM survey
pm_rc <- standardize_tubenumber(pm_rc)

# ---- 7. Select Final Columns ----
# Keep all relevant columns
# NOTE: All column names are now in snake_case thanks to janitor::clean_names()
cols_to_keep <- c("id", "daterated", "datepart", "timepart", "b4noon")

# Add TubeNumber if it exists (for hormone data matching)
if ("TubeNumber" %in% names(pm_rc)) {
  cols_to_keep <- c(cols_to_keep, "TubeNumber")
}

# Add bdefs, DRSP, and craving variables using vectorized approach
all_scale_vars <- c(bdefs_vars, drsp_vars, 
                    "crave_alc_pm", "crave_nic_pm", "crave_mar_pm", "crave_coc_pm", 
                    "crave_benzo_pm", "crave_hall_pm", "crave_opiate_pm", "crave_oth_pm")
cols_to_keep <- c(cols_to_keep, intersect(all_scale_vars, names(pm_rc)))

# Add any other description columns that exist
desc_cols <- names(pm_rc)[grepl("_description$", names(pm_rc))]
if(length(desc_cols) > 0) {
  cols_to_keep <- c(cols_to_keep, desc_cols)
}

# Use unique() to ensure no duplicate columns
pm_rc <- pm_rc %>% select(all_of(unique(cols_to_keep)))

# ---- 8. Diagnostics ----
# Visualize Missing Data
cat("\n=== PM Survey Missing Data Pattern (Final) ===\n")
##vis_miss(pm_rc) 

# Print summary
cat("\n=== PM Survey Cleaning Complete ===\n")
cat("Variables: All in snake_case format\n")
cat(sprintf("DRSP variables (%d): dep, hopeless, anxious, mood_swing, etc. (1-6 scale)\n", length(drsp_cols_present)))
cat("Craving variables: crave_alc_pm, crave_nic_pm, crave_mar_pm, etc. (0-10 scale)\n\n")

count_rows_ids(pm_rc)

#vis_miss(pm_rc)

#View(pm_rc)
```

```{r create-merged-pmdf}
# Prepare PM data for merge: remove duplicates, validate structure

pmdf <- pm_rc

# Check for duplicates and remove
pmdf <- pmdf %>%
  distinct()

# Print variable summary
cat("\n=== PM Survey Data Ready for Merge ===\n")
cat("All variables in snake_case with _pm suffix on cravings\n")
cat("bdefs variables (snake_case): diff_inhib, diff_stop, diff_change, etc.\n")
cat("DRSP variables (snake_case): dep, hopeless, anxious, mood_swing, etc.\n")
cat("Craving variables: crave_alc_pm, crave_nic_pm, crave_mar_pm, etc.\n\n")

count_rows_ids(pmdf)

print.variable.names(pmdf)

#vis_miss(pmdf)
```

## Remove Duplicates & Apply Exclusions

```{r manual-exclusions-changes}
# Remove ~55 confirmed duplicate PM surveys (identified by study staff via datepart/timepart)
# Exclude all data from IDs 125, 127 (insufficient data quality)
# Drop temporary date variables (datepart, timepart, b4noon) after removal

#ID 106
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-08" & pmdf$timepart == "10H 0M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-13" & pmdf$timepart == "4H 57M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-19" & pmdf$timepart == "4H 52M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-22" & pmdf$timepart == "10H 32M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-23" & pmdf$timepart == "13H 8M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 106 & pmdf$datepart == "2021-05-24" & pmdf$timepart == "7H 18M 0S"), ]
#ID 107
pmdf <- pmdf[!(pmdf$id == 107 & pmdf$datepart == "2021-05-29" & pmdf$timepart == "7H 18M 0S"), ]

#ID 109 
pmdf <- pmdf[!(pmdf$id == 109 & pmdf$datepart == "2021-06-22" & pmdf$timepart == "20H 30M 0S"), ]

#ID 113 
pmdf <- pmdf[!(pmdf$id == 113 & pmdf$datepart == "2021-06-29" & pmdf$timepart == "21H 25M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 113 & pmdf$datepart == "2021-07-02" & pmdf$timepart == "23H 03M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 113 & pmdf$datepart == "2021-07-05" & pmdf$timepart == "0H 21M 0S"), ]
#ID 114
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-03" & pmdf$timepart == "5H 57M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-05" & pmdf$timepart == "4H 46M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-09" & pmdf$timepart == "21H 43M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-09" & pmdf$timepart == "21H 44M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-11" & pmdf$timepart == "20H 46M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 114 & pmdf$datepart == "2021-07-14" & pmdf$timepart == "7H 48M 0S"), ]
#ID 115 
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-06-26" & pmdf$timepart == "8H 20M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-06-28" & pmdf$timepart == "22H 1M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-01" & pmdf$timepart == "19H 52M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-03" & pmdf$timepart == "3H 5M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-09" & pmdf$timepart == "21H 52M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-09" & pmdf$timepart == "21H 54M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-09" & pmdf$timepart == "21H 57M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-10" & pmdf$timepart == "8H 37M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-10" & pmdf$timepart == "8H 40M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "6H 48M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "6H 56M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "6H 59M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "7H 3M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "7H 6M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "7H 8M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "7H 10M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 115 & pmdf$datepart == "2021-07-19" & pmdf$timepart == "7H 12M 0S"), ]

#ID 118 
pmdf <- pmdf[!(pmdf$id == 118 & pmdf$datepart == "2021-08-08" & pmdf$timepart == "12H 26M 0S"), ]
#ID 119 
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-03" & pmdf$timepart == "21H 58M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-08" & pmdf$timepart == "20H 50M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-13" & pmdf$timepart == "5H 37M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-20" & pmdf$timepart == "23H 28M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-22" & pmdf$timepart == "5H 40M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-24" & pmdf$timepart == "16H 23M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 119 & pmdf$datepart == "2021-08-27" & pmdf$timepart == "21H 40M 0S"), ]
#ID 136
pmdf <- pmdf[!(pmdf$id == 136 & pmdf$datepart == "2021-11-29" & pmdf$timepart == "17H 10M 0S"), ]
#ID 138
pmdf <- pmdf[!(pmdf$id == 138 & pmdf$datepart == "2022-02-06" & pmdf$timepart == "12H 21M 0S"), ]
#ID 143
pmdf <- pmdf[!(pmdf$id == 143 & pmdf$datepart == "2022-01-31" & pmdf$timepart == "17H 04M 0S"), ]
#ID 149
pmdf <- pmdf[!(pmdf$id == 149 & pmdf$datepart == "2022-05-25" & pmdf$timepart == "0H 42M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 149 & pmdf$datepart == "2022-06-10" & pmdf$timepart == "12H 20M 0S"), ]
#ID 150
pmdf <- pmdf[!(pmdf$id == 150 & pmdf$datepart == "2022-06-07" & pmdf$timepart == "21H 34M 0S"), ]
#ID 153
pmdf <- pmdf[!(pmdf$id == 153 & pmdf$datepart == "2022-06-23" & pmdf$timepart == "8H 24M 0S"), ]
#ID 159
pmdf <- pmdf[!(pmdf$id == 159 & pmdf$datepart == "2022-07-17" & pmdf$timepart == "6H 40M 0S"), ]
#ID 168
pmdf <- pmdf[!(pmdf$id == 168 & pmdf$datepart == "2022-09-25" & pmdf$timepart == "20H 23M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 168 & pmdf$datepart == "2022-10-09" & pmdf$timepart == "18H 47M 0S"), ]
#ID 171
pmdf <- pmdf[!(pmdf$id == 171 & pmdf$datepart == "2022-10-07" & pmdf$timepart == "0H 10M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 171 & pmdf$datepart == "2022-10-13" & pmdf$timepart == "12H 34M 0S"), ]
pmdf <- pmdf[!(pmdf$id == 171 & pmdf$datepart == "2022-10-31" & pmdf$timepart == "20H 36M 0S"), ]
#ID 180
pmdf <- pmdf[!(pmdf$id == 180 & pmdf$datepart == "2023-04-15" & pmdf$timepart == "22H 7M 0S"), ]
#ID 181
pmdf <- pmdf[!(pmdf$id == 181 & pmdf$datepart == "2023-03-22" & pmdf$timepart == "19H 13M 0S"), ]
#ID 185
pmdf <- pmdf[!(pmdf$id == 185 & pmdf$datepart == "2023-05-29" & pmdf$timepart == "23H 34M 0S"), ]

# ---- Drop id 125 and 127 from both datasets ----
amdf_complete <- amdf_complete %>%
  filter(id != 125, id != 127)
pmdf <- pmdf %>%
  filter(id != 125, id != 127)

##vis_miss(pmdf)

# Check for duplicates and remove
pmdf <- pmdf %>%
  distinct()

# ---- Drop temporary date variables before merge ----
# datepart, timepart, and b4noon were needed for duplicate removal above
# Now drop them before the AM/PM merge to keep only daterated
cat("\n=== Dropping temporary date variables (datepart, timepart, b4noon) before merge ===\n")

# Validation: Check which temporary variables exist before dropping
temp_vars <- c("datepart", "timepart", "b4noon")
existing_temp_vars <- intersect(temp_vars, names(pmdf))
missing_temp_vars <- setdiff(temp_vars, names(pmdf))

if (length(existing_temp_vars) > 0) {
  cat(sprintf("Dropping %d temporary variable(s): %s\n", 
              length(existing_temp_vars), 
              paste(existing_temp_vars, collapse = ", ")))
  pmdf <- pmdf %>%
    select(-any_of(existing_temp_vars))
} else {
  cat("No temporary variables to drop (already cleaned)\n")
}

if (length(missing_temp_vars) > 0) {
  cat(sprintf("Note: %d expected variable(s) not found: %s\n", 
              length(missing_temp_vars), 
              paste(missing_temp_vars, collapse = ", ")))
}

cat("\nVariables remaining in pmdf:\n")
print(names(pmdf))

count_rows_ids(pmdf)

# Note: Lots of replacements of duplicates etc happen here. 

```

## Visualize Missing Data Pre-Merge

```{r visualize-missingness-pre-am-pm-merge}
# Visualize missingness by AM/PM after cleaning 

# create separate #vis_miss for am and pm 
cat("\n=== AM Survey Missing Data Pattern ===\n")
vis_miss(amdf_complete)
cat("\n=== PM Survey Missing Data Pattern ===\n")
vis_miss(pmdf)

#View(amdf_complete)

```

## Merge AM/PM Surveys

```{r merge_am_pm, warning=FALSE}
# Full join of AM and PM surveys on id + daterated
# Preserves AM-only days (alcohol data), PM-only days (mood data), and complete days
# Validates merge keys, documents missing data patterns

# ---- 1. Pre-Merge Validation ----
cat("\n=== Pre-Merge Validation ===\n")

# Check for duplicate keys in AM data
am_dups <- amdf_complete %>%
  group_by(id, daterated) %>%
  filter(n() > 1) %>%
  ungroup()

if(nrow(am_dups) > 0) {
  warning(sprintf("Found %d duplicate id-date combinations in AM data!", nrow(am_dups)))
  print(am_dups %>% select(id, daterated) %>% distinct())
} else {
  cat("✓ No duplicate keys in AM data\n")
}

# Check for duplicate keys in PM data
pm_dups <- pmdf %>%
  group_by(id, daterated) %>%
  filter(n() > 1) %>%
  ungroup()

if(nrow(pm_dups) > 0) {
  warning(sprintf("Found %d duplicate id-date combinations in PM data!", nrow(pm_dups)))
  print(pm_dups %>% select(id, daterated) %>% distinct())
} else {
  cat("✓ No duplicate keys in PM data\n")
}

# merge duplicates on id and daterated preferring non-NA values
amdf_complete <- amdf_complete %>%
  group_by(id, daterated) %>%
  summarise(across(everything(), ~ ifelse(all(is.na(.)), NA, first(na.omit(.)))), .groups = "drop")

#View(amdf_complete)

# ---- 2. Perform Full Join ----
cat("\n=== Performing Full Join ===\n")
cat(sprintf("AM data: %d rows, %d unique IDs\n", nrow(amdf_complete), n_distinct(amdf_complete$id)))
cat(sprintf("PM data: %d rows, %d unique IDs\n", nrow(pmdf), n_distinct(pmdf$id)))

df <- amdf_complete %>%
  full_join(pmdf, by = c("id", "daterated"), suffix = c("_am_dup", "_pm_dup"))

cat(sprintf("\nMerged data: %d rows, %d unique IDs\n", nrow(df), n_distinct(df$id)))

# ---- 3. Analyze Missing Data Patterns ----
cat("\n=== Missing Data Pattern Analysis ===\n")

# Identify data completeness patterns
df <- df %>%
  mutate(
    has_am_data = !is.na(num_drinks) | !is.na(crave_alc_am),
    has_pm_data = !is.na(dep) | !is.na(crave_alc_pm),
    data_pattern = case_when(
      has_am_data & has_pm_data ~ "Both_AM_PM",
      has_am_data & !has_pm_data ~ "AM_only",
      !has_am_data & has_pm_data ~ "PM_only",
      TRUE ~ "Neither"
    )
  )

# Summarize patterns
pattern_summary <- df %>%
  group_by(data_pattern) %>%
  summarise(
    n_rows = n(),
    n_ids = n_distinct(id),
    pct_rows = 100 * n() / nrow(df)
  ) %>%
  arrange(desc(n_rows))

cat("\nData Completeness Patterns:\n")
print(pattern_summary)

# ---- 4. Visualize Missing Data ----
cat("\n=== Missing Data Visualization ===\n")
#vis_miss(df)

# ---- 5. Identify IDs with Sufficient Data ----
# Create list of IDs with at least 10 days of survey data (any type)
ids_with_survey_data <- df %>%
  filter(data_pattern != "Neither") %>%
  group_by(id) %>%
  summarise(n_survey_days = n()) %>%
  filter(n_survey_days >= 10) %>%
  pull(id)

cat(sprintf("\n%d participants have ≥10 days of survey data\n", length(ids_with_survey_data)))

# ---- 6. Final Diagnostics ----
cat("\n=== Merge Complete ===\n")
count_rows_ids(df)


only_in_amdf <- setdiff(names(amdf), names(pmdf))
only_in_pmdf <- setdiff(names(pmdf), names(amdf))

print(list(only_in_amdf = only_in_amdf, only_in_pmdf = only_in_pmdf))

# Final diagnostics: crave_alc_am and crave_alc_pm (AM/PM) min/max and NA counts
crave_cols <- grep("^crave_.*_(am|pm)$", names(df), value = TRUE)
if (length(crave_cols) > 0) {
  crave_summary <- df %>% 
    summarise(across(all_of(crave_cols), 
                     list(N = ~ sum(!is.na(.)), 
                          NA_count = ~ sum(is.na(.)), 
                          min = ~ suppressWarnings(min(., na.rm=TRUE)), 
                          max = ~ suppressWarnings(max(., na.rm=TRUE)))))
  cat("\nFinal crave_* (AM/PM) Summary after merge:\n")
  print(crave_summary)
}


# Create a list of ids with 10 or more days of survey data in this dataset
ids_with_survey_data <- df %>%
  group_by(id) %>%
  summarize(n = n()) %>%
  filter(n >= 10) %>%
  pull(id) %>% unique()

ids_with_survey_data # This is a list that can be used to cross-reference missing data later on

print.variable.names(df)

# sort by id and daterated
df <- df %>%
  arrange(id, daterated)

count_rows_ids(df)

#vis_miss(df)

#remove id 123 observations without any AM survey data
df <- df %>%
  filter(!(id == 123 & is.na(num_drinks)))

#vis_miss(df)

#View(df)

# list number of observations by id
obs_by_id <- df %>%
  group_by(id) %>%
  summarise(n_obs = n()) %>%
  arrange(id)

print(obs_by_id)

# Visualize observations by id
p_obs_by_id <- ggplot(obs_by_id, aes(x = factor(id), y = n_obs)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Observations by Participant ID",
       x = "Participant ID",
       y = "Number of Observations") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

print(p_obs_by_id)

# save to data prep subfolder
save_png(plot = p_obs_by_id, 
         filename = "obs_by_id_barplot.png",
         folder = data_prep_folder)
```


```{r merge-am-pm-deduplicate}
df <- amdf_complete %>%
  full_join(pmdf, by = c("id", "daterated"), suffix = c("_am_dup", "_pm_dup"))

cat(sprintf("\nMerged data: %d rows, %d unique IDs\n", nrow(df), n_distinct(df$id)))

# ===== ADD THIS CHECKPOINT AND DEDUPLICATION =====
cat("\n=== Checking for duplicates after AM/PM merge ===\n")
df_dupes <- df %>%
  group_by(id, daterated) %>%
  filter(n() > 1) %>%
  ungroup()

if (nrow(df_dupes) > 0) {
  cat(sprintf("WARNING: Found %d duplicate (id, daterated) rows after AM/PM merge!\n", nrow(df_dupes)))
  cat("Deduplicating by keeping first row per (id, daterated)...\n")
  
  df <- df %>%
    distinct(id, daterated, .keep_all = TRUE)
  
  cat(sprintf("Rows after deduplication: %d\n", nrow(df)))
} else {
  cat("✓ No duplicates found\n")
}
# ===== END OF FIX =====

# ---- 3. Analyze Missing Data Patterns ----
cat("\n=== Missing Data Pattern Analysis ===\n")
```


```{r import-merge-visits}
# Merge in study visit dates by id: '/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_ALLDATES/03_cleaned_data/Alcohol Session Dates.xlsx'

visit_dates <- read_excel("/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_ALLDATES/03_cleaned_data/Alcohol Session Dates.xlsx")

names(visit_dates)

#View(visit_dates)

# ensure id is in the same format
visit_dates <- visit_dates %>%
  mutate(id = as.integer(id))
df <- df %>%
  mutate(id = as.integer(id))
# merge visit dates into df
df <- df %>%
  left_join(visit_dates, by = "id")

# when s1_date or s2_date is equal to daterated, set new variable session_today to 1, else 0
df <- df %>%
  mutate(session_today = case_when(
    daterated == s1_date | daterated == s2_date ~ 1,
    TRUE ~ 0
  ))

# check merge
table(df$session_today, useNA = "ifany")

```







##### TRAIT DATA IMPORT AND MERGE - have to import this here because it's used in next chunk

```{r import-trait-data}

trait_data <- read.csv('/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_TRAIT/03_cleaned_data/ukalc_trait.csv')

#View(trait_data)

#str(trait_data)

# set (id) as character, set (age, audit_score, all vars starting with tlfb or pdhq) as number, set (all variables that are 0 or 1) as factors, and set (bev_pref, race, sexorient) as factor, set trait_firststudyperiod as date
trait_data <- trait_data %>%
  mutate(id = as.integer(id),
         age = as.integer(age),
         audit_score = as.integer(audit_score),
         tlfb30d_daysdrank = as.integer(tlfb30d_daysdrank),
         tlfb30d_cumuldrinks = as.numeric(tlfb30d_cumuldrinks),
         tlfb30d_daysbinge = as.integer(tlfb30d_daysbinge),
         tlfb30d_daysdrunk = as.integer(tlfb30d_daysdrunk),
         pdhq_daysdrinkperwk = as.numeric(pdhq_daysdrinkperwk),
         pdhq_numtypicaldrinks = as.numeric(pdhq_numtypicaldrinks),
         pdhq_hrstypdrinking = as.numeric(pdhq_hrstypdrinking),
         med_user = as.factor(med_user),
         sedative_user = as.factor(sedative_user),
         cannabis_user = as.factor(cannabis_user),
         stimulant_user = as.factor(stimulant_user),
         opioid_user = as.factor(opioid_user),
         hallucinogen_user = as.factor(hallucinogen_user),
         pcp_user = as.factor(pcp_user),
         inhalant_user = as.factor(inhalant_user),
         otherdrug_user = as.factor(otherdrug_user),
         bev_pref = as.factor(bev_pref)
         )

# Add AUDIT categories based on audit_score
trait_data <- trait_data %>%
  mutate(audit_category = case_when(
    audit_score >= 0 & audit_score <= 7 ~ "Low Risk",
    audit_score >= 8 & audit_score <= 15 ~ "Hazardous",
    audit_score >= 16 & audit_score <= 19 ~ "Harmful",
    audit_score >= 20 ~ "Possible Dependence",
    TRUE ~ NA_character_
  ))

# Output appropriate descriptives to the output_folder for all and also print below
trait_summary <- trait_data %>%
  summarise(
    n_participants = n(),
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    mean_audit_score = mean(audit_score, na.rm = TRUE),
    sd_audit_score = sd(audit_score, na.rm = TRUE),
    mean_weight_kg = mean(weight_kg, na.rm = TRUE),
    sd_weight_kg = sd(weight_kg, na.rm = TRUE),
    mean_height_cm = mean(height_cm, na.rm = TRUE),
    sd_height_cm = sd(height_cm, na.rm = TRUE)
  )

cat("\n=== Trait Data Summary ===\n")
print(trait_summary)

# Merge trait data into df by id
df <- df %>%
  left_join(trait_data, by = "id")

# Print histograms below and save each to output folder

hist_age <- ggplot(trait_data, aes(x = age)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  labs(title = "Age Distribution", x = "Age", y = "Count") +
  theme_minimal()



print(hist_age)

hist_audit <- ggplot(trait_data, aes(x = audit_score)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "black") +
  labs(title = "AUDIT Score Distribution", x = "AUDIT Score", y = "Count") +
  theme_minimal()

print(hist_audit)

hist_weight <- ggplot(trait_data, aes(x = weight_kg)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "black") +
  labs(title = "Weight (kg) Distribution", x = "Weight (kg)", y = "Count") +
  theme_minimal()

print(hist_weight)

hist_height <- ggplot(trait_data, aes(x = height_cm)) +
  geom_histogram(binwidth = 2, fill = "steelblue", color = "black") +
  labs(title = "Height (cm) Distribution", x = "Height (cm)", y = "Count") +
  theme_minimal()

print(hist_height)

# Save histograms to output folder
save_png(plot = hist_age, filename = "trait_age_histogram.png", folder = data_prep_folder)
save_png(plot = hist_audit, filename = "trait_audit_histogram.png", folder = data_prep_folder)
save_png(plot = hist_weight, filename = "trait_weight_histogram.png", folder = data_prep_folder)
save_png(plot = hist_height, filename = "trait_height_histogram.png", folder = data_prep_folder)



```

```{r import-merge-weights}
# =============================================================================
# VALIDATE AND IMPUTE PARTICIPANT WEIGHTS
# =============================================================================
# Validate weight data already present in the dataset and impute missing values
# for BAC calculation.
#
# WHAT IT DOES:
# 1. Checks existing weight_kg column
# 2. Reports statistics in kg
# 3. Imputes missing data with population averages (in kg)
#
# EXPECTED COLUMNS IN DF:
# - id: Participant identifier (numeric)
# - weight_kg: Weight in kilograms (numeric)
#
# OUTPUT: Fills NAs in weight_kg column
# =============================================================================

# Report Initial Stats based on existing data
valid_weights <- df %>% filter(!is.na(weight_kg))

if (nrow(valid_weights) > 0) {
  cat(sprintf("Weight range (observed): %.1f - %.1f kg\n",
              min(valid_weights$weight_kg, na.rm = TRUE), 
              max(valid_weights$weight_kg, na.rm = TRUE)))
} else {
  warning("No valid weight data found in dataframe before imputation.")
}

# Check for participants without weights
ids_no_weight <- df %>%
  filter(is.na(weight_kg)) %>%
  pull(id) %>%
  unique()

n_missing_before <- length(ids_no_weight)

if (n_missing_before > 0) {
  warning(sprintf("%d participants missing weight data: %s", 
                  n_missing_before, 
                  paste(ids_no_weight, collapse=", ")))
  
  # Use population average weight for females 18-35 when weight is missing
  # Source: CDC/NHANES data for adult women aged 18-39
  POP_AVG_WEIGHT_KG <- 77.1 
  
  # Impute missing values
  df <- df %>%
    mutate(
      weight_kg = ifelse(is.na(weight_kg), POP_AVG_WEIGHT_KG, weight_kg)
    )
  
  cat(sprintf("\nUsing population average weight (%.1f kg) for %d participants with missing data\n",
              POP_AVG_WEIGHT_KG, n_missing_before))
  cat("Source: CDC/NHANES data for adult women aged 18-39\n")
  
} else {
  cat("\nAll participants have weight data. No imputation needed.\n")
}

```

```{r rename-alcohol-vars}
# =============================================================================
# ALCOHOL VARIABLE CREATION
# =============================================================================
# Creates derived alcohol use variables from the morning survey's report of
# prior evening drinking. These are the PRIMARY OUTCOMES for the study.
#
# NEW VARIABLE NAMES:
# - num_drinks (was Alc_Amt): Number of standard drinks consumed
# - drank_alcohol (was Alc_Drunk): Binary indicator of whether they drank
# - time_start_alcohol (was Alc_Time): Time drinking started
# - dur_alcohol (was Alc_Dur): Duration of drinking episode
#
# WHAT IT DOES:
# 1. Fills missing num_drinks with 0 when participant answered craving question
# 2. Creates lagged versions (prior evening → current day alignment)
# 3. Creates binary indicators for any drinking and binge drinking
#
# KEY CONCEPT - TEMPORAL ALIGNMENT:
# AM surveys ask about PREVIOUS evening (e.g., morning of Jan 2 asks about Jan 1 evening)
# PM surveys ask about CURRENT day (e.g., evening of Jan 1 asks about Jan 1)
# NEED: Align drinking with the day it occurred, not the day it was reported
# SOLUTION: Use lead() to shift num_drinks forward one day
#
# EXAMPLE:
# Row 1: id=101, date=Jan 1, num_drinks=NA (no AM survey yet)
# Row 2: id=101, date=Jan 2, num_drinks=3 (reported drinking 3 drinks on Jan 1)
# After lead(): 
# Row 1: id=101, date=Jan 1, num_drinks_today=3 (drinking happened on Jan 1)
# Row 2: id=101, date=Jan 2, num_drinks_today=NA (will be filled with next day's AM report)
#
# VARIABLES CREATED:
# 1. num_drinks_today: Number of drinks consumed on THIS day (numeric, continuous)
# 2. drink_today: Binary indicator for any drinking (Yes/No factor)
# 3. drink_today_bin: Numeric version of drink_today (0/1)
# 4. fourplustoday: Binary indicator for 4+ drinks (Yes/No factor) 
# 5. fourplustoday_bin: Numeric version of fourplustoday (0/1)
#
# WHY 4+ DRINKS:
# - Standard definition of "binge drinking" for women (NIAAA guidelines)
# - Important threshold for increased risk of negative consequences
# - Primary outcome for examining hormonal effects on heavy drinking
#
# MISSING DATA LOGIC:
# "If crave_alc_am is not missing but num_drinks is missing, set num_drinks to 0"
# RATIONALE: Craving question comes AFTER drinking question in survey
# If participant answered craving, they saw the drinking question
# If num_drinks is missing, they likely skipped it (meaning no drinking)
#
# INPUTS: df (merged AM/PM data)
# OUTPUTS: df with num_drinks_today, drink_today, fourplustoday variables
# =============================================================================

```


```{r align-next-day-alcohol}
library(dplyr)

df <- df %>%
  # Impute 0 for 'num_drinks' if 'crave_alc_am' exists but 'num_drinks' is NA
  mutate(
    num_drinks = if_else(!is.na(crave_alc_am) & is.na(num_drinks), 0, num_drinks)
  ) %>%
  group_by(id) %>%
  arrange(daterated) %>%
  mutate(
    # Lead variables to align "next day" or "today" values
    # NOTE: lead() shifts values forward by 1 row, so:
    # - "num_drinks" from tomorrow's AM survey → becomes "num_drinks_today"
    # - This aligns drinking reports with the EVENING they occurred
    # - IMPORTANT: The last observation in each ID's time series will be NA
    #   because there's no "next day" to pull from. This is expected behavior.
    num_drinks_today = lead(num_drinks, n = 1),
    dur_alcohol_hours_today = lead(dur_alcohol_hours, n = 1),
    time_start_alcohol_clean_today = lead(time_start_alcohol_clean, n = 1),
    felt_drunk_today = lead(felt_drunk, n = 1),
    
    # --- Create Numeric Binary (0/1) Variables ---
    
    # 1. Any drinking (Threshold > 0)
    drink_today_bin = case_when(
      is.na(num_drinks_today) ~ NA_real_,
      num_drinks_today > 0 ~ 1,
      TRUE ~ 0
    ),
    
    # 2. Binge drinking (Threshold >= 4)
    fourplustoday_bin = case_when(
      is.na(num_drinks_today) ~ NA_real_,
      num_drinks_today >= 4 ~ 1, 
      TRUE ~ 0
    ),
    
    # 3. Felt Drunk (already 0/1, but cleaned to match drinking data)
    felt_drunk_today_bin = case_when(
      is.na(num_drinks_today) ~ NA_real_,
      num_drinks_today == 0 ~ NA_real_,     # No drinking means felt drunk is NA
      felt_drunk_today == 1 ~ 1,
      felt_drunk_today == 0 ~ 0,
      TRUE ~ NA_real_                  # Keep NA if drinking occurred but feeling is unknown
    )
  ) %>%
  ungroup() %>% 
  # --- Create Factor Versions (Yes/No) ---
  mutate(
    drink_today = factor(
      drink_today_bin, 
      levels = c(0, 1), 
      labels = c("No", "Yes")
    ),
    fourplustoday = factor(
      fourplustoday_bin, 
      levels = c(0, 1), 
      labels = c("No", "Yes")
    ),
    felt_drunk_today = factor( # Suffix _fac used to distinguish from raw lagged variable
      felt_drunk_today_bin,
      levels = c(0, 1),
      labels = c("No", "Yes")
    )
  )


# Print narrative of script actions
cat("\n=== Alcohol Variable Creation Summary ===\n")
cat("1. Imputed 0 for 'num_drinks' when 'crave_alc_am' exists but 'num_drinks' is missing.\n")
cat("2. Created lagged variables to align drinking data with the day it occurred.\n")
cat("3. Created binary indicators for any drinking and binge drinking (4+ drinks).\n")
cat("4. Created both numeric (0/1) and factor (Yes/No) versions of drinking indicators.\n")
```


```{r calculate-bac}
# =============================================================================
# BAC (BLOOD ALCOHOL CONCENTRATION) ESTIMATION
# =============================================================================
# Calculate estimated peak BAC using Widmark formula for each drinking episode
#
# ASSUMPTIONS AND LIMITATIONS:
# - All participants are female (r = 0.55 for Widmark constant)
# - Standard drink = 14g ethanol (US definition)
# - Elimination rate = 0.015% per hour (population average)
# - Absorption lag = 0.5 hours after end of drinking
# - Drink size/type unknown (assumes standard drinks)
# - This is an ESTIMATE, not a full pharmacokinetic model
#
# WIDMARK FORMULA:
# BAC = (Alcohol_g / (r × Weight_g)) × 100
# Where: r = 0.55 for females, Weight_g = weight in grams
#
# PEAK BAC TIMING:
# Peak occurs at: start_time + duration + 0.5 hours (absorption lag)
# By peak time, some alcohol has been eliminated: 0.015% × hours_elapsed
#
# INPUTS:
# - num_drinks: Number of standard drinks consumed
# - dur_hours: Duration of drinking episode in hours
# - weight_kg: Participant weight in kilograms (uses pop avg 77.1kg if missing)
# - start_time: Time drinking started (HH:MM format, optional)
#
# OUTPUTS:
# - alcohol_g: Total alcohol consumed in grams
# - bac_peak_percent: Estimated peak BAC in percent (e.g., 0.08 for 0.08%)
# - bac_peak_mgdl: Estimated peak BAC in mg/dL
# - bac_peak_time: Estimated time of peak BAC (if start_time provided)
#
# NOTE: Population average weight (170 lbs / 77.1 kg) is automatically used
# when individual weight data is missing. Source: CDC/NHANES for women 18-39.
# =============================================================================

#' Calculate Peak Blood Alcohol Concentration (BAC)
#'
#' Estimates peak BAC using the Widmark formula with elimination adjustment.
#' Designed for female participants consuming alcohol over a specified duration.
#'
#' @param num_drinks Numeric. Number of standard drinks consumed (14g ethanol each)
#' @param dur_hours Numeric. Duration of drinking episode in hours
#' @param weight_kg Numeric. Body weight in kilograms
#' @param start_time Character. Time drinking started in HH:MM format (24-hour)
#' @param date Date. Date of drinking episode (for calculating peak time)
#'
#' @return List with elements:
#'   - alcohol_g: Total alcohol in grams
#'   - bac_peak_percent: Peak BAC in percent
#'   - bac_peak_mgdl: Peak BAC in mg/dL
#'   - bac_peak_time: POSIXct datetime of peak (if start_time and date provided)
#'
#' @details
#' Uses Widmark constant r=0.55 for females, elimination rate of 0.015%/hr,
#' and assumes peak occurs 0.5 hours after drinking ends (absorption lag).
#' Note: The pipeline uses population average weight (77.1 kg) when individual
#' weight data is missing, based on CDC/NHANES data for women aged 18-39.
#'
#' @examples
#' calc_peak_bac(3, 2, 64)   # 3 drinks over 2 hours, 64 kg
#' calc_peak_bac(0, 0, 64)   # No drinking
#' calc_peak_bac(3, 2, 77.1) # Using population average weight
calc_peak_bac <- function(num_drinks, dur_hours, weight_kg, start_time = NA, date = NA) {
  
  # Initialize output list
  result <- list(
    alcohol_g = NA_real_,
    bac_peak_percent = NA_real_,
    bac_peak_mgdl = NA_real_,
    bac_peak_time = as.POSIXct(NA)
  )
  
  # Handle edge case: no drinks or zero drinks
  if (is.na(num_drinks) || num_drinks == 0) {
    result$alcohol_g <- 0
    result$bac_peak_percent <- 0
    result$bac_peak_mgdl <- 0
    return(result)
  }
  
  # Handle edge case: missing weight
  if (is.na(weight_kg)) {
    result$alcohol_g <- num_drinks * 14
    # BAC values remain NA (cannot calculate without weight)
    return(result)
  }
  
  # Constants
  GRAMS_PER_DRINK <- 14        # US standard drink = 14g ethanol
  WIDMARK_R_FEMALE <- 0.55     # Body water constant for females
  ELIMINATION_RATE <- 0.015    # BAC elimination: 0.015% per hour
  ABSORPTION_LAG <- 0.5        # Hours after drinking ends until peak
  
  # Calculate total alcohol consumed (grams)
  alcohol_g <- num_drinks * GRAMS_PER_DRINK
  
  # Handle edge case: missing or zero duration
  if (is.na(dur_hours)) {
    dur_hours <- 0  # Assume instantaneous consumption if duration unknown
  }
  
  # Convert weight to grams
  weight_g <- weight_kg * 1000
  
  # Calculate raw BAC at end of drinking (before elimination)
  # Widmark formula: BAC = (alcohol_g / (r × weight_g)) × 100
  bac_raw <- (alcohol_g / (WIDMARK_R_FEMALE * weight_g)) * 100
  
  # Time from start of drinking to peak (duration + absorption lag)
  t_peak <- dur_hours + ABSORPTION_LAG
  
  # Apply elimination: subtract BAC eliminated by peak time
  bac_peak_percent <- bac_raw - (ELIMINATION_RATE * t_peak)
  
  # Clamp to non-negative (can't have negative BAC)
  bac_peak_percent <- max(bac_peak_percent, 0)
  
  # Convert to mg/dL
  bac_peak_mgdl <- bac_peak_percent * 1000
  
  # Calculate peak time if start time and date are provided
  bac_peak_time <- as.POSIXct(NA)
  if (!is.na(start_time) && !is.na(date)) {
    tryCatch({
      # Parse start time (HH:MM format)
      time_parts <- as.numeric(strsplit(start_time, ":")[[1]])
      if (length(time_parts) == 2) {
        # Combine date and time
        start_datetime <- as.POSIXct(paste(date, start_time), 
                                      format = "%Y-%m-%d %H:%M",
                                      tz = "America/New_York")
        # Add duration and absorption lag
        bac_peak_time <- start_datetime + (t_peak * 3600)  # Convert hours to seconds
      }
    }, error = function(e) {
      # If time parsing fails, leave as NA
    })
  }
  
  # Return results
  result$alcohol_g <- alcohol_g
  result$bac_peak_percent <- bac_peak_percent
  result$bac_peak_mgdl <- bac_peak_mgdl
  result$bac_peak_time <- bac_peak_time
  
  return(result)
}

# Apply BAC calculation to dataset
cat("\n=== Calculating Peak BAC Estimates ===\n")

# Check if weight data is available
has_weights <- "weight_kg" %in% names(df) && sum(!is.na(df$weight_kg)) > 0

if (has_weights) {
  
  # Calculate BAC for each row
  # Use pmap to pass multiple columns to calc_peak_bac function
  bac_results <- df %>%
    select(num_drinks_today, dur_alcohol_hours_today, weight_kg, 
           time_start_alcohol_clean_today, daterated) %>%
    pmap(function(num_drinks_today, dur_alcohol_hours_today, weight_kg, 
                  time_start_alcohol_clean_today, daterated) {
      calc_peak_bac(
        num_drinks = num_drinks_today,
        dur_hours = dur_alcohol_hours_today,
        weight_kg = weight_kg,
        start_time = time_start_alcohol_clean_today,
        date = daterated
      )
    })
  
  # Extract results into separate columns
  df <- df %>%
    mutate(
      alcohol_g = map_dbl(bac_results, ~ .x$alcohol_g),
      bac_peak_percent = map_dbl(bac_results, ~ .x$bac_peak_percent),
      bac_peak_mgdl = map_dbl(bac_results, ~ .x$bac_peak_mgdl),
      bac_peak_time = map(bac_results, ~ .x$bac_peak_time) %>% 
        map(as.POSIXct, origin = "1970-01-01") %>%
        do.call(c, .)
    )
  
  # Report summary statistics
  n_drinking_days <- sum(df$num_drinks_today > 0, na.rm = TRUE)
  n_bac_calculated <- sum(!is.na(df$bac_peak_percent) & df$bac_peak_percent > 0, na.rm = TRUE)
  
  cat(sprintf("Drinking days: %d\n", n_drinking_days))
  cat(sprintf("BAC calculated: %d (%.1f%% of drinking days)\n", 
              n_bac_calculated, 100 * n_bac_calculated / max(n_drinking_days, 1)))
  
  if (n_bac_calculated > 0) {
    bac_summary <- df %>%
      filter(!is.na(bac_peak_percent) & bac_peak_percent > 0) %>%
      summarise(
        mean_bac = mean(bac_peak_percent, na.rm = TRUE),
        median_bac = median(bac_peak_percent, na.rm = TRUE),
        min_bac = min(bac_peak_percent, na.rm = TRUE),
        max_bac = max(bac_peak_percent, na.rm = TRUE)
      )
    
    cat(sprintf("\nBAC Summary (drinking days only):\n"))
    cat(sprintf("  Mean: %.3f%% (%.0f mg/dL)\n", 
                bac_summary$mean_bac, bac_summary$mean_bac * 1000))
    cat(sprintf("  Median: %.3f%% (%.0f mg/dL)\n", 
                bac_summary$median_bac, bac_summary$median_bac * 1000))
    cat(sprintf("  Range: %.3f%% - %.3f%%\n", 
                bac_summary$min_bac, bac_summary$max_bac))
    
    # Report binge drinking threshold (0.08% BAC)
    n_binge_bac <- sum(df$bac_peak_percent >= 0.08, na.rm = TRUE)
    cat(sprintf("  Days with BAC ≥ 0.08%%: %d (%.1f%% of drinking days)\n",
                n_binge_bac, 100 * n_binge_bac / max(n_drinking_days, 1)))
  }
  
  cat("\nNOTE: BAC estimates assume:\n")
  cat("  - Female participants (r=0.55)\n")
  cat("  - Standard drinks (14g ethanol each)\n")
  cat("  - Elimination rate: 0.015%/hour\n")
  cat("  - Peak at drinking_end + 0.5 hours\n")
  cat("  - Population average weight (77.1 kg) used when individual weight missing\n")
  
} else {
  warning("Weight data not available. BAC calculations skipped.")
  cat("\nTo calculate BAC, ensure weights.csv is in the data directory with columns: id, weight\n")
  
  # Add placeholder BAC columns
  df <- df %>%
    mutate(
      alcohol_g = ifelse(!is.na(num_drinks_today), num_drinks_today * 14, NA_real_),
      bac_peak_percent = NA_real_,
      bac_peak_mgdl = NA_real_,
      bac_peak_time = as.POSIXct(NA)
    )
}


# -------------------------------------------------------------------------
# VALIDATE BINARY VARIABLES
# -------------------------------------------------------------------------
# Binary variables (drink_today_bin, fourplustoday_bin, felt_drunk_today_bin)
# were correctly created in the create-alc-vars chunk above.
# This section validates that they contain only 0, 1, or NA values.

# Validate binary variables are correctly coded (0/1 only)
cat("\n=== Binary Variable Validation ===\n")
for (var in c("drink_today_bin", "fourplustoday_bin", "felt_drunk_today_bin")) {
  unique_vals <- unique(df[[var]])
  unique_vals <- unique_vals[!is.na(unique_vals)]
  if (!all(unique_vals %in% c(0, 1))) {
    stop(sprintf("ERROR: %s contains non-binary values: %s", 
                 var, paste(unique_vals, collapse=", ")))
  }
  cat(sprintf("✓ %s: Valid binary (0/1) with %d values\n", 
              var, sum(!is.na(df[[var]]))))
}

# Display structure and distribution
str(df$drink_today_bin)
hist(df$drink_today_bin, main = "Distribution of drink_today_bin", xlab = "Value")
str(df$fourplustoday_bin)
hist(df$fourplustoday_bin, main = "Distribution of fourplustoday_bin", xlab = "Value")

#vis_miss(df)
```

```{r score-bdefs}
# =============================================================================
# BDEFS COMPOSITE SCORE CALCULATION (FOR UKALC DIAGNOSTIC PLOTS)
# =============================================================================
# NOTE: These subscales are calculated here for UKALC diagnostic histograms,
# but they are ALSO calculated in UKALC_03_processing_shared.Rmd for the 
# merged ADHD+UKALC dataset to ensure both studies receive identical processing.
#
# Creates a single composite score from Barkley's Deficits in Executive 
# Functioning Scale (BDEFS). This measures executive function and impulse control.
#
# WHAT IT DOES:
# - Calculates row-wise mean of 7 bdefs items
# - Handles missing data by using na.rm=TRUE (calculates mean of available items)
# - Creates histogram of distribution
#
# bdefs ITEMS (defined in bdefs_vars list):
# 1. diff_inhib: Difficulty inhibiting responses
# 2. diff_stop: Difficulty stopping ongoing behavior
# 3. diff_change: Difficulty changing course
# 4. immed_payoff: Preference for immediate payoff
# 5. res_urge: Difficulty resisting urges
# 6. diff_long_rew: Difficulty waiting for long-term rewards
# 7. diff_res_rew: Difficulty resisting rewarding stimuli
#
# SCALE PROPERTIES:
# - Range: 1-4 (Never/Rarely=1, Sometimes=2, Often=3, Very Often=4)
# - Higher scores = Greater deficits in executive functioning (worse impulse control)
# - Composite range: 1-4 (mean of item scores)
#
# WHY THIS MATTERS:
# - Executive dysfunction predicts alcohol use and problems
# - May vary across menstrual cycle phases
# - Key mediator in hormonal effects on drinking
#
# MISSING DATA HANDLING:
# - na.rm=TRUE means: Calculate mean of available items
# - If participant has 5/7 items, mean is based on those 5
# - If ALL items missing, result is NaN (not a number)
# - Decision: Include participants with partial data (not dropping if <7 items)
# - Assumption: Missing items are missing at random (MAR)
#
# INPUTS: df with bdefs items (from PM surveys)
# OUTPUTS: df with bdefs composite score + histogram PNG
# =============================================================================

# Variables needed for THIS chunk only
bdefs_vars <- c(
  "diff_inhib", "diff_stop", "diff_change", "immed_payoff", 
  "res_urge", "diff_long_rew", "diff_res_rew"
)

# Create a mean score for these daily items across each row and make sure to account for missing values (diff_inhib, diff_stop, diff_change, immed_payoff, res_urge, diff_long_rew, diff_res_rew)

df <- df %>%
  mutate(bdefs_total = rowMeans(select(., all_of(bdefs_vars)), na.rm = TRUE))

# Create and save histogram for bdefs_total using ggplot2
p_bdefs_total <- ggplot(df, aes(x = bdefs_total)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 30) +
  labs(title = "BDEFS Total Distribution", x = "BDEFS Total", y = "Count") +
  theme_minimal(base_size = 14)
save_png(plot = p_bdefs_total, 
         filename = "bdefs_total_histogram.png",
         folder = data_prep_folder)

# score the mean of bdefs_impulsiveaction
df <- df %>%
  mutate(bdefs_impulsiveaction = rowMeans(select(., diff_inhib, diff_stop, diff_change), na.rm = TRUE))

# Create and save histogram for bdefs_impulsiveaction
p_bdefs_impulsiveaction <- ggplot(df, aes(x = bdefs_impulsiveaction)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 30) +
  labs(title = "BDEFS Impulsive Action Distribution", x = "BDEFS Impulsive Action", y = "Count") +
  theme_minimal(base_size = 14)
save_png(plot = p_bdefs_impulsiveaction, 
         filename = "bdefs_impulsiveaction_histogram.png",
         folder = data_prep_folder)

# score the mean of bdefs_delayaversion
df <- df %>%
  mutate(bdefs_delayaversion = rowMeans(select(., immed_payoff, res_urge, diff_long_rew, diff_res_rew), na.rm = TRUE))

# Create and save histogram for bdefs_delayaversion
p_bdefs_delayaversion <- ggplot(df, aes(x = bdefs_delayaversion)) +
  geom_histogram(fill = "lightblue", color = "black", bins = 30) +
  labs(title = "BDEFS Delay Aversion Distribution", x = "BDEFS Delay Aversion", y = "Count") +
  theme_minimal(base_size = 14)
save_png(plot = p_bdefs_delayaversion, 
         filename = "bdefs_delayaversion_histogram.png",
         folder = data_prep_folder)

```





```{r cycle-dates-import-clean}
# =============================================================================
# MENSTRUAL CYCLE DATA IMPORT & CLEANING
# =============================================================================
# Imports comprehensive participant tracking data including RTB-recoded menses 
# dates, multiple ovulation dates, and exclusion flags. This is essential for 
# menstrual cycle phase analysis.
#
# WHAT IT DOES:
# 1. Loads UKALC_Participant_Data_Tracker with all cycle dates and metadata
# 2. Creates TWO versions of menses dates:
#    - period1/2/3 (non-RTB): Original MP dates only
#    - period1/2/3_rtb (RTB): RTB-recoded dates, falls back to MP if no RTB
# 3. Creates event flags for both versions (menses/ovtoday and menses_rtb/ovtoday_rtb)
# 4. Applies exclusion flags (Excl_all_id, Excl_CycData, Excl_Horm)
# 5. Calculates cycle length for both versions
# 6. Later: Runs menstrualcycleR for both versions (see run-menstrualcycler chunk)
#
# DATA SOURCE:
# File: UKALC_Participant_Data_Tracker_20251225.csv
# Format: One row per participant with comprehensive cycle and metadata
# Variables:
#   - id: Participant ID
#   - MP1, MP2, MP3: Original menses onset dates (1st, 2nd, 3rd)
#   - RTBMP1, RTBMP2, RTBMP3: RTB-recoded menses dates (RTB = Return To Baseline of progesterone)
#   - EDO1, EDO2, EDO3: Estimated day of ovulation dates (1st, 2nd, 3rd)
#   - Ov_Status: Ovulatory status (categorical)
#   - Assay_Batch: Which assay batch participant was in
#   - Excl_all_id: Exclude from ALL datasets (logical)
#   - Excl_Horm: Exclude from hormone analyses (logical)
#   - Excl_CycData: Exclude cycle data but keep survey data (logical)
#   - Assay_Plate: List of assay plates for participant
#
# ASSAY PLATE RECODING:
# - Raw Assay_Plate is comma-separated list (e.g., "1,2,3" or "3")
# - Created variables for analysis:
#   * n_assay_plates: Count of plates participant was on (0 if NA)
#   * in_plate_X: Binary indicators for each plate (dynamically created based on actual plate numbers)
#   * Assay_Plate_orig: Original string preserved
# - These allow testing for plate effects and batch corrections
#
# WHY MENSTRUAL CYCLE MATTERS:
# - Hormones (e2, p4) fluctuate dramatically across cycle
# - Different phases have different hormone profiles:
#   * Follicular: Low p4, rising e2
#   * Ovulatory: e2 peak, lh surge
#   * Luteal: High p4, moderate e2
#   * Menstrual: Low e2 and p4
# - Alcohol responses may vary by phase
#
# RTB RECODING - DUAL VERSION APPROACH:
# RTB = "Return To Baseline" - period dates recoded based on progesterone return to baseline
# Creates TWO parallel versions of all cycle-related variables:
# 1. NON-RTB VERSION (original variable names): Uses only original MP dates
#    - period1, period2, period3 = MP1, MP2, MP3
#    - menses, ovtoday = event flags from MP dates
#    - cycle_length = period2 - period1
#    - menstrualcycleR outputs: cyclic_time_impute, cyclic_time_fwd/bwd, etc.
# 2. RTB VERSION (_rtb suffix): Uses RTB-recoded dates when available, falls back to MP
#    - period1_rtb, period2_rtb, period3_rtb = coalesce(RTBMP*, MP*)
#    - menses_rtb, ovtoday_rtb = event flags from RTB dates (or MP if no RTB)
#    - cycle_length_rtb = period2_rtb - period1_rtb
#    - menstrualcycleR outputs: cyclic_time_impute_rtb, cyclic_time_fwd/bwd_rtb, etc.
# This allows analyses comparing RTB-recoded vs original dates
# Track which dates were recoded with period*_recoded boolean flags
#
# EXCLUSION FLAGS:
# - Excl_all_id = TRUE: Remove ALL data for this participant
# - Excl_CycData = TRUE: Keep survey data, but set cycle variables to NA
# - Excl_Horm = TRUE: Flag for later exclusion from hormone analyses
#
# CYCLE LENGTH CALCULATION:
# - cycle_length = period2 - period1 (in days)
# - Valid range: Typically 21-38 days (biological normal)
# - Participants outside this range are filtered from cycle data
# - BUT: They are NOT removed from survey data
#
# IMPORTANT DESIGN DECISION:
# OLD approach: Filtered survey data to only include participants with valid cycles
# NEW approach: Keep ALL survey data, even if cycle data is invalid
# Rationale: Survey data is still valuable for non-cycle analyses
# Impact: Some participants will have NA for cycle-related variables
#
# EVENT FLAGS:
# - menses: Binary indicator (1=period onset day, 0=other) - includes ALL menses dates
# - ovtoday: Binary indicator (1=ovulation day, 0=other) - includes ALL ovulation dates
# - Used for: Plotting cycle events, defining cycle phases
# - Merged to survey data by id and daterated
#
# CYCLE PHASE VISUALIZATION:
# - Creates plots showing period and ovulation events across study
# - Red vertical lines = period onset
# - Green vertical lines = ovulation
# - Helps identify data quality issues (irregular cycles, missing events)
#
# INPUTS: 
#   - UKALC_Participant_Data_Tracker_20251225.csv
#   - Survey data (df, amdf, pmdf) for date imputation
# OUTPUTS:
#   - df with menses and ovtoday columns (replacing TAEMPeriodStart and TAEMOvPos)
#   - cycle_length variable
#   - Participant metadata (Ov_Status, Assay_Batch, exclusion flags)
#   - Diagnostic output and plots
# =============================================================================

# ---- 1. Import the comprehensive participant tracking data ----
raw_cycle_df <- read_csv("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_ALLDATES/03_cleaned_data/UKALC_Participant_Data_Tracker_20251231.csv")

cat("\n=== Imported UKALC Participant Data Tracker ===\n")


# select only relevant columns
raw_cycle_df <- raw_cycle_df %>%
  select(id, MP1, MP2, MP3, RTBMP1, RTBMP2, RTBMP3, EDO1, EDO2,
         Ov_Status, Assay_Batch, Assay_Plate,
         Excl_all_id, Excl_Horm, Excl_CycData,
         e_spike_date_remove, e_spike_date_remove_2, p_spike_date_remove, p_spike_date_remove_2)

#View(raw_cycle_df)

# Parse dates and create "best" menses dates prioritizing RTB-recoded dates
wide_cycles <- raw_cycle_df %>%
  mutate(
    # Ensure id is character
    id = as.character(id),
    
    # Parse all date columns (in M/D/YY format)
    MP1 = mdy(MP1),
    MP2 = mdy(MP2),
    MP3 = mdy(MP3),
    RTBMP1 = mdy(RTBMP1),
    RTBMP2 = mdy(RTBMP2),
    RTBMP3 = mdy(RTBMP3),
    EDO1 = mdy(EDO1),
    EDO2 = mdy(EDO2),
    espike1 = mdy(e_spike_date_remove),
    espike2 = mdy(e_spike_date_remove_2),
    pspike1 = mdy(p_spike_date_remove),
    pspike2 = mdy(p_spike_date_remove_2),
    
    # Convert exclusion flags to logical
    Excl_all_id = as.logical(Excl_all_id),
    Excl_Horm = as.logical(Excl_Horm),
    Excl_CycData = as.logical(Excl_CycData),
    
    # Create RTB-prioritized menses dates (with _rtb suffix)
    period1_rtb = coalesce(RTBMP1, MP1),
    period2_rtb = coalesce(RTBMP2, MP2),
    period3_rtb = coalesce(RTBMP3, MP3),
    
    # Create non-RTB menses dates (original MP dates only)
    period1 = MP1,
    period2 = MP2,
    period3 = MP3,
    
    # Track which dates were recoded
    period1_recoded = !is.na(RTBMP1),
    period2_recoded = !is.na(RTBMP2),
    period3_recoded = !is.na(RTBMP3)
  ) %>%
  # Select relevant columns
  select(id, period1, period2, period3, 
         period1_rtb, period2_rtb, period3_rtb,
         period1_recoded, period2_recoded, period3_recoded,
         EDO1, EDO2, Ov_Status, Assay_Batch, Assay_Plate,
         Excl_all_id, Excl_Horm, Excl_CycData, espike1, espike2, pspike1, pspike2)

#View(wide_cycles)


# for ids where Excl_CycData is TRUE, set all period* and EDO* dates to NA
wide_cycles <- wide_cycles %>%
  mutate(
    period1 = if_else(Excl_CycData %in% TRUE, as.Date(NA), period1),
    period2 = if_else(Excl_CycData %in% TRUE, as.Date(NA), period2),
    period3 = if_else(Excl_CycData %in% TRUE, as.Date(NA), period3),
    period1_rtb = if_else(Excl_CycData %in% TRUE, as.Date(NA), period1_rtb),
    period2_rtb = if_else(Excl_CycData %in% TRUE, as.Date(NA), period2_rtb),
    period3_rtb = if_else(Excl_CycData %in% TRUE, as.Date(NA), period3_rtb),
    EDO1 = if_else(Excl_CycData %in% TRUE, as.Date(NA), EDO1),
    EDO2 = if_else(Excl_CycData %in% TRUE, as.Date(NA), EDO2)
  )





# ---- Parse Assay_Plate for analysis ----
# Assay_Plate can be comma-separated list (e.g., "1,2,3" or "3")
# Create indicator variables and count variable
wide_cycles <- wide_cycles %>%
  mutate(
    # Convert Assay_Plate to character and handle NA
    Assay_Plate_orig = as.character(Assay_Plate),
    
    # Count number of plates (count commas + 1, or 0 if NA)
    n_assay_plates = if_else(
      is.na(Assay_Plate_orig), 
      0L, 
      as.integer(str_count(Assay_Plate_orig, ",") + 1)
    )
  )

# ---- Parse Assay Plate Indicators Dynamically ----
# Extract all unique plate numbers from the Assay_Plate column
all_plates <- wide_cycles %>%
  filter(!is.na(Assay_Plate_orig)) %>%
  pull(Assay_Plate_orig) %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  unique() %>%
  as.numeric() %>%
  sort()

# Create binary indicator for each plate dynamically
for (plate_num in all_plates) {
  col_name <- paste0("in_plate_", plate_num)
  wide_cycles <- wide_cycles %>%
    mutate(!!col_name := str_detect(Assay_Plate_orig, paste0("\\b", plate_num, "\\b")) & !is.na(Assay_Plate_orig))
}

# ---- Diagnostic Output: RTB Recoding Statistics ----
cat("\n=== RTB RECODING STATISTICS ===\n")
cat("Period 1 - Recoded:", sum(wide_cycles$period1_recoded, na.rm = TRUE), 
    "/ Original:", sum(!wide_cycles$period1_recoded & !is.na(wide_cycles$period1), na.rm = TRUE), "\n")
cat("Period 2 - Recoded:", sum(wide_cycles$period2_recoded, na.rm = TRUE), 
    "/ Original:", sum(!wide_cycles$period2_recoded & !is.na(wide_cycles$period2), na.rm = TRUE), "\n")
cat("Period 3 - Recoded:", sum(wide_cycles$period3_recoded, na.rm = TRUE), 
    "/ Original:", sum(!wide_cycles$period3_recoded & !is.na(wide_cycles$period3), na.rm = TRUE), "\n")

cat("\n=== OVULATION DATE STATISTICS ===\n")
cat("Participants with EDO1:", sum(!is.na(wide_cycles$EDO1)), "\n")
cat("Participants with EDO2:", sum(!is.na(wide_cycles$EDO2)), "\n")

cat("\n=== EXCLUSION FLAG STATISTICS ===\n")
cat("Excl_all_id (exclude from all datasets):", sum(wide_cycles$Excl_all_id, na.rm = TRUE), "\n")
cat("Excl_CycData (exclude cycle data only):", sum(wide_cycles$Excl_CycData, na.rm = TRUE), "\n")
cat("Excl_Horm (exclude from hormone analyses):", sum(wide_cycles$Excl_Horm, na.rm = TRUE), "\n")

cat("\n=== ASSAY PLATE STATISTICS ===\n")
cat("Participants by number of plates:\n")
print(table(wide_cycles$n_assay_plates, useNA = "ifany"))
cat("\nPlate numbers found:", paste(all_plates, collapse = ", "), "\n")
cat("Participants in each plate:\n")
for (plate_num in all_plates) {
  col_name <- paste0("in_plate_", plate_num)
  cat(sprintf("Plate %s: %d\n", plate_num, sum(wide_cycles[[col_name]], na.rm = TRUE)))
}

# ---- 2. Calculate Cycle Length (No Imputation) ----
# Only use dates from the tracker - do NOT impute missing dates
# Calculate for both RTB and non-RTB versions
wide_cycles_with_length <- wide_cycles %>%
  filter(!Excl_CycData | is.na(Excl_CycData)) %>%  # Keep only non-excluded participants
  mutate(
    cycle_length = as.numeric(period2 - period1),
    cycle_length_rtb = as.numeric(period2_rtb - period1_rtb)
  ) %>%
  # Filter to ensure at least ONE version has valid cycle data
  # Design choice: Allow participants with EITHER version (RTB or non-RTB) to be included
  # This means some participants may have NA for one version but valid data for the other
  # If both versions are required, change to: filter(!is.na(cycle_length) & !is.na(cycle_length_rtb))
  filter(!is.na(cycle_length) | !is.na(cycle_length_rtb))

# ---- 3. Apply Exclusion Flags ----
if (!exists("df")) stop("df (survey data) must be loaded first.")

# First, remove participants with Excl_all_id = TRUE from main dataset
ids_excl_all <- wide_cycles %>% 
  filter(Excl_all_id == TRUE) %>% 
  pull(id)

cat("\n=== APPLYING EXCLUSIONS TO MAIN DATASET ===\n")
cat("Removing", length(ids_excl_all), "participants with Excl_all_id = TRUE:", 
    paste(ids_excl_all, collapse = ", "), "\n")

clean_df <- df %>%
  mutate(id = as.character(id), daterated = as_date(daterated)) %>%
  filter(!id %in% ids_excl_all) %>%  # Remove Excl_all_id participants
  filter(!is.na(daterated))


# print what's been done
cat("Main dataset now has", nrow(clean_df), "observations after applying Excl_all_id exclusions.\n")
cat("Participants with valid cycle data (non-RTB or RTB) retained for cycle analyses:", 
    n_distinct(wide_cycles_with_length$id), "\n")

# ---- 4. Create Event Flags for Menses and Ovulation ----
# Create TWO versions: RTB-prioritized (_rtb suffix) and non-RTB (original names)

# RTB VERSION: Uses period*_rtb dates (RTB dates with MP fallback)
events_long_rtb <- bind_rows(
  # Menses: Include ALL three period dates (RTB version)
  wide_cycles_with_length %>% 
    select(id, period1_rtb, period2_rtb, period3_rtb) %>% 
    pivot_longer(cols = c(period1_rtb, period2_rtb, period3_rtb), 
                 values_to = "daterated", 
                 names_to = "period_num") %>% 
    filter(!is.na(daterated)) %>%
    mutate(menses_rtb = 1) %>% 
    select(id, daterated, menses_rtb),
  
  # Ovulation: Include BOTH ovulation dates (same for both versions)
  wide_cycles_with_length %>% 
    select(id, EDO1, EDO2) %>%
    pivot_longer(cols = c(EDO1, EDO2), 
                 values_to = "daterated", 
                 names_to = "ov_num") %>% 
    filter(!is.na(daterated)) %>%
    mutate(ovtoday_rtb = 1) %>%
    select(id, daterated, ovtoday_rtb)
) %>%
  filter(!is.na(daterated)) %>%
  group_by(id, daterated) %>%
  summarize(
    menses_rtb = max(menses_rtb, na.rm = TRUE),
    ovtoday_rtb = max(ovtoday_rtb, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    menses_rtb = if_else(is.finite(menses_rtb), menses_rtb, NA_real_),
    ovtoday_rtb = if_else(is.finite(ovtoday_rtb), ovtoday_rtb, NA_real_)
  )

# NON-RTB VERSION: Uses original period* dates (MP only, no RTB)
events_long <- bind_rows(
  # Menses: Include ALL three period dates (non-RTB version)
  wide_cycles_with_length %>% 
    select(id, period1, period2, period3) %>% 
    pivot_longer(cols = c(period1, period2, period3), 
                 values_to = "daterated", 
                 names_to = "period_num") %>% 
    filter(!is.na(daterated)) %>%
    mutate(menses = 1) %>% 
    select(id, daterated, menses),
  
  # Ovulation: Include BOTH ovulation dates (same for both versions)
  wide_cycles_with_length %>% 
    select(id, EDO1, EDO2) %>%
    pivot_longer(cols = c(EDO1, EDO2), 
                 values_to = "daterated", 
                 names_to = "ov_num") %>% 
    filter(!is.na(daterated)) %>%
    mutate(ovtoday = 1) %>%
    select(id, daterated, ovtoday)
) %>%
  filter(!is.na(daterated)) %>%
  group_by(id, daterated) %>%
  summarize(
    menses = max(menses, na.rm = TRUE),
    ovtoday = max(ovtoday, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    menses = if_else(is.finite(menses), menses, NA_real_),
    ovtoday = if_else(is.finite(ovtoday), ovtoday, NA_real_)
  )

clean_df <- clean_df %>%
  mutate(id = as.character(id))

# Also ensure wide_cycles has character id (already done on line 1361, but double-check)
wide_cycles <- wide_cycles %>%
  mutate(id = as.character(id))


# CHECKPOINT A: Right BEFORE the cycle data merge (before line 1605)
cat("\n=== CHECKPOINT A: Before Cycle Merge ===\n")
clean_df_dupes <- clean_df %>%
  group_by(id, daterated) %>%
  filter(n() > 1) %>%
  summarise(n_rows = n(), .groups = "drop")
cat("Duplicate (id, daterated) in clean_df:", nrow(clean_df_dupes), "\n")

#View(clean_df)

# ---- 5. Merge Event Flags and Metadata into Main Dataframe ----
final_df <- clean_df %>%
  # Merge NON-RTB cycle events
  full_join(events_long, by = c("id", "daterated")) %>%
  # Merge RTB cycle events
  full_join(events_long_rtb, by = c("id", "daterated")) %>%
  # Merge cycle lengths from both versions
  left_join(wide_cycles_with_length %>%
              select(id, cycle_length, cycle_length_rtb),
            by = "id") %>%
  # Merge participant metadata from ALL participants (not just those with valid cycles)
  left_join(
    wide_cycles %>%
      select(
        id,
        Ov_Status,
        Assay_Batch,
        Assay_Plate,
        Assay_Plate_orig,
        espike1,
        espike2,
        pspike1,
        pspike2,
        n_assay_plates,
        starts_with("in_plate_"),
        Excl_Horm,
        Excl_CycData
      ),
    by = "id"
    ) %>%
  mutate(
    # Fill NAs for event flags with 0 (missing = no event on that day)
    menses = replace_na(menses, 0),
    ovtoday = replace_na(ovtoday, 0),
    menses_rtb = replace_na(menses_rtb, 0),
    ovtoday_rtb = replace_na(ovtoday_rtb, 0),
    id = as.character(id),
  ) %>%
  arrange(id, daterated)


# ADD CHECKPOINT RIGHT HERE (after line ~1650):
cat("\n=== CHECKPOINT:  After All Cycle Merges ===\n")
final_df_dupes <- final_df %>%
  group_by(id, daterated) %>%
  filter(n() > 1) %>%
  summarise(n_rows = n(), .groups = "drop")
cat("Duplicate (id, daterated) in final_df:", nrow(final_df_dupes), "\n")

if (nrow(final_df_dupes) > 0) {
  cat("WARNING: Found duplicate (id, daterated) combinations!\n")
  cat("Showing first 20 duplicates:\n")
  print(head(final_df_dupes, 20))
  
  # Show example of what's duplicated for ID 118
  cat("\n=== Example:  ID 118 duplicates ===\n")
  final_df %>%
    filter(id == "118") %>%
    group_by(daterated) %>%
    filter(n() > 1) %>%
    select(id, daterated, Excl_Horm, e2, p4, lh, menses, ovtoday) %>%
    arrange(daterated) %>%
    print(n = 20)
}



# ---- 7. Final Output and Diagnostics ----
cat("\n=== FINAL DATASET SUMMARY ===\n")
cat("Final Rows:", nrow(final_df), "\n")
cat("Unique IDs:", n_distinct(final_df$id), "\n")

cat("\n=== NON-RTB VERSION (Original MP dates) ===\n")
cat("IDs with menses events:", n_distinct(final_df$id[final_df$menses == 1]), "\n")
cat("IDs with ovulation events:", n_distinct(final_df$id[final_df$ovtoday == 1]), "\n")
cat("Total menses events:", sum(final_df$menses, na.rm = TRUE), "\n")
cat("Total ovulation events:", sum(final_df$ovtoday, na.rm = TRUE), "\n")

cat("\n=== RTB VERSION (RTB-prioritized dates) ===\n")
cat("IDs with menses_rtb events:", n_distinct(final_df$id[final_df$menses_rtb == 1]), "\n")
cat("IDs with ovulation_rtb events:", n_distinct(final_df$id[final_df$ovtoday_rtb == 1]), "\n")
cat("Total menses_rtb events:", sum(final_df$menses_rtb, na.rm = TRUE), "\n")
cat("Total ovulation_rtb events:", sum(final_df$ovtoday_rtb, na.rm = TRUE), "\n")

count_rows_ids(final_df)

# ---- 8. Track Excluded IDs for Documentation ----
# Track IDs that were excluded due to various reasons for later reference
# Note: ids_excl_all was already defined earlier when applying exclusions

ids_excl_cycdata <- wide_cycles %>% 
  filter(Excl_CycData == TRUE) %>% 
  pull(id)

# Manual exclusions (hardcoded removals)
ids_manual_excl <- c() # Add any manual exclusions here if needed

# Combined list of IDs with no menses data (excluded or missing cycle data)
ids_w_no_menses_onset <- unique(c(ids_excl_all, ids_excl_cycdata, ids_manual_excl))

cat("\n=== IDs EXCLUDED FROM CYCLE DATA ===\n")
cat("IDs with Excl_all_id:", paste(ids_excl_all, collapse = ", "), "\n")
cat("IDs with Excl_CycData:", paste(ids_excl_cycdata, collapse = ", "), "\n")
cat("IDs with manual exclusions:", paste(ids_manual_excl, collapse = ", "), "\n")
cat("All IDs without menses data:", paste(ids_w_no_menses_onset, collapse = ", "), "\n")

# Assign final_df back to df for consistency with later code
df <- final_df

```


```{r verify-column-names}
#print column names to verify
cat("\n=== FINAL DATAFRAME COLUMNS ===\n")
print(colnames(df))
```


```{r calc-day-of-week}
#Create day of week variables including categorical, sin, and cos variables
df <- df %>%
  mutate(day_of_week = weekdays(as.Date(daterated)),
         day_of_week = factor(day_of_week, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
         day_of_week_sin = sin(2 * pi * as.numeric(day_of_week) / 7),
         day_of_week_cos = cos(2 * pi * as.numeric(day_of_week) / 7))

```

```{r visualize-cycle-events}
# Step 5: Visualize Period Start Dates Over Time with Color-Coded Tiles
periodstartplot <- df %>%
  ggplot(aes(x = daterated, y = id, fill = factor(menses))) +
  geom_tile() +
  labs(
    title = "Period Start Dates Over Time",
    x = "Date",
    y = "ID",
    fill = "Period Start"
  ) +
  scale_fill_manual(values = c("grey", "red")) +
  theme_minimal(base_size = 14)

periodstartplot

# Save to data prep subfolder, specify periodstartplot
save_png(plot = periodstartplot, 
         filename = "period_start_dates_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap

# print location of saved file
cat("Period start dates heatmap saved\n")

# Step 6: Visualize Ovulation Dates Over Time
ovulationplot <- df %>%
  ggplot(aes(x = daterated, y = id, fill = factor(ovtoday))) +
  geom_tile() +
  labs(
    title = "Ovulation Dates Over Time",
    x = "Date",
    y = "ID",
    fill = "Ovulation Date"
  ) +
  scale_fill_manual(values = c("grey", "blue")) +
  theme_minimal(base_size = 14)

# Save to data prep subfolder, specify ovulationplot
save_png(plot = ovulationplot, 
         filename = "ovulation_dates_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap

ovulationplot

#df %>% select(id, daterated, menses, ovtoday) %>% View()

##vis_miss(df)

```

```{r horm-import-merge}
# =============================================================================
# HORMONE DATA IMPORT & MERGE
# =============================================================================
# Imports hormone assay results (e2, p4, lh) from Excel file and merges with
# daily survey data. This is a complex merge due to date format inconsistencies.
#
# WHAT IT DOES:
# 1. Loads hormone Excel file with mixed date formats
# 2. Consolidates duplicate columns (date, TubeNumber, hormones)
# 3. Assigns calendar dates to samples using written dates OR calculated dates
# 4. Deduplicates hormone data (one row per person per day)
# 5. Merges with survey data using FULL JOIN
#
# HORMONE VARIABLES:
# - e2: Estradiol (pg/mL) - primary female sex hormone
# - p4: Progesterone (ng/mL) - maintains uterine lining, rises after ovulation
# - lh: Luteinizing Hormone (mIU/mL) - triggers ovulation
#
# DATA COLLECTION METHOD:
# - Participants self-collected saliva samples daily
# - Samples numbered sequentially (TubeNumber: 1, 2, 3...)
# - Some tubes had handwritten dates, others did not
# - All samples batch-analyzed for hormone concentrations
#
# COMPLEX DATE ASSIGNMENT LOGIC (Lines 1274-1301):
# Challenge: Not all tubes have written dates, need to infer collection date
# Strategy: Use TWO methods and prioritize written dates
#
# Method A (parsed_date):
#   - Parse any written date from tube label
#   - Try multiple formats: ymd, mdy, dmy
#   - Example: "05/15/2021" → 2021-05-15
#
# Method B (calculated_date):
#   - Calculate from TubeNumber and study start date
#   - Logic: TubeNumber 1 = Day 1, TubeNumber 5 = Day 5 (start + 4 days)
#   - Requires: study_start date from first survey for each participant
#   - Example: If start = Jan 1, TubeNumber 5 → Jan 5
#
# FINAL DECISION RULE (Line 1298):
#   - Use parsed_date if available (handwritten date)
#   - Fall back to calculated_date if no written date
#   - Rationale: Written dates are more reliable but not always present
#
# DEDUPLICATION (Lines 1307-1314):
# Problem: Multiple hormone measurements for same person-day
# Causes: Re-runs, duplicate entries in lab file
# Solution: Keep first row after sorting by:
#   1. Valid e2 values first (is.na(e2))
#   2. Rows with explicit dates first (is.na(raw_date))
# Rationale: Prioritize complete, explicitly dated measurements
#
# MERGE STRATEGY (Lines 1316-1333):
# - Uses FULL JOIN (not inner or left join)
# - Preserves survey-only rows (have survey but no hormone data)
# - Preserves hormone-only rows (have hormone but no survey data)
# - Creates flag: in_original_survey (TRUE/FALSE)
#
# WHY FULL JOIN:
# - Some days have survey but no saliva sample collected
# - Some days have saliva sample but no survey completed
# - Need to keep both types for maximum data retention
# - Can analyze survey-only or hormone-only subsets later
#
# "GHOST ROWS" (Lines 1345-1346):
# - Rows added solely for hormone data (no survey data that day)
# - Created when participant collected saliva but didn't complete survey
# - Have e2/p4/lh values but NA for all survey variables
# - Useful for hormone trajectory plots even without survey data
#
# DATA RECOVERY (Lines 1341-1342):
# - Counts rows saved by calculated_date method
# - Shows how many samples would have been lost without calculation approach
# - Validates that calculation strategy is adding value
#
# INTERMEDIATE SAVE (Lines 1357-1358):
# - Saves checkpoint: df_with_hormones.rds
# - Allows restarting analysis from this point
# - Useful if subsequent steps fail
#
# KNOWN ISSUES:
# - TODO Line 1377: Need to update pipeline with better cycle date integration
# - Missing hormone data for some ID ranges (noted in comments)
# - Some date calculations may be off by 1 day (check TubeNumber=1 should be start date)
#
# INPUTS:
#   - Excel: ukalc_all_horm_data_20251218.xlsx
#   - df: Survey data with study_start dates
# OUTPUTS:
#   - df: Now includes e2, p4, lh, TubeNumber
#   - Intermediate file: df_with_hormones.rds
# NEW ROWS: May add "ghost rows" for hormone-only days
# =============================================================================

# ==============================================================================
# 1. Load and Clean Hormone Data
# ==============================================================================
# Use read_excel with col_types="text" to safely read mixed data without crashing
rawhormdf <- read_excel(
  "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/02_datasets/UKALC_HORM/02_data_prep_workspace/ukalc_all_horm_data_20251218.xlsx",
  col_types = "text",           
  .name_repair = "unique_quiet" 
) %>%
  select(-matches("^\\.\\.\\."))

# --- Helper: Safely Merge Duplicate Columns ---
coalesce_pattern <- function(df, pattern, new_name) {
  # Find columns matching the regex pattern
  cols_to_fix <- grep(pattern, names(df), value = TRUE, ignore.case = TRUE)
  
  if (length(cols_to_fix) == 0) return(df)
  
  # Create new column by taking the first non-NA value from the found columns
  df <- df %>% mutate(!!new_name := coalesce(!!!syms(cols_to_fix)))
  
  # Remove the old columns (unless it's the same name as the new one)
  cols_to_drop <- setdiff(cols_to_fix, new_name)
  if (length(cols_to_drop) > 0) df <- df %>% select(-all_of(cols_to_drop))
  
  return(df)
}

hormdf <- rawhormdf %>%
  # 1. Consolidate Tube Numbers (e.g., "Tube Number", "Tube_Number")
  coalesce_pattern("Tube[_ ]?Number", "TubeNumber") %>%
  
  # 2. Consolidate Dates (Crucial Fix for your issue)
  # This regex "^date$|^daterated$" looks specifically for "date" OR "daterated"
  coalesce_pattern("^date$|^daterated$|Collection[_ ]?Date", "raw_date") %>%
  
  # 3. Consolidate Hormone Columns
  coalesce_pattern("^e2", "e2") %>%
  coalesce_pattern("^p4", "p4") %>%
  coalesce_pattern("^lh", "lh") %>%
  
  # 4. Clean & Format Values
  mutate(
    id = as.character(gsub("[^0-9]", "", id)),
    TubeNumber = suppressWarnings(as.numeric(TubeNumber)),
    across(any_of(c("e2", "p4", "lh")), ~ suppressWarnings(as.numeric(.)))
  ) %>%
  
  # 5. Filter: Must have ID and TubeNumber to be useful
  filter(!is.na(id), !is.na(TubeNumber))

# ==============================================================================
# 2. Assign Calendar Dates to Tubes
# ==============================================================================
if (!exists("df")) stop("Main dataframe 'df' must be loaded to calculate dates.")

# A. Calculate Study Start Date (Day 1) for each person from main survey data
study_start_dates <- df %>%
  mutate(daterated = as_date(daterated)) %>%
  group_by(id) %>%
  summarize(
    study_start = min(daterated, na.rm = TRUE),
    .groups = "drop"
  )

# B. Apply Logic: "Written Date" > "Calculated Date"
hormdf <- hormdf %>%
  left_join(study_start_dates, by = "id") %>%
  mutate(
    # Strategy A: Use the written date from the tube (if it exists)
    parsed_date = as_date(parse_date_time(raw_date, orders = c("ymd", "mdy", "dmy"), quiet = TRUE)),
    
    # Strategy B: Calculate it based on the Tube Index
    # Logic: If Start Date is Jan 1st, Tube 5 is Jan 5th (Start + 4 days)
    calculated_date = study_start + days(TubeNumber - 1),
    
    # Final Decision: Use parsed if available, otherwise default to calculated
    daterated = coalesce(parsed_date, calculated_date)
  ) %>%
  # Drop rows where we failed to find ANY date (likely ID missing from df)
  filter(!is.na(daterated))

# ==============================================================================
# 3. Deduplicate & Merge
# ==============================================================================

# A. Deduplicate: One hormone row per ID/Day
hormdf_clean <- hormdf %>%
  # Sort preference: Valid e2 values first, then rows that had explicit raw_dates
  arrange(id, daterated, is.na(e2), is.na(raw_date)) %>% 
  group_by(id, daterated) %>%
  slice(1) %>%
  ungroup() %>%
  select(id, daterated, TubeNumber, any_of(c("e2", "p4", "lh")))

# B. Prepare Main Data (Flagging original rows)
df_ready <- df %>%
  mutate(
    id = as.character(id), 
    daterated = as_date(daterated),
    in_original_survey = TRUE
  ) %>%
  # Clean out old hormone columns to prevent duplicates
  select(-any_of(c("e2", "p4", "lh", "TubeNumber_Hormone")))

# C. FULL JOIN (Preserves Surveys AND Hormone-Only rows)
final_df <- df_ready %>%
  full_join(hormdf_clean, by = c("id", "daterated")) %>%
  mutate(
    # If 'in_original_survey' is NA, it means this row came from the hormone file
    in_original_survey = replace_na(in_original_survey, FALSE)
  ) %>%
  arrange(id, daterated)


# ===== ADD THIS FIX =====
# Fill Excl_Horm for hormone-only rows from the participant's other rows
cat("\n=== Filling Excl_Horm for hormone-only rows ===\n")

excl_before <- final_df %>%
  group_by(id) %>%
  summarise(n_na = sum(is.na(Excl_Horm))) %>%
  summarise(total_na = sum(n_na)) %>%
  pull(total_na)

cat("Rows with Excl_Horm = NA before fill:", excl_before, "\n")

final_df <- final_df %>%
  group_by(id) %>%
  mutate(
    Excl_Horm = first(na.omit(Excl_Horm)) # Use the participant's Excl_Horm value for all their rows
  ) %>%
  ungroup()

excl_after <- final_df %>%
  group_by(id) %>%
  summarise(n_na = sum(is.na(Excl_Horm))) %>%
  summarise(total_na = sum(n_na)) %>%
  pull(total_na)

cat("Rows with Excl_Horm = NA after fill:", excl_after, "\n")
# ===== END OF FIX =====


# CHECKPOINT 2: After merging hormones into final_df (around line 1950)
cat("\n=== CHECKPOINT 2: After Hormone Merge (Before Excl_Horm) ===\n")
hormone_check_2 <- final_df %>%
  group_by(id) %>%
  summarise(
    n_e2 = sum(!is.na(e2)),
    has_Excl_Horm = any(! is.na(Excl_Horm)),
    Excl_Horm_value = first(Excl_Horm)
  ) %>%
  filter(n_e2 > 0)
cat("IDs with any hormone data:", n_distinct(hormone_check_2$id), "\n")
cat("IDs with >10 hormone obs:", sum(hormone_check_2$n_e2 > 10), "\n")
cat("IDs with Excl_Horm = TRUE:", sum(hormone_check_2$Excl_Horm_value == TRUE, na.rm = TRUE), "\n")





# Set e2, p4, lh to NA for participants with Excl_Horm = TRUE
final_df <- final_df %>%
  mutate(
    e2 = if_else(Excl_Horm %in% TRUE, NA_real_, e2),
    p4 = if_else(Excl_Horm %in% TRUE, NA_real_, p4),
    lh = if_else(Excl_Horm %in% TRUE, NA_real_, lh)
  )







# CHECKPOINT 3: After applying Excl_Horm (around line 1962)
cat("\n=== CHECKPOINT 3: After Excl_Horm Applied ===\n")
hormone_check_3 <- final_df %>%
  group_by(id) %>%
  summarise(n_e2 = sum(!is.na(e2))) %>%
  filter(n_e2 > 0)
cat("IDs with any hormone data:", n_distinct(hormone_check_3$id), "\n")
cat("IDs with >10 hormone obs:", sum(hormone_check_3$n_e2 > 10), "\n")






# Set e2 to NA for id-date combinations indicated by espike1 date or espike2 date
# Note: espike1, espike2, pspike1, pspike2 are already in final_df from earlier join at line 1617-1618
final_df <- final_df %>%
  mutate(
    e2 = if_else(
      (! is.na(espike1) & daterated == espike1) | 
      (!is.na(espike2) & daterated == espike2), 
      NA_real_, 
      e2
    ),
    p4 = if_else(
      (! is.na(pspike1) & daterated == pspike1) | 
      (!is.na(pspike2) & daterated == pspike2), 
      NA_real_, 
      p4
    )
  ) %>%
  select(-espike1, -espike2, -pspike1, -pspike2)




# CHECKPOINT 4: After applying spike date filtering (around line 1970)
cat("\n=== CHECKPOINT 4: After Spike Date Filtering ===\n")
hormone_check_4 <- final_df %>%
  group_by(id) %>%
  summarise(n_e2 = sum(!is.na(e2))) %>%
  filter(n_e2 > 0)
cat("IDs with any hormone data:", n_distinct(hormone_check_4$id), "\n")
cat("IDs with >10 hormone obs:", sum(hormone_check_4$n_e2 > 10), "\n")

# CHECKPOINT: Verify Excl_Horm is still working (no leakage from spike filtering)
cat("\n=== CHECKPOINT: Verify Excl_Horm After Spike Filtering ===\n")
excl_horm_diagnostic <- final_df %>%
  filter(Excl_Horm == TRUE) %>%
  summarise(
    n_rows = n(),
    n_with_e2 = sum(!is.na(e2)),
    n_with_p4 = sum(!is.na(p4)),
    n_with_lh = sum(!is.na(lh))
  )
print(excl_horm_diagnostic)
if (any(excl_horm_diagnostic[c("n_with_e2", "n_with_p4", "n_with_lh")] > 0)) {
  warning("WARNING: Excl_Horm=TRUE participants still have hormone values after spike filtering!")
}





# ==============================================================================
# 4. Diagnostics
# ==============================================================================
cat("Final Rows:", nrow(final_df), "\n")

# Count rows recovered by calculation
recovered_rows <- hormdf %>% filter(is.na(parsed_date) & !is.na(calculated_date))
cat("Rows with NO written date that were saved by calculation:", nrow(recovered_rows), "\n")

# Count Ghost Rows (Hormone data with no survey)
ghost_rows <- final_df %>% filter(in_original_survey == FALSE)
cat("New rows added solely for hormone data (no survey):", nrow(ghost_rows), "\n")

# Cleanup flag
final_df <- final_df %>% select(-in_original_survey)

# Assign final_df back to df for consistency with later code
df <- final_df

# ==============================================================================
# 5. View Results
# ==============================================================================
#View(df)
count_rows_ids(df)



names(df)

```

# Calculate aggregate DRSP scores
```{r calc-drsp-aggregates}
# =============================================================================
# DRSP COMPOSITE SCORE CALCULATION (FOR UKALC DIAGNOSTIC DATA)
# =============================================================================
# NOTE: These subscales are calculated here for the UKALC dataset specifically,
# but they are ALSO calculated in UKALC_03_processing_shared.Rmd for the 
# merged ADHD+UKALC dataset to ensure both studies receive identical processing.
# =============================================================================

# set drsp variables to numeric
df <- df %>%
  mutate(across(c("dep", "hopeless", "worthless", "anxious", "mood_swing", "rej_sens", 
                  "angry", "conflicts", "loss_interest", "diff_conc", "tired", 
                  "over_eat", "sp_food_crav", "high_sleep", "diff_sleep", "overwhelm", 
                  "loss_control", "breast_tend", "breast_swell", "headache", 
                  "jm_pain", "loss_prod", "int_relat"), as.numeric))

# Create total DRSP symptom score (means of all items)
df <- df %>%
  mutate(drsp_total = rowMeans(select(., 
      dep, hopeless, worthless, anxious, mood_swing, rej_sens, 
      angry, conflicts, loss_interest, diff_conc, tired, 
      over_eat, sp_food_crav, high_sleep, diff_sleep, overwhelm, 
      loss_control, breast_tend, breast_swell, headache, 
      jm_pain, loss_prod, int_relat
  ), na.rm = TRUE))

# Create "drsp_distress" - dep hope worth anx ang loss_interest overwhelm loss_control
df <- df %>%
  mutate(drsp_distress = rowMeans(select(., 
      dep, hopeless, worthless, anxious, overwhelm, loss_control), na.rm = TRUE))

# create drsp_labilsocsx - mood_swing, rej_sens, conflicts,int_relat, angry
df <- df %>%
  mutate(drsp_labilsocsx = rowMeans(select(., mood_swing, rej_sens, conflicts, int_relat, angry), na.rm = TRUE))

# create drsp_breastpain - breast_tend, breast_swell
df <- df %>%
  mutate(drsp_breastpain = rowMeans(select(., breast_tend, breast_swell), na.rm = TRUE))

# create drsp_eatsx - over_eat, sp_food_crav
df <- df %>%
  mutate(drsp_eatsx = rowMeans(select(., over_eat, sp_food_crav), na.rm = TRUE))

# create drsp_lowarousal - tired, diff_conc, diff_sleep, high_sleep
df <- df %>%
  mutate(drsp_lowarousal = rowMeans(select(., tired, diff_conc, diff_sleep, high_sleep, loss_interest, loss_prod, headache), na.rm = TRUE))

```

```{r winsorize-hormones}
library(dplyr)

# 1. Define a simple helper function to cap the top 5%
winsor_upper <- function(x, prob = 0.95) {
  # If all data is missing, return as is to avoid errors
  if (all(is.na(x))) return(x)
  
  # Calculate the cutoff
  limit <- quantile(x, probs = prob, na.rm = TRUE)
  
  # Replace values above the limit with the limit
  x[x > limit & !is.na(x)] <- limit
  return(x)
}

# 2. Apply it within-person
df <- df %>%
  group_by(id) %>%
  mutate(
    e2_w = winsor_upper(e2, 0.95),
    p4_w = winsor_upper(p4, 0.95),
    lh_w = winsor_upper(lh, 0.95)
  ) %>%
  ungroup()

# Enforce Excl_Horm exclusion on winsorized hormone variables (e2_w, p4_w, lh_w)
df <- df %>%
  mutate(
    e2_w = if_else(Excl_Horm %in% TRUE, NA_real_, e2_w),
    p4_w = if_else(Excl_Horm %in% TRUE, NA_real_, p4_w),
    lh_w = if_else(Excl_Horm %in% TRUE, NA_real_, lh_w)
  )

# Verify exclusion enforcement
cat("\n=== CHECKPOINT: After Winsorization ===\n")
excl_horm_check <- df %>%
  filter(Excl_Horm == TRUE) %>%
  summarise(
    n_with_e2 = sum(!is.na(e2)),
    n_with_e2_w = sum(!is.na(e2_w)),
    n_with_p4 = sum(!is.na(p4)),
    n_with_p4_w = sum(!is.na(p4_w)),
    n_with_lh = sum(!is.na(lh)),
    n_with_lh_w = sum(!is.na(lh_w))
  )
print(excl_horm_check)
if (any(excl_horm_check > 0)) {
  warning("WARNING: Excl_Horm=TRUE participants still have hormone values after winsorization!")
}
```

```{r note-derived-metrics-moved}
# =============================================================================
# NOTE: DERIVED METRICS CALCULATION MOVED TO UKALC_03_processing_shared.Rmd
# =============================================================================
# 
# Previous versions of this file calculated derived variables here (.d, .zd, 
# .szd, .roll, .log, .m, .sd). This caused problems when merging with ADHD data
# because ADHD observations never received these transformations.
#
# NOW: This file outputs CLEAN data only (no derived variables)
# - Output: data/ukalc_YYYYMMDD_clean.rds
# - Next step: UKALC_02 merges clean datasets
# - Final step: UKALC_03 applies ALL processing to merged data
#
# This ensures both UKALC and ADHD observations receive identical processing.
# =============================================================================
```

```{r remove-dupes}
df <- df %>%
  distinct(id, daterated, .keep_all = TRUE)
```

```{r how-many-drinking-ids}
#How many unique ids are in the dataset? 
ids_in_am_dataset <- n_distinct(df$id) # 83

# How many unique participants (factor var id) reported drinking at least once? 
unique_drinkers <- n_distinct(df$id[df$drink_today == "Yes"]) # 79

# Print out each id with the number of times that each id reported drinking
df %>%
  filter(drink_today == "Yes") %>%
  count(id) %>%
  arrange(desc(n))

table(df$drink_today)

# use cat to print out the number of unique ids and unique drinkers
cat("Number of unique ids in the dataset: ", ids_in_am_dataset, "\n")


```

```{r visualize-drinks-fourplus}

p_drink_today <- df %>% 
  filter(!is.na(drink_today)) %>% 
  # CHANGE 1: Wrap drink_today in as.factor() to make it discrete
  ggplot(aes(x = daterated, y = id, fill = as.factor(drink_today))) +
  geom_tile() +
  labs(title = "Drinking Across the Study by ID",
       x = "Date",
       y = "ID",
       fill = "Drinking Today") +
  # CHANGE 2: Quote the numbers ("1" and "0") so R recognizes them as names
  scale_fill_manual(values = c("Yes" = "blue", "No" = "lightgray")) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(size = 8))

print(p_drink_today)

# Save the plot
save_png(plot = p_drink_today, 
         filename = "drink_today_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap



p_fourplustoday <- df %>% 
  filter(!is.na(fourplustoday)) %>% 
  # CHANGE 1: Wrap fourplustoday in as.factor() to make it discrete
  ggplot(aes(x = daterated, y = id, fill = as.factor(fourplustoday))) +
  geom_tile() +
  labs(title = "Drinking 4+ Drinks/Day Across the Study by ID",
       x = "Date",
       y = "ID",
       fill = "Drinking 4+ Today") +
  # CHANGE 2: Quote the numbers ("1" and "0") so R recognizes them as names
  scale_fill_manual(values = c("Yes" = "blue", "No" = "lightgray")) + 
  theme_minimal(base_size = 14) +
  theme(axis.text.y = element_text(size = 8))

print(p_fourplustoday)

# Save the plot
save_png(plot = p_fourplustoday, 
         filename = "fourplustoday_heatmap.png",
         folder = data_prep_folder,
         height = 8)  # Keep taller height for heatmap

```

# Save cycledata and save as traits
```{r merge-cycledata-traits}

availablecycledata <- cycledata_check(df, c("dep"))

# format id as same format for merging 
availablecycledata <- availablecycledata$by_id %>%
  mutate(id = as.character(id))

# merge drink_cycledata into df
df <- df %>%
  left_join(availablecycledata, by = "id")

# check merge
cat("Cycledata merged into df. Here are the new columns from availablecycledata:\n")
new_cycledata_cols <- setdiff(names(availablecycledata), names(df))
print(new_cycledata_cols)

```




```{r count rows and ids}
count_rows_ids(df)
```

```{r data N details}

# Filter the data for non-missing e2 values and count unique ids
non_missing_E2 <- df %>%
  filter(!is.na(e2))

non_missing_E2_count <- n_distinct(non_missing_E2$id)
non_missing_E2_ids <- unique(non_missing_E2$id)

# Get IDs with missing e2 values
missing_E2_ids <- setdiff(unique(df$id), non_missing_E2_ids)

# Print results for non-missing and missing e2 values
cat("Number of people WITH usable hormones:", non_missing_E2_count, "\n",
    "The id values are:", paste(non_missing_E2_ids, collapse = ", "), "\n\n",
    "Number of people WITHOUT usable hormone data:", length(missing_E2_ids), "\n",
    "Their id values are:", paste(missing_E2_ids, collapse = ", "), "\n\n",
    "In addition, these ids were removed due to a lack of complete data:", 
    paste(ids_w_no_menses_onset, collapse = ", "), "\n\n")

# Filter the data for non-missing survey values and count unique ids
non_missing_survey <- df %>%
  filter(!is.na(drank_alcohol))

non_missing_survey_count <- n_distinct(non_missing_survey$id)
non_missing_survey_ids <- unique(non_missing_survey$id)

# Get IDs with missing survey values
missing_survey_ids <- setdiff(unique(df$id), non_missing_survey_ids)

# Print results for non-missing and missing survey values
cat("Number of people WITH any surveys:", non_missing_survey_count, "\n",
    "Their id values are:", paste(non_missing_survey_ids, collapse = ", "), "\n\n",
    "Number of people WITHOUT any surveys:", length(missing_survey_ids), "\n",
    "Their id values are:", paste(missing_survey_ids, collapse = ", "), "\n\n")

# Find the intersection of available hormone data and available survey data
intersection_ids <- intersect(non_missing_E2_ids, non_missing_survey_ids)
intersection_count <- length(intersection_ids)

# Print result for the intersection of hormone and survey data
cat("Number of people WITH both usable hormone and any survey data:", intersection_count, "\n",
    "Their id values are:", paste(intersection_ids, collapse = ", "), "\n\n")

# Count the non-missing observations of drink_today and e2 for people with both hormone and survey data
non_missing_intersection <- df %>%
  filter(id %in% intersection_ids, !is.na(drink_today), !is.na(e2))

non_missing_intersection_count <- nrow(non_missing_intersection)

# Print the result for non-missing observations of drink_today and e2
cat("Number of non-missing observations of drink_today and e2 for people with both hormone and survey data:", non_missing_intersection_count, "\n")

# Create a list of IDs with the number of days with both e2 and DRSP
usable_days_list <- non_missing_intersection %>%
  group_by(id) %>%
  summarise(usable_days = n()) %>%
  arrange(usable_days)

# Print the list of IDs with the number of days in words
cat("List of IDs with the number of days having both e2 and DRSP (sorted by number of observations ascending):", "\n")
apply(usable_days_list, 1, function(row) {
  cat("id", row["id"], "has", row["usable_days"], "hormone-survey pairs.", "\n")
})

```

```{r create-days-since-enrollment}
# Create a variable called days_since_enrollment that counts the number of days since the first day of data collection (daterated) for each participant (id) in df
df <- df %>%
  group_by(id) %>%
  mutate(days_since_enrollment = daterated - min(daterated))

# Create a variable that counts the number of rows per id and saves it as a trait variable (repeating within each id)
df <- df %>%
  group_by(id) %>%
  mutate(id_daycount = n())

hist(df$id_daycount)
```

```{r data-overlap-horm-surv-cyc, message=FALSE, warning=FALSE}


# ---- Plot Overlap of Available Data  ----        
# Plot days in study, nonmissing survey observations, and e2 observations per id 
#TODO need to add hormone data into this and redo
library(gridExtra)
# Count the total number of "Days in Study" (total observations per ID)
df_count <- df %>%
  group_by(id) %>%
  summarize(id_daycount = n())

# Count the number of non-missing observations per ID for the variable drink_today_am
df_nonmissing_count <- df %>%
  group_by(id) %>%
  summarize(number_of_observations = sum(!is.na(drink_today)))

# Count the number of non-missing observations for the variable e2
df_e2_nonmissing_count <- df %>%
  group_by(id) %>%
  summarize(e2_nonmissing = sum(!is.na(e2)))

# Count the number of non-missing observations for the variable dep
df_dep_nonmissing_count <- df %>%
  group_by(id) %>%
  summarize(dep_nonmissing = sum(!is.na(dep)))

#### Handle Outliers ONLY for "Days in Study" ####

# Outlier removal for "Days in Study"
Q1 <- quantile(df_count$id_daycount, 0.25)
Q3 <- quantile(df_count$id_daycount, 0.75)
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

df_no_outliers <- df_count %>%
  filter(id_daycount >= lower_bound & id_daycount <= upper_bound)

# remove id 123 from df_dep_nonmissing_count 
df_dep_nonmissing_count <- df_dep_nonmissing_count %>%
  filter(id != 123)

#### Determine Y-Axis Limits ####
y_min <- min(df_no_outliers$id_daycount, df_nonmissing_count$number_of_observations, df_e2_nonmissing_count$e2_nonmissing, df_dep_nonmissing_count$dep_nonmissing, na.rm = TRUE)
y_max <- max(df_no_outliers$id_daycount, df_nonmissing_count$number_of_observations, df_e2_nonmissing_count$e2_nonmissing, df_dep_nonmissing_count$dep_nonmissing, na.rm = TRUE)

#### Plots ####

# Plot "Days in Study" per ID (with outliers removed)
plot1_no_outliers <- ggplot(df_no_outliers, aes(x = factor(1), y = id_daycount, label = id)) +
  geom_violin(fill = "lightblue", alpha = 0.5) +  # Violin plot to show distribution
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +  # Boxplot inside the violin
  geom_jitter(width = 0.15, size = 2, color = "blue", alpha = 0.7) +  # Dots for raw data
  geom_text(aes(label = id), hjust = -0.1, size = 3, check_overlap = TRUE) +  # Add ID labels
  labs(x = NULL, y = "Days in Study", title = "Days in Study per ID") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  ylim(y_min, y_max)

# Plot "Number of Observations" for drink_today_am (without removing outliers)
plot2_no_outliers <- ggplot(df_nonmissing_count, aes(x = factor(1), y = number_of_observations, label = id)) +
  geom_violin(fill = "lightgreen", alpha = 0.5) +  # Violin plot to show distribution
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +  # Boxplot inside the violin
  geom_jitter(width = 0.15, size = 2, color = "darkgreen", alpha = 0.7) +  # Dots for raw data
  geom_text(aes(label = id), hjust = -0.1, size = 3, check_overlap = TRUE) +  # Add ID labels
  labs(x = NULL, y = "Alcohol Survey Days", title = "AM Surveys per ID") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  ylim(y_min, y_max)

# Plot "Number of non-missing e2 observations" (without removing outliers)
plot3_no_outliers <- ggplot(df_e2_nonmissing_count, aes(x = factor(1), y = e2_nonmissing, label = id)) +
  geom_violin(fill = "lightcoral", alpha = 0.5) +  # Violin plot to show distribution
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +  # Boxplot inside the violin
  geom_jitter(width = 0.15, size = 2, color = "red", alpha = 0.7) +  # Dots for raw data
  geom_text(aes(label = id), hjust = -0.1, size = 3, check_overlap = TRUE) +  # Add ID labels
  labs(x = NULL, y = "Number of Hormone Days", title = "Hormone Days per ID") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  ylim(y_min, y_max)

# Plot "Number of non-missing dep observations" (without removing outliers)
plot4_no_outliers <- ggplot(df_dep_nonmissing_count, aes(x = factor(1), y = dep_nonmissing, label = id)) +
  geom_violin(fill = "orange", alpha = 0.5) +  # Violin plot to show distribution
  geom_boxplot(width = 0.1, outlier.shape = NA, alpha = 0.7) +  # Boxplot inside the violin
  geom_jitter(width = 0.15, size = 2, color = "red", alpha = 0.7) +  # Dots for raw data
  geom_text(aes(label = id), hjust = -0.1, size = 3, check_overlap = TRUE) +  # Add ID labels
  labs(x = NULL, y = "Number of DRSP Days", title = "PM Surveys per ID") +
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  ylim(y_min, y_max)

# Combine the three plots on one page using grid.arrange from gridExtra
observations_plot <- grid.arrange(plot1_no_outliers, plot2_no_outliers, plot3_no_outliers, plot4_no_outliers, ncol = 4)

# Save the combined plot to a file with the current date
save_png(plot = observations_plot, 
         filename = paste0("N_observations_plot_", current_date, ".png"),
         folder = data_prep_folder,
         width = 20,  # Keep wider for 3-column layout
         height = 8)



```




```{r print variable names}

names(df)

```





##### END OF DATA PREP 

```{r save-prepped-dataset-box}

write.csv(df, paste0("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/03_analytic_projects/UKALC-Primary ms/02_datasets_output/ukalc_", current_date, "_clean.csv"), row.names = FALSE)

# Save as .rds
saveRDS(df, paste0("~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/UKALC/03_analytic_projects/UKALC-Primary ms/02_datasets_output/ukalc_", current_date, "_clean.rds"))
```


```{r save-prepped-dataset-data-folder}
# output a copy to the data folder
write.csv(df, paste0("data/ukalc_", current_date, "_clean.csv"), row.names = FALSE)
# Save as .rds
saveRDS(df, paste0("data/ukalc_", current_date, "_clean.rds"))

cat("\n✓ Clean dataset saved (no derived variables).\n")
cat("  Output: data/ukalc_", current_date, "_clean.rds\n", sep="")
cat("  Next step: Run UKALC_02_merge_adhd.Rmd to merge with ADHD data.\n")
cat("  Then: Run UKALC_03_processing_shared.Rmd to create derived variables.\n")
```

```{r placeholder-empty}

```

