---
title: "HERA_GAMMs_cyclic"
author: "Tory Eisenlohr-Moul, PhD"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_caption: true
    code_folding: show
    self_contained: true
---




# =============================================================================
# GENERALIZED ADDITIVE MIXED MODELS (GAMMs) FOR CYCLICAL ANALYSIS
# =============================================================================
# This script runs GAMMs to model cyclical patterns in alcohol use and other
# outcomes across the menstrual cycle. It processes both binary and continuous
# outcome variables using menses-centered and ovulation-centered time frames.
#
# WHAT IT DOES:
# 1. Sources setup file for packages and output folder configuration
# 2. Loads preprocessed hormone/daily survey data
# 3. Runs GAMMs for binary outcomes (e.g., drink_today, felt_drunk)
# 4. Runs GAMMs for continuous outcomes (e.g., num_drinks, cravings)
# 5. Creates visualizations with rolling averages and model predictions
# 6. Saves all outputs (models, summaries, plots) to dated output folder
#
# OUTPUT STRUCTURE:
# - All outputs saved to: output_folder/[outcome_name]_gamm/
# - Each outcome gets: models (.rds), summaries (.txt), and plots (.png)
#
# MODELS:
# - Binary outcomes: Logistic regression with cyclic splines
# - Continuous outcomes: Gaussian regression on log-transformed data
# - Both include: random intercepts, cyclic trends, random cyclic slopes
# - Covariates: 
#   * day_of_week: controls for weekday effects (e.g., weekend drinking)
#   * session_today: controls for laboratory visit effects
#
# NOTE: Core GAMM structure (cyclic splines, random effects) is preserved.
#       session_today was added to continuous-outcome models for consistency
#       with the binary-outcome models.
# =============================================================================

```{r setup-packages}
# Source R File with Setup, Package Loading, and Output Folder Creation
# This loads all required packages and creates the dated output folder
source("HERA_00_setup.R")

# Create subfolder for GAMM outputs (keep root-level output_folder unchanged)
gamms_folder <- file.path(output_folder, "03_gamms")
if (!dir.exists(gamms_folder)) {
  dir.create(gamms_folder, recursive = TRUE)
  cat("Created GAMMs subfolder:", gamms_folder, "\n")
}

```

# =============================================================================
# DATA IMPORT
# =============================================================================
# Load the preprocessed data file containing hormone measurements and daily
# survey responses. This dataset should include:
# - Participant ID (id)
# - Cyclic time variables (cyclic_time, cyclic_time_ov)
# - Binary outcome variables (drink_today, fourplustoday, felt_drunk)
# - Continuous outcome variables (num_drinks, cravings, symptoms)
# - Hormone measurements (e2, p4, lh)
# - Covariates (day_of_week, session_today)
# =============================================================================

```{r import-data}
# -------------------------------------------------------------------------
# Configure Data File Path
# -------------------------------------------------------------------------
# LOAD LATEST CLEAN DATA FILE 

# Load the dataset
df <- readRDS('/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_20260319_clean.rds')

message(paste0("Data loaded successfully: ", nrow(df), " rows, ", ncol(df), " columns"))
```

# =============================================================================
# CUSTOM FUNCTIONS CHECK
# =============================================================================
# This script uses custom plotting functions for rolling averages and data
# availability visualization. If these functions are not available, those
# specific plots will be skipped while GAMMs still run.
#
# REQUIRED CUSTOM FUNCTIONS:
# - cycle_plot(): Creates rolling average plots centered on menses or ovulation
# - cycledata_check(): Visualizes data availability across the cycle
#
# If you have these functions in a separate file:
# source("path/to/custom_functions.R")
# =============================================================================

```{r check-custom-functions}
# Check if custom functions are available
has_cycle_plot <- exists("cycle_plot") && is.function(cycle_plot)
has_cycledata_check <- exists("cycledata_check") && is.function(cycledata_check)

if (!has_cycle_plot) {
  message("WARNING: cycle_plot() function not found. Rolling average plots will be skipped.")
}

if (!has_cycledata_check) {
  message("WARNING: cycledata_check() function not found. Data availability plots will be skipped.")
}
```

```{r}
# set saturday as the reference level 

df$day_of_week <- factor(
  df$day_of_week,
  levels = c(
    "Saturday",
    "Sunday",
    "Monday",
    "Tuesday",
    "Wednesday",
    "Thursday",
    "Friday"
  )
)
```


```{r}
names(df)
```



<!-- # ============================================================================= -->
<!-- # BINARY OUTCOMES: GAMMs for Dichotomous Variables -->
<!-- # ============================================================================= -->
<!-- # This section analyzes binary outcomes (0/1) using logistic GAMMs. -->
<!-- # Each outcome is modeled with both menses-centered and ovulation-centered -->
<!-- # time frames to capture different aspects of cyclical patterns. -->
<!-- # -->
<!-- # BINARY OUTCOMES ANALYZED: -->
<!-- # - drink_today: Whether participant drank alcohol (yes/no) -->
<!-- # - fourplustoday: Whether participant had 4+ drinks (binge drinking) -->
<!-- # - felt_drunk: Whether participant reported feeling drunk -->
<!-- # -->
<!-- # MODEL SPECIFICATION: -->
<!-- # - Family: Binomial with logit link -->
<!-- # - Fixed effects: day_of_week, session_today -->
<!-- # - Cyclic smooth: s(cyclic_time, bs='cc') - captures average cycle pattern -->
<!-- # - Random intercept: s(id, bs='re') - accounts for baseline differences -->
<!-- # - Random cyclic slope: s(cyclic_time, id, bs=c('re','cc')) - individual variation -->
<!-- # -->
<!-- # OUTPUTS FOR EACH VARIABLE: -->
<!-- # - Rolling average plots (menses and ovulation centered) -->
<!-- # - Data availability check plot -->
<!-- # - Model objects (.rds files) -->
<!-- # - Model summaries (.txt files) -->
<!-- # - Prediction plots with confidence bands -->
<!-- # ============================================================================= -->

<!-- ```{r binary-outcomes-gamms} -->
<!-- # ------------------------------------------------------------------------- -->
<!-- # 1. SETUP: Define Binary Outcome Variables -->
<!-- # ------------------------------------------------------------------------- -->

<!-- # List of binary outcome variable names to analyze -->
<!-- binary_vars <- c("drink_today") #, "felt_drunk")  -->

<!-- # Helper function to convert p-values to significance stars -->
<!-- # *** p < 0.001, ** p < 0.01, * p < 0.05, . p < 0.1 -->
<!-- get_stars <- function(p) { -->
<!--   stats::symnum(p, corr = FALSE, na = FALSE,  -->
<!--                 cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),  -->
<!--                 symbols = c("***", "**", "*", ".", " ")) -->
<!-- } -->

<!-- # ------------------------------------------------------------------------- -->
<!-- # 2. LOOP THROUGH EACH BINARY OUTCOME -->
<!-- # ------------------------------------------------------------------------- -->

<!-- for (outcome_name in binary_vars) { -->

<!--   message(paste0("\n===== Processing Binary Outcome: ", outcome_name, " =====")) -->

<!--   # Create outcome-specific subdirectory in the gamms_folder -->
<!--   # This keeps all outputs for each variable organized together -->
<!--   save_dir <- file.path(gamms_folder, paste0(outcome_name, "_gamm")) -->

<!--   if (!dir.exists(save_dir)) { -->
<!--     dir.create(save_dir, recursive = TRUE) -->
<!--     message(paste0("Created directory: ", save_dir)) -->
<!--   } -->

<!--   # ----------------------------------------------------------------------- -->
<!--   # 3. RAW DATA VISUALIZATION: Rolling Averages -->
<!--   # ----------------------------------------------------------------------- -->
<!--   # These plots show the observed data patterns before model fitting. -->
<!--   # Rolling averages smooth out day-to-day noise to reveal trends. -->

<!--   # --- Menses-Centered Rolling Average Plot --- -->
<!--   # Time axis: ovulation (-1) to menses (0) to ovulation (+1) -->
<!--   if (has_cycle_plot) { -->
<!--     message("  Creating menses-centered rolling average plot...") -->
<!--     tryCatch({ -->
<!--       rollplot_menses <- cycle_plot( -->
<!--         df, -->
<!--         outcome_name, -->
<!--         centering = "menses", -->
<!--         include_impute = TRUE, -->
<!--         y_scale = "person-centered_roll", -->
<!--         se = TRUE -->
<!--       ) -->

<!--       save_png( -->
<!--         plot = rollplot_menses$plot, -->
<!--         filename = "rollplot_menses.png", -->
<!--         folder = save_dir -->
<!--       ) -->
<!--     }, error = function(e) { -->
<!--       message(paste0("  Error creating menses-centered plot: ", e$message)) -->
<!--     }) -->
<!--   } else { -->
<!--     message("  Skipping menses-centered rolling average plot (cycle_plot not available)") -->
<!--   } -->

<!--   # --- Ovulation-Centered Rolling Average Plot --- -->
<!--   # Time axis: menses (-1) to ovulation (0) to menses (+1) -->
<!--   if (has_cycle_plot) { -->
<!--     message("  Creating ovulation-centered rolling average plot...") -->
<!--     tryCatch({ -->
<!--       rollplot_ov <- cycle_plot( -->
<!--         df, -->
<!--         outcome_name, -->
<!--         centering = "ovulation", -->
<!--         include_impute = TRUE, -->
<!--         y_scale = "person-centered_roll",  -->
<!--         se = TRUE -->
<!--       ) -->

<!--       save_png( -->
<!--         plot = rollplot_ov$plot, -->
<!--         filename = "rollplot_ovulation.png", -->
<!--         folder = save_dir -->
<!--       ) -->
<!--     }, error = function(e) { -->
<!--       message(paste0("  Error creating ovulation-centered plot: ", e$message)) -->
<!--     }) -->
<!--   } else { -->
<!--     message("  Skipping ovulation-centered rolling average plot (cycle_plot not available)") -->
<!--   } -->

<!--   # --- Data Availability Check --- -->
<!--   # Visualizes which participants have data at which cycle times -->
<!--   if (has_cycledata_check) { -->
<!--     message("  Creating data availability plot...") -->
<!--     tryCatch({ -->
<!--       checkdata <- cycledata_check(df, symptom_columns = c(outcome_name)) -->

<!--       save_png( -->
<!--         plot = checkdata$data_symptom_plots, -->
<!--         filename = paste0("data_availability_", outcome_name, ".png"), -->
<!--         folder = save_dir -->
<!--       ) -->
<!--     }, error = function(e) { -->
<!--       message(paste0("  Error creating data availability plot: ", e$message)) -->
<!--     }) -->
<!--   } else { -->
<!--     message("  Skipping data availability plot (cycledata_check not available)") -->
<!--   } -->

<!--   # ----------------------------------------------------------------------- -->
<!--   # 4. GAMM: MENSES-CENTERED MODEL -->
<!--   # ----------------------------------------------------------------------- -->
<!--   # Fits a logistic GAMM with menses onset as the reference point (time = 0). -->
<!--   # This model estimates the probability of the outcome across the cycle. -->

<!--   message("  Fitting menses-centered GAMM...") -->

<!--   # Prepare data: keep only complete cases for required variables -->
<!--   selected_vars <- c("cyclic_time", "day_of_week", "session_today", outcome_name, "id") -->
<!--   dat_menses <- df[complete.cases(df[selected_vars]), ] -->
<!--   dat_menses$id <- as.factor(dat_menses$id) -->

<!--   # Construct model formula -->
<!--   # - day_of_week: controls for weekday effects -->
<!--   # - session_today: controls for laboratory visit effects -->
<!--   # - s(cyclic_time, bs='cc'): cyclic smooth for average effect -->
<!--   # - s(id, bs='re'): random intercept for each person -->
<!--   # - s(cyclic_time, id, bs=c('re','cc')): random cyclic slope -->
<!--   f_menses <- as.formula(paste(outcome_name, "~ day_of_week + session_today + s(cyclic_time, bs = 'cc') + s(id, bs = 're') + s(cyclic_time, id, bs = c('re', 'cc'))")) -->

<!--   # Fit the model -->
<!--   gamm_menses <- mgcv::gam( -->
<!--     formula = f_menses, -->
<!--     knots = list(cyclic_time = c(-1, 1)), -->
<!--     data = dat_menses,  -->
<!--     method = 'REML', -->
<!--     family = binomial(link = "logit") -->
<!--   ) -->

<!--   # Save model object and text summary -->
<!--   saveRDS(gamm_menses, file = file.path(save_dir, "model_menses.rds")) -->

<!--   sink(file.path(save_dir, "summary_menses.txt")) -->
<!--   summ_menses <- summary(gamm_menses) -->
<!--   print(summ_menses) -->
<!--   sink() -->

<!--   # --- Extract Statistics for Plot Annotation --- -->
<!--   edf_fixed <- round(summ_menses$s.table[1, "edf"], 2) -->
<!--   stars_fixed <- get_stars(summ_menses$s.table[1, "p-value"]) -->

<!--   edf_rand <- round(summ_menses$s.table[3, "edf"], 2) -->
<!--   stars_rand <- get_stars(summ_menses$s.table[3, "p-value"]) -->

<!--   stat_label_menses <- paste0("Average Effect: edf=", edf_fixed, stars_fixed, "\n", -->
<!--                               "Individual Differences: edf=", edf_rand, stars_rand) -->

<!--   # --- Generate Predictions for Plotting --- -->
<!--   message("  Generating predictions for menses-centered model...") -->

<!--   # Set reference values for covariates (held constant during prediction) -->
<!--   ref_day_menses <- dat_menses$day_of_week[1] -->
<!--   ref_session_menses <- dat_menses$session_today[1] -->

<!--   # Create prediction grid -->
<!--   grid_list <- list( -->
<!--     cyclic_time = seq(-1, 1, by = 0.05), -->
<!--     id = 0, -->
<!--     day_of_week = ref_day_menses, -->
<!--     session_today = ref_session_menses -->
<!--   ) -->
<!--   plotdat_menses <- do.call(expand.grid, grid_list) -->

<!--   # Get predictions with confidence intervals -->
<!--   pred_menses <- marginaleffects::predictions( -->
<!--     gamm_menses, -->
<!--     newdata = plotdat_menses, -->
<!--     type = "response" -->
<!--   ) -->

<!--   plotdat_menses$estimate <- pred_menses$estimate -->
<!--   plotdat_menses$conf.low <- pred_menses$conf.low -->
<!--   plotdat_menses$conf.high <- pred_menses$conf.high -->

<!--   # --- Create Prediction Plot --- -->
<!--   message("  Creating menses-centered prediction plot...") -->
<!--   gammplot_menses <- ggplot(plotdat_menses, aes(x = cyclic_time, y = estimate)) + -->
<!--     scale_x_continuous( -->
<!--       limits = c(-1, 1), -->
<!--       breaks = seq(-1, 1, by = 0.50), -->
<!--       labels = c("Ovulation", "50%L", "Menses Onset", "50%F", "Ovulation") -->
<!--     ) + -->
<!--     labs(x = "", y = paste("Probability of", outcome_name, "(Menses-Centered)"), -->
<!--          caption = paste("Predictions holding day_of_week at:", ref_day_menses,  -->
<!--                         "and session_today at:", ref_session_menses)) + -->
<!--     geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) + -->
<!--     geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) + -->
<!--     geom_line(linewidth = 1) + -->
<!--     geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) + -->
<!--     annotate("text", x = -1, y = Inf, label = stat_label_menses,  -->
<!--              hjust = 0, vjust = 1.1, size = 5, fontface = "italic") + -->
<!--     theme_minimal(base_size = 14) + -->
<!--     theme(text = element_text(size = 16), -->
<!--           axis.text = element_text(size = 14)) -->

<!--   save_png( -->
<!--     plot = gammplot_menses, -->
<!--     filename = "gamm_plot_menses.png", -->
<!--     folder = save_dir -->
<!--   ) -->

<!--   # ----------------------------------------------------------------------- -->
<!--   # 5. GAMM: OVULATION-CENTERED MODEL -->
<!--   # ----------------------------------------------------------------------- -->
<!--   # Fits a logistic GAMM with ovulation as the reference point (time = 0). -->

<!--   message("  Fitting ovulation-centered GAMM...") -->

<!--   # Prepare data: keep only complete cases for required variables -->
<!--   selected_vars_ov <- c("cyclic_time_ov", "day_of_week", "session_today", outcome_name, "id") -->
<!--   dat_ov <- df[complete.cases(df[selected_vars_ov]), ] -->
<!--   dat_ov$id <- as.factor(dat_ov$id) -->

<!--   # Construct model formula -->
<!--   f_ov <- as.formula(paste(outcome_name, "~ day_of_week + session_today + s(cyclic_time_ov, bs = 'cc') + s(id, bs = 're') + s(cyclic_time_ov, id, bs = c('re', 'cc'))")) -->

<!--   # Fit the model -->
<!--   gamm_ov <- mgcv::gam( -->
<!--     formula = f_ov, -->
<!--     knots = list(cyclic_time_ov = c(-1, 1)), -->
<!--     data = dat_ov,  -->
<!--     method = 'REML', -->
<!--     family = binomial(link = "logit") -->
<!--   ) -->

<!--   # Save model object and text summary -->
<!--   saveRDS(gamm_ov, file = file.path(save_dir, "model_ovulation.rds")) -->

<!--   sink(file.path(save_dir, "summary_ovulation.txt")) -->
<!--   summ_ov <- summary(gamm_ov) -->
<!--   print(summ_ov) -->
<!--   sink() -->

<!--   # --- Extract Statistics for Plot Annotation --- -->
<!--   edf_fixed_ov <- round(summ_ov$s.table[1, "edf"], 2) -->
<!--   stars_fixed_ov <- get_stars(summ_ov$s.table[1, "p-value"]) -->

<!--   edf_rand_ov <- round(summ_ov$s.table[3, "edf"], 2) -->
<!--   stars_rand_ov <- get_stars(summ_ov$s.table[3, "p-value"]) -->

<!--   stat_label_ov <- paste0("Average Effect: edf=", edf_fixed_ov, stars_fixed_ov, "\n", -->
<!--                           "Individual Differences: edf=", edf_rand_ov, stars_rand_ov) -->

<!--   # --- Generate Predictions for Plotting --- -->
<!--   message("  Generating predictions for ovulation-centered model...") -->

<!--   # Set reference values for covariates -->
<!--   ref_day_ov <- dat_ov$day_of_week[1] -->
<!--   ref_session_ov <- dat_ov$session_today[1] -->

<!--   # Create prediction grid -->
<!--   grid_list_ov <- list( -->
<!--     cyclic_time_ov = seq(-1, 1, by = 0.05),  -->
<!--     id = 0,  -->
<!--     day_of_week = ref_day_ov, -->
<!--     session_today = ref_session_ov -->
<!--   ) -->
<!--   plotdat_ov <- do.call(expand.grid, grid_list_ov) -->

<!--   # Get predictions with confidence intervals -->
<!--   pred_ov <- marginaleffects::predictions( -->
<!--     gamm_ov, -->
<!--     newdata = plotdat_ov, -->
<!--     type = "response" -->
<!--   ) -->

<!--   plotdat_ov$estimate <- pred_ov$estimate -->
<!--   plotdat_ov$conf.low <- pred_ov$conf.low -->
<!--   plotdat_ov$conf.high <- pred_ov$conf.high -->

<!--   # --- Create Prediction Plot --- -->
<!--   message("  Creating ovulation-centered prediction plot...") -->
<!--   gammplot_ov <- ggplot(plotdat_ov, aes(x = cyclic_time_ov, y = estimate)) + -->
<!--     scale_x_continuous( -->
<!--       limits = c(-1, 1), -->
<!--       breaks = seq(-1, 1, by = 0.50), -->
<!--       labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses") -->
<!--     ) + -->
<!--     labs(x = "", y = paste("Probability of", outcome_name, "(Ovulation-Centered)"), -->
<!--          caption = paste("Predictions holding day_of_week at:", ref_day_ov, -->
<!--                         "and session_today at:", ref_session_ov)) + -->
<!--     geom_rect(xmin = -0.04, xmax = 0.04, ymin = -Inf, ymax = Inf, fill = "grey70", alpha = 0.2, color = NA) + -->
<!--     geom_rect(xmin = 0.92, xmax = 1, ymin = -Inf, ymax = Inf, fill = "grey87", alpha = 0.2, color = NA) + -->
<!--     geom_line(linewidth = 1) + -->
<!--     geom_ribbon(aes(ymin = conf.low, ymax = conf.high), fill = "lightgrey", alpha = 0.3) + -->
<!--     annotate("text", x = -1, y = Inf, label = stat_label_ov,  -->
<!--              hjust = 0, vjust = 1.1, size = 5, fontface = "italic") + -->
<!--     theme_minimal(base_size = 14) + -->
<!--     theme(text = element_text(size = 16), -->
<!--           axis.text = element_text(size = 14)) -->

<!--   save_png( -->
<!--     plot = gammplot_ov, -->
<!--     filename = "gamm_plot_ovulation.png", -->
<!--     folder = save_dir -->
<!--   ) -->

<!--   message(paste0("===== Completed: ", outcome_name, " =====\n")) -->
<!-- } -->
<!-- ``` -->



# =============================================================================
# CONTINUOUS OUTCOMES: GAMMs for Quantitative Variables
# =============================================================================
# This section analyzes continuous outcomes using Gaussian GAMMs on
# log-transformed data. Log transformation handles skewness and zeros.
#
# CONTINUOUS OUTCOMES ANALYZED:
# - Alcohol: num_drinks_today, crave_alc_am/pm, crave_nic_am/pm, crave_mar_am/pm
# - Impulsivity: diff_stop, diff_change, immed_payoff, res_urge, diff_long_rew, diff_res_rew
# - Mood: dep, hopeless, worthless, anxious, mood_swing, rej_sens, angry, conflicts
# - Cognition: loss_interest, diff_conc, tired
# - Appetite: over_eat, sp_food_crav
# - Sleep: high_sleep, diff_sleep
# - Other symptoms: overwhelm, loss_control, breast_tend, breast_swell, headache, jm_pain
# - Functioning: loss_prod, int_relat
#
# MODEL SPECIFICATION:
# - Family: Gaussian (suitable for log-transformed continuous data)
# - Data transformation: log(x + 1) to handle zeros
# - Fixed effects: day_of_week, session_today (for consistency with binary models)
# - Cyclic smooth: s(cyclic_time, bs='cc') - captures average cycle pattern
# - Random intercept: s(id, bs='re') - accounts for baseline differences
# - Random cyclic slope: s(cyclic_time, id, bs=c('re','cc')) - individual variation
#
# OUTPUTS FOR EACH VARIABLE:
# - Rolling average plots on original scale (menses and ovulation centered)
# - Model objects fitted on log scale (.rds files)
# - Model summaries (.txt files)
# - Prediction plots back-transformed to original scale
#
# NOTE: Models are fitted on log scale but predictions are back-transformed
# for interpretability using exp(y) - 1
# =============================================================================

  
```{r continuous-outcomes-gamms}

# -------------------------------------------------------------------------
# 1. SETUP: Define Continuous Outcome Variables
# -------------------------------------------------------------------------

continuous_vars <- c(
 # "numdrinkstoday",
  "drsp_total", "drsp_distress", "drsp_labilsocsx", "drsp_breastpain",
  "drsp_eatsx", "drsp_lowarousal", "drsp_NA", "drsp_NAmax",
  paste0("drsp_", 1:21)
)


get_stars <- function(p) {
  stats::symnum(
    p,
    corr = FALSE,
    na = FALSE,
    cutpoints = c(0, 0.001, 0.01, 0.05, 0.1, 1),
    symbols = c("***", "**", "*", ".", " ")
  )
}

# -------------------------------------------------------------------------
# Helper: Extract Smooth Statistics Safely
# -------------------------------------------------------------------------

extract_smooth_stats <- function(summary_obj, term_pattern) {
  s_table <- summary_obj$s.table
  term_row <- grep(term_pattern, rownames(s_table))

  if (length(term_row) == 0) {
    return(list(edf = NA, stars = ""))
  }

  edf <- round(s_table[term_row[1], "edf"], 2)
  stars <- get_stars(s_table[term_row[1], "p-value"])

  list(edf = edf, stars = stars)
}

# -------------------------------------------------------------------------
# Helper: Set Reference Day to Saturday if Available
# -------------------------------------------------------------------------

get_reference_day <- function(day_vector, preferred_day = "Saturday") {
  day_vector <- day_vector[!is.na(day_vector)]

  if (length(day_vector) == 0) {
    return(NA)
  }

  # Character case
  if (is.character(day_vector)) {
    if (preferred_day %in% day_vector) {
      return(preferred_day)
    } else {
      warning(paste0(
        "Preferred reference day '", preferred_day,
        "' not found. Using first observed value instead: ", day_vector[1]
      ))
      return(day_vector[1])
    }
  }

  # Factor case
  if (is.factor(day_vector)) {
    day_levels <- levels(day_vector)

    if (preferred_day %in% day_levels) {
      return(factor(preferred_day, levels = day_levels))
    } else {
      warning(paste0(
        "Preferred reference day '", preferred_day,
        "' not found in factor levels. Using first observed value instead: ",
        as.character(day_vector[1])
      ))
      return(day_vector[1])
    }
  }

  # Fallback for unexpected types
  warning("day_of_week is neither character nor factor. Using first observed value.")
  day_vector[1]
}

# -------------------------------------------------------------------------
# 2. LOOP THROUGH EACH CONTINUOUS OUTCOME
# -------------------------------------------------------------------------

for (i in seq_along(continuous_vars)) {

  outcome_name <- continuous_vars[i]

  message("\n--------------------------------------------------")
  message("Outcome ", i, " of ", length(continuous_vars), ": ", outcome_name)
  message("--------------------------------------------------")

  if (!outcome_name %in% names(df)) {
    message(paste0("  WARNING: Variable '", outcome_name, "' not found. Skipping."))
    next
  }

  outcome_data <- df[[outcome_name]]

  if (length(outcome_data) == 0 || all(is.na(outcome_data))) {
    message(paste0("  WARNING: Variable '", outcome_name, "' has no data. Skipping."))
    next
  }

  log_outcome_name <- paste0(outcome_name, "_log")
  df[[log_outcome_name]] <- log(df[[outcome_name]] + 1)

  save_dir <- file.path(gamms_folder, paste0(outcome_name, "_gamm"))

  if (!dir.exists(save_dir)) {
    dir.create(save_dir, recursive = TRUE)
  }

  # -----------------------------------------------------------------------
  # 3. RAW DATA VISUALIZATION
  # -----------------------------------------------------------------------

  if (has_cycle_plot) {

    tryCatch({

      rollplot_menses <- cycle_plot(
        df,
        outcome_name,
        centering = "menses",
        include_impute = TRUE,
        y_scale = "person-centered_roll",
        se = TRUE
      )

      save_png(
        plot = rollplot_menses$plot,
        filename = "rollplot_menses_raw.png",
        folder = save_dir
      )

    }, error = function(e) message(e$message))

    tryCatch({

      rollplot_ov <- cycle_plot(
        df,
        outcome_name,
        centering = "ovulation",
        include_impute = TRUE,
        y_scale = "person-centered_roll",
        se = TRUE
      )

      save_png(
        plot = rollplot_ov$plot,
        filename = "rollplot_ovulation_raw.png",
        folder = save_dir
      )

    }, error = function(e) message(e$message))
  }

  # -----------------------------------------------------------------------
  # 4. MENSES-CENTERED GAMM
  # -----------------------------------------------------------------------

  message("  Fitting menses-centered GAMM...")

  selected_vars <- c("cyclic_time", "day_of_week", log_outcome_name, "id")

  dat_menses <- df[complete.cases(df[selected_vars]), ]
  dat_menses$id <- as.factor(dat_menses$id)

  f_menses <- as.formula(
    paste0(
      log_outcome_name,
      " ~ day_of_week + s(cyclic_time, bs='cc') + ",
      "s(id, bs='re') + ",
      "s(cyclic_time, id, bs=c('re','cc'))"
    )
  )

  gamm_menses <- mgcv::gam(
    formula = f_menses,
    knots = list(cyclic_time = c(-1, 1)),
    data = dat_menses,
    method = "REML",
    family = gaussian()
  )

  saveRDS(gamm_menses, file.path(save_dir, "model_menses_log.rds"))

  sink(file.path(save_dir, "summary_menses_log.txt"))
  summ_menses <- summary(gamm_menses)
  print(summ_menses)
  sink()

  fixed_stats <- extract_smooth_stats(summ_menses, "^s\\(cyclic_time\\)$")
  rand_stats  <- extract_smooth_stats(summ_menses, "^s\\(cyclic_time,id\\)$")

  stat_label_menses <- paste0(
    "Average Effect: edf=", fixed_stats$edf, fixed_stats$stars,
    "\nIndividual Differences: edf=", rand_stats$edf, rand_stats$stars
  )

  message("  Generating predictions...")

  ref_day_menses <- get_reference_day(dat_menses$day_of_week, preferred_day = "Saturday")

  plotdat_menses <- expand.grid(
    cyclic_time = seq(-1, 1, by = 0.05),
    id = 0,
    day_of_week = ref_day_menses
  )

  pred_menses <- marginaleffects::predictions(
    gamm_menses,
    newdata = plotdat_menses,
    type = "response"
  )

  plotdat_menses$estimate  <- exp(pred_menses$estimate) - 1
  plotdat_menses$conf.low  <- exp(pred_menses$conf.low) - 1
  plotdat_menses$conf.high <- exp(pred_menses$conf.high) - 1

  gammplot_menses <- ggplot(plotdat_menses, aes(x = cyclic_time, y = estimate)) +
    geom_line(linewidth = 1) +
    geom_ribbon(
      aes(ymin = conf.low, ymax = conf.high),
      fill = "lightgrey",
      alpha = 0.3
    ) +
    annotate(
      "text",
      x = -1,
      y = Inf,
      label = stat_label_menses,
      hjust = 0,
      vjust = 1.1,
      size = 5,
      fontface = "italic"
    ) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, 0.5),
      labels = c("Ovulation", "50%L", "Menses Onset", "50%F", "Ovulation")
    ) +
    labs(
      x = "",
      y = paste(outcome_name, "(Original Scale)"),
      caption = paste("Predictions holding day_of_week at:", as.character(ref_day_menses))
    ) +
    theme_minimal(base_size = 14)

  save_png(
    plot = gammplot_menses,
    filename = "gamm_plot_menses_backtrans.png",
    folder = save_dir
  )

  # -----------------------------------------------------------------------
  # 5. OVULATION-CENTERED GAMM
  # -----------------------------------------------------------------------

  message("  Fitting ovulation-centered GAMM...")

  selected_vars_ov <- c("cyclic_time_ov", "day_of_week", log_outcome_name, "id")

  dat_ov <- df[complete.cases(df[selected_vars_ov]), ]
  dat_ov$id <- as.factor(dat_ov$id)

  f_ov <- as.formula(
    paste0(
      log_outcome_name,
      " ~ day_of_week + s(cyclic_time_ov, bs='cc') + ",
      "s(id, bs='re') + ",
      "s(cyclic_time_ov, id, bs=c('re','cc'))"
    )
  )

  gamm_ov <- mgcv::gam(
    formula = f_ov,
    knots = list(cyclic_time_ov = c(-1, 1)),
    data = dat_ov,
    method = "REML",
    family = gaussian()
  )

  saveRDS(gamm_ov, file.path(save_dir, "model_ovulation_log.rds"))

  sink(file.path(save_dir, "summary_ovulation_log.txt"))
  summ_ov <- summary(gamm_ov)
  print(summ_ov)
  sink()

  fixed_stats_ov <- extract_smooth_stats(summ_ov, "^s\\(cyclic_time_ov\\)$")
  rand_stats_ov  <- extract_smooth_stats(summ_ov, "^s\\(cyclic_time_ov,id\\)$")

  stat_label_ov <- paste0(
    "Average Effect: edf=", fixed_stats_ov$edf, fixed_stats_ov$stars,
    "\nIndividual Differences: edf=", rand_stats_ov$edf, rand_stats_ov$stars
  )

  ref_day_ov <- get_reference_day(dat_ov$day_of_week, preferred_day = "Saturday")

  plotdat_ov <- expand.grid(
    cyclic_time_ov = seq(-1, 1, by = 0.05),
    id = 0,
    day_of_week = ref_day_ov
  )

  pred_ov <- marginaleffects::predictions(
    gamm_ov,
    newdata = plotdat_ov,
    type = "response"
  )

  plotdat_ov$estimate  <- exp(pred_ov$estimate) - 1
  plotdat_ov$conf.low  <- exp(pred_ov$conf.low) - 1
  plotdat_ov$conf.high <- exp(pred_ov$conf.high) - 1

  gammplot_ov <- ggplot(plotdat_ov, aes(x = cyclic_time_ov, y = estimate)) +
    geom_line(linewidth = 1) +
    geom_ribbon(
      aes(ymin = conf.low, ymax = conf.high),
      fill = "lightgrey",
      alpha = 0.3
    ) +
    annotate(
      "text",
      x = -1,
      y = Inf,
      label = stat_label_ov,
      hjust = 0,
      vjust = 1.1,
      size = 5,
      fontface = "italic"
    ) +
    scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, 0.5),
      labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses")
    ) +
    labs(
      x = "",
      y = paste(outcome_name, "(Original Scale)"),
      caption = paste("Predictions holding day_of_week at:", as.character(ref_day_ov))
    ) +
    theme_minimal(base_size = 14)

  save_png(
    plot = gammplot_ov,
    filename = "gamm_plot_ovulation_backtrans.png",
    folder = save_dir
  )

  message(paste0("===== Completed: ", outcome_name, " =====\n"))
}
```


```{r}

```




