---
title: "HERA Individual GAM Reports"
author: "Tory Eisenlohr-Moul, PhD"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

> **Script Purpose:** Automated, participant-centric analysis of symptom cyclicity using cyclic GAMs.
> 1) **Setup & Preparation** — load data, define outcomes/layouts, precompute transforms.
> 2) **Automated Reporting** — one **landscape** PDF per participant with compact cover and paged plots.
> 3) **CSV Audit** — per-ID × outcome × time-var stats for triage/QC.

---

# Libraries
```{r}
library(dplyr)
library(ggplot2)
library(mgcv)
library(patchwork)
library(gridExtra)
library(grid)
library(tibble)
library(tidyr)
library(knitr)
library(broom)
library(scales)
library(glue)

options(scipen = 999)
set.seed(12345)
```

# Data
```{r}

df <- readRDS('/Users/toryeisenlohr-moul/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/HERA_20260414_clean.rds')
```

# TEST CHUNK - KEEP COMMENTED OUT UNLESS TESTING

```{r}
set.seed(12345)
test_ids <- sample(unique(df$id), 5)
df <- df %>% filter(id %in% test_ids)
```

# Section 1: Set output folder & time vars
```{r}
# --- Path to the main output directory ---
gam_output_base <- "~/Library/CloudStorage/Box-Box/00 - CLEAR Lab (Locked Folders)/02 - Data Management, Analysis, and Papers/Studies_Projects/HERA/03_analytic_projects/HERA-Primary ms/02_datasets_output/06_Single_GAMs/"

# --- Define all Time Variables to iterate over (must be scaled to [-1, 1]) ---
time_vars_to_run <- c("cyclic_time_impute")

  #"cyclic_time", "cyclic_time_ov", "cyclic_time_impute", "cyclic_time_imp_ov")

# --- Table Formatting ---
table_cols_per_page <- 6

# --- Plot Layout (grid per page) ---
plots_per_col <- 3
plots_per_row <- 3

# --- Minimum observations per cycle-half to run GAMs ---
# Per-outcome gate: after filtering to complete cases (outcome + time var),
# if either cycle half has fewer than this many observations,
# the GAM is skipped for that specific outcome × time-var combination.
min_obs_per_half <- 3
```

# Section 2: Create output folders & check time vars
```{r}
# --- Create output folder with current date ---
current_date <- format(Sys.Date(), "%Y%m%d")
output_folder <- file.path(gam_output_base, paste0("GAM_Reports_", current_date))
dir.create(output_folder, showWarnings = FALSE, recursive = TRUE)

# --- Robustness: Verify time variables exist ---
time_vars <- intersect(time_vars_to_run, names(df))
if (length(time_vars) < length(time_vars_to_run)) {
  missing_vars <- setdiff(time_vars_to_run, time_vars)
  warning("The following time_vars were not found and will be skipped: ", paste(missing_vars, collapse = ", "))
}
```

# Section 3: Define outcomes
```{r}
# --- Define Outcomes to Model ---
outcomes <- c(
  "drsp_NA"          = "DRSP Negative Affect Mean",
  "drsp_distress"    = "Distress (DRSP)",
  "drsp_labilsocsx"  = "Social/Affect Lability (DRSP)",
  "drsp_breastpain"  = "Breast Sx (DRSP)",
  "drsp_eatsx"       = "Eating Sx",
  "drsp_lowarousal"  = "Low Arousal Sx (DRSP)",
  "drsp_1"           = "Depression",
  "drsp_2"           = "Hopelessness",
  "drsp_3"           = "Worthlessness",
  "drsp_4"           = "Anxiety",
  "drsp_5"           = "Mood Swings",
  "drsp_6"           = "Rejection Sensitivity",
  "drsp_7"           = "Anger/Irritability",
  "drsp_8"           = "Interpersonal Conflicts",
  "drsp_9"           = "Loss of Interest",
  "drsp_16"          = "Feeling Overwhelmed",
  "drsp_17"          = "Feeling Out of Control",
  "drsp_10"          = "Difficulty Concentrating",
  "drsp_11"          = "Fatigue/Lethargy",
  "drsp_14"          = "Hypersomnia",
  "drsp_15"          = "Insomnia",
  "drsp_12"          = "Overeating",
  "drsp_13"          = "Specific Food Cravings",
  "drsp_18"          = "Breast Tenderness",
  "drsp_19"          = "Breast Swelling",
  "drsp_20"          = "Headache",
  "drsp_21"          = "Joint/Muscle Pain"
)

# --- Physical vs Affect symptom classification (raw DRSP items only) ---
physical_symptoms <- c("drsp_12", "drsp_13", "drsp_14", "drsp_15",
                       "drsp_18", "drsp_19", "drsp_20", "drsp_21")

# --- Pattern constants for variable matching ---
DRSP_VAR_PATTERN <- "^drsp_\\d+$"

# --- Max possible counts per category (for density-plot Y-axis scaling) ---
all_raw_drsp <- names(outcomes)[grepl(DRSP_VAR_PATTERN, names(outcomes))]
n_physical_max <- length(intersect(all_raw_drsp, physical_symptoms))
n_affect_max   <- length(setdiff(all_raw_drsp, physical_symptoms))

# --- Define colors for plots (each outcome stays consistent) ---
## Use a colorblind friendly palette
drsp_color_map <- c(
  # DRSP mood/affect individual items
  "drsp_1"  = "#0072B2",  # blue
  "drsp_2"  = "#E69F00",  # orange
  "drsp_3"  = "#009E73",  # green
  "drsp_4"  = "#D55E00",  # vermillion
  "drsp_5"  = "#CC79A7",  # reddish purple
  "drsp_6"  = "#56B4E9",  # sky blue
  "drsp_7"  = "#F0E442",  # yellow
  "drsp_8"  = "#000000",  # black
  "drsp_9"  = "#999999",  # gray
  "drsp_10" = "#004488",  # dark blue
  "drsp_11" = "#BB5566",  # muted red
  "drsp_12" = "#228833",  # dark green
  "drsp_13" = "#EE6677",  # soft red
  "drsp_14" = "#AA3377",  # magenta
  "drsp_15" = "#66CCEE",  # light cyan
  "drsp_16" = "#332288",  # deep purple
  "drsp_17" = "#117733",  # dark green 2
  # DRSP physical items
  "drsp_18" = "#44AA99",  # teal
  "drsp_19" = "#88CCEE",  # light blue
  "drsp_20" = "#DDCC77",  # sand
  "drsp_21" = "#CC6677",  # rose
  # DRSP composite outcomes (Okabe-Ito colorblind-friendly palette)
  "drsp_NA"         = "#0072B2",  # blue
  "drsp_distress"   = "#E69F00",  # orange
  "drsp_labilsocsx" = "#009E73",  # green
  "drsp_breastpain" = "#D55E00",  # vermillion
  "drsp_eatsx"      = "#CC79A7",  # reddish purple
  "drsp_lowarousal" = "#56B4E9"   # sky blue
)

# Safety check: warn (do not stop) if any outcome is missing a color
missing_colors <- setdiff(names(outcomes), names(drsp_color_map))
if (length(missing_colors) > 0) {
  warning("Missing colors for: ", paste(missing_colors, collapse = ", "),
          ". Falling back to #0072B2 for these outcomes.")
}

# --- Per-outcome y-axis limits and reference-line settings ---
# y_axis_mode "fixed" uses a hard y_min/y_max (Likert 1-6 scale).
# y_axis_mode "data"  uses data-driven limits; no fixed reference line.
outcome_plot_settings <- tribble(
  ~variable,          ~y_axis_mode, ~y_min, ~y_max, ~ref_line,
  # DRSP composite outcomes — default 1-4 scale; expands to 5 or 6 if predictions exceed 4
  "drsp_NA",          "fixed",       1,      4,     NA,
  "drsp_distress",    "fixed",       1,      4,     NA,
  "drsp_labilsocsx",  "fixed",       1,      4,     NA,
  "drsp_breastpain",  "fixed",       1,      4,     NA,
  "drsp_eatsx",       "fixed",       1,      4,     NA,
  "drsp_lowarousal",  "fixed",       1,      4,     NA,
  # DRSP individual items — fixed 1-6 Likert scale
  "drsp_1",           "fixed",       1,      6,      4,
  "drsp_2",           "fixed",       1,      6,      4,
  "drsp_3",           "fixed",       1,      6,      4,
  "drsp_4",           "fixed",       1,      6,      4,
  "drsp_5",           "fixed",       1,      6,      4,
  "drsp_6",           "fixed",       1,      6,      4,
  "drsp_7",           "fixed",       1,      6,      4,
  "drsp_8",           "fixed",       1,      6,      4,
  "drsp_9",           "fixed",       1,      6,      4,
  "drsp_16",          "fixed",       1,      6,      4,
  "drsp_17",          "fixed",       1,      6,      4,
  "drsp_10",          "fixed",       1,      6,      4,
  "drsp_11",          "fixed",       1,      6,      4,
  "drsp_14",          "fixed",       1,      6,      4,
  "drsp_15",          "fixed",       1,      6,      4,
  "drsp_12",          "fixed",       1,      6,      4,
  "drsp_13",          "fixed",       1,      6,      4,
  "drsp_18",          "fixed",       1,      6,      4,
  "drsp_19",          "fixed",       1,      6,      4,
  "drsp_20",          "fixed",       1,      6,      4,
  "drsp_21",          "fixed",       1,      6,      4
)

# --- Normalization scale ranges for normalized overlay pages ---
# Fixed-scale outcomes use their configured y_min/y_max.
# Data-mode outcomes fall back to per-curve GAM min/max at plot time.
norm_scale_ranges <- outcome_plot_settings %>%
  dplyr::mutate(
    norm_min = dplyr::if_else(!is.na(y_min), y_min, NA_real_),
    norm_max = dplyr::if_else(!is.na(y_max), y_max, NA_real_)
  ) %>%
  dplyr::select(variable, norm_min, norm_max)

# --- Define Categories for organization ---
outcome_categories <- tribble(
  ~category,                ~variable,
  "DRSP Composites",        "drsp_NA",
  "DRSP Composites",        "drsp_distress",
  "DRSP Composites",        "drsp_labilsocsx",
  "DRSP Composites",        "drsp_breastpain",
  "DRSP Composites",        "drsp_eatsx",
  "DRSP Composites",        "drsp_lowarousal",
  "Mood & Physical (DRSP)", "drsp_1",
  "Mood & Physical (DRSP)", "drsp_2",
  "Mood & Physical (DRSP)", "drsp_3",
  "Mood & Physical (DRSP)", "drsp_4",
  "Mood & Physical (DRSP)", "drsp_5",
  "Mood & Physical (DRSP)", "drsp_6",
  "Mood & Physical (DRSP)", "drsp_7",
  "Mood & Physical (DRSP)", "drsp_8",
  "Mood & Physical (DRSP)", "drsp_9",
  "Mood & Physical (DRSP)", "drsp_16",
  "Mood & Physical (DRSP)", "drsp_17",
  "Mood & Physical (DRSP)", "drsp_10",
  "Mood & Physical (DRSP)", "drsp_11",
  "Mood & Physical (DRSP)", "drsp_14",
  "Mood & Physical (DRSP)", "drsp_15",
  "Mood & Physical (DRSP)", "drsp_12",
  "Mood & Physical (DRSP)", "drsp_13",
  "Mood & Physical (DRSP)", "drsp_18",
  "Mood & Physical (DRSP)", "drsp_19",
  "Mood & Physical (DRSP)", "drsp_20",
  "Mood & Physical (DRSP)", "drsp_21"
) %>%
  right_join(enframe(outcomes, name = "variable", value = "outcome_label"), by = "variable")

# --- Define custom page layouts ---
page_definitions <- c(
  list(
    Composites_Page = outcome_categories %>%
      filter(category == "DRSP Composites") %>%
      pull(outcome_label)
  ),
  split(
    outcome_categories %>%
      filter(category == "Mood & Physical (DRSP)") %>%
      pull(outcome_label),
    ceiling(seq_along(
      outcome_categories %>%
        filter(category == "Mood & Physical (DRSP)") %>%
        pull(outcome_label)
    ) / 9)
  )
)
names(page_definitions)[2:length(page_definitions)] <-
  paste0("DRSP_Page_", seq_len(length(page_definitions) - 1))

# --- Pre-calculate log-transformations ---
for (var in names(outcomes)) {
  if (var %in% names(df)) {
    log_var_name <- paste0(var, "_Log")
    df[[log_var_name]] <- log(df[[var]] + 1)
  }
}
```

# Section 4: Per ID outcomes (fit functions + helpers)
```{r}
# --- Prediction grids reused per time var ---
pred_grid_list <- setNames(lapply(time_vars, function(tv) {
  pg_df <- data.frame(seq(-1, 1, by = 0.05))
  names(pg_df) <- tv
  pg_df
}), time_vars)

# --- Axis labels + reference lines per time var ---
axis_meta <- lapply(time_vars, function(tv) {
  if (grepl("ov", tv, ignore.case = TRUE)) {
    list(labels = c("Menses", "50%F", "Ovulation", "50%L", "Menses"),
         menses = c(-1, 1), ov = 0)
  } else {
    list(labels = c("Ovulation", "50%L", "Menses", "50%F", "Ovulation"),
         menses = 0, ov = c(-1, 1))
  }
})
names(axis_meta) <- time_vars

# --- Assess cycle coverage for a given time variable ---
# Divides the cycle variable range into 6 fixed bins and counts observations per bin.
# Returns a list with total_n, bins_present, bins_with_2plus, empty_bins, and bin_counts.
assess_cycle_coverage <- function(data, cycle_var) {
  n_bins <- 6L
  vals <- data[[cycle_var]]
  vals <- vals[!is.na(vals)]

  # Use fixed breakpoints across the full theoretical cycle range [-1, 1]
  # so that bins are comparable across participants and outcomes
  cycle_bin <- cut(vals,
                   breaks = seq(-1, 1, length.out = n_bins + 1),
                   labels = FALSE,
                   include.lowest = TRUE)

  bin_counts <- as.data.frame(table(cycle_bin = factor(cycle_bin, levels = seq_len(n_bins))),
                              stringsAsFactors = FALSE)
  names(bin_counts) <- c("cycle_bin", "n_per_bin")
  bin_counts$n_per_bin <- as.integer(bin_counts$n_per_bin)

  total_n <- length(vals)
  bins_present <- sum(bin_counts$n_per_bin > 0)
  bins_with_2plus <- sum(bin_counts$n_per_bin >= 2)
  empty_bins <- n_bins - bins_present

  list(
    total_n = total_n,
    bins_present = bins_present,
    bins_with_2plus = bins_with_2plus,
    empty_bins = empty_bins,
    bin_counts = bin_counts
  )
}

# --- Decide GAM specification based on cycle coverage ---
# Returns a list with run_gam (logical), k (integer or NA), and mode (character).
decide_gam_spec <- function(coverage) {
  # Good coverage → full model
  if (
    coverage$total_n >= 15 &&
    coverage$bins_present >= 5 &&
    coverage$bins_with_2plus >= 4 &&
    coverage$empty_bins <= 1
  ) {
    return(list(
      run_gam = TRUE,
      k = 6L,
      mode = "full_cycle"
    ))
  }

  # Moderate coverage → reduced complexity
  if (
    coverage$total_n >= 12 &&
    coverage$bins_present >= 4
  ) {
    return(list(
      run_gam = TRUE,
      k = 4L,
      mode = "reduced_k"
    ))
  }

  # Poor coverage → do not run GAM
  list(
    run_gam = FALSE,
    k = NA_integer_,
    mode = "insufficient_coverage"
  )
}

# --- Fit + predict one outcome for one time var (bam + predict) ---
fit_outcome <- function(dat, outcome_var, tv, outcomes_map, k_max = 10L) {
  log_out <- paste0(outcome_var, "_Log")
  keep <- c(log_out, tv)
  if (!all(keep %in% names(dat))) return(NULL)
  dat <- dat[complete.cases(dat[, keep]), keep, drop = FALSE]
  if (nrow(dat) < 10 || var(dat[[log_out]], na.rm = TRUE) < 1e-6) return(NULL)

  # Count observations in each positional half for adequacy flagging.
  # These are orientation-neutral: n_neg = values in [-1, 0), n_nonneg = values in [0, 1].
  # For menses-centered vars (cyclic_time, cyclic_time_impute): neg = luteal, nonneg = follicular.
  # For ov-centered vars      (cyclic_time_ov, cyclic_time_imp_ov): neg = follicular, nonneg = luteal.
  n_neg    <- sum(dat[[tv]] <  0, na.rm = TRUE)
  n_nonneg <- sum(dat[[tv]] >= 0, na.rm = TRUE)

  # Skip if either cycle half has fewer than min_obs_per_half observations.
  if (min(n_neg, n_nonneg) < min_obs_per_half) {
    message(glue::glue(
      "Skipping GAM for {outcome_var}: only {min(n_neg, n_nonneg)} obs in one ",
      "cycle half (min_obs_per_half={min_obs_per_half})"
    ))
    return(NULL)
  }

  # --- Cycle coverage check & adaptive k ---
  coverage <- assess_cycle_coverage(data = dat, cycle_var = tv)
  decision <- decide_gam_spec(coverage)

  message(glue::glue(
    "[{outcome_var}] Total N: {coverage$total_n}, Bins present: {coverage$bins_present}, ",
    "Bins with >=2: {coverage$bins_with_2plus}, Empty bins: {coverage$empty_bins}, ",
    "Mode: {decision$mode}"
  ))

  if (!decision$run_gam) {
    message(glue::glue("Skipping GAM for {outcome_var}: insufficient cycle coverage"))
    return(NULL)
  }

  # Use the coverage-based k, but still respect k_max and data-size constraints
  k <- min(decision$k, k_max, nrow(dat) - 1L)
  if (k < 3L) return(NULL)

  fml <- as.formula(paste0(log_out, " ~ s(", tv, ", bs = 'cc', k = ", k, ")"))
  knots <- setNames(list(c(-1, 1)), tv)

  mf <- try(mgcv::bam(fml, data = dat, knots = knots, method = "fREML",
                      discrete = TRUE, select = TRUE),
            silent = TRUE)
  if (inherits(mf, "try-error")) return(NULL)

  s  <- summary(mf)
  p  <- suppressWarnings(as.numeric(s$s.pv[1]))
  st <- s$s.table
  edf <- suppressWarnings(as.numeric(st[1, "edf"]))
  Fval <- suppressWarnings(as.numeric(st[1, "F"]))

  pg <- pred_grid_list[[tv]]
  pr <- predict(mf, newdata = pg, se.fit = TRUE, type = "link")  # link scale for log(y+1)
  est <- exp(pr$fit) - 1
  conf.low  <- exp(pr$fit - 1.96 * pr$se.fit) - 1
  conf.high <- exp(pr$fit + 1.96 * pr$se.fit) - 1

  # Record cycle time at peak (max) and trough (min) of predicted curve
  peak_time   <- pg[[tv]][which.max(est)]
  trough_time <- pg[[tv]][which.min(est)]

  pred <- cbind(pg, estimate = est, conf.low = conf.low, conf.high = conf.high)
  pred$outcome_var   <- outcome_var
  pred$outcome_label <- outcomes_map[[outcome_var]]
  pred$p.value <- p
  pred$raw_range <- diff(range(est, na.rm = TRUE))
  pred$n <- nrow(dat)
  pred$edf <- edf

  list(
    pred = pred,
    sig  = tibble::tibble(
      variable = outcome_var,
      outcome_label = outcomes_map[[outcome_var]],
      time_var = tv,
      p.value = p,
      edf = edf,
      F = Fval,
      n = nrow(dat),
      n_neg = n_neg,
      n_nonneg = n_nonneg,
      range = diff(range(est, na.rm = TRUE)),
      peak_time = peak_time,
      trough_time = trough_time,
      coverage_mode = decision$mode
    )
  )
}

# --- Plot one outcome curve with annotations ---
# y_limits = NULL means: use per-outcome defaults from outcome_plot_settings
#   (fixed outcomes → c(y_min, y_max); data-mode outcomes → data-driven, no clamp)
# y_limits = c(lo, hi) means: use those limits as-is (e.g. from shared-y section)
create_annotated_plot <- function(plot_df, tv, axis_meta_list, y_limits = NULL,
                                  raw_data = NULL) {
  if (is.null(plot_df) || nrow(plot_df) == 0) return(NULL)

  max_pt <- dplyr::slice_max(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  min_pt <- dplyr::slice_min(plot_df, order_by = estimate, n = 1, with_ties = FALSE)
  rng   <- dplyr::first(plot_df$raw_range)
  p_val <- dplyr::first(plot_df$p.value)
  n_obs <- dplyr::first(plot_df$n)
  edf_val <- dplyr::first(plot_df$edf)

  outcome_var <- dplyr::first(plot_df$outcome_var)
  line_col <- drsp_color_map[[outcome_var]]
  if (is.null(line_col) || is.na(line_col)) line_col <- "#0072B2"

  # Look up per-outcome y-axis and reference-line settings
  ps        <- outcome_plot_settings[outcome_plot_settings$variable == outcome_var, ]
  mode      <- if (nrow(ps) > 0) ps$y_axis_mode[1] else "fixed"
  ref_line  <- if (nrow(ps) > 0) ps$ref_line[1]    else 4
  y_fix_min <- if (nrow(ps) > 0) ps$y_min[1]       else 1
  y_fix_max <- if (nrow(ps) > 0) ps$y_max[1]       else 6

  # Resolve y-axis limits when not explicitly supplied by the caller
  if (is.null(y_limits)) {
    if (mode == "fixed" && !is.na(y_fix_min) && !is.na(y_fix_max)) {
      y_limits <- c(y_fix_min, y_fix_max)
    }
    # data mode: leave y_limits = NULL → ggplot auto-scales
  }

  # Dynamic expansion for composite outcomes: expand y_max to 5 or 6 if predictions exceed it
  is_composite <- !grepl(DRSP_VAR_PATTERN, outcome_var)
  if (is_composite && !is.null(y_limits)) {
    max_pred <- max(plot_df$estimate, na.rm = TRUE)
    expanded_ymax <- expand_composite_ymax(y_limits[2], max_pred)
    if (expanded_ymax > y_limits[2]) {
      y_limits[2] <- expanded_ymax
      y_fix_max <- expanded_ymax
    }
  }

  cpass_flag <- !is.null(rng) && is.finite(rng) && rng >= 1.5
  stars <- dplyr::case_when(
    is.finite(p_val) & p_val < 0.001 ~ "***",
    is.finite(p_val) & p_val < 0.01  ~ "**",
    is.finite(p_val) & p_val < 0.05  ~ "*",
    TRUE ~ ""
  )
  cpass_label <- if (cpass_flag) "\u25B2 Range \u22651.5 | " else ""
  edf_str <- if (!is.null(edf_val) && !is.na(edf_val)) sprintf(", EDF = %.2f", edf_val) else ""
  ann <- if (!is.null(n_obs) && !is.na(n_obs)) {
    sprintf("%s%s Range = %.2f%s, N = %d", cpass_label, stars, rng, edf_str, as.integer(n_obs))
  } else {
    sprintf("%s%s Range = %.2f%s", cpass_label, stars, rng, edf_str)
  }
  meta  <- axis_meta_list[[tv]]

  # Prepare raw data overlay using the same inclusion rule as fit_outcome:
  # require non-missing log-transformed outcome and time var, then plot on original scale.
  raw_pts <- NULL
  log_out <- paste0(outcome_var, "_Log")
  if (!is.null(raw_data) && outcome_var %in% names(raw_data) &&
      log_out %in% names(raw_data) && tv %in% names(raw_data)) {
    raw_pts <- raw_data[complete.cases(raw_data[, c(log_out, tv)]), c(outcome_var, tv), drop = FALSE]
    if (nrow(raw_pts) == 0) raw_pts <- NULL
  }

  p <- ggplot2::ggplot(plot_df, ggplot2::aes(x = .data[[tv]], y = estimate))

  # Reference line only for outcomes that define one (Likert-scale outcomes)
  if (!is.na(ref_line)) {
    p <- p + ggplot2::geom_hline(yintercept = ref_line, color = "red",
                                 linewidth = 0.4, alpha = 0.4)
  }

  # Overlay raw data points behind the GAM curve
  if (!is.null(raw_pts)) {
    p <- p + ggplot2::geom_point(
      data = raw_pts,
      ggplot2::aes(x = .data[[tv]], y = .data[[outcome_var]]),
      color = line_col, alpha = 0.25, size = 1.2, shape = 16,
      inherit.aes = FALSE
    )
  }

  p <- p +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red") +
    ggplot2::geom_vline(xintercept = meta$ov,     color = "forestgreen") +
    ggplot2::geom_line(linewidth = 1, color = line_col) +
    ggplot2::geom_point(data = max_pt, size = 2.5, color = line_col) +
    ggplot2::geom_point(data = min_pt, size = 2.5, color = line_col) +
    ggplot2::annotate("text", x = -1, y = Inf, label = ann,
                      hjust = -0.05, vjust = 1.5, size = 2.5, fontface = "bold") +
    ggplot2::labs(title = dplyr::first(plot_df$outcome_label), x = "Cycle Phase", y = NULL) +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::theme_minimal() +
    ggplot2::theme(
      plot.title = ggplot2::element_text(size = 9, face = "bold", hjust = 0.5),
      plot.margin = ggplot2::margin(t = 10, r = 5, b = 5, l = 5, unit = "pt"),
      axis.text = ggplot2::element_text(size = 7)
    )

  # For fixed-scale outcomes, show all integer breaks on the y-axis
  if (mode == "fixed" && !is.na(y_fix_min) && !is.na(y_fix_max)) {
    p <- p + ggplot2::scale_y_continuous(
      breaks = seq(y_fix_min, y_fix_max, by = 1)
    )
  }

  # Always disable clipping so top annotations (y = Inf) are never cut off.
  # When y_limits is defined, also lock the y-axis range.
  if (!is.null(y_limits)) {
    p <- p + ggplot2::coord_cartesian(ylim = y_limits, clip = "off")
  } else {
    p <- p + ggplot2::coord_cartesian(clip = "off")
  }

  # Colored effect-size borders for DRSP individual items
  is_drsp <- grepl(DRSP_VAR_PATTERN, outcome_var)

  if (is_drsp) {
    border_col <- dplyr::case_when(
      is.finite(rng) && rng >= 1.5 ~ "red",
      is.finite(rng) && rng >= 1.0 ~ "salmon",
      is.finite(rng) && rng >= 0.5 ~ "pink",
      TRUE                         ~ NA_character_
    )

    if (!is.na(border_col)) {
      p <- p + ggplot2::theme(
        panel.border = ggplot2::element_rect(color = border_col, linewidth = 1.5, fill = NA)
      )
    }
  }

  p
}

# --- Compute shared per-ID y-limits for a plot, clamped to [1, 6] ---
make_shared_id_y_limits <- function(plot_df, shared_total_range, lower_bound = 1, upper_bound = 6) {
  this_ymin <- min(plot_df$estimate, na.rm = TRUE)
  this_ymax <- max(plot_df$estimate, na.rm = TRUE)
  this_mid  <- (this_ymin + this_ymax) / 2

  half_span <- shared_total_range / 2
  y_low  <- this_mid - half_span
  y_high <- this_mid + half_span

  # Clamp to allowed symptom scale
  if (y_low < lower_bound) {
    shift_up <- lower_bound - y_low
    y_low  <- y_low + shift_up
    y_high <- y_high + shift_up
  }
  if (y_high > upper_bound) {
    shift_down <- y_high - upper_bound
    y_low  <- y_low - shift_down
    y_high <- y_high - shift_down
  }

  # Final safeguard in case shared_total_range is very large
  y_low  <- max(y_low, lower_bound)
  y_high <- min(y_high, upper_bound)

  c(y_low, y_high)
}

# --- Dynamic Y-axis expansion for composite outcomes ---
# Composites default to y_max = 4 but expand to 5 or 6 if predicted values exceed 4.
expand_composite_ymax <- function(y_max_current, max_predicted, absolute_max = 6) {
  if (is.na(max_predicted) || !is.finite(max_predicted) || max_predicted <= y_max_current) {
    return(y_max_current)
  }
  min(ceiling(max_predicted), absolute_max)
}
```

# Section 5: Per ID outcome reports (main PDF loop)
```{r}
report_folder <- file.path(output_folder, "Participant_Reports_PDF_byID")
dir.create(report_folder, showWarnings = FALSE, recursive = TRUE)

ids_to_run    <- unique(df$id)   # set to subset for testing, e.g., c("203","109")
skip_existing <- TRUE            # set FALSE to overwrite

# collector for a single CSV at the end
all_ids_signif <- list()

# fixed layout controls
max_plots_per_page <- plots_per_col * plots_per_row
table_theme <- gridExtra::ttheme_default(
  base_size = 7,
  core    = list(fg_params = list(hjust = 0, x = 0.05)),
  colhead = list(fg_params = list(hjust = 0, x = 0.05)),
  padding = grid::unit(c(3, 2), "mm")
)

for (current_id in ids_to_run) {
  message(sprintf("--- Processing ID: %s ---", current_id))
  datSX_id <- subset(df, id == current_id)

  # ---- Open a single LANDSCAPE PDF per participant ----
  pdf_filename <- file.path(report_folder, sprintf("Report_ID_%s.pdf", current_id))
  if (isTRUE(skip_existing) && file.exists(pdf_filename)) {
    message("    -> File exists; skipping. Set skip_existing=FALSE to overwrite.")
    next
  }
  grDevices::pdf(pdf_filename, width = 11, height = 8.5)   # LANDSCAPE

  # ---- Page 0: Raw data coverage (Joint/Muscle Pain + DRSP Negative Affect) ----
  # Shows raw data points across the cycle variable so the reader can quickly
  # judge whether there is enough data across phases for interpretation.
  # This page is always produced, even if GAMs cannot be run.
  if (length(time_vars) > 0) {
    tv0 <- time_vars[1]
    meta0 <- axis_meta[[tv0]]

    # Shared helper: count observations in each cycle-phase half and label
    # them as luteal vs follicular, accounting for time-variable orientation.
    # Uses the same non-overlapping split as fit_outcome(): tv < 0 vs tv >= 0.
    #   non-"ov" vars: -1=Ov, 0=Menses, 1=Ov  → tv<0 is luteal, tv>=0 is follicular
    #   "ov" vars:     -1=Men, 0=Ov,    1=Men  → tv<0 is follicular, tv>=0 is luteal
    count_phase_halves <- function(tv_vals, meta) {
      n_lo <- sum(tv_vals < 0)
      n_hi <- sum(tv_vals >= 0)
      is_ov_centered <- identical(meta$ov, 0)
      if (is_ov_centered) {
        list(n_luteal = n_hi, n_follicular = n_lo)
      } else {
        list(n_luteal = n_lo, n_follicular = n_hi)
      }
    }

    # Returns TRUE when either half of the cycle has fewer than min_obs_per_half observations.
    is_half_sparse <- function(halves) {
      halves$n_luteal < min_obs_per_half || halves$n_follicular < min_obs_per_half
    }

    make_raw_panel <- function(dat, tv, meta, var, y_lab, color) {
      if (!var %in% names(dat)) {
        p <- ggplot2::ggplot() +
          ggplot2::annotate("text", x = 0, y = 0, label = paste0(var, " not in data"),
                            size = 5, color = "grey50") +
          ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = "n = 0 daily observations") +
          ggplot2::theme_minimal(base_size = 12)
        return(p)
      }
      vals <- dat[[var]]
      tv_vals <- dat[[tv]]
      valid <- !is.na(vals) & !is.na(tv_vals)
      n_obs <- sum(valid)
      has_data <- n_obs > 0
      halves <- count_phase_halves(tv_vals[valid], meta)
      sub_label <- paste0(
        "n = ", n_obs, " daily observations",
        "  (luteal: ", halves$n_luteal, ", follicular: ", halves$n_follicular, ")"
      )
      p <- ggplot2::ggplot(dat, ggplot2::aes(x = .data[[tv]], y = .data[[var]])) +
        ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                            linewidth = 0.6, linetype = "dashed") +
        ggplot2::scale_x_continuous(
          limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
          labels = meta$labels
        ) +
        ggplot2::labs(x = "Cycle Phase", y = y_lab, subtitle = sub_label) +
        ggplot2::theme_minimal(base_size = 12)
      if (has_data) {
        p <- p + ggplot2::geom_point(color = color, alpha = 0.45, size = 1.8)
      } else {
        p <- p + ggplot2::annotate("text", x = 0, y = 0, label = "No data",
                                   size = 5, color = "grey50")
      }
      p
    }

    p_joint  <- make_raw_panel(datSX_id, tv0, meta0, "drsp_21",
                               "Joint/Muscle Pain (DRSP 21)", "#CC6677")
    p_drsp_na <- make_raw_panel(datSX_id, tv0, meta0, "drsp_NA",
                                "DRSP Negative Affect Mean", "#D55E00")

    tv0_vals <- datSX_id[[tv0]]
    tv0_vals <- tv0_vals[!is.na(tv0_vals)]
    tv0_halves <- count_phase_halves(tv0_vals, meta0)
    tv0_sparse <- is_half_sparse(tv0_halves)

    # Subtitle is informational only; the skip decision is made per time variable.
    coverage_subtitle <- paste0("Time variable: ", tv0)
    if (tv0_sparse) {
      coverage_subtitle <- paste0(
        coverage_subtitle,
        "\n\u26A0 Fewer than ", min_obs_per_half,
        " observations in one or both cycle-halves for ", tv0
      )
    }

    p_cover <- p_joint / p_drsp_na +
      patchwork::plot_annotation(
        title    = paste0("Raw Data Coverage \u2014 ID: ", current_id),
        subtitle = coverage_subtitle,
        theme    = ggplot2::theme(
          plot.title    = ggplot2::element_text(face = "bold", size = 12),
          plot.subtitle = ggplot2::element_text(
            size = if (tv0_sparse) 10 else 9,
            color = if (tv0_sparse) "red" else "grey30"
          )
        )
      )
    print(p_cover)
  } else {
    grid::grid.newpage()
    grid::grid.text(
      label = paste0(
        "Raw Data Coverage \u2014 ID: ", current_id,
        "\n\nNo eligible time variables were available for GAM fitting.",
        "\nA coverage-only placeholder report was generated."
      ),
      x = 0.5, y = 0.55,
      gp = grid::gpar(fontsize = 12, fontface = "bold")
    )
    grDevices::dev.off()
    next
  } # end Page 0

  # ---- Pass 1: fit & collect (coverage gate applied per time variable) ----
  preds_by_tv <- list()
  sig_by_tv   <- list()

  for (tv in time_vars) {
    message(sprintf("  ... time variable: %s", tv))

    # Per-tv coverage gate — consistent with fit_outcome(): tv < 0 vs tv >= 0.
    tv_vals_all <- datSX_id[[tv]]
    tv_vals_all <- tv_vals_all[!is.na(tv_vals_all)]
    tv_meta     <- axis_meta[[tv]]
    tv_halves   <- count_phase_halves(tv_vals_all, tv_meta)
    if (is_half_sparse(tv_halves)) {
      message(sprintf("    -> Skipping tv=%s: insufficient cycle-half coverage (luteal=%d, follicular=%d, min=%d)",
                      tv, tv_halves$n_luteal, tv_halves$n_follicular, min_obs_per_half))
      next
    }

    all_preds <- list()
    sig_list  <- list()

    for (outcome_var in names(outcomes)) {
      res <- fit_outcome(datSX_id, outcome_var, tv, outcomes)
      if (is.null(res)) next
      all_preds[[length(all_preds) + 1]] <- res$pred
      sig_list[[length(sig_list) + 1]]  <- res$sig
    }

    if (length(all_preds) > 0) {
      preds_by_tv[[tv]] <- dplyr::bind_rows(all_preds)
      sig_by_tv[[tv]]   <- dplyr::bind_rows(sig_list)
    }
  }

  if (!length(preds_by_tv)) {
    message("    -> No successful models for this ID; skipping.")
    grDevices::dev.off()
    next
  }

  # Build per-ID significance object for CSV
  sig_all <- dplyr::bind_rows(sig_by_tv)
  sig_all$id <- current_id
  all_ids_signif[[as.character(current_id)]] <- sig_all

  # ---- COVER: Compact counts by time_var ----
  sig_all_clean <- sig_all %>%
    dplyr::mutate(
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    )

  cover_counts <- sig_all_clean %>%
    dplyr::group_by(time_var) %>%
    dplyr::summarise(
      tested = dplyr::n(),
      n_sig  = sum(stars != "", na.rm = TRUE),
      n_cpass = sum(is.finite(range) & range >= 1.5 & stars != "", na.rm = TRUE),
      .groups = "drop"
    ) %>%
    tidyr::unite("count", n_sig, tested, sep = " / ") %>%
    dplyr::rename(`Sig / Tested` = count, `Range ≥1.5 sig` = n_cpass)

  cover_theme <- gridExtra::ttheme_default(
    base_size = 9,
    core = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"),
      padding  = grid::unit(c(2, 2), "mm")
    ),
    colhead = list(
      fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
    )
  )

  title_grob    <- grid::textGrob(paste("Compact Summary — ID:", current_id),
                                  gp = grid::gpar(fontsize = 16, fontface = "bold"))
  subtitle_grob <- grid::textGrob("Rows show: # significant / # tested (p < .05) for each time variable",
                                  gp = grid::gpar(fontsize = 9))
  cover_grob    <- gridExtra::tableGrob(cover_counts, rows = NULL, theme = cover_theme)
  gridExtra::grid.arrange(title_grob, subtitle_grob, cover_grob,
                          ncol = 1, heights = grid::unit(c(0.12, 0.08, 0.80), "npc"))

  # ---- OPTIONAL: compact stars matrix for outcomes with any significance ----
  compact_stars <- sig_all_clean %>%
    dplyr::group_by(variable) %>%
    dplyr::mutate(any_sig = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE)) %>%
    dplyr::ungroup() %>%
    dplyr::filter(any_sig) %>%
    dplyr::mutate(Outcome = dplyr::coalesce(outcome_label, variable)) %>%
    dplyr::mutate(
      Outcome = ifelse(nchar(Outcome) > 22, variable, Outcome),
      stars = dplyr::case_when(
        is.finite(p.value) & p.value < 0.001 ~ "***",
        is.finite(p.value) & p.value < 0.01  ~ "**",
        is.finite(p.value) & p.value < 0.05  ~ "*",
        TRUE ~ ""
      )
    ) %>%
    dplyr::select(Outcome, time_var, stars) %>%
    tidyr::pivot_wider(names_from = time_var, values_from = stars, values_fill = "") %>%
    dplyr::arrange(Outcome)

  if (nrow(compact_stars) > 0) {
    compact_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(
        fg_params = list(hjust = 0, x = 0.02),
        padding = grid::unit(c(1.5, 2), "mm")
      ),
      colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
    )
    header <- grid::textGrob("Significant Outcomes Only (★ = p < .05)",
                             gp = grid::gpar(fontsize = 12, fontface = "bold"))
    tbl <- gridExtra::tableGrob(compact_stars, rows = NULL, theme = compact_theme)
    gridExtra::grid.arrange(header, tbl, ncol = 1, heights = grid::unit(c(0.10, 0.90), "npc"))
  }

  # ---- PAGE 2: Significant EDFs ranked by Range (across all time vars) ----
  sig_ranked <- sig_all_clean %>%
    dplyr::filter(is.finite(p.value) & p.value < 0.05) %>%
    dplyr::mutate(
      Outcome = dplyr::coalesce(outcome_label, variable),
      Outcome = ifelse(nchar(Outcome) > 40, variable, Outcome)
    ) %>%
    dplyr::transmute(
      Outcome,
      TimeVar  = time_var,
      EDF      = round(edf, 2),
      Range    = round(range, 2),
      F        = round(F, 2),
      N        = n
    ) %>%
    dplyr::arrange(dplyr::desc(Range), Outcome, TimeVar)

  if (nrow(sig_ranked) == 0) {
    header <- grid::textGrob(
      paste("Significant Smooths — ID:", current_id, "(none at p < .05)"),
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    sub <- grid::textGrob("No significant cyclic effects detected across any time variable.",
                          gp = grid::gpar(fontsize = 10, fontface = "italic"))
    gridExtra::grid.arrange(header, sub, ncol = 1,
                            heights = grid::unit(c(0.2, 0.8), "npc"))
  } else {
    ranked_theme <- gridExtra::ttheme_default(
      base_size = 8,
      core = list(
        fg_params = list(hjust = 0, x = 0.02),
        padding = grid::unit(c(1.2, 2), "mm")
      ),
      colhead = list(
        fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")
      )
    )
    title_grob <- grid::textGrob(
      paste("Significant Smooths (p < .05), Ranked by Range — ID:", current_id),
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    note_grob <- grid::textGrob(
      "Columns: EDF = effective degrees of freedom; Range = max(pred) − min(pred) on original scale.",
      gp = grid::gpar(fontsize = 9, fontface = "italic")
    )
    ranked_tbl <- gridExtra::tableGrob(sig_ranked, rows = NULL, theme = ranked_theme)
    gridExtra::grid.arrange(
      title_grob, note_grob, ranked_tbl,
      ncol = 1,
      heights = grid::unit(c(0.12, 0.06, 0.82), "npc")
    )
  }

  # ---- Helper: generate an overlay page for qualifying GAM curves ----
  # @param var_pattern    regex to select variables; used if var_list is NULL
  # @param var_list       character vector of variable names; overrides var_pattern if provided
  # @param range_thresh   minimum range for inclusion alongside p < .05
  # @param title_prefix   text before " — ID: ..." in plot title
  # @param subtitle_label description of items shown
  # @param ylim_strategy  "fixed" uses fixed_ylim arg; "data" derives from conf intervals
  # @param fixed_ylim     c(lo, hi) used when ylim_strategy == "fixed"
  # @param show_hline     whether to show the horizontal reference line at hline_y
  # @param hline_y        y-intercept for the reference line
  # @param legend_ncol    number of columns in the legend
  plot_overlay_page <- function(sig_tv, pred_tv, tv, meta,
                                range_thresh, title_prefix, subtitle_label,
                                var_pattern = NULL, var_list = NULL,
                                ylim_strategy = "fixed", fixed_ylim = c(1, 6),
                                show_hline = TRUE, hline_y = 4,
                                legend_ncol = 3,
                                show_range_in_labels = FALSE,
                                normalize_01 = FALSE,
                                ref_lines_per_var = NULL) {
    if (!is.null(var_list)) {
      qualifying <- sig_tv %>%
        dplyr::filter(variable %in% var_list)
    } else if (!is.null(var_pattern)) {
      qualifying <- sig_tv %>%
        dplyr::filter(grepl(var_pattern, variable))
    } else {
      return(invisible(NULL))
    }
    qualifying <- qualifying %>%
      dplyr::filter(
        (is.finite(p.value) & p.value < 0.05) |
        (is.finite(range) & range >= range_thresh)
      )
    if (nrow(qualifying) == 0) return(invisible(NULL))

    ov_vars <- unique(qualifying$variable)
    ov_data <- pred_tv %>% dplyr::filter(outcome_var %in% ov_vars)
    if (nrow(ov_data) == 0) return(invisible(NULL))

    n_curves <- length(ov_vars)

    ov_colors <- drsp_color_map[ov_vars]
    ov_colors[is.na(ov_colors)] <- "#0072B2"
    names(ov_colors) <- ov_vars

    range_lookup <- qualifying %>%
      dplyr::distinct(variable, .keep_all = TRUE) %>%
      dplyr::select(variable, range)

    if (show_range_in_labels) {
      ov_labels <- sapply(ov_vars, function(v) {
        r <- range_lookup$range[range_lookup$variable == v]
        if (length(r) > 0 && is.finite(r)) {
          paste0(outcomes[v], " (", sprintf("%.2f", r), ")")
        } else {
          outcomes[v]
        }
      })
    } else {
      ov_labels <- outcomes[ov_vars]
    }

    if (normalize_01) {
      ov_data <- ov_data %>%
        dplyr::left_join(
          norm_scale_ranges,
          by = c("outcome_var" = "variable")
        ) %>%
        dplyr::group_by(outcome_var) %>%
        dplyr::mutate(
          scale_min = dplyr::coalesce(norm_min, min(estimate, na.rm = TRUE)),
          scale_max = dplyr::coalesce(norm_max, max(estimate, na.rm = TRUE)),
          scale_range = scale_max - scale_min,
          estimate  = ifelse(scale_range > 0, (estimate  - scale_min) / scale_range, 0.5),
          conf.low  = ifelse(scale_range > 0, (conf.low  - scale_min) / scale_range, 0.5),
          conf.high = ifelse(scale_range > 0, (conf.high - scale_min) / scale_range, 0.5)
        ) %>%
        dplyr::select(-norm_min, -norm_max, -scale_min, -scale_max, -scale_range) %>%
        dplyr::ungroup()
      ylim_strategy <- "fixed"
      fixed_ylim <- c(0, 1)
    }

    if (ylim_strategy == "data") {
      y_min <- min(ov_data$conf.low, na.rm = TRUE)
      y_max <- max(ov_data$conf.high, na.rm = TRUE)
      y_pad <- (y_max - y_min) * 0.05
      plot_ylim <- c(max(0, y_min - y_pad), y_max + y_pad)
    } else {
      plot_ylim <- fixed_ylim
    }

    range_label <- format(range_thresh, nsmall = 1)
    y_axis_label <- if (normalize_01) "Normalized Score (0\u20131)" else "Predicted Score"

    if (is.infinite(range_thresh)) {
      filter_label <- "p < .05"
    } else {
      filter_label <- paste0("p < .05 or range \u2265 ", range_label)
    }

    p <- ggplot2::ggplot(ov_data,
           ggplot2::aes(x = .data[[tv]], y = estimate, color = outcome_var))

    if (show_hline) {
      p <- p + ggplot2::geom_hline(yintercept = hline_y, color = "red",
                                   linewidth = 0.4, alpha = 0.4)
    }

    if (!is.null(ref_lines_per_var)) {
      for (rv in names(ref_lines_per_var)) {
        if (rv %in% ov_vars) {
          ref_col <- ov_colors[[rv]]
          if (is.null(ref_col) || is.na(ref_col)) ref_col <- "#0072B2"
          p <- p + ggplot2::geom_hline(
            yintercept = ref_lines_per_var[[rv]],
            color = ref_col, linewidth = 0.5, linetype = "solid", alpha = 0.7
          )
        }
      }
    }

    p <- p +
      ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                          linewidth = 0.6, linetype = "dashed") +
      ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                          linewidth = 0.6, linetype = "dashed") +
      ggplot2::geom_line(linewidth = 0.9) +
      ggplot2::scale_color_manual(
        values = ov_colors,
        breaks = ov_vars,
        labels = ov_labels
      ) +
      ggplot2::scale_x_continuous(
        limits = c(-1, 1),
        breaks = seq(-1, 1, by = 0.5),
        labels = meta$labels
      ) +
      ggplot2::coord_cartesian(ylim = plot_ylim) +
      ggplot2::labs(
        title = paste0(title_prefix, " \u2014 ID: ", current_id),
        subtitle = paste0(
          subtitle_label, " with ", filter_label,
          "  |  N curves = ", n_curves, "  |  Time variable: ", tv
        ),
        x = "Cycle Phase",
        y = y_axis_label,
        color = "Outcome"
      ) +
      ggplot2::theme_minimal(base_size = 10) +
      ggplot2::theme(
        legend.position = "bottom",
        legend.text = ggplot2::element_text(size = 6),
        plot.title = ggplot2::element_text(face = "bold", size = 12),
        plot.subtitle = ggplot2::element_text(size = 8),
        plot.margin = ggplot2::margin(t = 8, r = 8, b = 8, l = 8, unit = "pt")
      ) +
      ggplot2::guides(
        color = ggplot2::guide_legend(ncol = legend_ncol)
      )

    print(p)
  }

  # ---- Overlaid GAM pages by outcome category ----
  for (tv in names(preds_by_tv)) {

    # (1) Individual DRSP items: sig OR range >= 0.5, Y scale 1-6
    plot_overlay_page(
      sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
      tv = tv, meta = axis_meta[[tv]],
      var_list       = all_raw_drsp,
      range_thresh   = 0.5,
      title_prefix   = "Overlaid GAM Curves (Individual DRSP Items)",
      subtitle_label = "Individual DRSP items",
      ylim_strategy  = "fixed", fixed_ylim = c(1, 6),
      show_hline = TRUE, hline_y = 4,
      legend_ncol = 4,
      show_range_in_labels = TRUE
    )

    # (2) DRSP composite outcomes: sig only, 1-4 Y scale (expands to 5 or 6 if needed)
    composite_vars <- names(outcomes)[!grepl(DRSP_VAR_PATTERN, names(outcomes))]
    # Compute dynamic Y-max for composites: default 4, expand if predictions exceed
    composite_pred <- preds_by_tv[[tv]] %>% dplyr::filter(outcome_var %in% composite_vars)
    composite_ymax <- 4
    if (nrow(composite_pred) > 0) {
      max_est <- max(composite_pred$estimate, na.rm = TRUE)
      composite_ymax <- expand_composite_ymax(4, max_est)
    }
    plot_overlay_page(
      sig_tv = sig_by_tv[[tv]], pred_tv = preds_by_tv[[tv]],
      tv = tv, meta = axis_meta[[tv]],
      var_list       = composite_vars,
      range_thresh   = Inf,
      title_prefix   = "Overlaid GAM Curves (DRSP Composites)",
      subtitle_label = "DRSP composite outcomes",
      ylim_strategy  = "fixed", fixed_ylim = c(1, composite_ymax),
      show_hline = FALSE,
      legend_ncol = 3,
      show_range_in_labels = TRUE
    )
  }

  # ---- Sections per time variable: compact table + pages of plots ----
  for (tv in names(preds_by_tv)) {

    summary_df <- sig_by_tv[[tv]] %>%
      dplyr::left_join(dplyr::select(outcome_categories, variable, category), by = "variable") %>%
      dplyr::mutate(
        stars = dplyr::case_when(
          is.finite(p.value) & p.value < 0.001 ~ "***",
          is.finite(p.value) & p.value < 0.01  ~ "**",
          is.finite(p.value) & p.value < 0.05  ~ "*",
          TRUE ~ ""
        ),
        Outcome = dplyr::coalesce(outcome_label, variable)
      ) %>%
      dplyr::select(Category = category, Outcome, N = n, EDF = edf, F = F, Range = range, Sig = stars) %>%
      dplyr::arrange(Category, Outcome)

    tv_title <- grid::textGrob(paste("Significance — ID:", current_id, "| Time Variable:", tv),
                               gp = grid::gpar(fontsize = 14, fontface = "bold"))
    tv_table <- gridExtra::tableGrob(summary_df, rows = NULL, theme = table_theme)
    gridExtra::grid.arrange(tv_title, tv_table, ncol = 1,
                            heights = grid::unit(c(0.12, 0.88), "npc"))

    # Plots organized by outcome category (using page_definitions)
    combined_plot_data <- preds_by_tv[[tv]]

    for (page_name in names(page_definitions)) {
      page_outcome_labels <- page_definitions[[page_name]]
      page_data <- dplyr::filter(combined_plot_data, outcome_label %in% page_outcome_labels)
      if (nrow(page_data) == 0) next

      plot_list <- lapply(page_outcome_labels, function(lab) {
        plot_df <- dplyr::filter(page_data, outcome_label == lab)
        if (nrow(plot_df) == 0) return(NULL)
        create_annotated_plot(plot_df, tv, axis_meta, raw_data = datSX_id)
      })
      plot_list <- Filter(Negate(is.null), plot_list)

      if (length(plot_list) > 0) {
        if (length(plot_list) < max_plots_per_page) {
          plot_list <- c(plot_list,
                         rep(list(patchwork::plot_spacer()), max_plots_per_page - length(plot_list)))
        }
        page_title_label <- gsub("_", " ", page_name)
        page_title <- paste(page_title_label, "\u2014 ID:", current_id, "| Time Variable:", tv)
        final_page <- patchwork::wrap_plots(plotlist = plot_list, ncol = plots_per_col) +
          patchwork::plot_annotation(
            title = page_title,
            theme = ggplot2::theme(plot.title = ggplot2::element_text(size = 12, face = "bold"))
          )
        print(final_page)
      }
    }
  }

  # ---- Symptom Trajectories with ID-specific shared y-range ----
  # For each ID and time variable:
  #   - find the largest predicted range per y_axis_mode group (fixed / data)
  #   - add 0.25, then apply that shared total y-range to every outcome in the group
  #   - fixed-scale outcomes (DRSP individual items): clamped to [1, 6]
  #   - data-scale outcomes (DRSP composites): unclamped, data-driven bounds
  for (tv in names(preds_by_tv)) {

    combined_plot_data <- preds_by_tv[[tv]]
    if (is.null(combined_plot_data) || nrow(combined_plot_data) == 0) next

    id_range_df <- combined_plot_data %>%
      dplyr::group_by(outcome_var, outcome_label) %>%
      dplyr::summarise(
        ymin = min(estimate, na.rm = TRUE),
        ymax = max(estimate, na.rm = TRUE),
        yrange = ymax - ymin,
        .groups = "drop"
      ) %>%
      dplyr::left_join(
        dplyr::select(outcome_plot_settings, variable, y_axis_mode, y_min, y_max),
        by = c("outcome_var" = "variable")
      ) %>%
      dplyr::mutate(
        y_axis_mode = dplyr::coalesce(y_axis_mode, "fixed"),
        is_composite = !grepl(DRSP_VAR_PATTERN, outcome_var),
        lower_bound = dplyr::if_else(
          y_axis_mode == "fixed", dplyr::coalesce(y_min, 1), ymin - 0.5
        ),
        # For composites (y_max=4), expand upper_bound to 5 or 6 if predictions exceed 4
        upper_bound = dplyr::if_else(
          y_axis_mode == "fixed",
          pmax(dplyr::coalesce(y_max, 6), pmin(ceiling(ymax), 6)),
          ymax + 0.5
        )
      )

    if (nrow(id_range_df) == 0) next

    shared_range_by_mode <- id_range_df %>%
      dplyr::group_by(y_axis_mode) %>%
      dplyr::summarise(
        shared_total_range = max(yrange, na.rm = TRUE) + 0.25,
        .groups = "drop"
      ) %>%
      dplyr::mutate(
        shared_total_range = dplyr::if_else(
          y_axis_mode == "fixed",
          pmin(shared_total_range, 5),
          shared_total_range
        )
      )

    for (page_name in names(page_definitions)) {
      page_outcome_labels <- page_definitions[[page_name]]
      page_outcome_labels <- page_outcome_labels[page_outcome_labels %in% id_range_df$outcome_label]
      if (length(page_outcome_labels) == 0) next

      page_data <- dplyr::filter(combined_plot_data, outcome_label %in% page_outcome_labels)
      if (nrow(page_data) == 0) next

      plot_list_shared <- lapply(page_outcome_labels, function(lab) {
        plot_df <- dplyr::filter(page_data, outcome_label == lab)
        if (nrow(plot_df) == 0) return(NULL)

        ov        <- dplyr::first(plot_df$outcome_var)
        mode_row  <- id_range_df[id_range_df$outcome_var == ov, ]
        mode_str  <- if (nrow(mode_row) > 0) mode_row$y_axis_mode[1] else "fixed"
        lb        <- if (nrow(mode_row) > 0) mode_row$lower_bound[1] else 1
        ub        <- if (nrow(mode_row) > 0) mode_row$upper_bound[1] else 6
        rng_row   <- shared_range_by_mode[shared_range_by_mode$y_axis_mode == mode_str, ]
        str_range <- if (nrow(rng_row) > 0) rng_row$shared_total_range[1] else 5

        shared_y_limits <- make_shared_id_y_limits(
          plot_df = plot_df,
          shared_total_range = str_range,
          lower_bound = lb,
          upper_bound = ub
        )

        create_annotated_plot(
          plot_df = plot_df,
          tv = tv,
          axis_meta_list = axis_meta,
          y_limits = shared_y_limits,
          raw_data = datSX_id
        )
      })

      plot_list_shared <- Filter(Negate(is.null), plot_list_shared)

      if (length(plot_list_shared) > 0) {
        if (length(plot_list_shared) < max_plots_per_page) {
          plot_list_shared <- c(
            plot_list_shared,
            rep(list(patchwork::plot_spacer()), max_plots_per_page - length(plot_list_shared))
          )
        }

        page_title_label <- gsub("_", " ", page_name)
        page_title_shared <- paste(
          "Symptom Trajectories (ID-Scaled Y Axis) \u2014", page_title_label,
          "\u2014 ID:", current_id, "| Time Variable:", tv
        )
        page_subtitle_shared <- paste0(
          "Shared y-range per scale group (items=Likert [1-6]; composites=[1-4, expands if needed]). ",
          "Time variable: ", tv
        )

        final_page_shared <- patchwork::wrap_plots(
          plotlist = plot_list_shared,
          ncol = plots_per_col
        ) +
          patchwork::plot_annotation(
            title = page_title_shared,
            subtitle = page_subtitle_shared,
            theme = ggplot2::theme(
              plot.title = ggplot2::element_text(size = 12, face = "bold"),
              plot.subtitle = ggplot2::element_text(size = 9)
            )
          )

        print(final_page_shared)
      }
    }
  }

  grDevices::dev.off()
  message(sprintf("    --- \u2705 Saved Report: %s ---", pdf_filename))
}

# ---- Write long CSV for evaluation across IDs ----
if (length(all_ids_signif)) {
  eval_csv <- file.path(output_folder, paste0("participant_significance_summary_", current_date, ".csv"))
  write.csv(dplyr::bind_rows(all_ids_signif), eval_csv, row.names = FALSE)
  message(sprintf("Wrote summary CSV: %s", eval_csv))
}
```

# Section 5b: Overall Sample Summary Report
```{r overall-summary-report}
# =========================================================
# OVERALL SAMPLE SUMMARY REPORT
# Generates a landscape PDF summarizing the distributions and
# descriptives of cyclical symptom change across the full sample.
# =========================================================

if (length(all_ids_signif) > 0) {
  gam_summary_all <- dplyr::bind_rows(all_ids_signif)

  overall_pdf <- file.path(output_folder, paste0("Overall_Sample_Summary_", current_date, ".pdf"))
  grDevices::pdf(overall_pdf, width = 11, height = 8.5)

  # --- Classify outcomes for affective calculations ---
  # Affective raw DRSP items: all individual DRSP items EXCEPT physical symptoms
  affective_raw_drsp <- setdiff(all_raw_drsp, physical_symptoms)

  # Affective composites: composites capturing mood/affect constructs
  affective_composites <- c("drsp_NA", "drsp_distress", "drsp_labilsocsx", "drsp_lowarousal")

  # All affective outcomes (raw + composite)
  all_affective_outcomes <- c(affective_raw_drsp, affective_composites)

  # --- Per-ID summary: best result per outcome (one row per id × variable) ---
  # Use finite-safe aggregation to avoid Inf/-Inf when all values are NA.
  id_outcome_summary <- gam_summary_all %>%
    dplyr::group_by(id, variable) %>%
    dplyr::summarise(
      best_p      = ifelse(any(is.finite(p.value)), min(p.value[is.finite(p.value)]), NA_real_),
      best_range  = ifelse(any(is.finite(range)),   max(range[is.finite(range)]),     NA_real_),
      best_n      = ifelse(any(is.finite(n)),        max(n[is.finite(n)]),             NA_real_),
      sig         = any(is.finite(p.value) & p.value < 0.05, na.rm = TRUE),
      .groups     = "drop"
    )

  all_ids <- unique(gam_summary_all$id)
  n_gam_valid <- length(all_ids)
  n_total_sample <- length(unique(df$id))

  # --- KEY METRIC 1: % of GAM-valid sample with at least one symptom/composite range >= 1.5 ---
  ids_any_range15 <- id_outcome_summary %>%
    dplyr::filter(is.finite(best_range) & best_range >= 1.5) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_any_range15 <- length(ids_any_range15) / n_gam_valid * 100

  # --- KEY METRIC 2: % of GAM-valid sample with at least one affective symptom/composite range >= 1.5 ---
  ids_affective_range15 <- id_outcome_summary %>%
    dplyr::filter(variable %in% all_affective_outcomes,
                  is.finite(best_range) & best_range >= 1.5) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_affective_range15 <- length(ids_affective_range15) / n_gam_valid * 100

  # --- KEY METRIC 3: % of GAM-valid sample with at least one significant GAM (any outcome) ---
  ids_any_sig <- id_outcome_summary %>%
    dplyr::filter(sig) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_any_sig <- length(ids_any_sig) / n_gam_valid * 100

  # --- KEY METRIC 4: % of GAM-valid sample with at least one significant affective GAM ---
  ids_affective_sig <- id_outcome_summary %>%
    dplyr::filter(variable %in% all_affective_outcomes, sig) %>%
    dplyr::distinct(id) %>%
    dplyr::pull(id)
  pct_affective_sig <- length(ids_affective_sig) / n_gam_valid * 100

  # ---- PAGE 1: Overall Descriptive Summary ----
  # Per-outcome descriptive statistics across all participants
  # Percentages use n_gam_valid (participants with valid GAM data) as denominator
  outcome_desc <- id_outcome_summary %>%
    dplyr::group_by(variable) %>%
    dplyr::summarise(
      N           = dplyr::n(),
      mean_range  = mean(best_range, na.rm = TRUE),
      sd_range    = sd(best_range, na.rm = TRUE),
      median_range = median(best_range, na.rm = TRUE),
      q25_range   = quantile(best_range, 0.25, na.rm = TRUE),
      q75_range   = quantile(best_range, 0.75, na.rm = TRUE),
      min_range   = min(best_range, na.rm = TRUE),
      max_range   = max(best_range, na.rm = TRUE),
      n_sig       = sum(sig, na.rm = TRUE),
      pct_sig     = sum(sig, na.rm = TRUE) / n_gam_valid * 100,
      n_range15   = sum(is.finite(best_range) & best_range >= 1.5, na.rm = TRUE),
      pct_range15 = sum(is.finite(best_range) & best_range >= 1.5, na.rm = TRUE) / n_gam_valid * 100,
      .groups     = "drop"
    ) %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(label = dplyr::coalesce(label, variable)) %>%
    dplyr::arrange(dplyr::desc(median_range))

  # Key metrics summary table
  key_metrics <- data.frame(
    Metric = c(
      "Total Participants in Data",
      "Participants with Valid GAM Data",
      "",
      "N with \u2265 1 significant GAM (any outcome)",
      "% with \u2265 1 significant GAM (any outcome)",
      "N with \u2265 1 significant affective GAM",
      "% with \u2265 1 significant affective GAM",
      "",
      "N with \u2265 1 symptom/composite range \u2265 1.5",
      "% with \u2265 1 symptom/composite range \u2265 1.5",
      "N with \u2265 1 affective symptom/composite range \u2265 1.5",
      "% with \u2265 1 affective symptom/composite range \u2265 1.5"
    ),
    Value = c(
      as.character(n_total_sample),
      as.character(n_gam_valid),
      "",
      as.character(length(ids_any_sig)),
      sprintf("%.1f%%", pct_any_sig),
      as.character(length(ids_affective_sig)),
      sprintf("%.1f%%", pct_affective_sig),
      "",
      as.character(length(ids_any_range15)),
      sprintf("%.1f%%", pct_any_range15),
      as.character(length(ids_affective_range15)),
      sprintf("%.1f%%", pct_affective_range15)
    ),
    stringsAsFactors = FALSE
  )

  key_theme <- gridExtra::ttheme_default(
    base_size = 11,
    core    = list(fg_params = list(hjust = 0, x = 0.05)),
    colhead = list(fg_params = list(hjust = 0, x = 0.05, fontface = "bold"))
  )

  title_grob <- grid::textGrob(
    "Overall Sample Summary \u2014 Cyclical Symptom Change",
    gp = grid::gpar(fontsize = 18, fontface = "bold")
  )
  subtitle_grob <- grid::textGrob(
    paste0("N = ", n_total_sample, " total | N = ", n_gam_valid,
           " with valid GAM data (percentages use GAM-valid N as denominator)"),
    gp = grid::gpar(fontsize = 11, fontface = "italic")
  )
  key_tbl <- gridExtra::tableGrob(key_metrics, rows = NULL, theme = key_theme)
  note_grob <- grid::textGrob(
    paste0("Affective outcomes include: raw DRSP emotional/affective items (excl. physical) ",
           "and composites (Neg. Affect, Distress, Social/Affect Lability, Low Arousal). ",
           "All percentages use GAM-valid N = ", n_gam_valid, " as denominator."),
    gp = grid::gpar(fontsize = 8, fontface = "italic")
  )

  gridExtra::grid.arrange(
    title_grob, subtitle_grob, key_tbl, note_grob,
    ncol = 1,
    heights = grid::unit(c(0.10, 0.06, 0.72, 0.08), "npc")
  )

  # ---- PAGE 2: Per-Outcome Range Descriptives Table ----
  display_desc <- outcome_desc %>%
    dplyr::transmute(
      Outcome       = label,
      N             = N,
      `Median Range` = sprintf("%.2f", median_range),
      `Mean Range`  = sprintf("%.2f", mean_range),
      `SD Range`    = sprintf("%.2f", sd_range),
      `IQR`         = sprintf("%.2f-%.2f", q25_range, q75_range),
      `Min-Max`     = sprintf("%.2f-%.2f", min_range, max_range),
      `N Sig`       = n_sig,
      `% Sig`       = sprintf("%.1f%%", pct_sig),
      `N ≥1.5`  = n_range15,
      `% ≥1.5`  = sprintf("%.1f%%", pct_range15)
    )

  desc_theme <- gridExtra::ttheme_default(
    base_size = 7,
    core    = list(fg_params = list(hjust = 0, x = 0.02)),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold")),
    padding = grid::unit(c(2, 1.5), "mm")
  )

  title2 <- grid::textGrob(
    "Per-Outcome Range Descriptives (sorted by median range)",
    gp = grid::gpar(fontsize = 14, fontface = "bold")
  )
  note2 <- grid::textGrob(
    paste0("Range = max(predicted) \u2013 min(predicted) on original scale; Sig = p < .05; ",
           "% Sig and % \u22651.5 use GAM-valid N = ", n_gam_valid, " as denominator"),
    gp = grid::gpar(fontsize = 9, fontface = "italic")
  )
  desc_tbl <- gridExtra::tableGrob(display_desc, rows = NULL, theme = desc_theme)
  gridExtra::grid.arrange(
    title2, note2, desc_tbl,
    ncol = 1,
    heights = grid::unit(c(0.08, 0.05, 0.87), "npc")
  )

  # ---- PAGE 3: Distribution of Ranges (histogram faceted by outcome) ----
  # Show range distributions for all outcomes, faceted
  range_plot_data <- id_outcome_summary %>%
    dplyr::left_join(
      tibble::enframe(outcomes, name = "variable", value = "label"),
      by = "variable"
    ) %>%
    dplyr::mutate(label = dplyr::coalesce(label, variable))

  p_range_dist <- ggplot2::ggplot(range_plot_data,
                                   ggplot2::aes(x = best_range, fill = sig)) +
    ggplot2::geom_histogram(bins = 20, alpha = 0.7, position = "identity") +
    ggplot2::geom_vline(xintercept = 1.5, color = "red", linetype = "dashed",
                        linewidth = 0.6) +
    ggplot2::facet_wrap(~ label, scales = "free_y", ncol = 4) +
    ggplot2::scale_fill_manual(
      values = c("TRUE" = "#0072B2", "FALSE" = "#999999"),
      labels = c("TRUE" = "Significant", "FALSE" = "Not Significant"),
      name = NULL
    ) +
    ggplot2::labs(
      title = "Distribution of GAM Ranges Across Participants",
      subtitle = "Red dashed line = range \u2265 1.5 threshold",
      x = "Range (max \u2013 min predicted)",
      y = "Count"
    ) +
    ggplot2::theme_minimal(base_size = 8) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 12),
      strip.text = ggplot2::element_text(size = 6),
      legend.position = "bottom"
    )
  print(p_range_dist)

  # ---- PAGE 4: Per-person count of outcomes meeting range >= 1.5 ----
  ids_range15_count <- id_outcome_summary %>%
    dplyr::filter(is.finite(best_range) & best_range >= 1.5) %>%
    dplyr::group_by(id) %>%
    dplyr::summarise(n_outcomes_range15 = dplyr::n(), .groups = "drop")

  # Include IDs with zero qualifying outcomes
  ids_range15_all <- tibble::tibble(id = all_ids) %>%
    dplyr::left_join(ids_range15_count, by = "id") %>%
    dplyr::mutate(n_outcomes_range15 = dplyr::coalesce(n_outcomes_range15, 0L))

  p_count_hist <- ggplot2::ggplot(ids_range15_all,
                                   ggplot2::aes(x = n_outcomes_range15)) +
    ggplot2::geom_histogram(binwidth = 1, fill = "#0072B2", color = "white", alpha = 0.8) +
    ggplot2::labs(
      title = "Number of Outcomes with Range \u2265 1.5 per Participant",
      subtitle = paste0("GAM-valid N = ", n_gam_valid, " participants; ",
                        sprintf("%.1f%%", pct_any_range15),
                        " have at least one outcome with range \u2265 1.5"),
      x = "Number of outcomes with range \u2265 1.5",
      y = "Number of participants"
    ) +
    ggplot2::scale_x_continuous(breaks = seq(0, max(c(0, ids_range15_all$n_outcomes_range15), na.rm = TRUE))) +
    ggplot2::theme_minimal(base_size = 12) +
    ggplot2::theme(plot.title = ggplot2::element_text(face = "bold", size = 14))
  print(p_count_hist)

  # ---- PAGE 5: Median range by outcome (bar chart) ----
  p_median_bar <- ggplot2::ggplot(outcome_desc,
                                   ggplot2::aes(x = reorder(label, median_range),
                                                y = median_range)) +
    ggplot2::geom_col(fill = "#0072B2", alpha = 0.8) +
    ggplot2::geom_text(ggplot2::aes(label = sprintf("%.2f", median_range)),
                       hjust = -0.15, size = 2.8) +
    ggplot2::geom_hline(yintercept = 1.5, color = "red", linetype = "dashed",
                        linewidth = 0.6) +
    ggplot2::coord_flip() +
    ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
    ggplot2::labs(
      title = "Median GAM Range by Outcome",
      subtitle = "Red dashed line = range \u2265 1.5 threshold",
      x = NULL,
      y = "Median Range"
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(plot.title = ggplot2::element_text(face = "bold", size = 14))
  print(p_median_bar)

  # ---- PAGE 6: % Significant and % Range >= 1.5 by outcome (side-by-side bars) ----
  # Precompute ordering by % Range >= 1.5 before pivoting so both metrics
  # share a consistent outcome order (rather than ordering by the mean of both).
  pct_data <- outcome_desc %>%
    dplyr::mutate(label = factor(label, levels = label[order(pct_range15)])) %>%
    dplyr::select(label, pct_sig, pct_range15) %>%
    tidyr::pivot_longer(cols = c(pct_sig, pct_range15),
                        names_to = "metric", values_to = "pct") %>%
    dplyr::mutate(
      metric = dplyr::case_when(
        metric == "pct_sig"     ~ "% Significant (p < .05)",
        metric == "pct_range15" ~ "% Range \u2265 1.5",
        TRUE ~ metric
      ),
      pct_label = sprintf("%.1f%%", pct)
    )

  p_pct_bars <- ggplot2::ggplot(pct_data,
                                 ggplot2::aes(x = label,
                                              y = pct, fill = metric)) +
    ggplot2::geom_col(position = "dodge", alpha = 0.8) +
    ggplot2::geom_text(
      ggplot2::aes(label = pct_label),
      position = ggplot2::position_dodge(width = 0.9),
      hjust = -0.1, size = 2.5
    ) +
    ggplot2::coord_flip() +
    ggplot2::scale_y_continuous(expand = ggplot2::expansion(mult = c(0, 0.15))) +
    ggplot2::scale_fill_manual(values = c("% Significant (p < .05)" = "#0072B2",
                                          "% Range \u2265 1.5" = "#D55E00")) +
    ggplot2::labs(
      title = "Percentage of GAM-Valid Sample: Significant and Range \u2265 1.5 by Outcome",
      subtitle = paste0("Denominator: GAM-valid N = ", n_gam_valid),
      x = NULL,
      y = "Percentage",
      fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 10) +
    ggplot2::theme(
      plot.title = ggplot2::element_text(face = "bold", size = 13),
      legend.position = "bottom"
    )
  print(p_pct_bars)

  grDevices::dev.off()
  message(sprintf("Wrote overall summary report: %s", overall_pdf))
} else {
  message("No significance data available; skipping overall summary report.")
}
```

---

# Section 6: Per outcome exploration (analysis functions)

```{r exploration-analysis-functions}
# =========================================================
# EXPLORATION / THRESHOLD ANALYSIS FUNCTIONS
# These power the per-outcome exploration PDFs below.
# =========================================================

# --- Basic cleaning: add numeric significance flag, coerce types ---
clean_gam_summary <- function(df) {
  df %>%
    dplyr::mutate(
      significant     = ifelse(p.value < 0.05, "Significant", "Not Significant"),
      significant_num = ifelse(significant == "Significant", 1, 0),
      significant_f   = factor(significant_num,
                               levels = c(0, 1),
                               labels = c("Not Significant", "Significant")),
      range    = as.numeric(range),
      n        = as.numeric(n),
      variable = as.character(variable),
      id       = as.character(id)
    ) %>%
    dplyr::filter(!is.na(range), !is.na(significant_num),
                  !is.na(variable), !is.na(id))
}

# --- Filter to one outcome variable, one row per ID ---
# Aggregates across potential multiple time_var rows per id/variable.
# peak_time, trough_time, and best_time_var come from the row with the
# smallest p.value (best significance), ensuring internal consistency.
get_outcome_df <- function(df, target_var) {
  df %>%
    dplyr::filter(variable == target_var) %>%
    dplyr::group_by(id, variable) %>%
    dplyr::summarise(
      p.value         = min(p.value, na.rm = TRUE),
      range           = max(range,   na.rm = TRUE),
      n               = max(n,       na.rm = TRUE),
      significant_num = ifelse(any(significant_num == 1, na.rm = TRUE), 1, 0),
      peak_time       = peak_time[which.min(p.value)][1],
      trough_time     = trough_time[which.min(p.value)][1],
      best_time_var   = time_var[which.min(p.value)][1],
      .groups         = "drop"
    ) %>%
    dplyr::mutate(
      significant_f = factor(
        significant_num,
        levels = c(0, 1),
        labels = c("Not Significant", "Significant")
      )
    )
}

# --- Descriptive summary (overall + by significance) ---
summarize_cyclicity <- function(df_outcome) {
  overall <- df_outcome %>%
    dplyr::summarise(
      n_participants   = dplyr::n(),
      n_significant    = sum(significant_num == 1, na.rm = TRUE),
      prop_significant = mean(significant_num == 1, na.rm = TRUE)
    )
  by_sig <- df_outcome %>%
    dplyr::group_by(significant_f) %>%
    dplyr::summarise(
      n            = dplyr::n(),
      mean_range   = mean(range, na.rm = TRUE),
      sd_range     = sd(range, na.rm = TRUE),
      median_range = median(range, na.rm = TRUE),
      q25_range    = quantile(range, .25, na.rm = TRUE),
      q75_range    = quantile(range, .75, na.rm = TRUE),
      min_range    = min(range, na.rm = TRUE),
      max_range    = max(range, na.rm = TRUE),
      mean_n       = mean(n, na.rm = TRUE),
      .groups      = "drop"
    )
  list(overall = overall, by_significance = by_sig)
}

# --- Proportion significant at various range thresholds ---
threshold_table <- function(df_outcome,
                            thresholds = c(0.10, 0.25, 0.36, 0.50, 0.75, 1.00)) {
  do.call(rbind, lapply(thresholds, function(th) {
    tmp <- df_outcome %>% dplyr::filter(range >= th)
    tibble::tibble(
      threshold        = th,
      n_at_or_above    = nrow(tmp),
      prop_of_sample   = ifelse(nrow(df_outcome) > 0,
                                nrow(tmp) / nrow(df_outcome), NA_real_),
      n_significant    = sum(tmp$significant_num == 1, na.rm = TRUE),
      prop_significant = ifelse(nrow(tmp) > 0,
                                mean(tmp$significant_num == 1, na.rm = TRUE),
                                NA_real_)
    )
  }))
}

# --- Logistic regression: significance ~ range (& + n) ---
fit_logistic_models <- function(df_outcome) {
  model_range   <- glm(significant_num ~ range,     family = binomial, data = df_outcome)
  model_range_n <- glm(significant_num ~ range + n, family = binomial, data = df_outcome)
  list(
    model_range   = model_range,
    model_range_n = model_range_n,
    tidy_range    = broom::tidy(model_range,   exponentiate = TRUE),
    tidy_range_n  = broom::tidy(model_range_n, exponentiate = TRUE)
  )
}

# --- Predicted probability of significance across range values ---
make_prediction_df <- function(df_outcome, model_range, model_range_n = NULL,
                               n_value = NULL) {
  xseq <- seq(min(df_outcome$range, na.rm = TRUE),
              max(df_outcome$range, na.rm = TRUE),
              length.out = 300)
  pred1 <- tibble::tibble(range = xseq) %>%
    dplyr::mutate(pred_prob = predict(model_range, newdata = ., type = "response"),
                  model     = "Range only")
  if (!is.null(model_range_n)) {
    if (is.null(n_value)) n_value <- median(df_outcome$n, na.rm = TRUE)
    pred2 <- tibble::tibble(range = xseq, n = n_value) %>%
      dplyr::mutate(pred_prob = predict(model_range_n, newdata = ., type = "response"),
                    model     = paste0("Range + n (n fixed at ", round(n_value, 1), ")")) %>%
      dplyr::select(range, pred_prob, model)
    dplyr::bind_rows(pred1, pred2)
  } else {
    pred1
  }
}

# --- Histogram of range, colored by significance ---
plot_range_histogram <- function(df_outcome) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, fill = significant_f)) +
    ggplot2::geom_histogram(position = "identity", alpha = 0.65, bins = 30) +
    ggplot2::labs(
      title = paste0("Histogram of ", unique(df_outcome$variable), " Range by Significance"),
      x = "Cyclical range", y = "Count", fill = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- Logistic curve with jittered observed points ---
plot_significance_probability <- function(df_outcome, pred_df,
                                          thresholds = c(0.25, 0.36, 0.50),
                                          jitter_height = 0.04) {
  ggplot2::ggplot(df_outcome, ggplot2::aes(x = range, y = significant_num)) +
    ggplot2::geom_jitter(ggplot2::aes(color = significant_f),
                         width = 0, height = jitter_height,
                         alpha = 0.45, size = 2) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2, inherit.aes = FALSE) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(-0.02, 1.02), breaks = seq(0, 1, by = 0.1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title = paste0("Probability of Significant Cyclicity: ", unique(df_outcome$variable)),
      x = "Cyclical range",
      y = "Predicted probability of significant GAM",
      color = NULL, linetype = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- Binned empirical proportions ---
make_empirical_bins <- function(df_outcome, n_bins = 10) {
  df_outcome %>%
    dplyr::mutate(range_bin = dplyr::ntile(range, n_bins)) %>%
    dplyr::group_by(range_bin) %>%
    dplyr::summarise(
      mean_range       = mean(range, na.rm = TRUE),
      prop_significant = mean(significant_num, na.rm = TRUE),
      n                = dplyr::n(),
      .groups          = "drop"
    )
}

# --- Empirical bins + model overlay ---
plot_empirical_and_model <- function(df_outcome, pred_df, empirical_df,
                                     thresholds = c(0.25, 0.36, 0.50)) {
  ggplot2::ggplot() +
    ggplot2::geom_point(data = empirical_df,
                        ggplot2::aes(x = mean_range, y = prop_significant, size = n),
                        alpha = 0.8) +
    ggplot2::geom_line(data = pred_df,
                       ggplot2::aes(x = range, y = pred_prob, linetype = model),
                       linewidth = 1.2) +
    ggplot2::geom_vline(xintercept = thresholds, linetype = "dashed", alpha = 0.6) +
    ggplot2::scale_y_continuous(limits = c(0, 1),
                                labels = scales::percent_format(accuracy = 1)) +
    ggplot2::labs(
      title    = paste0("Empirical and Modeled Probability: ", unique(df_outcome$variable)),
      x        = "Cyclical range",
      y        = "Probability of significant GAM",
      size     = "Bin n",
      linetype = NULL
    ) +
    ggplot2::theme_minimal(base_size = 16)
}

# --- All-in-one analysis wrapper for one outcome variable ---
analyze_one_outcome <- function(df, target_var,
                                thresholds = c(0.25, 0.36, 0.50),
                                n_bins = 10) {
  df_clean   <- clean_gam_summary(df)
  df_outcome <- get_outcome_df(df_clean, target_var)
  if (nrow(df_outcome) == 0) stop(paste("No rows found for variable:", target_var))

  desc      <- summarize_cyclicity(df_outcome)
  thresh_tab <- threshold_table(df_outcome, thresholds = thresholds)
  mods      <- fit_logistic_models(df_outcome)
  pred_df   <- make_prediction_df(df_outcome  = df_outcome,
                                  model_range   = mods$model_range,
                                  model_range_n = mods$model_range_n)
  emp_df    <- make_empirical_bins(df_outcome, n_bins = n_bins)

  p_hist <- plot_range_histogram(df_outcome)
  p_prob <- plot_significance_probability(df_outcome, pred_df, thresholds = thresholds)
  p_emp  <- plot_empirical_and_model(df_outcome, pred_df, emp_df, thresholds = thresholds)

  list(
    data                        = df_outcome,
    descriptives_overall        = desc$overall,
    descriptives_by_significance = desc$by_significance,
    threshold_table             = thresh_tab,
    model_range                 = mods$model_range,
    model_range_n               = mods$model_range_n,
    model_range_tidy_OR         = mods$tidy_range,
    model_range_n_tidy_OR       = mods$tidy_range_n,
    predictions                 = pred_df,
    empirical_bins              = emp_df,
    plot_histogram              = p_hist,
    plot_probability            = p_prob,
    plot_empirical              = p_emp
  )
}

# --- Extract DRSP variable names from the summary data ---
get_drsp_variables <- function(df) {
  df %>%
    clean_gam_summary() %>%
    dplyr::distinct(variable) %>%
    dplyr::pull(variable) %>%
    .[grepl("^drsp", ., ignore.case = TRUE)]
}

# --- Combined overlay GAM plot for a group of participants ---
# Overlays individual GAM fitted curves on a single plot.
# fit_results_list: named list (keyed by id) of fit_outcome() results
# ids: character vector of participant IDs to include
# tv: time variable name
# axis_meta_list: axis metadata (labels, reference lines)
# outcome_label: display label for the outcome
# title_prefix: e.g., "Significant GAMs" or "Non-Significant GAMs"
plot_overlay_gam <- function(fit_results_list, ids, tv, axis_meta_list,
                             outcome_label, title_prefix,
                             outcome_var = NULL) {
  meta <- axis_meta_list[[tv]]

  overlay_data <- do.call(rbind, lapply(ids, function(qid) {
    res <- fit_results_list[[as.character(qid)]]
    if (is.null(res)) return(NULL)
    pred_df <- res$pred
    pred_df$id <- qid
    pred_df
  }))

  if (is.null(overlay_data) || nrow(overlay_data) == 0) return(NULL)

  n_people <- length(unique(overlay_data$id))

  p <- ggplot2::ggplot(overlay_data,
                       ggplot2::aes(x = .data[[tv]], y = estimate, group = id)) +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red",
                        linewidth = 0.8) +
    ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen",
                        linewidth = 0.8) +
    ggplot2::geom_line(alpha = 0.4, linewidth = 0.6, color = "#0072B2") +
    ggplot2::scale_x_continuous(
      limits = c(-1, 1),
      breaks = seq(-1, 1, by = 0.5),
      labels = meta$labels
    ) +
    ggplot2::labs(
      title    = paste0(title_prefix, ": ", outcome_label),
      subtitle = paste0("N = ", n_people, " participants | Time variable: ", tv),
      x        = "Cycle Phase",
      y        = outcome_label
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::theme(
      plot.title    = ggplot2::element_text(size = 16, face = "bold"),
      plot.subtitle = ggplot2::element_text(size = 12)
    )

  # Apply composite-specific Y-axis limits (1-4, expanding to 5 or 6 if needed)
  if (!is.null(outcome_var) && !grepl(DRSP_VAR_PATTERN, outcome_var)) {
    max_pred <- max(overlay_data$estimate, na.rm = TRUE)
    comp_ymax <- expand_composite_ymax(4, max_pred)
    p <- p +
      ggplot2::scale_y_continuous(breaks = seq(1, comp_ymax, by = 1)) +
      ggplot2::coord_cartesian(ylim = c(1, comp_ymax))
  }

  p
}

# --- Peak and trough timing distribution plot for significant participants ---
plot_peak_trough_timing <- function(df_sig, tv, axis_meta_list, outcome_label) {
  meta <- axis_meta_list[[tv]]

  tl_peak   <- df_sig %>%
    dplyr::filter(!is.na(peak_time)) %>%
    dplyr::transmute(time = peak_time, type = "Peak (Max)")
  tl_trough <- df_sig %>%
    dplyr::filter(!is.na(trough_time)) %>%
    dplyr::transmute(time = trough_time, type = "Trough (Min)")

  if (nrow(tl_peak) < 2 || nrow(tl_trough) < 2) return(NULL)

  tl_all <- dplyr::bind_rows(tl_peak, tl_trough)

  ggplot2::ggplot() +
    ggplot2::geom_density(data = tl_peak,
                          ggplot2::aes(x = time,
                                       y = ggplot2::after_stat(ndensity),
                                       fill = "Peak (Max)", color = "Peak (Max)"),
                          alpha = 0.3, adjust = 1.2) +
    ggplot2::geom_density(data = tl_trough,
                          ggplot2::aes(x = time,
                                       y = -ggplot2::after_stat(ndensity),
                                       fill = "Trough (Min)", color = "Trough (Min)"),
                          alpha = 0.3, adjust = 1.2) +
    ggplot2::geom_rug(data = tl_all, ggplot2::aes(x = time, color = type),
                      alpha = 0.5, length = grid::unit(0.03, "npc")) +
    ggplot2::geom_hline(yintercept = 0, linewidth = 0.5, color = "grey40") +
    ggplot2::geom_vline(xintercept = meta$menses, color = "red", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::geom_vline(xintercept = meta$ov, color = "forestgreen", linewidth = 0.8,
                        linetype = "dashed") +
    ggplot2::scale_x_continuous(limits = c(-1, 1), breaks = seq(-1, 1, by = 0.5),
                                labels = meta$labels) +
    ggplot2::scale_y_continuous(limits = c(-1, 1),
                                breaks = seq(-1, 1, by = 0.25),
                                labels = abs) +
    ggplot2::labs(
      title = paste("Peak & Trough Timing:", outcome_label),
      subtitle = paste0("Among participants with significant GAMs (p < .05); N = ",
                        nrow(df_sig)),
      x = "Cycle Phase",
      y = "Peaks \u2191 / Troughs \u2193 (normalized density)",
      fill = NULL, color = NULL
    ) +
    ggplot2::theme_minimal(base_size = 14) +
    ggplot2::scale_fill_manual(values = c("Peak (Max)" = "#D55E00",
                                          "Trough (Min)" = "#0072B2")) +
    ggplot2::scale_color_manual(values = c("Peak (Max)" = "#D55E00",
                                           "Trough (Min)" = "#0072B2"))
}
```

# Section 7: Per outcome reports (exploration PDF loop)

```{r generate-explore-pdfs, results='asis'}
# =========================================================
# GENERATE PER-OUTCOME EXPLORATION PDF REPORTS
# =========================================================
# For each DRSP outcome, generates a landscape PDF containing:
#   1. Summary page: descriptive stats + threshold table + peak/trough timing
#   2. Combined overlay GAM plot (significant participants only)
#   3. Combined overlay GAM plot (non-significant participants only)
#   4. Significance probability plot (logistic curve)
#   5. Histogram of range by significance
#   6. Empirical + modeled probability plot
#   Per time variable:
#   7. Peak & trough timing density (significant participants)
#   8+. Individual GAM trajectory pages (all IDs, sorted by descending range)
# =========================================================

if (length(all_ids_signif) > 0) {
  gam_summary_df <- dplyr::bind_rows(all_ids_signif)

  # Create output sub-folder for exploration PDFs
  explore_folder <- file.path(output_folder, "Outcome_Exploration_PDFs")
  dir.create(explore_folder, showWarnings = FALSE, recursive = TRUE)

  drsp_vars <- get_drsp_variables(gam_summary_df)
  all_outcome_vars <- drsp_vars
  cat("Found", length(drsp_vars), "DRSP variables to process\n")

  # Thresholds for the exploration
  explore_thresholds <- c(0.10, 0.20, 0.25, 0.36, 0.50, 0.75, 1.00)

  # Plot layout for individual GAM pages
  plots_per_col_exp <- 3
  plots_per_row_exp <- 3
  max_plots_per_page_exp <- plots_per_col_exp * plots_per_row_exp

  # Table themes for summary pages
  summary_table_theme <- gridExtra::ttheme_default(
    base_size = 11,
    core    = list(fg_params = list(hjust = 0, x = 0.05)),
    colhead = list(fg_params = list(hjust = 0, x = 0.05, fontface = "bold"))
  )
  threshold_table_theme <- gridExtra::ttheme_default(
    base_size = 10,
    core    = list(fg_params = list(hjust = 0.5, x = 0.5)),
    colhead = list(fg_params = list(hjust = 0.5, x = 0.5, fontface = "bold"))
  )

  for (target_var in all_outcome_vars) {
    cat(sprintf("\n=== Processing: %s ===\n", target_var))

    # 1. Run exploration analysis
    result <- tryCatch(
      analyze_one_outcome(gam_summary_df, target_var,
                          thresholds = explore_thresholds,
                          n_bins = 10),
      error = function(e) {
        message("  Skipping ", target_var, ": ", e$message)
        NULL
      }
    )
    if (is.null(result)) next

    # 2. Identify all IDs sorted by range descending, plus sig/nonsig splits for overlay pages
    sig_ids <- result$data %>%
      dplyr::filter(significant_num == 1) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    nonsig_ids <- result$data %>%
      dplyr::filter(significant_num == 0) %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    qualifying_ids <- result$data %>%
      dplyr::arrange(dplyr::desc(range)) %>%
      dplyr::pull(id)

    # 3. Get outcome label
    outcome_label <- if (target_var %in% names(outcomes)) {
      outcomes[[target_var]]
    } else {
      target_var
    }

    cat(sprintf("  N participants: %d | N significant: %d | N qualifying IDs: %d\n",
                result$descriptives_overall$n_participants,
                result$descriptives_overall$n_significant,
                length(qualifying_ids)))

    # 4. Pre-compute GAM fit results for all qualifying IDs (all time variables)
    all_fit_results_by_tv <- list()
    for (tv in time_vars) {
      tv_fits <- list()
      for (qid in qualifying_ids) {
        dat_id <- subset(df, as.character(id) == qid)
        if (nrow(dat_id) == 0) next
        res <- tryCatch(
          fit_outcome(dat_id, target_var, tv, outcomes),
          error = function(e) NULL
        )
        if (!is.null(res)) tv_fits[[as.character(qid)]] <- res
      }
      all_fit_results_by_tv[[tv]] <- tv_fits
    }

    # 5. Open PDF
    pdf_path <- file.path(explore_folder, paste0("Explore_", target_var, ".pdf"))
    grDevices::pdf(pdf_path, width = 11, height = 8.5)

    # ---- PAGE 1: Summary + Threshold Table ----
    title_grob <- grid::textGrob(
      paste("GAM Cyclicity Exploration:", outcome_label),
      gp = grid::gpar(fontsize = 18, fontface = "bold")
    )
    subtitle_grob <- grid::textGrob(
      paste("Variable:", target_var),
      gp = grid::gpar(fontsize = 12, fontface = "italic")
    )

    n_cpass <- sum(result$data$significant_num == 1 &
                    is.finite(result$data$range) &
                    result$data$range >= 1.5, na.rm = TRUE)

    n_total <- result$descriptives_overall$n_participants
    pct_cpass <- if (n_total > 0) n_cpass / n_total * 100 else NA_real_

    desc_df <- data.frame(
      Metric = c("Total Participants",
                 "N Significant (p < .05)",
                 "Proportion Significant",
                 "N Range \u22651.5 (p < .05 & range \u2265 1.5)",
                 "% Range \u22651.5 (of total sample)",
                 "Median Range",
                 "Mean Observations per ID"),
      Value = c(
        as.character(n_total),
        as.character(result$descriptives_overall$n_significant),
        sprintf("%.1f%%", result$descriptives_overall$prop_significant * 100),
        as.character(n_cpass),
        ifelse(is.na(pct_cpass), "N/A", sprintf("%.1f%%", pct_cpass)),
        sprintf("%.2f", median(result$data$range, na.rm = TRUE)),
        sprintf("%.1f", mean(result$data$n, na.rm = TRUE))
      ),
      stringsAsFactors = FALSE
    )

    sig_peak_data <- result$data %>%
      dplyr::filter(significant_num == 1, !is.na(peak_time), !is.na(trough_time))

    if (nrow(sig_peak_data) > 0) {
      timing_by_tv <- sig_peak_data %>%
        dplyr::group_by(best_time_var) %>%
        dplyr::summarise(
          med_peak   = median(peak_time, na.rm = TRUE),
          med_trough = median(trough_time, na.rm = TRUE),
          n_sig      = dplyr::n(),
          .groups    = "drop"
        )
      for (i in seq_len(nrow(timing_by_tv))) {
        tv_label <- timing_by_tv$best_time_var[i]
        suffix <- if (nrow(timing_by_tv) > 1) paste0(" [", tv_label, "]") else ""
        desc_df <- rbind(desc_df, data.frame(
          Metric = c(paste0("Median Peak Time (sig.)", suffix),
                      paste0("Median Trough Time (sig.)", suffix)),
          Value = c(
            sprintf("%.2f (N=%d)", timing_by_tv$med_peak[i], timing_by_tv$n_sig[i]),
            sprintf("%.2f (N=%d)", timing_by_tv$med_trough[i], timing_by_tv$n_sig[i])
          ),
          stringsAsFactors = FALSE
        ))
      }
    }

    desc_tbl <- gridExtra::tableGrob(desc_df, rows = NULL, theme = summary_table_theme)

    thresh_display <- result$threshold_table %>%
      dplyr::mutate(
        threshold        = sprintf("%.2f", threshold),
        n_at_or_above    = as.character(n_at_or_above),
        prop_of_sample   = sprintf("%.1f%%", prop_of_sample * 100),
        n_significant    = as.character(n_significant),
        prop_significant = ifelse(is.na(prop_significant), "N/A",
                                  sprintf("%.1f%%", prop_significant * 100))
      )
    names(thresh_display) <- c("Range Threshold", "N >= Threshold",
                               "% of Sample", "N Significant", "% Significant")

    thresh_tbl <- gridExtra::tableGrob(thresh_display, rows = NULL,
                                       theme = threshold_table_theme)
    thresh_header <- grid::textGrob(
      "Significance Rates at Range Thresholds",
      gp = grid::gpar(fontsize = 14, fontface = "bold")
    )
    note_grob <- if (length(qualifying_ids) > 0) {
      grid::textGrob(
        sprintf("Individual GAM pages follow for %d IDs (sorted by descending range)",
                length(qualifying_ids)),
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    } else {
      grid::textGrob(
        "No qualifying IDs for individual GAM pages",
        gp = grid::gpar(fontsize = 11, fontface = "italic")
      )
    }

    gridExtra::grid.arrange(
      title_grob, subtitle_grob, desc_tbl, thresh_header, thresh_tbl, note_grob,
      ncol = 1,
      heights = grid::unit(c(0.10, 0.05, 0.25, 0.08, 0.38, 0.08), "npc")
    )

    # ---- PAGE 2: Combined Overlay GAM - Significant Participants ----
    for (tv in time_vars) {
      p_sig_overlay <- plot_overlay_gam(
        fit_results_list = all_fit_results_by_tv[[tv]],
        ids              = sig_ids,
        tv               = tv,
        axis_meta_list   = axis_meta,
        outcome_label    = outcome_label,
        title_prefix     = "Significant GAMs (p < .05)",
        outcome_var      = target_var
      )
      if (!is.null(p_sig_overlay)) print(p_sig_overlay)
    }

    # ---- PAGE 3: Combined Overlay GAM - Non-Significant Participants ----
    for (tv in time_vars) {
      p_nonsig_overlay <- plot_overlay_gam(
        fit_results_list = all_fit_results_by_tv[[tv]],
        ids              = nonsig_ids,
        tv               = tv,
        axis_meta_list   = axis_meta,
        outcome_label    = outcome_label,
        title_prefix     = "Non-Significant GAMs (p \u2265 .05)",
        outcome_var      = target_var
      )
      if (!is.null(p_nonsig_overlay)) print(p_nonsig_overlay)
    }

    # ---- PAGE 4: Significance Probability Plot ----
    print(result$plot_probability)

    # ---- PAGE 5: Histogram ----
    print(result$plot_histogram)

    # ---- PAGE 6: Empirical + Model ----
    print(result$plot_empirical)

    # ---- PAGE 7+: Peak & Trough Timing Summary (per time variable) ----
    for (tv in time_vars) {
      sig_timing_data <- result$data %>%
        dplyr::filter(significant_num == 1, best_time_var == tv,
                      !is.na(peak_time), !is.na(trough_time))

      if (nrow(sig_timing_data) > 0) {
        p_timing <- plot_peak_trough_timing(sig_timing_data, tv, axis_meta,
                                            outcome_label)
        if (!is.null(p_timing)) print(p_timing)
      }
    }

    # ---- Individual GAM trajectory plots for all IDs (per time variable) ----
    # All IDs sorted by descending range
    for (tv in time_vars) {
      if (length(qualifying_ids) == 0) next

      all_fit_results <- all_fit_results_by_tv[[tv]]

      id_plots <- list()
      for (qid in qualifying_ids) {
        res <- all_fit_results[[as.character(qid)]]
        if (is.null(res)) next

        datSX_qid <- subset(df, as.character(id) == qid)
        p <- create_annotated_plot(res$pred, tv, axis_meta, raw_data = datSX_qid)
        if (!is.null(p)) {
          id_range <- round(res$sig$range, 2)
          id_n     <- res$sig$n
          id_pval  <- res$sig$p.value
          id_stars <- dplyr::case_when(
            is.finite(id_pval) & id_pval < 0.001 ~ "***",
            is.finite(id_pval) & id_pval < 0.01  ~ "**",
            is.finite(id_pval) & id_pval < 0.05  ~ "*",
            TRUE ~ ""
          )
          p <- p + ggplot2::labs(
            title = sprintf("ID %s %s (range=%.2f, N=%d)", qid, id_stars, id_range, id_n)
          )
          # Red title when either positional half has fewer than 5 observations
          if (isTRUE(res$sig$n_neg < 5) || isTRUE(res$sig$n_nonneg < 5)) {
            p <- p + ggplot2::theme(
              plot.title = ggplot2::element_text(size = 10, face = "bold",
                                                 hjust = 0.5, color = "red")
            )
          }
          id_plots[[length(id_plots) + 1]] <- p
        }
      }

      if (length(id_plots) > 0) {
        pages <- split(id_plots,
                       ceiling(seq_along(id_plots) / max_plots_per_page_exp))

        for (page_plots in pages) {
          if (length(page_plots) < max_plots_per_page_exp) {
            page_plots <- c(page_plots,
                           rep(list(patchwork::plot_spacer()),
                               max_plots_per_page_exp - length(page_plots)))
          }
          final_page <- patchwork::wrap_plots(plotlist = page_plots,
                                              ncol = plots_per_col_exp) +
            patchwork::plot_annotation(
              title    = paste("Individual GAMs:", outcome_label),
              subtitle = paste("Sorted by descending range | Time variable:", tv),
              theme    = ggplot2::theme(
                plot.title    = ggplot2::element_text(size = 12, face = "bold"),
                plot.subtitle = ggplot2::element_text(size = 9)
              )
            )
          print(final_page)
        }
      }
    }

    grDevices::dev.off()
    message(sprintf("  Saved: %s", pdf_path))
  }

  cat("\n=== All exploration PDFs generated ===\n")
  cat("Output folder:", explore_folder, "\n")
} else {
  message("No significance data available; skipping per-outcome exploration PDFs.")
}
```
